---
title: "TidyTemplate"
date: 2024-10-22
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(tidymodels)
library(ggrepel)
library(shiny)
library(ggtext)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2024-10-22")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

# bring in continent or region data for these countries from a package
cia_factbook <- tt$cia_factbook |>
  mutate(
    region = countrycode::countrycode(
      country,
      origin = "country.name",
      destination = "continent"
    )
  )

# list the columns in the cia_factbook data in the following comments with some example data
# this will help you remember what columns are available for plotting
# you can also use the glimpse function to see the columns
# glimpse(cia_factbook)

# Rows: 259
# Columns: 12
# $ country                 <chr> "Russia", "Canada", "United States", "China", "…
# $ area                    <dbl> 17098242, 9984670, 9826675, 9596960, 8514877, 7…
# $ birth_rate              <dbl> 11.87, 10.29, 13.42, 12.17, 14.72, 12.19, 19.89…
# $ death_rate              <dbl> 13.83, 8.31, 8.15, 7.44, 6.54, 7.07, 7.35, 7.34…
# $ infant_mortality_rate   <dbl> 7.08, 4.71, 6.17, 14.79, 19.21, 4.43, 43.19, 9.…
# $ internet_users          <dbl> 40853000, 26960000, 245000000, 389000000, 75982…
# $ life_exp_at_birth       <dbl> 70.16, 81.67, 79.56, 75.15, 73.28, 82.07, 67.80…
# $ maternal_mortality_rate <dbl> 34, 12, 21, 37, 56, 7, 200, 77, 51, 97, 540, NA…
# $ net_migration_rate      <dbl> 1.69, 5.66, 2.45, -0.32, -0.15, 5.74, -0.05, 0.…
# $ population              <dbl> 142470272, 34834841, 318892103, 1355692576, 202…
# $ population_growth_rate  <dbl> -0.03, 0.76, 0.77, 0.44, 0.80, 1.09, 1.25, 0.95…
# $ region                  <chr> "Europe", "Americas", "Americas", "Asia", "Amer…


```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

p <- cia_factbook |>
  ggplot(aes(x = internet_users, y = population, color = region)) +
  geom_point(
    # use circles instead of dots
    shape = 21
  ) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  scale_x_log10(labels = comma_format()) +
  scale_y_log10(labels = comma_format()) +
  coord_fixed() +
  expand_limits(x = 1, y = 1) +
  labs(
    title = "Internet Users vs Population by Region",
    # add a subtitle about what it means to be close or far from the dashed line
    subtitle = "Countries close to the dashed line have a similar proportion of internet users to population.",
    x = "Internet Users",
    y = "Population"
  ) +
  theme(
    legend.position = "bottom"
  )
```

# Model
Create an unsupervised clustering model on the cia_factbook data. This will add a new column to the data called cluster.

```{r Model}

# select only the numeric columns for clustering
features <- cia_factbook |>
  select(country, where(is.numeric)) |>
  na.omit() |>
  # scale the numeric data between 0 and 1
  mutate(
    across(where(is.numeric), scale),
    across(where(is.numeric), as.vector)
  )

k_test <- NbClust::NbClust(
  features |> select(-country),
  distance = "euclidean",
  min.nc = 2,
  max.nc = 9,
  method = "kmeans"
)

# create a tibble with the number of clusters we want to try
kclusts <- 
  tibble(k = 1:9) %>%
  mutate(
    kclust = map(k, ~kmeans(features |> select(-country), .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, features)
  )

# plot the clustering results
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))

# this will show the total within sum of squares for each number of clusters
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 1:9)

```

Based on the elbow plot, it looks like 2 clusters is a good choice. Let's visualize the clusters.

# Visualize Model

```{r Visualize Model}

cia_factbook_clust <- assignments |>
  filter(k == 2) |>
  transmute(country, cluster = .cluster) |>
  left_join(cia_factbook, by = "country")

# plot the number of countries by region and cluster
cia_factbook_clust |>
  count(region, cluster) |>
  ggplot(
    aes(
      x = n, 
      y = region, 
      fill = factor(cluster)
    )
  ) +
  geom_col() +
  labs(
    title = "Number of Countries by Region and Cluster",
    x = "Region",
    y = "Number of Countries",
    fill = "Cluster"
  ) +
  theme(
    legend.position = "bottom"
  )

# Using base R, create a bivariate plot colored by cluster
plot(
  cia_factbook_clust |> 
    select(-country, -cluster, -region), 
  col = cia_factbook_clust$cluster
)

cia_factbook_clust |>
  ggplot(aes(x = internet_users, y = population)) +
  geom_point(aes(color = cluster, size = area), shape = 21, alpha = 0.8) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_text_repel(aes(label = country), size = 2) +
  scale_size_continuous(labels = comma_format()) +
  expand_limits(x = 1000, y = 1000) +
  scale_x_log10(
    breaks = c(10000, 1000000, 100000000),
    labels = c("10K", "1M", "1B")
  ) +
  scale_y_log10(
    breaks = c(10000, 1000000, 100000000),
    labels = c("10K", "1M", "1B")
  ) +
  coord_fixed() +
  labs(
    title = "Internet Users vs Population by Cluster",
    x = "Internet Users",
    y = "Population",
    color = "Cluster"
  ) +
  theme(
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.margin = margin(0, 0, 0, 0),
    strip.text = element_text(size = 6, face = "bold")
  )

# create a plot comparing two other unused variables by cluster
cia_factbook_clust |>
  ggplot(aes(
    x = birth_rate,
    y = death_rate
  )) +
  geom_point(aes(color = cluster, size = population),shape = 21, alpha = 0.8) +
  geom_text_repel(aes(label = country), size = 2) +
  scale_color_manual(values = c("#161618", "#F32A53")) +
  scale_size_continuous(labels = comma_format()) +
  labs(
    title = "Countries: <span style='color:#161618'>**Haves**</span> vs <span style='color:#F32A53'>**Have Nots**</span>",
    subtitle = "Size of point represents population. Clusters determines by k-means.",
    x = "Birth Rate",
    y = "Death Rate",
    caption = "Source: CIA World Factbook  |  DataViz: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_markdown(size = 14),
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.margin = margin(0, 0, 0, 0)
  )

# plot infant mortality rate vs life expectancy at birth by cluster
cia_factbook_clust |>
  ggplot(aes(
    x = infant_mortality_rate,
    y = life_exp_at_birth
  )) +
  geom_point(aes(color = cluster, size = population), shape = 21, alpha = 0.8) +
  geom_text_repel(aes(label = country), size = 2) +
  scale_size_continuous(labels = comma_format()) +
  labs(
    title = "Infant Mortality Rate vs Life Expectancy at Birth by Cluster",
    x = "Infant Mortality Rate",
    y = "Life Expectancy at Birth",
    color = "Cluster",
    size = "Population"
  ) +
  theme(
    legend.position = "top",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.margin = margin(0, 0, 0, 0)
  )

# plot net migration rate vs population growth rate by cluster
cia_factbook_clust |>
  ggplot(aes(
    x = net_migration_rate,
    y = population_growth_rate
  )) +
  geom_point(aes(color = cluster, size = population), shape = 21, alpha = 0.8) +
  geom_text_repel(aes(label = country), size = 2) +
  scale_size_continuous(labels = comma_format()) +
  labs(
    title = "Net Migration Rate vs Population Growth Rate by Cluster",
    x = "Net Migration Rate",
    y = "Population Growth Rate",
    color = "Cluster",
    size = "Population"
  ) +
  theme(
    legend.position = "top",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.margin = margin(0, 0, 0, 0)
  )

# create the same plot as above, but filter out Syria and Lebanon
cia_factbook_clust |>
  filter(!country %in% c("Syria", "Lebanon")) |>
  ggplot(aes(
    x = net_migration_rate,
    y = population_growth_rate
  )) +
  geom_point(aes(color = cluster, size = population), shape = 21, alpha = 0.8) +
  geom_text_repel(aes(label = country), size = 2) +
  scale_size_continuous(labels = comma_format()) +
  labs(
    title = "Net Migration Rate vs Population Growth Rate by Cluster",
    x = "Net Migration Rate",
    y = "Population Growth Rate",
    color = "Cluster",
    size = "Population"
  ) +
  theme(
    legend.position = "top",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.margin = margin(0, 0, 0, 0)
  )

```


```{r ShinyApp}

# create a shiny app that creates a scatter plot of two variables and colors by cluster; provide dropdowns for the x and y axis variables and checkboxes to log scale the x and/or y axis

ui <- fluidPage(
  titlePanel("Cluster Analysis"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        "x", 
        "X-axis variable",
        selected = "population",
        choices = names(cia_factbook_clust |> select(where(is.numeric)))
      ),
      selectInput(
        "y", 
        "Y-axis variable",
        selected = "internet_users",
        choices = names(cia_factbook_clust |> select(where(is.numeric)))
      ),
      checkboxInput("log_x", "Log scale X-axis", value = TRUE),
      checkboxInput("log_y", "Log scale Y-axis", value = TRUE)
    ),
    mainPanel(
      plotOutput("plot")
    )
  )
)

server <- function(input, output) {
  output$plot <- renderPlot({
    label_x = str_to_title(str_replace_all(input$x, "_", " "))
    label_y = str_to_title(str_replace_all(input$y, "_", " "))
    
    p <- cia_factbook_clust |>
      ggplot(aes(
        x = !!sym(input$x),
        y = !!sym(input$y),
        color = factor(cluster)
      )) +
      geom_point(
        aes(size = population), 
        shape = 21, 
        alpha = 0.8
      ) +
      geom_text_repel(aes(label = country), size = 2) +
      scale_size_continuous(labels = comma_format()) +
      scale_color_manual(values = c("#161618", "#F32A53")) +
      labs(
        # using the ggtext package make a title about the Haves (black) and Have Nots (red)
        title = "Countries: <span style='color:#161618'>**Haves**</span> vs <span style='color:#F32A53'>**Have Nots**</span>",
        subtitle = paste(label_x, "vs", label_y, "by k-means Cluster. Size of point represents population."),
        x = label_x,
        y = label_y,
        caption = "Source: CIA World Factbook  |  DataViz: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
      ) +
      theme(
        legend.position = "none",
        legend.margin = margin(0, 0, 0, 0),
        plot.title = element_markdown(size = 14),
      )
    
    if (input$log_x) {
      p <- p + scale_x_log10(
        labels = comma_format()
      )
    }
    
    if (input$log_y) {
      p <- p + scale_y_log10(
        labels = comma_format()
      )
    }
    
    print(p)
  })
}

shinyApp(ui = ui, server = server)

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2024_10_22_tidy_tuesday_cia_factbook.png",
  device = "png"
)
```
