---
title: "Rocks from Space: Mapping 45,000 Meteorite Landings Across Earth"
author: "Tony Galvan"
date: "6/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Every year, thousands of meteorites survive the fiery plunge through Earth's atmosphere and land on the surface. The Meteoritical Society has cataloged over 45,000 of these cosmic visitors — recording their mass, classification, and landing coordinates. This dataset lets us explore where meteorites fall, what they're made of, and how our ability to find them has changed over the centuries.

## Loading and Preparing the Data

We'll filter out records missing year or mass data and create a decade variable for temporal analysis.

```{r}
library(tidyverse)
theme_set(theme_light())

meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv") |>
  filter(!is.na(year) & !is.na(mass)) |>
  mutate(decade = 10 * (year %/% 10))
```

## When Were Meteorites Found?

Let's look at the temporal distribution of meteorite discoveries. The pattern tells us more about human search efforts than about actual fall rates.

```{r}
meteorites |>
  ggplot(aes(decade)) +
  geom_histogram(binwidth = 30)
```

There are some meteorites found a long time ago, but most were found after 1800. The explosion of discoveries in recent decades reflects systematic search programs in deserts and Antarctica, where meteorites are easier to spot against uniform backgrounds.

## Most Common Meteorite Classes

What classes of meteorites are found the most? The classification system groups meteorites by their mineral composition and structure.

```{r}
meteorites |>
  mutate(class = fct_lump(class, 24)) |>
  count(class, sort = TRUE) |>
  mutate(class = fct_reorder(class, n)) |>
  ggplot(aes(class, n, fill = class)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  labs(x = "Class",
       y = "# of meteorites",
       title = "Top meteorite classes",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: Meteoritical Society")
```

L5 and H5 chondrites dominate — these are "ordinary chondrites," the most common type of stony meteorite. They represent the building blocks of the early solar system.

## Mass Distribution

Meteorite masses span an enormous range — from tiny fragments to multi-ton behemoths. A log scale reveals the true distribution.

```{r}
meteorites |>
  mutate(log_mass = log10(mass)) |>
  ggplot(aes(log_mass)) +
  geom_histogram()
```

The distribution is roughly normal on a log scale, centered around 100 grams. Most meteorites are small enough to hold in your hand.

## Global Map of Meteorite Landings by Class

Where on Earth have meteorites been found? Let's map them all, faceted by classification, to see if different types cluster in different regions.

```{r}
meteorites |>
  mutate(class = fct_lump(class, 24)) |>
  ggplot(aes(long, lat, size = mass, color = class)) + 
  geom_point(alpha = 0.5) + 
  coord_map() + 
  theme_void() + 
  theme(legend.position = "none") +
  facet_wrap(~class)
```

The clustering patterns are fascinating — many finds concentrate in North Africa, the Middle East, and Antarctica, where dry conditions preserve meteorites and sparse vegetation makes them visible. The geographic distribution tells us as much about where humans look as where meteorites land.

## Mass Statistics by Decade

Let's summarize how meteorite finds have changed over time in terms of quantity and size.

```{r}
meteorites |>
  filter(!is.na(mass) & mass > 0) |>
  group_by(decade) |>
  summarise(n = n(),
            average = mean(mass),
            heaviest = max(mass),
            lightest = min(mass))
```

Recent decades show dramatically more finds but smaller average masses — consistent with systematic search programs that find many small specimens rather than relying on chance discoveries of large ones.
