---
title: "TidyTemplate"
date: 2023-08-22
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(camcorder)
library(ggtext)
library(showtext)
library(tidytext)

# Font selected from http://fonts.google.com
tt_family <- "archivo"
font_add_google("Archivo Black", tt_family)

theme_set(theme_light(base_size = 32))

font_add(family = "fa-brands", regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid", regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")
showtext_auto()

# Colors selected using https://coolors.co/
tt_source <- "UNHCR Refugee Data Finder"
bg_color <- "#FFFFFF"
txt_color <- "#000000"
dark_blue <- "#18375F"
blue <- "#0271BC"
light_blue <- "#8EBEFF"
aqua <- "#00B398"
red <- "#EF4A61"


tt_caption <- paste0("DataViz: Tony Galvan #TidyTuesday<span style='color:", bg_color, ";'>..</span><span style='font-family:fa-solid;color:", txt_color, ";'>&#xf0ce;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>", tt_source, "</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf099;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>@GDataScience1</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf09b;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>GDataScience</span><span style='color:", bg_color, ";'>..</span>")

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-08-22")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

refugees <- refugees::population |>
  summarize(
    total_refugees = sum(refugees, na.rm = TRUE),
    total_asylum_seekers = sum(asylum_seekers, na.rm = TRUE),
    total_oip = sum(oip, na.rm = TRUE),
    .by = year
  ) |>
  left_join(
    refugees::idmc |>
      summarize(
        total_idmc = sum(total, na.rm = TRUE),
        .by = year
      ),
    by = "year"
  ) |>
  left_join(
    refugees::unrwa |>
      summarize(
        total_unrwa = sum(total, na.rm = TRUE),
        .by = year
      ),
    by = "year"
  ) |> 
  mutate(
    total_idmc = coalesce(total_idmc, 0),
    total_unrwa = coalesce(total_unrwa, 0)
  ) |>
  pivot_longer(
    cols = -year,
    names_to = "population_type",
    values_to = "total_people"
  ) |> 
  filter(total_people > 0) |>
  mutate(
    population_type = case_when(
      population_type == "total_refugees" ~ "Refugees (under UNHCR's mandate)",
      population_type == "total_idmc" ~ "Internally Displaced People",
      population_type == "total_oip" ~ "Other people in need of international protection",
      population_type == "total_unrwa" ~ "Palestine refugees (under UNRWA's mandate)",
      TRUE ~ "Asylum-seekers"
    ),
    population_type = fct_relevel(
      population_type, 
      "Other people in need of international protection",
      "Asylum-seekers",
      "Internally Displaced People",
      "Palestine refugees (under UNRWA's mandate)"
    )
  )

```


# Start recording

Use the {camcorder} package to record all plots that are output to the console

```{r camcorder}

# start recording
gg_record(
  dir = "~/Downloads/camcorder/2023_08_22_tidy_tuesday_refugees", # where to save the recording
  device = "png", # device to use to save images
  width = 6, # width of saved image
  height = 6, # height of saved image
  units = "in", # units for width and height
  dpi = 300 # dpi to use when saving image
)

```

# Theme

Set up a custom data visualization theme

```{r VizTheme}

tt_theme <- function() {
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    plot.background = element_rect(fill = bg_color),
    panel.background = element_rect(fill = bg_color),
    legend.background = element_blank(),
    legend.key = element_rect(fill = bg_color),
    strip.background = element_rect(fill = bg_color), 
    plot.title = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      face = "bold",
      size = 72,
      margin = margin(t = 0, r = 0, b = 0, l = 0)
    ),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 50,
      lineheight = 0.3,
      margin = margin(t = 0, r = 0, b = 4, l = 0)
    ),
    plot.caption = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 20,
      margin = margin(t = 10, r = 0, b = 0, l = 0)
    ),
    plot.caption.position = "plot",
    strip.text = element_text(
      family = tt_family,
      color = txt_color,
      face = "bold",
      size = 36, 
      lineheight = 0.3,
      margin = margin(t = 0, r = 0, b = 0, l = 0)
    ),
    legend.position = c(-0.05, 1.1),
    legend.justification = c(0, 1.1),
    legend.direction = "vertical",
    legend.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    legend.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    axis.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    axis.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
  )
}

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

refugees |>
  ggplot(aes(year, total_people, fill = population_type)) +
  geom_col(position = "stack") +
  scale_y_continuous(
    labels = comma_format(
      scale = 1/1000000,
      suffix = "M"
    )
  ) +
  scale_fill_manual(
    values = c(
      aqua, 
      red, 
      light_blue,
      dark_blue,
      blue
    )
  ) +
  tt_theme() +
  guides(
    fill = guide_legend(
      reverse = TRUE
    )
  ) +
  labs(
    x = NULL,
    y = "# of people",
    fill = NULL,
    title = "FORCIBLY DISPLACED PEOPLE",
    subtitle = "Over 100M people forcibly displaced in 2022",
    caption = tt_caption
  )
  
```

# Backcasting

```{r}

idp <- refugees |>
  filter(population_type == "Internally Displaced People") |>
  select(-population_type) 

idp_rev <- idp |>
  mutate(
    rev_year = rev(year),
    ds = ymd(paste0(as.character(rev_year), "-12-31"))
  ) |>
  transmute(
    ds,
    y = total_people
  )

idp |>
  ggplot(aes(year, total_people)) +
  geom_line(
    color = blue,
    linewidth = 1.5
  ) +
  theme_light()

library(prophet)

fit_idp_backcast <- prophet(idp_rev)

past_idp <- make_future_dataframe(fit_idp_backcast, periods = 38, freq = 'year')

backcast_idp <- predict(fit_idp_backcast, past_idp)

idp_backcast <- idp_rev |>
  full_join(backcast_idp, by = "ds") |>
  arrange(ds) |> 
  transmute(
    id = row_number(),
    year = 2022 - id + 1,
    total_people = coalesce(y, yhat),
    low_people = if_else(yhat_lower <= 0, 0, yhat_lower),
    upr_people = if_else(yhat_upper <= 0, 0, yhat_upper),
    color = if_else(is.na(y), "Backcast", "Actual")
  )

idp_backcast |>
  ggplot(aes(year, total_people, color = color)) +
  geom_ribbon(
    data = idp_backcast |> filter(color == "Backcast"),
    aes(
      x = year,
      ymin = low_people,
      ymax = upr_people
    ),
    color = NA,
    fill = "gray75",
    alpha = 0.8
  ) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = c(blue, red)) +
  scale_y_continuous(
    labels = comma_format(
      scale = 1/1000000,
      suffix = "M"
    )
  ) +
  tt_theme() +
  theme(legend.position = "top",
        legend.direction = "horizontal") +
  labs(
    x = NULL,
    y = "# of people",
    color = "",
    title = "Backcasting in R",
    subtitle = "Using the {prophet} package",
    caption = "Source: UNHCR Refugee Data Finder"
  )

```

# Add Backcast to original plot

```{r}

refugees |>
  bind_rows(
    idp_backcast |>
      filter(color == "Backcast") |>
      transmute(
        year,
        population_type = "Internally Displaced People (Backcast)",
        total_people
      )
  ) |>
  mutate(
    population_type = fct_relevel(
      population_type, 
      "Other people in need of international protection",
      "Asylum-seekers",
      "Internally Displaced People (Backcast)",
      "Internally Displaced People",
      "Palestine refugees (under UNRWA's mandate)"
    )
  ) |>
  ggplot(aes(year, total_people, fill = population_type)) +
  geom_col(position = "stack") +
  scale_y_continuous(
    labels = comma_format(
      scale = 1/1000000,
      suffix = "M"
    )
  ) +
  scale_fill_manual(
    values = c(
      aqua, 
      red, 
      "gray",
      light_blue,
      dark_blue,
      blue
    )
  ) +
  tt_theme() +
  guides(
    fill = guide_legend(
      reverse = TRUE
    )
  ) +
  labs(
    x = NULL,
    y = "# of people",
    fill = NULL,
    title = "FORCIBLY DISPLACED PEOPLE",
    subtitle = "Over 100M people forcibly displaced in 2022",
    caption = tt_caption
  )

```



# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2023_08_22_tidy_tuesday_refugees.png",
  device = "png")

gg_stop_recording()

gg_playback(
  name = "2023_08_22_tidy_tuesday_refugees.gif",
  first_image_duration = 4,
  last_image_duration = 20,
  frame_duration = 0.5
)

```
