---
title: "1,990 Documented Catholic Miracles: What the Church's Own Records Reveal"
date: "2026-08-24"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 6, fig.path = "outputs/"
)
```

The Catholic Church maintains an extensive — and surprisingly rigorous — apparatus for investigating miraculous claims. The Lourdes Medical Bureau has examined over 7,000 reported cures since 1858 and recognized only 72. Marian apparition claims since 1900 number in the hundreds, yet fewer than 25 received positive rulings. This is not an institution that approves everything.

But what does the full picture look like? By scraping three primary source websites — [MiracleHunter.com](https://www.miraclehunter.com), the [Carlo Acutis Eucharistic Miracles Exhibition](https://www.miracolieucaristici.org), and the [Official Lourdes Sanctuary](https://www.lourdes-france.com) — we compiled a dataset of **1,990 documented miraculous events** spanning 6 categories and nearly two millennia. Every entry is programmatically extracted from its source. Let's see what patterns emerge.

## Loading the Data

```{r load-libraries}
library(tidyverse)
library(scales)
library(showtext)
library(ggtext)

font_add_google("Source Sans 3", "source_sans")
showtext_auto()
showtext_opts(dpi = 300)

theme_set(theme_minimal(base_size = 14, base_family = "source_sans") +
          theme(
            plot.title.position = "plot",
            plot.title = element_text(face = "bold", size = 18),
            plot.subtitle = element_text(size = 13, color = "gray30"),
            panel.grid.minor = element_blank()
          ))
```

```{r load-data}
miracles <- read_csv("catholic_miracles.csv", show_col_types = FALSE)

cat(sprintf("Observations: %d\nColumns: %d\nCategories: %d\nCountries: %d\nYear range: %d – %d\n",
            nrow(miracles),
            ncol(miracles),
            n_distinct(miracles$category),
            n_distinct(miracles$country, na.rm = TRUE),
            min(miracles$year, na.rm = TRUE),
            max(miracles$year, na.rm = TRUE)))
```

## Profiling the Dataset

```{r eda-glimpse}
glimpse(miracles)
```

```{r eda-missing}
miracles |>
  summarise(across(everything(), ~ sum(is.na(.) | . == ""))) |>
  pivot_longer(everything(), names_to = "column", values_to = "missing") |>
  mutate(pct_missing = missing / nrow(miracles) * 100) |>
  arrange(desc(missing)) |>
  filter(missing > 0)
```

The `title` column is mostly NA because only Marian apparitions and miraculous images have devotional titles (like "Our Lady of Guadalupe"). The `person_involved` field is empty for Eucharistic miracles where no individual is named — the miracle pertains to the host itself.

## Category Distribution

```{r eda-category-counts}
miracles |>
  count(category, sort = TRUE) |>
  mutate(
    pct = n / sum(n),
    category = fct_reorder(category, n)
  ) |>
  ggplot(aes(x = category, y = n, fill = category)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = sprintf("%d (%.0f%%)", n, pct * 100)), hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.25))) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Marian apparitions dominate the dataset",
    subtitle = "1,990 events scraped from 3 primary source websites",
    x = NULL, y = "Number of documented events"
  ) +
  theme(panel.grid.major.y = element_blank())
```

Marian apparitions account for 74% of the dataset — partly because miraclehunter.com's apparitions list is the most comprehensive, including uninvestigated claims alongside approved events.

## Geographic Distribution

```{r eda-geography}
miracles |>
  filter(!is.na(country)) |>
  count(country, sort = TRUE) |>
  slice_head(n = 15) |>
  mutate(country = fct_reorder(country, n)) |>
  ggplot(aes(x = country, y = n)) +
  geom_col(fill = "#2E5090") +
  geom_text(aes(label = n), hjust = -0.1, size = 3.5) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Italy and France dominate miracle geography",
    subtitle = "Top 15 countries by number of documented events",
    x = NULL, y = "Number of events"
  ) +
  theme(panel.grid.major.y = element_blank())
```

```{r eda-geography-by-category}
miracles |>
  filter(!is.na(country)) |>
  filter(country %in% (miracles |> count(country, sort = TRUE) |> slice_head(n = 8) |> pull(country))) |>
  count(country, category) |>
  ggplot(aes(x = fct_reorder(country, -n, .fun = sum), y = n, fill = category)) +
  geom_col() +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Different countries specialize in different miracle types",
    subtitle = "Italy leads in Eucharistic miracles and incorruptibles; France in Lourdes cures",
    x = NULL, y = "Count", fill = "Category"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Temporal Patterns

```{r eda-timeline}
miracles |>
  count(century) |>
  ggplot(aes(x = century, y = n)) +
  geom_col(fill = "#6B4226", width = 0.8) +
  geom_text(aes(label = n), vjust = -0.3, size = 3) +
  scale_x_continuous(breaks = seq(1, 21, 2)) +
  labs(
    title = "Documented miracles by century",
    subtitle = "A massive spike in the 19th-20th centuries reflects improved documentation, not increased frequency",
    x = "Century", y = "Number of documented events"
  ) +
  theme(panel.grid.major.x = element_blank())
```

```{r eda-timeline-by-category}
miracles |>
  count(century, category) |>
  ggplot(aes(x = century, y = n, fill = category)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1, 21, 2)) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Which categories drive each century's count?",
    subtitle = "Marian apparition claims explode after 1900; Eucharistic miracles are more evenly distributed",
    x = "Century", y = "Count", fill = "Category"
  ) +
  theme(panel.grid.major.x = element_blank())
```

## The Church's Approval Rate

This is the most interesting analytical question. The dataset includes both approved and rejected/uninvestigated claims — which lets us compute actual approval rates.

```{r eda-approval-status}
miracles |>
  filter(!is.na(approval_status)) |>
  mutate(
    status_group = case_when(
      str_detect(approval_status, "(?i)approved|medically inexplicable") ~ "Approved",
      str_detect(approval_status, "(?i)negative|not established") ~ "Rejected",
      str_detect(approval_status, "(?i)no decision|no investigation") ~ "No Decision",
      str_detect(approval_status, "(?i)under investigation") ~ "Under Investigation",
      TRUE ~ "Other"
    )
  ) |>
  count(status_group, sort = TRUE) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = fct_reorder(status_group, n), y = n, fill = status_group)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = sprintf("%d (%.0f%%)", n, pct * 100)), hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    "Approved" = "#2E5090", "Rejected" = "#d73027",
    "No Decision" = "#CCCCCC", "Under Investigation" = "#fdae61", "Other" = "#999999"
  )) +
  labs(
    title = "Most modern claims receive no formal ruling",
    subtitle = "Approval status distribution across all 1,990 events",
    x = NULL, y = "Number of events"
  ) +
  theme(panel.grid.major.y = element_blank())
```

```{r eda-approval-by-category}
miracles |>
  filter(!is.na(approval_status)) |>
  mutate(
    approved = str_detect(approval_status, "(?i)approved|medically inexplicable|traditional"),
    category = fct_reorder(category, approved, .fun = mean)
  ) |>
  group_by(category) |>
  summarise(
    n = n(),
    n_approved = sum(approved),
    approval_rate = mean(approved),
    .groups = "drop"
  ) |>
  ggplot(aes(x = category, y = approval_rate, fill = category)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = sprintf("%.0f%%\n(%d/%d)", approval_rate * 100, n_approved, n)),
            vjust = -0.1, size = 3.5, lineheight = 0.9) +
  scale_y_continuous(labels = percent, limits = c(0, 1.15)) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Approval rates vary wildly by category",
    subtitle = "Lourdes and incorruptibles have near-100% approval (pre-filtered);\nMarian apparitions show the Church's selectivity",
    x = NULL, y = "Approval Rate"
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

**Key insight:** The Marian apparitions approval rate (~50%) masks a crucial difference. The historical/traditional apparitions (pre-1900) are overwhelmingly approved because they only entered the record after centuries of veneration. The *modern* claims since 1900 — where we have the full pipeline of claim → investigation → ruling — show a rejection rate over 75%.

## Deep Dive: Marian Apparitions Since 1900

```{r marian-modern-approval}
miracles |>
  filter(category == "Marian Apparition", year >= 1900) |>
  filter(!is.na(approval_status)) |>
  mutate(
    ruling = case_when(
      str_detect(approval_status, "(?i)negative|not established") ~ "Rejected",
      str_detect(approval_status, "(?i)no decision|no investigation") ~ "No Formal Ruling",
      str_detect(approval_status, "(?i)approved") ~ "Approved",
      str_detect(approval_status, "(?i)under") ~ "Under Investigation",
      TRUE ~ "Other"
    )
  ) |>
  count(ruling, sort = TRUE) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = fct_reorder(ruling, n), y = n, fill = ruling)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = sprintf("%d (%.0f%%)", n, pct * 100)), hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    "Approved" = "#2E5090", "Rejected" = "#d73027",
    "No Formal Ruling" = "#CCCCCC", "Under Investigation" = "#fdae61", "Other" = "#999999"
  )) +
  labs(
    title = "Since 1900: most apparition claims get no ruling at all",
    subtitle = "Of 808 claims since 1900, the Church has explicitly approved fewer than 20",
    x = NULL, y = "Count"
  ) +
  theme(panel.grid.major.y = element_blank())
```

## Deep Dive: Eucharistic Miracles Across Centuries

```{r eucharistic-timeline}
miracles |>
  filter(category == "Eucharistic Miracle") |>
  count(century) |>
  ggplot(aes(x = century, y = n)) +
  geom_col(fill = "#B2182B", width = 0.8) +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +
  scale_x_continuous(breaks = 1:21) +
  labs(
    title = "Eucharistic miracles peaked in the 13th-16th centuries",
    subtitle = "147 documented events from the Carlo Acutis catalog and MiracleHunter.com",
    x = "Century", y = "Count"
  ) +
  theme(panel.grid.major.x = element_blank())
```

```{r eucharistic-geography}
miracles |>
  filter(category == "Eucharistic Miracle", !is.na(country)) |>
  count(country, sort = TRUE) |>
  slice_head(n = 10) |>
  mutate(country = fct_reorder(country, n)) |>
  ggplot(aes(x = country, y = n)) +
  geom_col(fill = "#B2182B") +
  geom_text(aes(label = n), hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Italy and Spain lead in Eucharistic miracles",
    subtitle = "Concentration in Catholic heartland regions with dense parish networks",
    x = NULL, y = "Count"
  ) +
  theme(panel.grid.major.y = element_blank())
```

## Key Findings

### Finding 1: The Church rejects the vast majority of modern claims

Of 808 Marian apparition claims since 1900, the overwhelming majority receive either "No decision" or an explicit negative ruling. Fewer than 3% are formally approved. This is not a credulous institution.

### Finding 2: Geography reflects Church infrastructure, not miracle frequency

Italy leads in 4 of 6 categories — not because Italy is more miraculous, but because it has the densest Catholic parish network, the most preserved relics, and the longest continuous institutional history.

### Finding 3: The documentation explosion is modern

The 19th-20th century spike in recorded events reflects the formalization of investigation processes (Lourdes Medical Bureau 1883, codified canonization 1588, modern diocesan commissions) — not an increase in supernatural activity.

### Finding 4: Different categories have very different evidentiary profiles

Lourdes cures are medically verified with modern diagnostics. Eucharistic miracles range from medieval chronicle entries to 21st-century lab-tested tissue samples. Marian apparitions span credible multi-witness events to single-person unverified claims. The dataset captures this full spectrum.

## Data Sources

- **[MiracleHunter.com](https://www.miraclehunter.com)** — Comprehensive catalog of Marian apparitions (approved + unapproved), Eucharistic miracles, stigmata, incorruptibles, and miraculous images. Maintained by Michael O'Neill, author of *Exploring the Miraculous* (OSV, 2015).
- **[miracolieucaristici.org](https://www.miracolieucaristici.org)** — Carlo Acutis / Nicola Gori International Exhibition of Eucharistic Miracles. Cataloging 136+ Eucharistic miracles worldwide.
- **[lourdes-france.com](https://www.lourdes-france.com/en/miraculous-healings/)** — Official Sanctuary of Our Lady of Lourdes. List of 72 recognized cures declared medically inexplicable by the CMIL (Comité Médical International de Lourdes).
