---
title: "TidyTemplate"
date: 2021-11-09
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)

theme_set(theme_light())

tt_caption <- "Source: afrilearndata  |  Analysis: Tony Galvan (@GDataScience1)"

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

#tt <- tt_load("2021-11-09")

#remotes::install_github("afrimapr/afrilearndata")
    
library(afrilearndata)

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

library(afrilearndata)
library(sf)

# polygons
plot(sf::st_geometry(africountries))

```

```{r}

# lines
plot(sf::st_geometry(afrihighway))

```


```{r}

# points
plot(sf::st_geometry(africapitals))

```



```{r}

library(afrilearndata)

# install.packages("tmap") # if not already installed
library(tmap)

tm_shape(africountries) +
     tm_borders("grey", lwd = .5) +
     tm_text("name_fr", auto.placement=FALSE, remove.overlap=FALSE, just='centre', col='red4', size=0.7 ) 

```


```{r}

# points
plot(sf::st_geometry(afriairports))

```


```{r}

#install.packages("mapview") # if not already installed

library(mapview)
mapview::mapview(africountries, zcol="name")
#here to show all airports on the continent
mapview(afriairports, zcol='type', label='name', cex=2)


```


```{r}

str(afriairports)

```

```{r}

glimpse(afriairports)

```


```{r}

df_afriairports <- tibble(afriairports)

glimpse(df_afriairports)

```

```{r}

df_afriairports %>%
  count(type) %>%
  mutate(type = str_to_title(str_remove(type, "_airport"))) %>%
  ggplot(aes(n, type, fill = type)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("#BA2D0B", "#AB92BF", "#3587A4")) +
  scale_x_log10() +
  labs(x = "# of airports",
       y = "",
       title = "African Airports",
       caption = tt_caption)

```


```{r}

library(ggtext)

df_afriairports %>%
  ggplot() +
  borders("world", fill = "grey95", xlim = c(-30, 60), ylim = c(-40, 35)) +
  geom_sf(aes(geometry = geometry, 
              color = type, 
              size = fct_rev(type)),
          alpha = 0.3) +
  scale_color_manual(values = c("#BA2D0B", "#AB92BF", "#3587A4")) +
  scale_size_manual(values = c(0.5, 5, 10)) +
  theme_void() +
  theme(legend.position = "none",
        plot.subtitle = element_markdown()) +
  labs(title = "African Airports",
       subtitle = "<b style='color:#3587A4'>Small</b> airports are bunched up in the southern part of the continent<br><b style='color:#BA2D0B'>Large</b> airports are mostly by the coasts<br><b style='color:#AB92BF'>Medium</b> airports are spread all over <br><i style='font-size:8pt'>**size of marks corresponds to size of airport</i>",
       color = "",
       caption = tt_caption)

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2021_11_09_tidy_tuesday.png",
  device = "png")

```
