---
title: "TidyTemplate"
date: 2022-09-07
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(rayshader)

theme_set(theme_light())

tt_caption <- "Source: rebrickable  |  DataViz: Tony Galvan (@GDataScience1) #TidyTuesday"

```

# Load the weekly Data


```{r Load}

inventories <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/inventories.csv.gz')
inventory_parts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/inventory_parts.csv.gz')
sets <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/sets.csv.gz')
themes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/themes.csv.gz')
colors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/colors.csv.gz')

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

glimpse(inventories)
glimpse(inventory_parts)
glimpse(sets)
glimpse(themes)
glimpse(colors)

```

# Wrangle

Explore the data and process it into a nice format for plotting!

```{r Wrangle}

legos <- sets %>%
  inner_join(themes %>%
               select(id, theme_name = name), 
             by = c(theme_id = "id")) %>%
  inner_join(inventories %>%
               group_by(set_num) %>%
               filter(version == max(version)) %>%
               ungroup() %>%
               rename(inventory_id = id),
             by = "set_num") %>%
  inner_join(inventory_parts, by = "inventory_id", suffix = c("", "_inventory")) %>%
  inner_join(colors %>%
              rename(color = name), by = c(color_id = "id")) %>%
  mutate(rgb = paste0("#", rgb))

```

# Rayshader example

Copied from Joris Chau's GitHub - https://github.com/JorisChau/tidytuesday/blob/master/2022/wk36/lego.R

```{r Example}

## data
sets <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/sets.csv.gz')
sets1 <- subset(sets, num_parts > 1) 

## 2d histogram
gg <- ggplot(sets1, aes(x = num_parts, y = year)) + 
  geom_bin2d(bins = 30, color = NA) +
  coord_fixed(ratio = 1 / 19) +
  labs(x = "Number of parts per set", y = "Year", title = "Evolution of LEGO number of parts per set") + 
  scale_fill_gradientn(
    colours = c("#6c98c9", "#0A69AE", "#328349", "#A5BC45", "#E4CD9E", "#F2CD37", "#C91A09"),
    values = scales::rescale(c(0, 0.05, 0.1, 0.2, 0.35, 0.5, 1)),
    guide = guide_colorbar(title = "Count", barwidth = 0.5, barheight = 2, nbin = 5, label.position = "left")) +
  scale_x_log10()

gg <- gg + theme(    
  text = element_text(family = "Palatino"),
  axis.title.y = element_text(size = 6),
  axis.title.x = element_text(size = 6),
  axis.text.y = element_text(size = 5),
  axis.text.x = element_text(size = 5),
  plot.title = element_text(size = 10),
  plot.background = element_rect(color = NA, fill = "#ffffff"),
  panel.background = element_rect(color = NA, fill = "#ffffff"),
  plot.margin = margin(t = 0.5, r = 0.5, l = 0.75, b = 0.5, unit = "cm"),
  panel.grid.major = element_line(color = "#efefef", size= 0.25),
  panel.grid.minor = element_blank(),
  legend.title = element_text(size = 6),
  legend.text = element_text(size = 5)
)

## 3d image
rayshader::plot_gg(
  gg,
  multicore = TRUE,
  shadow_intensity = 0.4,
  width = 4,
  height = 4,
  scale = 50,
  preview = TRUE,
  raytrace = TRUE,
  triangulate = FALSE,
  offset_edges = TRUE
)
Sys.sleep(0.2)
render_snapshot(clear = TRUE)

```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

p <- legos %>%
  count(year, color, rgb, wt = quantity, name = "color_total") %>%
  add_count(year, wt = color_total, name = "year_total") %>%
  group_by(year) %>%
  top_n(5, wt = color_total) %>%
  ungroup() %>%
  mutate(pct_total = color_total / year_total) %>%
  ggplot(aes(factor(year), pct_total, fill = I(rgb))) +
  geom_col(color = "gray75") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010, 2020)) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        plot.caption = element_text(size = 5),
        text = element_text(family = "Tahoma")) +
  labs(x = "",
       y = "% of LEGO parts",
       title = "LEGO Part Colors",
       subtitle = "More color diversity over time",
       caption = tt_caption)

## 3d image
rayshader::plot_gg(
  p,
  height_aes = "y",
  multicore = TRUE,
  shadow_intensity = 0.2,
  width = 4,
  height = 4,
  scale = 25,
  preview = TRUE,
  raytrace = TRUE,
  triangulate = FALSE,
  offset_edges = TRUE
)
Sys.sleep(0.2)
render_snapshot(clear = TRUE)

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2022_09_06_tidy_tuesday.png",
  device = "png")

```
