---
title: "TidyTemplate"
date: 2022-08-30
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(geofacet)
library(ggtext)

theme_set(theme_light(base_family = "Avenir Next"))

tt_caption <- "Source: US Dept of Education  |  DataViz: Tony Galvan (@GDataScience1) #TidyTuesday"

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-08-30")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

pell <- tt$pell %>%
  janitor::clean_names() %>%
  select(-session)

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

pell %>%
  group_by(year) %>%
  summarize(total_award = sum(award, na.rm = TRUE),
            total_recipient = sum(recipient)) %>%
  ungroup() %>%
  mutate(award_per_recipient = total_award / total_recipient) %>%
  pivot_longer(cols = -year, names_to = "metric", values_to = "value") %>%
  mutate(metric = case_when(metric == "total_award" ~ "Total Awarded ($)",
                            metric == "total_recipient" ~ "Total Recipients",
                            TRUE ~ "Award per Recipient ($)"),
         metric = fct_relevel(metric, "Total Awarded ($)", "Total Recipients")) %>%
  ggplot(aes(year, value, fill = metric)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~metric, scales = "free_y", nrow = 3) +
  scale_y_continuous(labels = comma_format()) +
  labs(x = "",
       y = "",
       title = "Pell Grants Awarded",
       subtitle = "What caused the jump in awards/recipients in 2009?",
       caption = tt_caption)
  
```

## Answer to what happened in 2009
In 2008 and 2009, both the number of Pell Grant recipients and the cost of the program rose rapidly. This growth can be attributed to a combination of large increases in college enrollment and Pell Grant applications, the effect of a weak economy on applicants’ incomes and grant eligibility, broader eligibility rules that Congress passed in 2007 and 2008, and a substantial increase in the maximum grant that Congress included in the American Recovery and Reinvestment Act of 2009 (the economic stimulus bill). Source - https://www.newamerica.org/education-policy/topics/higher-education-funding-and-financial-aid/federal-student-aid/federal-pell-grants/pell-grant-funding/

# Visualize

Using your processed dataset, create your unique visualization.

```{r GeoFacet}

pell %>%
  group_by(year, state) %>%
  summarize(total_award = sum(award, na.rm = TRUE),
            total_recipient = sum(recipient)) %>%
  ungroup() %>%
  mutate(award_per_recipient = total_award / total_recipient) %>%
  ggplot(aes(year, award_per_recipient, color = state == "NH")) +
  geom_line(size = 2) +
  facet_geo(~state, grid = "us_state_grid2", label = "name") +
  scale_y_continuous(labels = dollar_format(scale = 0.001, suffix = "K"), 
                     breaks = c(2000, 3000, 4000)) +
  scale_x_continuous(breaks = c(2000, 2005, 2010, 2015), 
                     labels = c("'00", "'05", "'10", "'15")) +
  scale_color_manual(values = c("gray75", "#456990")) +
  theme(legend.position = "none",
        plot.title = element_markdown(size = 24),
        plot.subtitle = element_markdown(size = 18)) +
  labs(x = "",
       y = "Amount awarded per recipient",
       title = "<b>Pell Grants</b>",
       subtitle = "Why did the award per recipient in <span style = 'color:#456990;'><b>New Hampshire</b></span> decrease from 2010 to 2016?",
       caption = tt_caption)

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2022_08_30_tidy_tuesday.png",
  device = "png")

```
