---
title: "TidyTemplate"
date: 2022-12-06
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(elevators)

theme_set(theme_light())

tt_caption <- "Source: Hvitfeldt E (2022). elevators: Data Package Containing Information About Elevators in NYC. \nhttps://github.com/EmilHvitfeldt/elevators, https://emilhvitfeldt.github.io/elevators/  \n  DataViz: Tony Galvan  |  Twitter: @GDataScience1  |  #TidyTuesday"

```

# Load the weekly Data

```{r Load}

elevators <- elevators

glimpse(elevators)

```


# Wrangle

Explore the data and process it into a nice format for plotting!

```{r Wrangle}

elevators <- elevators

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

# SAMPLE PLOT - provided in package documentation
# https://emilhvitfeldt.github.io/elevators/
elevators |>
  ggplot(aes(speed_fpm, capacity_lbs)) +
  geom_point(alpha = 0.1)

elevators |>
  count(device_type, sort = TRUE)

elevators |>
  mutate(device_type = fct_relevel(device_type, "Passenger Elevator")) |>
  ggplot(aes(speed_fpm, capacity_lbs, color = device_type)) +
  geom_point(alpha = 0.5, show.legend = FALSE) +
  expand_limits(x = 0.1, y = 0.1) +
  scale_x_log10(labels = comma_format()) +
  scale_y_log10(labels = comma_format()) +
  facet_wrap(~device_type) +
  labs(x = "Speed (fpm)",
       y = "Capacity (lbs)",
       title = "New York City Registered Elevator Devices",
       subtitle = "Speed vs. Capacity by Device Type",
       caption = tt_caption)
  
```


```{r Visualize2}

# SAMPLE PLOT - provided in package documentation
# https://emilhvitfeldt.github.io/elevators/
elevators |>
  ggplot(aes(longitude, latitude, color = borough)) +
  coord_map() +
  geom_point(alpha = 0.1) +
  theme_minimal()


library(maps)

xmin = min(elevators$longitude)
xmax = max(elevators$longitude)
ymin = min(elevators$latitude)
ymax = max(elevators$latitude)

# Get the world polygon and extract UK
USA <- map_data("world") %>% filter(region == "USA")

elevators %>%
  mutate(device_type = fct_lump(device_type, 3),
         device_type = fct_relevel(device_type, "Passenger Elevator"),
         borough = fct_relevel(borough, "Manhattan")) %>%
  ggplot() +
  geom_polygon(data = USA, 
               aes(x=long, y = lat, group = group), 
               fill="grey95") +
  coord_map(xlim = c(xmin - ((xmax - xmin)*0.015), xmax + ((xmax - xmin)*0.015)), 
            ylim = c(ymin - ((ymax - ymin)*0.015), ymax + ((ymax - ymin)*0.015))) +
  geom_point(aes(x = longitude, y = latitude, 
                 color = borough),
             alpha = 0.75, shape = 0, size = 0.75) +
  theme_void() +
  theme(legend.position = "top") +
  facet_wrap(~device_type) +
  labs(color = "",
       title = "New York City Registered Elevator Devices",
       subtitle = "Types by Borough",
       caption = tt_caption)

library(sp)
library(rgdal)
library(rgeos)
library(ggthemes)

url <- "https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/nybb_22c.zip"
fil <- basename(url)
if (!file.exists(fil)) download.file(url, fil)

fils <- unzip(fil)

nyc <- readOGR(fils[1], ogrListLayers(fils[1])[1], stringsAsFactors=FALSE)

# simplifying the polygons speeds up ggplot2 a bit
nyc_map <- fortify(gSimplify(nyc, 0.05))

nyc %>%
  ggplot() +
  geom_map(map = nyc_map,
           aes(x = long, y = lat, map_id = id),
           color = "black", fill = "gray75", size = 0.25) +
  coord_equal() +
  theme_map()

```


```{r Visualize3}

# SAMPLE PLOT - provided in package documentation
# https://emilhvitfeldt.github.io/elevators/
elevators |>
  filter(borough == "Manhattan") |>
  mutate(floors = str_extract(floor_to, "\\d+"),
         floors = as.numeric(floors)) |>
  filter(!is.na(floors), floors < 100, floors > 0) |>
  ggplot(aes(longitude, latitude, color = floors)) +
  geom_point(alpha = 0.05) +
  scale_color_viridis_c() +
  theme_minimal() +
  facet_wrap(~ cut_width(floors, width = 10, boundary = 0))

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2022_12_06_tidy_tuesday_elevators.png",
  device = "png")

```
