---
title: "TidyTemplate"
date: 2023-08-30
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(camcorder)
library(ggtext)
library(showtext)
library(tidytext)

# Font selected from http://fonts.google.com
tt_family <- "playfair"
font_add_google("Playfair Display", tt_family)

theme_set(theme_light(base_size = 32))

font_add(family = "fa-brands", regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid", regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")
showtext_auto()

# Colors selected using https://coolors.co/
tt_source <- "U.S. Copyright Office Fair Use Index"
bg_color <- "#132E51"
txt_color <- "#EAECEF"
green <- "#79B473"  
red <- "#66101F"
blue <- "#4576C1"

tt_caption <- paste0("DataViz: Tony Galvan #TidyTuesday<span style='color:", bg_color, ";'>..</span><span style='font-family:fa-solid;color:", txt_color, ";'>&#xf0ce;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>", tt_source, "</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xe61b;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>@GDataScience1</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf09b;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>GDataScience</span><span style='color:", bg_color, ";'>..</span>")

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-08-29")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt |> 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

cariou_key_facts <- "Plaintiff Patrick Cariou published Yes Rasta, a book of portraits and landscape photographs taken in Jamaica. Defendant Richard Prince was an appropriation artist who altered and incorporated several of plaintiff’s photographs into a series of paintings and collages called Canal Zone that was exhibited at a gallery and in the gallery’s exhibition catalog. Plaintiff filed an infringement claim, and the district court ruled in his favor, stating that to qualify as fair use, a secondary work must “comment on, relate to the historical context of, or critically refer back to the original works.” Defendant appealed."

fair_use <- tt$fair_use_cases |>
  rename(court_short = court) |>
  select(-categories) |>
  bind_cols(
    tt$fair_use_findings |>
      select(-year, -outcome)
  ) |>
  mutate(
    tags = if_else(str_detect(case, "Cariou"), holding, tags),
    holding = if_else(str_detect(case, "Cariou"), issue, holding),
    issue = if_else(str_detect(case, "Cariou"), key_facts, issue),
    key_facts = if_else(str_detect(case, "Cariou"), cariou_key_facts, key_facts),
    tags = str_to_lower(tags),
    tags = case_when(
      str_detect(tags, "educational") ~ str_replace_all(tags, "educational", "education"),
      str_detect(tags, "photography") ~ str_replace_all(tags, "photography", "photograph"),
      str_detect(tags, "films") ~ str_replace_all(tags, "films", "film"),
      TRUE ~ tags
    )
  )

```


# Start recording

Use the {camcorder} package to record all plots that are output to the console

```{r camcorder}

# start recording
gg_record(
  dir = "~/Downloads/camcorder/2023_08_29_tidy_tuesday_fair_use", # where to save the recording
  device = "png", # device to use to save images
  width = 6, # width of saved image
  height = 6, # height of saved image
  units = "in", # units for width and height
  dpi = 300 # dpi to use when saving image
)

```

# Theme

Set up a custom data visualization theme

```{r VizTheme}

tt_theme <- function() {
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    plot.background = element_rect(fill = bg_color),
    panel.background = element_rect(fill = bg_color),
    legend.background = element_rect(fill = bg_color),
    legend.key = element_rect(fill = bg_color),
    strip.background = element_rect(fill = bg_color), 
    plot.title = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      face = "bold",
      size = 150,
      margin = margin(t = 0, r = 0, b = 0, l = 0)
    ),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 68,
      lineheight = 0.3,
      margin = margin(t = 0, r = 0, b = 4, l = 0)
    ),
    plot.caption = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 24,
      margin = margin(t = 10, r = 0, b = 0, l = 0)
    ),
    plot.caption.position = "plot",
    strip.text = element_text(
      family = tt_family,
      color = txt_color,
      face = "bold",
      size = 32, 
      lineheight = 0.3,
      margin = margin(t = 0, r = 0, b = 0, l = 0)
    ),
    legend.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    legend.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    axis.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    axis.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
  )
}

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

fair_use |>
  mutate(decade = 10 * (year %/% 10)) |>
  count(decade, fair_use_found) |>
  ggplot(aes(decade, n, fill = fair_use_found)) +
  geom_col() +
  scale_fill_manual(
    values = c(txt_color, blue)
  ) +
  tt_theme() +
  theme(legend.position = "top") +
  labs(
    x = "Decade",
    y = "# of cases",
    fill = "Fair Use Found?",
    title = "Fair Use Cases",
    subtitle = "When did these cases happen?",
    caption = tt_caption
  )
  
```


```{r}

fair_use |>
  separate_rows(
    tags,
    sep = "[,;](?=\\s)|\\/"
  ) |>
  mutate(
    tags = str_to_lower(tags),
    tags = str_trim(tags),
    tags = str_remove_all(tags, "\\.")
  ) |>
  filter(
    !str_detect(tags, "circuit"),
    !str_detect(tags, "court")
  ) |>
  count(tags) |>
  mutate(
    tags = str_to_title(tags),
    tags = fct_reorder(tags, n)
  ) |>
  ggplot(aes(n, tags)) +
  geom_col(fill = blue) +
  tt_theme() +
  labs(
    x = "# of cases",
    y = NULL,
    title = "Fair Use Cases",
    subtitle = "What are the most common categories?",
    caption = tt_caption
  )

```

# Text Features
As seen in Julia Silge's blog post - https://juliasilge.com/blog/last-airbender/

```{r TextFeatures}

gg_resize_film(
  height = 9, 
  width = 12,
  units = "in",
  dpi = 300
)

library(textfeatures)

tf <- textfeatures(
  fair_use |> rename(text = key_facts),
  sentiment = FALSE,
  word_dims = 0,
  normalize = FALSE,
  verbose = FALSE
)

tf |>
  bind_cols(fair_use) |>
  group_by(fair_use_found) |>
  summarize(across(starts_with("n_"), mean)) |>
  pivot_longer(starts_with("n_"), names_to = "text_feature") |>
  filter(value > 0.01) |>
  mutate(text_feature = fct_reorder(text_feature, -value)) |>
  ggplot(aes(fair_use_found, value, fill = fair_use_found)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  facet_wrap(~text_feature, scales = "free", ncol = 6) +
  scale_fill_manual(values = c(blue, txt_color)) +
  tt_theme() +
  labs(
    x = NULL, 
    y = "Mean text features per case",
    title = "Fair Use Cases",
    subtitle = "Can it be determined if fair use was found based on text features of case key facts?",
    caption = tt_caption
  )

```

Definitions - https://textfeatures.mikewk.com/reference/count_functions.html

# Modeling

Can we predict the outcome using text mining techniques?

```{r DataPrep}

fair_use_for_model <- fair_use |>
  transmute(
    case_number,
    fair_use_found = if_else(fair_use_found, "yes", "no"),
    year,
    key_facts,
    tags
  )

library(tidymodels)
library(textrecipes)

set.seed(123)

use_split <- initial_split(fair_use_for_model, strata = fair_use_found)
use_train <- training(use_split)
use_test <- testing(use_split)

```

Create cross-validation resamples of the training data, to evaluate our models

```{r cvfolds}

set.seed(234)
use_folds <- vfold_cv(use_train, strata = fair_use_found)
use_folds

```

Create bootstrap resamples of the training data, to tune our models

```{r bootfolds}

set.seed(567)
use_boots <- bootstraps(use_train, strata = fair_use_found)
use_boots

```


# Preprocessing

```{r preproc}

use_rec <- recipe(fair_use_found ~ key_facts, data = use_train) |>
  step_textfeature(key_facts) |>
  step_zv(all_predictors()) |>
  step_normalize(all_predictors())

use_prep <- prep(use_rec)
use_prep

juice(use_prep)

use_tr_rec <- recipe(fair_use_found ~ key_facts, data = use_train) |>
  step_tokenize(key_facts) |>
  step_stopwords(key_facts) |>
  step_tokenfilter(key_facts, max_tokens = 100) |>
  step_tfidf(key_facts) |>
  step_normalize(all_predictors())

use_tr_prep <- prep(use_tr_rec)
use_tr_prep

glimpse(juice(use_tr_prep))

```

Create model specs: random forest model and a support vector machine model

```{r RFSpec}

rf_spec <- rand_forest(trees = 1000) |>
  set_engine("ranger") |>
  set_mode("classification")

rf_spec

```

```{r SVMSpec}

svm_spec <- svm_rbf(cost = 0.5) |>
  set_engine("kernlab") |>
  set_mode("classification")

svm_spec

```


```{r LassoSpec}

lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet")

lasso_wf <- workflow() |>
  add_recipe(use_tr_rec) |>
  add_model(lasso_spec)

lasso_wf

```


Create a workflow

```{r WF}

use_wf <- workflow() |>
  add_recipe(use_rec)

use_wf

```

Fit to the resamples

```{r}

doParallel::registerDoParallel()

set.seed(1234)
rf_rs <- use_wf |>
  add_model(rf_spec) |>
  fit_resamples(
    resamples = use_folds,
    metrics = metric_set(roc_auc, accuracy, sens, spec),
    control = control_grid(save_pred = TRUE)
  )

set.seed(2345)
svm_rs <- use_wf |>
  add_model(svm_spec) |>
  fit_resamples(
    resamples = use_folds,
    metrics = metric_set(roc_auc, accuracy, sens, spec),
    control = control_grid(save_pred = TRUE)
  )

set.seed(3456)
lambda_grid <- grid_regular(penalty(), levels = 40)
lasso_grid <- tune_grid(
  lasso_wf,
  resamples = use_boots,
  grid = lambda_grid,
  metrics = metric_set(roc_auc, ppv, npv)
)

```

# Evaluate Models

```{r RFMetrics}

collect_metrics(rf_rs)

conf_mat_resampled(rf_rs)

```

```{r SVMMetrics}

collect_metrics(svm_rs)

conf_mat_resampled(svm_rs)

```

```{r LassoMetrics}

lasso_grid |>
  collect_metrics()

lasso_grid |>
  collect_metrics() |>
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_line(size = 1.5, show.legend = FALSE) +
  facet_wrap(~.metric) +
  scale_x_log10() +
  tt_theme()

```



```{r AUC}

gg_resize_film(
  height = 6, 
  width = 6,
  units = "in",
  dpi = 300
)

svm_rs |>
  collect_predictions() |>
  group_by(id) |>
  roc_curve(fair_use_found, .pred_yes) |>
  ggplot(aes(1 - specificity, sensitivity, color = id)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(show.legend = FALSE, alpha = 0.6, size = 1.2) +
  coord_equal() +
  tt_theme()

```

```{r SVMVIP}

library(vip)

set.seed(345)
use_imp <- use_wf |>
  add_model(svm_spec) |>
  fit(use_train) |>
  extract_fit_parsnip() |>
  vi(
    method = "permute", nsim = 10,
    target = "fair_use_found", metric = "auc", reference_class = "yes",
    pred_wrapper = kernlab::predict, train = juice(use_prep)
  )

use_imp |>
  slice_max(Importance, n = 8) |>
  mutate(
    Variable = str_remove(Variable, "textfeature_key_facts_n_"),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(Importance, Variable, color = Variable)) +
  geom_vline(xintercept = 0, color = txt_color, lty = 2) +
  geom_errorbar(aes(xmin = Importance - StDev, xmax = Importance + StDev),
    alpha = 0.5, size = 1.3
  ) +
  geom_point(size = 3) +
  tt_theme() +
  theme(legend.position = "none") +
  labs(
    y = NULL,
    title = "Fair Use Cases",
    subtitle = "Key fact text feature importance",
    caption = tt_caption
  )

```


```{r LassoBest}

best_auc <- lasso_grid |>
  select_best("roc_auc")

best_auc

```

```{r BestWF}

final_lasso <- finalize_workflow(lasso_wf, best_auc)

final_lasso

```

```{r LassoVIP}

set.seed(456)
p <- final_lasso |>
  fit(use_train) |>
  extract_fit_parsnip() |>
  vip(num_features = 20)

p$data |>
  group_by(Sign) |>
  top_n(7, wt = abs(Importance)) |>
  ungroup() |>
  mutate(
    Importance = abs(Importance),
    Variable = str_remove(Variable, "tfidf_key_facts_"),
    Variable = fct_reorder(Variable, Importance),
    Sign = if_else(Sign == "POS", "Fair Use Found", "Fair Use Not Found")
  ) |>
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c(blue, txt_color)) +
  facet_wrap(~Sign, scales = "free_y") +
  tt_theme() +
  labs(
    y = NULL,
    title = "Fair Use Cases",
    subtitle = "Key fact word importance",
    caption = tt_caption
  )

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2023_08_29_tidy_tuesday_fair_use.png",
  device = "png")

gg_stop_recording()

gg_playback(
  name = "2023_08_29_tidy_tuesday_fair_use.gif",
  first_image_duration = 4,
  last_image_duration = 20,
  frame_duration = 0.5
)

```
