---
title: "TidyTemplate"
date: 2023-08-08
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(camcorder)
library(ggtext)
library(showtext)
library(tidytext)

# start recording
gg_record(
  dir = "~/Downloads/camcorder/2023_08_08_tidy_tuesday_hotones", # where to save the recording
  device = "png", # device to use to save images
  width = 6, # width of saved image
  height = 6, # height of saved image
  units = "in", # units for width and height
  dpi = 300 # dpi to use when saving image
)

# Font selected from http://fonts.google.com
tt_family <- "rale"
font_add_google("Raleway", tt_family)

theme_set(theme_light(base_size = 32))

font_add(family = "fa-brands", regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid", regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")
showtext_auto()

# Colors selected using https://coolors.co/
tt_source <- "Wikipedia"
bg_color <- "#000000"
txt_color <- "#FFFFFF"
yellow <- "#F6CC45"  
red <- "#C6352B"
green <- "#99C24D"

tt_caption <- paste0("DataViz: Tony Galvan #TidyTuesday<span style='color:", bg_color, ";'>..</span><span style='font-family:fa-solid;color:", txt_color, ";'>&#xf0ce;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>", tt_source, "</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf099;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>@GDataScience1</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf09b;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>GDataScience</span><span style='color:", bg_color, ";'>..</span>")

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-08-08")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt |>
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

episodes <- tt$episodes

seasons <- tt$seasons

sauces_raw <- tt$sauces |>
  group_by(sauce_name) |>
  summarize(scoville = max(scoville)) |>
  ungroup()

super_hot <- quantile(sauces_raw$scoville, 0.5)

sauces <- sauces_raw |>
  mutate(super_hot = if_else(scoville > super_hot, "yes", "no"))

```


# Machine Learning

Can I predict a super hot sauce for Hot Ones sauces? 

Methodology is based on Julia Silge's sentiment analysis with tidymodels and #TidyTuesday Animal Crossing reviews  https://juliasilge.com/blog/animal-crossing/

## Setup

Split our data into training and testing sets


```{r MLSplit}

library(tidymodels)
library(textrecipes)

set.seed(123)
sauce_split <- initial_split(sauces, strata = super_hot)
sauce_train <- training(sauce_split)
sauce_test <- testing(sauce_split)
  
```

## Preprocessing

Prepare the data for modeling

-   A recipe stores what our model is going to be (the formula) and what data to use (training set)
-   Tokenize the text (break into individual words)
-   Remove "stop" words (common useless words, such as "a", "the", etc.)
-   Filter down to only keep the top 500 most-used words
-   Weight the words by TF-IDF (<https://www.tidytextmining.com/tfidf.html> - measure how important a word is to a document in a collection (or corpus) of documents)
-   Center and scale (i.e. normalize) all the newly created TF-IDF values because the model we are going to use is sensitive to this

```{r Preprocessing}

sauce_rec <- recipe(super_hot ~ sauce_name, data = sauce_train) |>
  step_tokenize(sauce_name) |>
  step_stopwords(sauce_name) |>
  step_tokenfilter(sauce_name, max_tokens = 100) |>
  step_tfidf(sauce_name) |>
  step_normalize(all_predictors())

sauce_prep <- prep(sauce_rec)

sauce_prep

```

## Specify the model

Set up the model specification for lasso regression with penalty = tune() since we don't yet know the best value for the regularization parameter and mixture = 1 for lasso.

```{r Specify}

lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet")

lasso_wf <- workflow() |>
  add_recipe(sauce_rec) |>
  add_model(lasso_spec)

lasso_wf

```

## Tuning

Try different values for the "penalty" model hyper-parameter


```{r Tuning}

lambda_grid <- grid_regular(penalty(), levels = 40)

set.seed(456)
sauce_folds <- bootstraps(sauce_train, strata = super_hot)

doParallel::registerDoParallel()

set.seed(2020)
lasso_grid <- tune_grid(
  lasso_wf,
  resamples = sauce_folds,
  grid = lambda_grid,
  metrics = metric_set(roc_auc, ppv, npv)
)

lasso_grid |>
  collect_metrics() |>
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_line(size = 1.5, show.legend = FALSE) +
  facet_wrap(~.metric) +
  scale_x_log10()

```

## Choose the final model

Choose a final model based on AUC

```{r FinalModel}

best_auc <- lasso_grid |>
  select_best("roc_auc")

final_lasso <- finalize_workflow(lasso_wf, best_auc)

final_lasso

```

## Set up a custom data visualization theme

```{r VizTheme}

tt_theme <- function() {
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    plot.background = element_rect(fill = bg_color),
    panel.background = element_rect(fill = bg_color),
    legend.background = element_rect(fill = bg_color),
    legend.key = element_rect(fill = bg_color),
    strip.background = element_rect(fill = bg_color), 
    plot.title = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      face = "bold",
      size = 100
    ),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 48,
      margin = margin(t = 5, r = 0, b = 2, l = 0)
    ),
    plot.caption = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 28,
      margin = margin(t = 10, r = 0, b = 0, l = 0)
    ),
    plot.caption.position = "plot",
    strip.text = element_text(
      family = tt_family,
      color = txt_color,
      face = "bold",
      size = 36, lineheight = 0.3,
      margin = margin(t = 0, r = 0, b = 0, l = 0)
    ),
    legend.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 26
    ),
    legend.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 26
    ),
    axis.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 26
    ),
    axis.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 26
    ),
  )
}

```

## Model results

```{r Visualize}

library(vip)

final_lasso |>
  fit(sauce_train) |>
  extract_fit_parsnip() |>
  vi(lambda = best_auc$penalty) |>
  group_by(Sign) |>
  top_n(7, wt = abs(Importance)) |>
  ungroup() |>
  mutate(
    Importance = abs(Importance),
    Variable = str_remove(Variable, "tfidf_sauce_name_"),
    Variable = fct_reorder(Variable, Importance)
  ) |>
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~if_else(Sign == "POS", "Super Hot", "Hot"), 
             scales = "free") +
  scale_fill_manual(values = c(yellow, red)) +
  tt_theme() +
  labs(
    x = "Importance of word (based on a Lasso Regression model with a ROC AUC of 0.77)",
    y = NULL,
    title = "HOT ONES: Sauce Names",
    subtitle = paste0("Which words indicate the sauce is <b><span style='color:", yellow, ";'>HOT</span></b> vs <b><span style='color:", red, ";'>SUPER HOT</span></b>?"),
    caption = tt_caption
  )

```

## Performance on Test Data

Fit the model one last time on the training data and evaluate it on the testing data

```{r TestEval}

sauce_final <- last_fit(final_lasso, sauce_split)

sauce_final |>
  collect_metrics()

```

## Confusion Matrix

Compare our predictions against the truth

```{r ConfMat}

sauce_final |>
  collect_predictions() |>
  conf_mat(super_hot, .pred_class)

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2023_08_08_tidy_tuesday_hotones.png",
  device = "png")

gg_stop_recording()

gg_playback(
  name = "2023_08_08_tidy_tuesday_hotones.gif",
  first_image_duration = 4,
  last_image_duration = 20,
  frame_duration = 0.5
)

```
