---
title: "TidyTemplate"
date: 2024-10-07
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(scales)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2024-10-08")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

nps <- tt$most_visited_nps_species_data |>
  janitor::clean_names()

# download the full data from the NPSpecies package (https://github.com/frankiethull/NPSpecies)

species <- NPSpecies::species |>
  janitor::clean_names()

```


# Exploration

Reproduce tables from https://github.com/frankiethull/NPSpecies

```{r Visualize}

species |>
    filter(park_code == "GRSA") |>
    filter(category_name == "Reptile") |>
    select(
      park_name, category_name, family, sci_name, common_names, observations
    ) |>
    arrange(
      desc(observations)
    ) |>
    head(3) |>
    tinytable::tt()

species |>
    filter(park_code == "EVER") |> 
    filter(nativeness == "Non-native") |>
    select(
      park_name, category_name, family, sci_name, common_names, observations
    ) |>
    arrange(desc(observations)) |>
    head(5) |>
    tinytable::tt()

species |>
  filter(category_name == "Bird") |> 
  group_by(park_name) |>
  select(park_name, common_names, observations) |>
  arrange(desc(observations)) |>
  slice(1) |>
  ungroup() |>
  arrange(desc(observations)) |> 
  head(10) |> # only display 10
  tinytable::tt()
  
```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

species |>
  filter(category_name == "Mammal") |>
  count(park_name, wt = observations, name = "total_observations") |>
  top_n(10, wt = total_observations) |>
  mutate(park_name = fct_reorder(park_name, total_observations)) |>
  ggplot(
    aes(
      x = total_observations, 
      y = park_name,
      fill = park_name
    )
  ) +
  geom_col() +
  # add text at the end of the bars
  geom_text(
    aes(label = comma(total_observations)),
    hjust = 0,
    nudge_x = 200
  ) +
  # give room for the text at the end of the bars
  scale_x_continuous(
    labels = comma_format(),
    limits = c(0, 21000)
  ) +
  labs(
    title = "Total Observations of Mammals",
    subtitle = "Top 10 National Parks with most mammal observations",
    x = "Total Observations",
    y = NULL,
    caption = "Data Source: {NPSpecies} R Package  |  DataViz: Tony Galvan (@GDataScience1). |  #TidyTuesday"
  ) +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot",
    legend.position = "none"
  )

```


```{r}

library(ggrepel)

# using the species data visualize the parks where you can see a coyote and a roadrunner
p1 <- species |>
  filter(
    sci_name %in% c("Canis latrans", "Geococcyx californianus"),
    observations > 0
  ) |>
  count(
    park_name, sci_name, 
    wt = observations, 
    name = "total_observations"
  ) |>
  pivot_wider(
    names_from = sci_name, 
    values_from = total_observations, 
    values_fill = 0
  ) |>
  rename(
    coyote = `Canis latrans`,
    roadrunner = `Geococcyx californianus`
  ) |>
  filter(
    coyote > 0,
    roadrunner > 0
  ) |>
  ggplot(
    aes(
      x = coyote,
      y = roadrunner,
      label = park_name
    )
  ) +
  geom_abline(
    slope = 1,
    color = "grey50",
    size = 0.5,
    lty = 2
  ) +
  geom_point(
    color = "goldenrod",
    size = 3
  ) +
  geom_text_repel(
    box.padding = 0.5,
    point.padding = 0.5,
    segment.color = "grey50",
    segment.size = 0.25
  ) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "National Parks with Coyotes and Roadrunners",
    subtitle = "Parks where you can see both animals",
    x = "Coyote Observations (log scale)",
    y = "Roadrunner Observations (log scale)",
    caption = "Data Source: {NPSpecies} R Package  |  DataViz: Tony Galvan (@GDataScience1). |  #TidyTuesday"
  )

p1

library(webp)
path = "~/Documents/GitHub/tidytuesday/"
img_coyote <- read_webp(paste0(path, "coyote.webp"))
img_roadrunner <- read_webp(paste0(path, "roadrunner.webp"))

library(png)
writePNG(img_coyote, paste0(path, "coyote.png"))
writePNG(img_roadrunner, paste0(path, "roadrunner.png"))

library(magick)
img_coyote2 <- image_read(paste0(path, "coyote.png"))
img_coyote3 <- image_flop(img_coyote2)

library(cowplot)
ggdraw(p1) +
  draw_image(
    paste0(path, "roadrunner.png"),
    x = 0.02, y = 0.6, width = 0.25, height = 0.25
  ) + 
  draw_image(
    img_coyote3,
    x = 0.75, y = 0.55, width = 0.325, height = 0.325
  )

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2024_10_08_tidy_tuesday_nps.png",
  device = "png"
)
```
