---
title: "TidyTemplate"
date: 2024-10-15
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2024-10-15")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

orcas <- tt$orcas

# Install the {orcas} package
# devtools::install_github("jadeynryan/orcas")
library(orcas)

orcas_tidy <- orcas::cwr_tidy # This is the same as the #TidyTuesday version

```

# Example code 
From the GitHub page for the {orcas} package - https://github.com/jadeynryan/orcas

```{r Examples}
# Example code from the GitHub page for the {orcas} package
# Scrape the two most recent encounters from 2022 and 2023:
orcas_small <- orcas::make_encounter_df(years = 2022:2023, max_urls = 2)

orcas::cwr_tidy |> 
  subset(year == 2023) |> 
  orcas::make_dt()

orcas::cwr_tidy |> 
  subset(year == 2023) |> 
  head(2) |> 
  orcas::make_leaflet()

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

orcas |> 
  count(year, name = "count") |>
  ggplot(aes(x = year, y = count)) + #, fill = type)) + # Copilot error
  geom_col() +
  labs(
    title = "Orca Encounters",
    subtitle = "Encounters by year",
    caption = "Data Source: Center for Whale Research",
    x = "Year",
    y = "Count"
  )

orcas |>
  filter(
    !is.na(date),
    date < ymd("2024-10-01")
  ) |>
  mutate(
    month = month(date),
    fct_year = factor(year)
  ) |>
  count(month, fct_year, name = "count") |>
  ggplot(aes(x = month, y = count, color = fct_year)) +
  geom_point(
    alpha = 0.5
  ) +
  geom_smooth(
    method = "loess",
    se = FALSE
  ) +
  scale_x_continuous(
    breaks = 1:12, 
    labels = month.abb
  ) +
  labs(title = "Orca Encounters",
       subtitle = "Encounters by Month and Year",
       caption = "Data Source: Center for Whale Research",
       color = "Year",
       x = "Month",
       y = "Count")

# create the plot above but with a single smoothed line representing all years
orcas |>
  filter(
    !is.na(date),
    date < ymd("2024-10-01")
  ) |>
  mutate(
    month = month(date),
    fct_year = factor(year)
  ) |>
  count(month, fct_year, name = "count") |>
  ggplot(aes(x = month, y = count)) +
  geom_jitter(
    aes(color = fct_year),
    alpha = 0.5
  ) +
  geom_smooth(
    method = "loess",
    se = FALSE
  ) +
  scale_x_continuous(
    breaks = 1:12, 
    labels = month.abb
  ) +
  labs(title = "Orca Encounters",
       subtitle = "Encounters tend to peak in August and September",
       caption = "Data Source: Center for Whale Research",
       color = "Year",
       x = "Month",
       y = "Encounters")

# create a plot of the cumulative number of orcas encountered by month with a line for each year
orcas |>
  filter(
    !is.na(date),
    date < ymd("2024-10-01")
  ) |>
  mutate(
    month = month(date),
    fct_year = factor(year)
  ) |>
  count(month, fct_year, name = "count") |>
  group_by(fct_year) |>
  mutate(cumulative = cumsum(count)) |>
  ggplot(aes(x = month, y = cumulative, color = fct_year)) +
  geom_line() +
  scale_x_continuous(
    breaks = 1:12, 
    labels = month.abb
  ) +
  labs(title = "Orca Encounters",
       subtitle = "Cumulative Encounters by Month and Year",
       caption = "Data Source: Center for Whale Research",
       color = "Year",
       x = "Month",
       y = "Cumulative Count")

# recreate the cumulative plot above but contrast 2024 vs. 2020 by highlighting them and making them thicker than the rest of the lines
library(ggtext)

orcas |>
  filter(
    !is.na(date),
    date < ymd("2024-10-01")
  ) |>
  mutate(
    month = month(date),
    fct_year = factor(year)
  ) |>
  count(month, fct_year, name = "count") |>
  group_by(fct_year) |>
  mutate(cumulative = cumsum(count)) |>
  ggplot(aes(
    x = month, 
    y = cumulative, 
    group = fct_year,
    color = if_else(fct_year == "2020" | fct_year == "2024", fct_year, "Other"), 
    size = fct_year == "2020" | fct_year == "2024",
    lty = !(fct_year == "2020" | fct_year == "2024")
  )) +
  geom_line() +
  scale_x_continuous(
    breaks = 1:12, 
    labels = month.abb
  ) +
  scale_size_manual(values = c(0.5, 1, 1)) +
  scale_color_manual(
    values = c("2020" = "orange", "2024" = "royalblue", "Other" = "gray")
  ) +
  labs(
    title = "Orca Encounters from 2017 - 2024",
    # use the ggtext package to callout 2020 and 2024 in the subtitle and color the year to use as a color legend
    subtitle = "Cumulative Encounters by Month and Year<br><span style='color:royalblue'>**2024**</span>: Most encounters  **|**  <span style='color:orange'>**2020**</span>: Least encounters",
    caption = "Data Source: Center for Whale Research",
    x = "Month",
    y = "Cumulative Count"
  ) +
  theme(
    legend.position = "none",
    plot.subtitle = element_markdown()
  )

# use the prophet package to forecast the number of orca encounters in 2025
library(prophet)

forecast_model <- orcas |>
  filter(
    !is.na(date),
    date < ymd("2024-10-01")
  ) |>
  mutate(month = floor_date(date, unit = "month")) |>
  count(month, name = "count") |>
  transmute(ds = month, y = count) |>
  prophet()

# make predictions for 12 months after the last date in the data
forecast <- make_future_dataframe(forecast_model, periods = 12, freq = "month")

# combine forecasted values and actual historical values
forecast_combined <- forecast_model |> 
  predict(forecast) |>
  left_join(orcas |>
               filter(
                 !is.na(date),
                 date < ymd("2024-10-01")
               ) |>
               mutate(month = floor_date(date, unit = "month")) |>
               count(month, name = "count") |>
               transmute(ds = month, y = count), by = "ds"
  )

# using ggplot, plot the hitorical data and the forecast so that the line changes color once the forecast starts
forecast_combined |> 
  ggplot(aes(x = ds)) +
  geom_line(aes(y = coalesce(y, yhat), color = is.na(y))) +
  geom_ribbon(
    aes(
      ymin = if_else(is.na(y), yhat_lower, y), 
      ymax = if_else(is.na(y), yhat_upper, y)
    ), 
    alpha = 0.2
  ) +
  scale_color_manual(
    values = c("black", "red"),
    labels = c("Historical", "Forecast")
  ) +
  labs(
    title = "Orca Encounters",
    subtitle = "Forecasted Encounters for the next 12 months",
    caption = "Data Source: Center for Whale Research",
    x = "Month",
    y = "Count",
    color = NULL
  ) +
  theme(
    legend.position = "top",
    # decrease the space on top and bottom of the legend
    legend.margin = margin(0, 0, 0, 0)
  )

# plot the forecasted values with calendar month on the x-axis and a line for each year; all the lines should be gray except the forecasted periods which should be red
library(ggtext)

forecast_combined |>
  mutate(
    month = factor(month(ds)),
    fct_year = factor(year(ds))
  ) |>
  ggplot(
    aes(
      x = month,
      y = coalesce(y, yhat)
    )
  ) +
  geom_boxplot(
    outlier.shape = NA,
    fill = "gray",
    alpha = 0.5
  ) +
  geom_point(
    data = forecast_combined |> filter(is.na(y)),
    aes(
      x = factor(month(ds)),
      y = yhat
    ),
    color = "red"
  ) +
  scale_x_discrete(
    breaks = 1:12, 
    labels = month.abb
  ) +
  labs(
    title = "Orca Encounters: 12 Month Forecast",
    subtitle = "<span style='color:red'>**Forecasted Encounters**</span> for the next 12 months (using Prophet)",
    caption = "Data Source: Center for Whale Research  |  DataViz: Tony Galvan (@GDataScience1)  |  #TidyTuesday",
    x = "Month",
    y = "Count",
    color = NULL
  ) +
  theme(
    legend.position = "top",
    legend.margin = margin(0, 0, 0, 0),
    plot.subtitle = element_markdown()
  )
  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2024_10_15_tidy_tuesday_orcas.png",
  device = "png"
)
```
