---
title: "TidyTemplate"
date: 2025-11-10
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-11-11")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

tb <- tt$who_tb_data

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

tb |>
  count(g_whoregion, year, wt = e_inc_num, name = "estimated_cases") |>
  ggplot(
    aes(
      x = year,
      y = estimated_cases,
      color = g_whoregion
    )
  ) +
  geom_line(linewidth = 1.2) +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(
    title = "Estimated Tuberculosis Cases by WHO Region",
    x = "Year",
    y = "Estimated Cases",
    color = "WHO Region"
  )

tb |>
  group_by(g_whoregion, year) |>
  summarize(mean_e_inc_100k = mean(e_inc_100k)) |>
  ungroup() |>
  mutate(g_whoregion = fct_reorder(g_whoregion, -mean_e_inc_100k)) |>
  ggplot(
    aes(
      x = year,
      y = mean_e_inc_100k,
      fill = g_whoregion
    )
  ) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::comma_format()) +
  facet_wrap(~g_whoregion, ncol = 1, scales = "free_y") +
  labs(
    title = "Incidence per 100k population",
    x = "Year",
    y = "Estimated Cases"
  )

# use the {gt} package to make a table that contains a spark line for each WHO region showing the trend of incidence per 100k population over time. Order the regions by the mean incidence per 100k population.
library(gt)
library(gtExtras)

tb_summary <- tb |>
  group_by(g_whoregion, year) |>
  summarize(
    mean_e_inc_100k = mean(e_inc_100k)
  ) |>
  ungroup() |>
  group_by(g_whoregion) |>
  summarize(
    first_inc = first(mean_e_inc_100k),
    last_inc = last(mean_e_inc_100k),
    pct_change = (last_inc - first_inc) / first_inc * 100,
    sparkline = list(mean_e_inc_100k)
  ) |>
  ungroup() |>
  arrange(desc(first_inc))

tb_summary |>
  gt() |>
  fmt_number(
    columns = c(first_inc, last_inc),
    decimals = 1
  ) |>
  fmt_number(
    columns = c(pct_change),
    decimals = 1,
    pattern = "{x}%"
  ) |>
  gt_plt_sparkline(
    column = sparkline,
    type = "shaded",
    same_limit = TRUE
  ) |>
  cols_label(
    g_whoregion = "WHO Region",
    first_inc = "2000",
    last_inc = "2023",
    pct_change = "% Change",
    sparkline = "Trend"
  ) |>
  tab_header(
    title = "Tuberculosis Incidence per 100k Population by WHO Region"
  )

# Create a table similar to above, but with 3 sections: top for incidence over 100 per 100k, middle for non-HIV mortality per 100k, and bottom for HIV mortality per 100k.
tb_summary_inc <- tb |>
  group_by(g_whoregion, year) |>
  summarize(
    mean_e_inc_100k = mean(e_inc_100k),
    mean_e_mort_hiv_100k = mean(e_mort_tbhiv_100k, na.rm = TRUE),
    mean_e_mort_nhiv_100k = mean(e_mort_exc_tbhiv_100k, na.rm = TRUE)
  ) |>
  ungroup() |>
  pivot_longer(
    cols = c(mean_e_inc_100k, mean_e_mort_hiv_100k, mean_e_mort_nhiv_100k),
    names_to = "metric",
    values_to = "value"
  ) |>
  group_by(g_whoregion, metric) |>
  arrange(year) |>
  summarize(
    first_value = first(value),
    last_value = last(value),
    pct_change = (last_value - first_value) / first_value * 100,
    sparkline = list(value)
  ) |>
  ungroup() |>
  arrange(metric, desc(first_value))

tb_tab <- tb_summary_inc |>
  gt() |>
  fmt_number(
    columns = c(first_value, last_value),
    decimals = 1
  ) |>
  fmt_number(
    columns = c(pct_change),
    decimals = 1,
    pattern = "{x}%"
  ) |>
  gt_plt_sparkline(
    column = sparkline,
    type = "default",
    same_limit = FALSE,
    label = FALSE,
    palette = c(
      "#BB8588", # sparkline color
      "transparent", # final value color
      "transparent", # range color low
      "transparent", # range color high
      "transparent" # 'type' color (e.g. shading or reference lines)
    )
  ) |>
  cols_label(
    g_whoregion = "WHO Region",
    first_value = "2000",
    last_value = "2023",
    pct_change = "% Change",
    sparkline = "Trend"
  ) |>
  tab_header(
    title = "Tuberculosis Metrics by WHO Region"
  ) |>  
  tab_row_group(
    group = "HIV Mortality per 100k",
    rows = metric == "mean_e_mort_hiv_100k",
    id = "hiv_mortality"
  ) |>
  tab_row_group(
    group = "Non-HIV Mortality per 100k",
    rows = metric == "mean_e_mort_nhiv_100k",
    id = "non_hiv_mortality"
  ) |>  
  tab_row_group(
    group = "Incidence per 100k",
    rows = metric == "mean_e_inc_100k",
    id = "incidence"
  ) |>
  cols_hide(columns = metric) |>
  gt_theme_dark() |>
  tab_style(
    style = cell_fill(color = "#F4D06F"),
    locations = cells_row_groups(
      groups = "hiv_mortality"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#65AFFF"),
    locations = cells_row_groups(
      groups = "non_hiv_mortality"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#7FD8BE"),
    locations = cells_row_groups(
      groups = "incidence"
    )
  ) |>
  # add a caption
  tab_source_note(
    source_note = "Data Source: {getTBinR} package | Visualization: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) |>
  # make the caption font smaller
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_source_notes()
  )

tb_tab

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
# ggsave(
#   filename = "2025_11_11_tidy_tuesday_tb.png",
#   device = "png"
# )
# Save the {gt} table as a png
gtsave(
  tb_tab,
  filename = "2025_11_11_tidy_tuesday_tb.png"
)
```
