---
title: "TidyTemplate"
date: 2025-09-25
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event! Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data. While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format. The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community! As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-09-23")
```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset. This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```

# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt |> 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

chess <- tt$fide_ratings_august |>
  mutate(report_month = "August") |>
  bind_rows(
    tt$fide_ratings_september |>
      mutate(report_month = "September")
  ) |>
  mutate(
    age_bins = cut(
      2025 - bday,
      breaks = c(0, 18, 30, 50, 65, 200),
      labels = c("<18", "18-29", "30-49", "50-64", "65+"),
      right = FALSE
    )
  )

```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

chess |>
  # mutate(sex = fct_rev(sex)) |>
  ggplot(aes(rating, fill = sex)) +
  geom_histogram(
    alpha = 0.5, 
    position = "identity", 
    bins = 30
  ) +
  scale_fill_manual(
    values = c("red", "lightblue"),
    labels = c("M" = "Male", "F" = "Female")
  ) +
  labs(
    title = "FIDE Chess Ratings Distribution",
    subtitle = "Distribution of chess ratings by sex",
    x = "Rating",
    y = "Count",
    fill = "Sex"
  )

chess |>
  ggplot(aes(x = bday, y = rating, group = bday)) +
  geom_boxplot() +
  facet_wrap(~sex) +
  labs(
    title = "FIDE Chess Ratings Over Time",
    subtitle = "Average ratings by country and sex",
    x = "Year",
    y = "Average Rating"
  )

chess |>
  filter(!is.na(age_bins)) |>
  ggplot(aes(x = age_bins, y = rating, fill = age_bins)) +
  geom_violin(
    show.legend = FALSE,
    # add median line
    draw_quantiles = c(0.5)
  ) +
  facet_grid(
    cols = vars(report_month),
    rows = vars(sex)
  ) +
  labs(
    title = "FIDE Chess Ratings Over Time",
    subtitle = "Average ratings by country and sex",
    x = "Age",
    y = "Average Rating"
  )

chess |>
  filter(!is.na(age_bins)) |>
  count(sex, age_bins, report_month) |>
  group_by(report_month, sex) |>
  mutate(
    total = sum(n),
    prop = n / total,
    count_emoji = round(prop * 100)
  ) |>
  ungroup() |>
  ggplot(aes(x = age_bins, y = count_emoji, fill = age_bins)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(count_emoji, "%")), vjust = -0.5) +
  facet_grid(
    cols = vars(report_month),
    rows = vars(sex)
  ) +
  labs(
    title = "FIDE Chess Players by Age Group",
    subtitle = "Proportion of chess players by age group and report month",
    x = "Age Group",
    y = "Proportion (%)"
  ) +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, by = 10),
    labels = function(x) paste0(x, "%")
  )
  
```


```{r}
# Load necessary libraries
library(baffle)
library(withr)
library(fontawesome)
library(ggtext)

# Set a font that supports emojis
# install.packages("extrafont")
# extrafont::font_import()
# You may need to run this once to install fonts on your system

waffle(c(30,70))

par(mfrow=c(1, 3))
waffle(c(25, 75), col=c("red", "gray"))
waffle(c(25, 75), col=c("blue", "gray"), ncol=5)
waffle(c(25, 75), col=c("darkgreen", "gray"), nrow=5)

par(mar=c(4,0,0,0))
cyl = table(mtcars$cyl)
waffle(cyl, stacked=FALSE, gap=1)
legend("top", horiz=TRUE, bty="n", inset=0.9, xpd=TRUE,
    legend=names(cyl), cex=2,
    fill=palette.colors(length(cyl), "Set 1"), border=NA,
    title="Number of cylinders\n in 'mtcars'", title.font=2, title.cex=1.5)

# Group by age bucket and sex, then count the number of players
waffle_data <- chess |>
  filter(!is.na(age_bins)) |>
  count(sex, age_bins, report_month) |>
  group_by(report_month, sex) |>
  mutate(
    total = sum(n),
    prop = n / total,
    count_emoji = round(prop * 100)
  ) |>
  ungroup()

# Define a color palette for the age bins
age_bin_colors <- c("<18" = "#FF7F00", "18-29" = "#4DAF4A", "30-49" = "#377EB8", "50-64" = "#984EA3", "65+" = "#E41A1C")

# Get unique combinations of sex and month
plot_data_list <- waffle_data |>
  group_by(sex, report_month) |>
  group_split()

# Loop through each data subset to create a waffle chart
for (data_subset in plot_data_list) {
  sex_val <- unique(data_subset$sex)
  month_val <- unique(data_subset$report_month)
  
  # The baffle::waffle function accepts a vector of abundancies.
  # We use the 'n' column for the abundances.
  abundances <- data_subset$count_emoji
  names(abundances) <- data_subset$age_bins
  
  # Create a filename
  filename <- paste0("waffle_chart_", sex_val, "_", month_val, ".png")
  
  # Open PNG device, create plot, and then close device
  png(filename, width = 600, height = 600)
  plot_title <- paste("Report Month:", month_val, "| Sex:", sex_val)
  baffle::waffle(abundances,
                #  main = plot_title,
                 col = age_bin_colors[names(abundances)])
  dev.off()
}

# Combine the four PNG files into a 2x2 grid: waffle_chart_F_August.png, waffle_chart_M_August.png, waffle_chart_F_September.png, and waffle_chart_M_September.png
library(patchwork)
library(png)
library(cowplot)
img1 <- readPNG("waffle_chart_F_August.png")
img2 <- readPNG("waffle_chart_M_August.png")
img3 <- readPNG("waffle_chart_F_September.png")
img4 <- readPNG("waffle_chart_M_September.png")

p_blank_fa <- ggplot(
    data = tibble(x = c(0,1,2), y = c(0,1,2)),
    aes(x,y)
  ) + 
  theme_void() +
  ggtitle("Females - August") +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold")
  )
p1 <- ggdraw(p_blank_fa) + draw_image(img1, scale = 0.75)

p_blank_fs <- ggplot(
    data = tibble(x = c(0,1,2), y = c(0,1,2)),
    aes(x,y)
  ) + theme_void() + ggtitle("Females - September") +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold")
  )
p3 <- ggdraw(p_blank_fs) + draw_image(img3, scale = 0.75)

p_blank_ma <- ggplot(
    data = tibble(x = c(0,1,2), y = c(0,1,2)),
    aes(x,y)
  ) + 
  theme_void() +
  ggtitle("Males - August") +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold")
  )
p2 <- ggdraw(p_blank_ma) + draw_image(img2, scale = 0.75)

p_blank_ms <- ggplot(
    data = tibble(x = c(0,1,2), y = c(0,1,2)),
    aes(x,y)
  ) + theme_void() + ggtitle("Males - September") +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold")
  )
p4 <- ggdraw(p_blank_ms) + draw_image(img4, scale = 0.75)

p_legend <- tibble(
  age_bins = names(age_bin_colors),
  colors = age_bin_colors
) |>
  ggplot(aes(x = age_bins, y = 1, fill = age_bins)) +
  geom_tile() +
  scale_fill_manual(values = age_bin_colors) +
  geom_text(aes(y = 0.75, label = age_bins), vjust = -0.5) +
  theme_void() +
  theme(legend.position = "none")

library(ggtext)
combined_plot <- (p1 | p3) / (p2 | p4) / p_legend +
  plot_layout(heights = c(6, 6, 1)) +
  plot_annotation(
    title = "FIDE Chess Players by Age Group",
    subtitle = "Proportion of chess players by age group and report month\nEach square represents 1% of the total players in that group",
    caption = "Data: FIDE | Chart: Tony Galvan (@GDataScience1) | #TidyTuesday"
  )
combined_plot

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter!

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_09_23_tidy_tuesday_chess.png",
  device = "png",
  width = 5,
  height = 7
)
```