---
title: "TidyTemplate"
date: 2025-09-19
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event! Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data. While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format. The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community! As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-09-16")
```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset. This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```

# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

all_recipes <- tt$all_recipes

```

# Summary Statistics

Calculate key summary statistics for the numerical variables. This helps you understand the central tendency, spread, and potential outliers.

```{r skimr}

summary(all_recipes)

library(skimr)
skim(all_recipes)

```

# Missing Values

It's crucial to identify if there are any missing values (NA). The `{visdat}` package is great for visualizing this.

```{r missing}

# Install and load the visdat package
# install.packages("visdat")
library(visdat)

# Visualize missing values
vis_miss(all_recipes)

```

# Correlation Analysis

Investigate the relationship between different numerical variables

```{r}
# Select numerical variables for correlation
numerical_data <- all_recipes %>%
  select(avg_rating, total_ratings, reviews, prep_time, cook_time, total_time, calories, fat, carbs, protein, servings)

# Calculate the correlation matrix
cor_matrix <- cor(numerical_data, use = "pairwise.complete.obs")
print(cor_matrix)

# Visualize the correlation matrix with a heatmap
# install.packages("corrplot")
library(corrplot)
corrplot(cor_matrix, method = "color")
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

# Plot a boxplot of cook times
all_recipes |> 
  filter(!is.na(avg_rating)) |> 
  ggplot(aes(x = avg_rating)) +
  geom_histogram(
    fill = "lightblue"
    , color = "black"
    # , binwidth = 0.5
  ) +
  labs(
    title = "Distribution of Recipe Ratings",
    x = "Rating",
    y = "Count",
    caption = "Source: AllRecipes.com | Graphic: Your Name"
  )

# Plot a boxplot of cook times
all_recipes |>
  ggplot(aes(y = cook_time)) +
  geom_boxplot() +
  labs(title = "Distribution of Cook Times", y = "Cook Time (minutes)")
  
```

# Categorical Variable Analysis

```{r Authors}

# Find the top 10 most prolific authors
top_authors <- all_recipes %>%
  count(author, sort = TRUE) %>%
  top_n(10)

print(top_authors)

# Analyze the average rating of the top authors
author_ratings <- all_recipes %>%
  group_by(author) %>%
  summarise(
    avg_rating = mean(avg_rating, na.rm = TRUE),
    total_recipes = n()
  ) %>%
  filter(total_recipes > 5) %>% # Filter for authors with a minimum number of recipes
  arrange(desc(avg_rating))

print(head(author_ratings))

```

# Time Series Analysis

```{r TimeSeries}

# Count recipes published by year
recipes_by_year <- all_recipes %>%
  mutate(year = year(date_published)) %>%
  group_by(year) %>%
  summarise(count = n())

# Plot recipes published over time
ggplot(recipes_by_year, aes(x = year, y = count)) +
  geom_line() +
  labs(title = "Recipes Published Over Time", x = "Year", y = "Number of Recipes")

```

# Clustering

```{r Clustering}

# Load necessary libraries
library(tidymodels)
# install.packages("tidyclust")
library(tidyclust)

# Create a tidymodels recipe for preprocessing
clustering_recipe <- recipe(~., data = all_recipes) |>
  # Select the numerical variables for clustering
  step_select(
    calories, fat, carbs, protein, prep_time, cook_time, total_time, servings
  ) |>
  # Impute missing numerical data with the median
  step_impute_median(all_numeric_predictors()) |>
  # Normalize the data so all features have a mean of zero and a standard deviation of one
  step_normalize(all_numeric_predictors())

# Define the K-Means clustering model
k_means_spec <- k_means(num_clusters = 3) |>
  set_engine("stats")

# Create a tidymodels workflow
clustering_workflow <- workflow() |>
  add_recipe(clustering_recipe) |>
  add_model(k_means_spec)

# Fit the workflow to the data
clustering_fit <- clustering_workflow |>
  fit(data = all_recipes)

# Augment the original data with cluster assignments
all_recipes_clustered <- augment(clustering_fit, new_data = all_recipes)

# View the distribution of recipes in each cluster
print("Cluster Distribution:")
all_recipes_clustered |>
  count(.pred_cluster)

# View the average values for each cluster to understand the groupings
print("Cluster Characteristics:")
all_recipes_clustered |>
  group_by(.pred_cluster) |>
  summarise(
    avg_calories = mean(calories, na.rm = TRUE),
    avg_fat = mean(fat, na.rm = TRUE),
    avg_protein = mean(protein, na.rm = TRUE),
    avg_carbs = mean(carbs, na.rm = TRUE),
    avg_total_time = mean(total_time, na.rm = TRUE)
  )

# Create a triangular plot to visualize the clusters based on fat, carbs, and protein
all_recipes_clustered |>
  select(fat, carbs, protein, .pred_cluster) |>
  drop_na() |>
  mutate(
    # Calculate the sum of the three components
    S = fat + carbs + protein,
    # Normalize the data to proportions
    fat_norm = fat / S,
    carbs_norm = carbs / S,
    protein_norm = protein / S,
    # Convert the ternary proportions to Cartesian coordinates (x, y)
    x = carbs_norm + 0.5 * protein_norm,
    y = (sqrt(3) / 2) * protein_norm
  ) |>
  ggplot() +
  # Add the points to the plot
  geom_point(aes(x = x, y = y, color = .pred_cluster), size = 0.25, alpha = 0.4) +
  # Manually draw the triangle to serve as a guide for the axes
  geom_path(
    data = data.frame(
      x = c(0, 1, 0.5, 0),
      y = c(0, 0, sqrt(3)/2, 0)
    ),
    aes(x = x, y = y),
    color = "black"
  ) +
  # Add labels for the vertices
  annotate("text", x = 0, y = -0.05, label = "Fat", vjust = 1, size = 5) +
  annotate("text", x = 1, y = -0.05, label = "Carbs", vjust = 1, size = 5) +
  annotate("text", x = 0.5, y = sqrt(3)/2 + 0.02, label = "Protein", vjust = 0, size = 5) +
  # Set plot title and adjust theme for a cleaner look
  labs(title = "Ternary Plot of Recipe Nutritional Content") +
  coord_equal() + # Make sure the axes are equal to maintain the triangle shape
  theme_void()    # Remove all axis and background elements

# Create a facetted version with one triangle per cluster
all_recipes_clustered |>
  select(fat, carbs, protein, .pred_cluster) |>
  drop_na() |>
  mutate(
    # Calculate the sum of the three components
    S = fat + carbs + protein,
    # Normalize the data to proportions
    fat_norm = fat / S,
    carbs_norm = carbs / S,
    protein_norm = protein / S,
    # Convert the ternary proportions to Cartesian coordinates (x, y)
    x = carbs_norm + 0.5 * protein_norm,
    y = (sqrt(3) / 2) * protein_norm
  ) |>
  ggplot() +
  # Add the points to the plot
  geom_point(
    aes(x = x, y = y, color = .pred_cluster), 
    size = 0.25,
    alpha = 0.15
  ) +
  # Manually draw the triangle to serve as a guide for the axes
  geom_path(
    data = data.frame(
      x = c(0, 1, 0.5, 0),
      y = c(0, 0, sqrt(3)/2, 0)
    ),
    aes(x = x, y = y),
    color = "black"
  ) +
  # Add labels for the vertices
  annotate("text", x = 0, y = -0.05, label = "Fat", vjust = 1, size = 2, hjust = 0) +
  annotate("text", x = 1, y = -0.05, label = "Carbs", vjust = 1, size = 2, hjust = 1) +
  annotate("text", x = 0.5, y = sqrt(3)/2 + 0.02, label = "Protein", vjust = 0, size = 2) +
  facet_wrap(~.pred_cluster, nrow = 1) +
  # Set plot title and adjust theme for a cleaner look
  labs(
    title = "Ternary Plot of Recipe Nutritional Content",
    caption = ""
  ) +
  coord_equal() + # Make sure the axes are equal to maintain the triangle shape
  # theme_void() +   # Remove all axis and background elements
  theme(
    legend.position = "none",
    plot.title.position = "panel",
    plot.title = element_text(hjust = 0.5)
  )

```



# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter!

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_09_16_tidy_tuesday_recipes.png",
  device = "png"
)
```
