---
title: "TidyTemplate"
date: 2025-03-14
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(ggrepel)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-03-11")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
library(pixarfilms)
# This package contains six data sets provided mostly in part by Wikipedia.
# 
# pixar_films - released and upcoming films
# pixar_people - main people involved in creating films
# genres - movie genres for each film
# box_office - box office reception and budget information
# public_response - critical and public response
# academy - academy awards and nominations

films <- pixar_films |>
  mutate(
    film = if_else(is.na(film), "Elemental", film)
  )

people <- pixar_people |>
  mutate(
    name = case_when(
      name == "Stanton" ~ "Andrew Stanton",
      name == "Docter" ~ "Pete Docter",
      name == "Lasseter" ~ "John Lasseter",
      name == "Unkrich" ~ "Lee Unkrich",
      name == "Scanlon" ~ "Dan Scanlon",
      name == "Aldrich" ~ "Matthew Aldrich",
      name == "Bird" ~ "Brad Bird",
      name == "Chapman" ~ "Brenda Chapman",
      name == "Cooley" ~ "Josh Cooley",
      name == "Fee" ~ "Brian Fee",
      name == "Folsom" ~ "Stephany Folsom",
      name == "Klubien" ~ "Jorgen Klubien",
      name == "LeFauve" ~ "Meg LeFauve",
      name == "Molina" ~ "Adrian Molina",
      name == "Sohn" ~ "Peter Sohn",
      name == "Ranft" ~ "Joe Ranft",
      name == "Peterson" ~ "Bob Peterson",
      TRUE ~ name
    )
  )

genres |>
  View()

box_office |>
  View()

public_response |>
  View()

academy |>
  View()

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

people |>
  filter(!is.na(name)) |>
  distinct(film, name) |>
  count(name, sort = TRUE) |>
  top_n(10, wt = n) |>
  mutate(name = fct_reorder(name, n)) |>
  ggplot(
    aes(
      x = n,
      y = name,
      fill = name
    )
  ) +
  geom_col(show.legend = FALSE) +
  scale_x_continuous(
    breaks = seq(0, 10, 2)
  ) +
  labs(
    x = "# of films",
    y = "",
    title = "Top 11 People with the Most Pixar Film Credits",
    caption = "Data: Wikipedia  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )

films |>
  ggplot(aes(release_date, run_time)) +
  geom_smooth(
    method = "loess", 
    se = FALSE,
    alpha = 0.05,
    linetype = 2
  ) +
  geom_point() +
  geom_text_repel(aes(label = film)) +
  expand_limits(y = 0) +
  labs(
    x = "Release Date",
    y = "Run Time (minutes)",
    title = "Pixar Film Run Time Over Time",
    caption = "Data: Wikipedia  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )

genres |>
  count(genre) |>
  mutate(genre = fct_reorder(genre, n)) |>
  ggplot(aes(n, genre, fill = genre)) +
  geom_col(show.legend = FALSE) +
  labs(
    x = "# of Films",
    y = "",
    title = "Pixar Film Genre Distribution",
    caption = "Data: Wikipedia  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )

box_office |>
  na.omit() |>
  inner_join(films, by = "film") |>
  select(film, budget, box_office_worldwide, release_date) |>
  pivot_longer(
    cols = c(budget, box_office_worldwide),
    names_to = "metric",
    values_to = "value"
  ) |>
  mutate(metric = case_when(
    metric == "budget" ~ "Budget",
    metric == "box_office_worldwide" ~ "Box Office"
  )) |>
  ggplot(aes(release_date, value, color = metric)) +
  geom_smooth(
    method = "loess",
    se = FALSE,
    linetype = 2
  ) +
  geom_point(
    alpha = 0.25,
    show.legend = FALSE
  ) +
  geom_text_repel(
    aes(label = film),
    angle = 270,
    show.legend = FALSE
  ) +
  scale_y_log10(
    labels = scales::dollar_format(
      scale = 1e-6, 
      suffix = "M", 
      accuracy = 1
    )
  ) +
  labs(
    x = "Release Date",
    y = "Amount (USD)",
    color = "",
    title = "Pixar Film Budget and Box Office Worldwide Over Time",
    caption = "Data: Wikipedia  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  ) +
  theme(
    legend.position = "top"
  )

public_response |>
  select(-cinema_score) |>
  na.omit() |>
  pivot_longer(
    cols = -film,
    names_to = "metric",
    values_to = "score"
  ) |>
  group_by(film) |>
  mutate(med_score = mean(score)) |>
  ungroup() |>
  inner_join(films |> select(film, release_date), by = "film") |> 
  mutate(film = fct_reorder(film, score)) |> 
  ggplot(aes(release_date, score, group = release_date, fill = film)) +
  geom_boxplot(
    show.legend = FALSE
  ) +
  geom_text(
    aes(y = med_score, label = film, color = film),
    angle = 270,
    nudge_y = -7.5,
    hjust = 0,
    show.legend = FALSE
  ) +
  labs(
    x = "Release Date",
    y = "Score",
    title = "Pixar Film Public Response Over Time",
    caption = "Data: Wikipedia  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )
  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_03_11_tidy_tuesday_pixar.png",
  device = "png"
)
```
