---
title: "TidyTemplate"
date: 2026-01-05
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
# library(tidytuesdayR)
library(scales)
library(patchwork)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
# tt <- tt_load("2026-01-06")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
# tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
# tt %>% 
#   map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set. The data was downloaded from - https://datausa.io/profile/university/university-of-notre-dame

```{r Wrangle}

tuition <- read_csv("~/Downloads/tuition.csv") |>
  rename(tuition = `State Tuition`) |>
  left_join(
    read_csv("~/Downloads/net_price.csv") |>
      transmute(
        University,
        Year,
        net_price = `Median Average Net Price Grant Or Scholarship Aid`
      ),
    by = c("University", "Year")
  ) |>
  janitor::clean_names() |>
  mutate(
    grants_and_loans = tuition - net_price
  ) |>
  filter(!university_id %in% c("139755", "15", "DOC"))

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

# Combine the three plots telling the story about the Pathways to Notre Dame" initiative. This means the university offers a no-loan financial aid policy, covering full demonstrated need with grants, scholarships, and work-study, eliminating student loans from aid packages for admitted students, supported by need-blind admissions for all students. This commitment ensures affordability for low- and middle-income families, making a Notre Dame education debt-free for many.

# 1. Tuition Costs Plot
p1 <- tuition |>
  ggplot(
    aes(
      x = factor(year),
      y = tuition,
      color = university,
      linewidth = university == "University of Notre Dame",
      group = university
    )
  ) +
  geom_line() +
  scale_y_continuous(labels = dollar_format()) +
  scale_linewidth_manual(
    values = c(`TRUE` = 1.5, `FALSE` = 0.5),
    guide = "none"
  ) +
  scale_color_manual(
    values = c(
      "University of Notre Dame" = "navy",
      "Georgetown University" = "gray80",
      "Case Western Reserve University" = "lightblue",
      "Emory University" = "gold",
      "Northeastern University" = "red"
    )
  ) +
  labs(
    title = "Tuition Costs Over Time",
    x = NULL, # Remove x-axis label for top plots to reduce clutter
    y = "Tuition ($)",
    color = "University"
  ) +
  theme_minimal()

# 2. Net Price Plot
p2 <- tuition |>
  ggplot(
    aes(
      x = factor(year),
      y = net_price,
      color = university,
      linewidth = university == "University of Notre Dame",
      group = university
    )
  ) +
  geom_line() +
  scale_y_continuous(labels = dollar_format()) +
  scale_linewidth_manual(
    values = c(`TRUE` = 1.5, `FALSE` = 0.5),
    guide = "none"
  ) +
  scale_color_manual(
    values = c(
      "University of Notre Dame" = "navy",
      "Georgetown University" = "gray80",
      "Case Western Reserve University" = "lightblue",
      "Emory University" = "gold",
      "Northeastern University" = "red"
    )
  ) +
  labs(
    title = "Net Price Over Time",
    x = NULL,
    y = "Net Price ($)",
    color = "University"
  ) +
  theme_minimal()

# 3. Grants and Loans Plot
p3 <- tuition |>
  ggplot(
    aes(
      x = factor(year),
      y = grants_and_loans,
      color = university,
      linewidth = university == "University of Notre Dame",
      group = university
    )
  ) +
  geom_line() +
  scale_y_continuous(labels = dollar_format()) +
  scale_linewidth_manual(
    values = c(`TRUE` = 1.5, `FALSE` = 0.5),
    guide = "none"
  ) +
  scale_color_manual(
    values = c(
      "University of Notre Dame" = "navy",
      "Georgetown University" = "gray80",
      "Case Western Reserve University" = "lightblue",
      "Emory University" = "gold",
      "Northeastern University" = "red"
    )
  ) +
  labs(
    title = "Grants and Loans Over Time",
    x = "Year",
    y = "Grants/Loans ($)",
    color = "University"
  ) +
  theme_minimal()

# Combine plots using patchwork
combined_plot <- (p1 / p2 / p3) + 
  plot_layout(guides = "collect") + # One legend for all plots
  plot_annotation(
    title = "Pathways to Notre Dame: A Commitment to Debt-Free Education",
    subtitle = "By eliminating student loans and covering full demonstrated need through grants and scholarships,\nNotre Dame ensures affordability regardless of rising tuition costs.",
    caption = "Data Source: University Financial Records | *Highlighted line represents University of Notre Dame",
    theme = theme(
      plot.title = element_text(size = 16, face = "bold", color = "navy"),
      plot.subtitle = element_text(size = 11)
    )
  )

# Display the combined plot
combined_plot

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2026_01_06_tidy_tuesday_nd.png",
  device = "png"
)
```
