---
title: "bird_collisions"
author: "Tony Galvan"
date: "April 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chicago Bird Collisions
### Tidy Tuesday 2019 Week 18

```{r}
library(tidyverse)

theme_set(theme_minimal())

bird_collisions <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-30/bird_collisions.csv")
mp_light <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-30/mp_light.csv")
```

#### EDA

```{r}
glimpse(bird_collisions)
```

```{r}
summary(bird_collisions$date)
```

The "bird_collisions" data contains observations of birds colliding with various lighted structures in Chicago from 9/15/1978 to 11/30/2016.

```{r}
mp_light %>%
  summary()
```

The "mp_light" data contains light scores (proportion of the 17 window bays that were illuminated) at McCormic Place (locality = "MP" from the "bird_collisions" data) in Chicago between 3/6/2000 and 5/26/2018.

Let's plot this data to see if there is a pattern based on the day of the week.

```{r}
mp_light %>%
  mutate(weekday = lubridate::wday(date, label = TRUE)) %>%
  group_by(weekday) %>%
  summarize(pct_light_score = mean(light_score)/100) %>%
  ggplot(aes(weekday, pct_light_score, fill = weekday)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Weekday", 
       y = "Average % of Window Bays Illuminated",
       title = "Thursday was the most illuminated day of the week",
       subtitle = "McCormick Place Between 3/6/2000 and 5/26/2018")
```

Let's plot this data to see if there is a pattern based on the month.

```{r}
mp_light %>%
  mutate(month = lubridate::month(date, label = TRUE)) %>%
  group_by(month) %>%
  summarize(pct_light_score = mean(light_score)/100) %>%
  ggplot(aes(month, pct_light_score, fill = month)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Month", 
       y = "Average % of Window Bays Illuminated",
       title = "September was the most illuminated month",
       subtitle = "McCormick Place Between 3/6/2000 and 5/26/2018")
```

Let's join the data together.

```{r}
mp_bird_collisions <- bird_collisions %>%
  filter(locality == "MP") %>%
  left_join(mp_light) %>%
  filter(date > "2000-03-05")

summary(mp_bird_collisions$date)
```

```{r}
mp_bird_collisions %>%
  count(date, light_score, sort = TRUE)
```


