# Federal Research and Development Spending by Agency
```{r}
library(tidyverse)
library(plotly)

theme_set(theme_light())

fed_rd <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-12/fed_r_d_spending.csv")
```
## Total Federal R&D Spending by agency/deparment
First, let's look at the data from 'fed_r_d_spending.csv'
```{r}
fed_rd %>% 
    summary()
```
This data contains 588 total observations from 1976 to 2017.

Let's do some pre-processing of the data:

* Add a new variable to compute the percent rd_budget vs. GDP
* Add a new variable to compute the percent rd_budget vs. total_outlays
* Add a new variable to compute the percent rd_budget vs. discretionary_outlays
```{r}
fed_rd_processed <- fed_rd %>%
    mutate(rd_pct_gdp = rd_budget / gdp, 
            rd_pct_tot = rd_budget / total_outlays, 
        rd_pct_dis = rd_budget / discretionary_outlays)
```
First, let's plot the percentage of GDP made up of research and development colored by department
```{r}
p <- fed_rd_processed %>% 
    ggplot(aes(year, rd_pct_gdp, color = department)) +
    geom_line() + 
    scale_y_continuous(labels = scales::percent_format())
    
ggplotly(p) 
```
Next, let's plot the percentage of total outlays made up of research and development colored by department
```{r}
p <- fed_rd_processed %>% 
    ggplot(aes(year, rd_pct_dis, color = department)) +
    geom_line() + 
    scale_y_continuous(labels = scales::percent_format())
    
ggplotly(p) 
```
