---
title: "D'oh! 30 Seasons of Simpsons Guest Stars — Who Showed Up the Most?"
author: "Tony Galvan"
date: "8/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

*The Simpsons* has been on the air for over 30 seasons — and in that time, it's hosted an extraordinary parade of guest stars. From musicians to politicians to fellow actors, the show has become a cultural milestone that celebrities actively seek out. But has the guest star phenomenon grown over time? And which voice actors have played the most unique characters? Let's dig into the data from Wikipedia's comprehensive guest star records.

## Loading the Data

The data uses a pipe delimiter (`|`) rather than the typical comma, reflecting its Wikipedia table origins.

```{r}
# Load libraries, set the default theme & caption, and grab the data
library(tidyverse)
library(patchwork)
theme_set(theme_light())

default_caption <- "Source: Wikipedia  |  Designer: Tony Galvan @gdatascience1  |  #TidyTuesday 2019 Week 35"

simpsons <- readr::read_delim("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-27/simpsons-guests.csv", delim = "|", quote = "")

glimpse(simpsons)
```

## Guest Stars Per Season: A Growing Trend

Has the show relied more heavily on guest stars as it's aged? Let's plot the count per season with a trend line.

```{r}
p1 <- simpsons |>
  filter(season != "Movie") |>
  group_by(season) |>
  summarise(n = n()) |>
  mutate(season = as.integer(str_replace(season, "Movie", "0"))) |>
  ggplot(aes(season, n)) +
  geom_line(color = "#FED90F", size = 3) +
  scale_x_continuous(breaks = c(1, 10, 20, 30), labels = c(1, 10, 20, 30)) +
  geom_smooth(method = "lm", se = FALSE, 
              color = "#70D1FE", linetype = 5, size = 2) +
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),) +
  labs(x = "Season",
       y = "# of guest stars",
       title = "The Simpsons: guest stars per season",
       subtitle = "Seasons 1 - 30",
       caption = default_caption)

p1
```

The trend is unmistakable — guest star appearances have roughly tripled from the early seasons to the most recent ones. As the show's cultural cachet grew, so did the celebrity guest list.

## Who Played the Most Unique Roles?

Some guest stars appear once as themselves; others return repeatedly to voice different characters. Let's find the most versatile guest performers.

```{r}
roles <- simpsons |>
  unnest(role = strsplit(role, ";")) |>
  mutate(role_name = if_else(role %in% c("Himself", "Herself", "Themselves"), guest_star, role))

glimpse(roles)
```

```{r}
p2 <- roles |>
  group_by(role_name, guest_star) |>
  summarise(n = n()) |>
  group_by(guest_star) |>
  summarise(star_roles = n()) |>
  top_n(10, wt = star_roles) |>
  mutate(guest_star = fct_reorder(guest_star, star_roles)) |>
  ggplot(aes(guest_star, star_roles)) +
  geom_col(fill = "#FED90F") + 
  coord_flip() +
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),) +
  labs(x = "",
       y = "# of unique roles",
       title = "Top 10 guest stars with the most unique roles",
       subtitle = "The Simpsons: seasons 1 - 30",
       caption = default_caption)

p2
```

Maurice LaMarche leads the pack — a professional voice actor known for his versatility, he's voiced dozens of distinct characters across the show's run.

## Combined View: Growth and Versatility

Let's combine both visualizations into a single panel that tells the complete story.

```{r}
p1 + 
  labs(title = "The Simpsons: seasons 1 - 30",
       subtitle = "More guest stars per season; Maurice LaMarche has played the most unique roles",
       caption = "") +
  p2 +
  labs(title = "",
       subtitle = "")

#ggsave("outputs/2019_08_27_tidy_tuesday_simpsons.png", width = 10)
```

Together, these charts paint a picture of a show that has increasingly relied on celebrity guest appearances — and a handful of versatile voice actors who have become unofficial members of the extended cast.
