---
title: "TidyTemplate"
date: 2025-08-21
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(sf)
library(maps)
# library(rgl)
library(rayshader)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-08-19")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  purrr::map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

munros_raw <- tt$scottish_munros

munros <-
  munros_raw |>
  pivot_longer(
    cols = starts_with(c("1", "2")),
    names_to = "year",
    values_to = "type"
  ) |>
  filter(!is.na(type)) |>
  janitor::clean_names()

munros_map_df <- munros |>
  count(xcoord, ycoord, name, height_ft, name = "observations") |>
  arrange(height_ft)

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

two_d_plot <- munros_map_df |>
  ggplot(
    aes(
      x = xcoord,
      y = ycoord
    )
  ) +
  geom_point(
    aes(
      size = observations,
      color = height_ft
    )
  ) +
  scale_color_gradient(low = "darkgreen", high = "yellow") +
  labs(
    title = "Scottish Munros",
    x = NULL,
    y = NULL,
    size = "Observations",
    color = "Height (ft)",
    caption = "Source: Database of British and Irish hills  |  Created by: Tony Galvan (@GDataScience)  |  #TidyTuesday"
  ) +
  coord_fixed() +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot",
    plot.title = element_text(hjust = 0.5, size = 20),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    plot.caption = element_text(hjust = 0.5, size = 5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 8)
  )

two_d_plot

plot_gg(
  two_d_plot, 
  width = 4, 
  height = 5,
  multicore = TRUE, 
  windowsize = c(800, 800), 
  # zoom = 0.85, 
  # phi = 35, 
  # theta = 30, 
  # sunangle = 225, 
  # soliddepth = -1
)
  
```


```{r Map}

# Convert the data frame to an sf object
munros_sf <- st_as_sf(munros_map_df, coords = c("xcoord", "ycoord"), crs = 27700)

# Transform the CRS of your sf object to WGS84 (lat/long)
# This is the crucial step to match the map data
munros_sf_wgs84 <- st_transform(munros_sf, crs = 4326)

# Define the original OSGB36 bounding box coordinates
osgb_lon_min <- min(munros_map_df$xcoord)
osgb_lon_max <- max(munros_map_df$xcoord)
osgb_lat_min <- min(munros_map_df$ycoord)
osgb_lat_max <- max(munros_map_df$ycoord)

# Create an sf bounding box from these OSGB36 coordinates
bbox_sf <- st_bbox(c(xmin = osgb_lon_min, xmax = osgb_lon_max, ymin = osgb_lat_min, ymax = osgb_lat_max), crs = 27700)
bbox_polygon <- st_as_sfc(bbox_sf)

# Transform the bounding box from OSGB36 to WGS84 (longitude/latitude)
bbox_wgs84 <- st_transform(bbox_polygon, crs = 4326)
bbox_wgs84_coords <- st_bbox(bbox_wgs84)

# Get the new WGS84 longitude and latitude limits with a buffer
lon_min <- bbox_wgs84_coords$xmin - (bbox_wgs84_coords$xmax - bbox_wgs84_coords$xmin) * 0.5
lon_max <- bbox_wgs84_coords$xmax + (bbox_wgs84_coords$xmax - bbox_wgs84_coords$xmin) * 0.5
lat_min <- bbox_wgs84_coords$ymin - (bbox_wgs84_coords$ymax - bbox_wgs84_coords$ymin) * 0.05
lat_max <- bbox_wgs84_coords$ymax + (bbox_wgs84_coords$ymax - bbox_wgs84_coords$ymin) * 0.05

# Get the world polygon and extract the UK
uk <- map_data("world") |>
  filter(
    region == "UK",
  )

p <- munros_sf_wgs84 |>
  ggplot() +
  geom_polygon(
    data = uk,
    aes(x = long, y = lat, group = group), 
    fill="grey55"
  ) +
  geom_point(
    aes(
      x = st_coordinates(munros_sf_wgs84)[, 1], 
      y = st_coordinates(munros_sf_wgs84)[, 2],
      size = observations,
      color = height_ft
    )
  ) +
  scale_color_gradient(low = "darkgreen", high = "yellow") +
  # make the points smaller
  scale_size(range = c(0.1, 2.5)) +
  coord_map(
    xlim = c(lon_min, lon_max), 
    ylim = c(lat_min, lat_max)
  ) +
  ggthemes::theme_map() +
  labs(
    title = "Scottish Munros",
    x = NULL,
    y = NULL,
    size = "Observations",
    color = "Height (ft)",
    caption = "Source: Database of British and Irish hills  |  Created by: Tony Galvan (@GDataScience)  |  #TidyTuesday"
  ) +
  theme(
    legend.position = "left",
    panel.background = element_rect(
      fill = "lightblue",
      color = NA
    ),
    plot.background = element_rect(
      fill = "white",
      color = NA
    ),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    plot.title = element_text(hjust = 0.5, size = 20),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    plot.caption = element_text(hjust = 0.5, size = 8),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 8)
  )

p

# Convert to 3D
plot_gg(
  p,
  width = 6,
  height = 4,
  multicore = TRUE,
  windowsize = c(1200, 800),
)

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_08_19_tidy_tuesday_monros.png",
  device = "png"
)
```
