---
title: "Community Pulse: Tracking the R4DS Slack Community's Growth and Engagement"
author: "Tony Galvan"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

The R4DS (R for Data Science) Online Learning Community started as a Slack workspace for people working through Hadley Wickham's book — and grew into one of the most active R communities on the internet. With daily membership and activity stats from August 2017 through July 2019, we can trace the community's growth patterns and identify when members are most engaged. The seasonal rhythms tell a story about how people learn data science.

## Loading and Preparing the Data

We'll filter to the period after the community's founding, create time-based features, and assign seasons for seasonal analysis.

```{r}
library(tidyverse)
library(lubridate)
library(ggridges)
theme_set(theme_light())

r4ds_members <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-16/r4ds_members.csv") |>
  filter(date > '2017-08-31') |>
  select(-guests, -name, -messages_in_shared_channels) |>
  mutate(month = factor(month(date, label = TRUE)),
         day = factor(wday(date, label = TRUE)), 
         year = factor(year(date)),
         season = case_when(
           date < '2017-09-22' ~ "Summer",
           date >= '2017-09-22' & date < '2017-12-21' ~ "Fall",
           date >= '2017-12-21' & date < '2018-03-20' ~ "Winter",
           date >= '2018-03-20' & date < '2018-06-21' ~ "Spring",
           date >= '2018-06-21' & date < '2018-09-22' ~ "Summer",
           date >= '2018-09-22' & date < '2018-12-21' ~ "Fall",
           date >= '2018-12-21' & date < '2019-03-20' ~ "Winter",
           date >= '2019-03-20' & date < '2019-06-21' ~ "Spring",
           date >= '2019-06-21' ~ "Summer"))
```

## Data Overview

```{r}
r4ds_members |>
  summary()
```

This data covers the R4DS community from its founding through early July 2019 (8/27/2017 to 7/5/2019).

## Daily Active Members by Month

Let's explore engagement patterns — which months see the most activity?

```{r}
r4ds_members |>
  group_by(month, year) |>
  summarise(total_active_members = sum(daily_active_members)) |>
  ggplot(aes(month, total_active_members, fill = month)) +
  geom_col(show.legend = FALSE) +
  labs(x = "",
       y = "# of daily active members",
       title = "Daily active R4DS members by month",
       subtitle = "Fewer active members the second half of the year",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: R4DS Slack")
```

The first half of the year consistently shows higher engagement — perhaps because New Year's resolutions drive people to learn new skills, or because academic semesters align with peak activity.

## Ridgeline Plot: Activity Distribution by Month

A ridgeline plot shows the full distribution of daily active members for each month, revealing not just averages but the spread of activity levels.

```{r}
r4ds_members |>
  ggplot(aes(daily_active_members, fct_rev(month))) + 
  geom_density_ridges(aes(fill = year), alpha = 0.5) + 
  theme_ridges(grid = FALSE) + 
  labs(x = "",
       y = "",
       title = "R4DS daily active members by month",
       fill = "Year",
       subtitle = "Fewer active members the second half of the year",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: R4DS Slack") +
  theme(legend.position = "top")
```

The 2019 distributions (shown in a different color) are shifted right compared to 2018, confirming that the community grew year-over-year even as seasonal patterns persisted.

## Seasonal Engagement Patterns

Grouping by season makes the pattern even clearer — when do R learners engage most?

```{r}
r4ds_members |>
  group_by(season) |>
  summarise(avg_active_members = sum(daily_active_members) / n()) |>
  mutate(season = fct_reorder(season, avg_active_members)) |>
  ggplot(aes(season, avg_active_members, fill = season)) +
  geom_col(show.legend = FALSE) +
  coord_flip() + 
  labs(x = "",
       y = "Avg. daily active members",
       title = "R4DS daily active members by season",
       subtitle = "Fewer active members in Fall",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: R4DS Slack")
```

Fall shows the lowest engagement — perhaps because the initial motivation from earlier in the year has faded, or because people are busy with back-to-school and work ramp-ups.

## Total Membership Growth

Finally, let's look at the raw growth trajectory — how has total membership evolved since the community's founding?

```{r}
r4ds_members |>
  ggplot(aes(date, total_membership)) +
  geom_line()
```

The growth is remarkably steady and linear — the community adds members at a consistent rate, suggesting strong word-of-mouth and ongoing relevance of the R4DS book as a learning resource.
