---
title: "Who Shot First? Exploring Star Wars Fandom Through Data"
author: "Tony Galvan"
date: "8/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

In 2014, FiveThirtyEight surveyed over 1,000 Americans about their Star Wars opinions — which films they'd seen, how they'd rank them, and how they felt about every major character from Han Solo to Jar Jar Binks. The results paint a fascinating portrait of a fandom divided by generation, geography, and gender. Let's dig in.

## Setting Up and Loading the Data

We'll pull the survey data directly from the TidyTuesday GitHub repository and clean up the column names. The raw dataset has cryptic headers from the survey tool, so we rename them to something human-readable.

```{r}
library(tidyverse)
theme_set(theme_light())

star_wars_data <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-05-14/week7_starwars.csv", show_col_types = FALSE, locale = readr::locale(encoding = "latin1"))


star_wars <- star_wars_data |>
  slice(2:1187) |>
  rename_at(2, ~"Any_seen") |> 
  rename_at(3, ~"Star_Wars_fan") |> 
  rename_at(4, ~"Ep1_seen") |> 
  rename_at(5, ~"Ep2_seen") |> 
  rename_at(6, ~"Ep3_seen") |> 
  rename_at(7, ~"Ep4_seen") |> 
  rename_at(8, ~"Ep5_seen") |> 
  rename_at(9, ~"Ep6_seen") |> 
  rename_at(10, ~"Ep1_rating") |> 
  rename_at(11, ~"Ep2_rating") |> 
  rename_at(12, ~"Ep3_rating") |> 
  rename_at(13, ~"Ep4_rating") |> 
  rename_at(14, ~"Ep5_rating") |> 
  rename_at(15, ~"Ep6_rating") |> 
  rename_at(16, ~"Han_Solo") |> 
  rename_at(17, ~"Luke_Skywalker") |> 
  rename_at(18, ~"Princess_Leia") |> 
  rename_at(19, ~"Anakin_Skywalker") |> 
  rename_at(20, ~"Obi_Wan") |> 
  rename_at(21, ~"Emperor_Palpatine") |> 
  rename_at(22, ~"Darth_Vader") |> 
  rename_at(23, ~"Lando_Calrissian") |> 
  rename_at(24, ~"Boba_Fett") |> 
  rename_at(25, ~"C_3PO") |> 
  rename_at(26, ~"R2D2") |> 
  rename_at(27, ~"Jar_Jar_Binks") |> 
  rename_at(28, ~"Padme_Amidala") |> 
  rename_at(29, ~"Yoda") |> 
  rename_at(30, ~"Shot_first") |> 
  rename_at(31, ~"Exp_Uni_Aware") |> 
  rename_at(32, ~"Exp_Uni_Fan") |> 
  rename_at(33, ~"Star_Trek_Fan") |> 
  rename_at(36, ~"Income") |> 
  rename_at(38, ~"Location")

glimpse(star_wars)
```

## Where in America Are the Biggest Star Wars Fans?

First, let's look at fandom by geography. Are some regions of the country more likely to claim Star Wars allegiance than others?

```{r}
star_wars |>
  filter(!is.na(Location)) |>
  group_by(Location) |>
  summarise(n = n(),
            fans = sum(Star_Wars_fan == "Yes", na.rm = TRUE),
            pct_fans = fans / n) |>
  mutate(Location = fct_reorder(Location, pct_fans)) |>
  ggplot(aes(Location, pct_fans, fill = Location)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(x = "",
       y = "% of respondants",
       title = "Where are Star Wars fans located?",
       subtitle = "Over 60% of respondants in the Mountain region are fans",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FiveThirtyEight")
```

The Mountain region leads the pack with over 60% of respondents identifying as fans. The South Atlantic and Pacific regions aren't far behind. Fandom is remarkably widespread — even the lowest region still clocks in above 45%.

## Does Education Level Predict Fandom?

Next, let's see whether education correlates with Star Wars enthusiasm. Do people with more schooling tend to be bigger fans?

```{r}
star_wars |>
  filter(!is.na(Education)) |>
  group_by(Education) |>
  summarise(n = n(),
            fans = sum(Star_Wars_fan == "Yes", na.rm = TRUE),
            pct_fans = fans / n) |>
  mutate(Education = fct_reorder(Education, pct_fans)) |>
  ggplot(aes(Education, pct_fans, fill = Education)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(x = "",
       y = "% of respondants",
       title = "Does more education make you more likely to be a Star Wars fan?",
       subtitle = "Over 50% of respondants with at least some college are fans",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FiveThirtyEight")
```

There's a mild positive trend — respondents with at least some college education are more likely to identify as fans. But the differences are modest, suggesting Star Wars truly cuts across educational backgrounds.

## Which Episodes Have People Actually Seen?

Now let's look at viewership. The original trilogy came out decades before the prequels — does that show up in how many people have seen each film?

```{r}
star_wars |>
  select(RespondentID, Ep1_seen, Ep2_seen, Ep3_seen, Ep4_seen, Ep5_seen, Ep6_seen) |>
  gather(episode, seen, -RespondentID) |>
  mutate(episode = regmatches(episode, regexpr("[[:digit:]]+", episode)),
         seen = if_else(!is.na(seen),1,0)) |>
  group_by(episode) |>
  summarise(n = n(),
            pct_seen = sum(seen) / n) |>
  ggplot(aes(episode, pct_seen, fill = episode)) + 
  geom_col(show.legend = FALSE) +
  coord_flip()+
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Star Wars episode #",
       y = "% of respondants",
       title = "Which Star Wars episode was seen the most?",
       subtitle = "Over 60% of respondants saw Empire Strikes Back or Return of the Jedi",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FiveThirtyEight")
```

Episodes 5 and 6 — *The Empire Strikes Back* and *Return of the Jedi* — edge out the rest, with over 60% of respondents having seen them. The prequels (Episodes 1-3) lag behind, which makes sense given the survey's age demographics.

## How Do Fans Rate Each Episode?

Seeing a movie is one thing — liking it is another. Let's look at average ratings across the six films.

```{r}
star_wars |>
  select(RespondentID, Ep1_rating, Ep2_rating, Ep3_rating, Ep4_rating, Ep5_rating, Ep6_rating) |>
  gather(episode, rating, -RespondentID) |>
  mutate(episode = regmatches(episode, regexpr("[[:digit:]]+", episode)),
         rating = 7 - as.numeric(rating)) |>
  group_by(episode) |>
  summarise(avg_rating = mean(rating, na.rm = TRUE)) |>
  ggplot(aes(episode, avg_rating, fill = episode)) + 
  geom_col(show.legend = FALSE) +
  coord_flip()+
  labs(x = "Star Wars episode #",
       y = "Average rating",
       title = "Which Star Wars episode has the most favorable rating?",
       subtitle = "Episodes 4 - 6 are more highly rated than 1 - 3",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FiveThirtyEight")
```

The original trilogy dominates here. Episodes 4-6 are consistently rated higher than the prequels, with *The Empire Strikes Back* (Episode 5) taking the crown. The prequel trilogy's lower ratings reflect a well-known divide in the fanbase.

## The Character Popularity Contest

Star Wars is ultimately about its characters. Let's rank them all by average favorability rating — from the galaxy's most beloved to its most despised.

```{r}
star_wars |>
  select(RespondentID, Han_Solo, Luke_Skywalker, Princess_Leia, 
         Anakin_Skywalker, Obi_Wan, Emperor_Palpatine, Darth_Vader, 
         Lando_Calrissian, Boba_Fett, C_3PO, R2D2, Jar_Jar_Binks, Padme_Amidala, 
         Yoda) |>
  gather(character, rating, -RespondentID) |>
  filter(!is.na(rating) & rating != "Unfamiliar (N/A)") |>
  mutate(character = str_replace_all(character, "_", " "),
         rating = as.numeric(if_else(rating == "Very favorably", 5, 
                          if_else(rating == "Somewhat favorably", 4, 
                                  if_else(rating == "Neither favorably nor unfavorably (neutral)", 3, 
                                          if_else(rating == "Somewhat unfavorably", 2, 1)))))) |>
  group_by(character) |>
  summarise(avg_rating = mean(rating, na.rm = TRUE)) |>
  mutate(character = fct_reorder(character, avg_rating)) |>
  ggplot(aes(character, avg_rating, fill = character)) + 
  geom_col(show.legend = FALSE) +
  coord_flip()+
  labs(x = "",
       y = "Average rating",
       title = "Han Solo is the highest rated Star Wars character",
       subtitle = "Jar Jar Binks is the lowest rated character",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FiveThirtyEight")
```

Han Solo reigns supreme — the scruffy-looking nerf herder is America's favorite. Yoda, Obi-Wan, and the droids round out the top tier. At the bottom? Jar Jar Binks, unsurprisingly, with Emperor Palpatine not far above. Even villains like Darth Vader and Boba Fett earn more goodwill than poor Jar Jar.

## Do Men and Women See These Characters Differently?

Finally, let's explore whether gender shapes character preferences. We'll calculate the difference between male and female average ratings for each character — positive values mean men rate the character higher.

```{r}
star_wars |>
  select(RespondentID, Han_Solo, Luke_Skywalker, Princess_Leia, 
         Anakin_Skywalker, Obi_Wan, Emperor_Palpatine, Darth_Vader, 
         Lando_Calrissian, Boba_Fett, C_3PO, R2D2, Jar_Jar_Binks, Padme_Amidala, 
         Yoda, Gender) |>
  gather(character, rating, -RespondentID, -Gender) |>
  filter(!is.na(rating) & rating != "Unfamiliar (N/A)" & !is.na(Gender)) |>
  mutate(character = str_replace_all(character, "_", " "),
         rating = as.numeric(if_else(rating == "Very favorably", 5, 
                          if_else(rating == "Somewhat favorably", 4, 
                                  if_else(rating == "Neither favorably nor unfavorably (neutral)", 3, 
                                          if_else(rating == "Somewhat unfavorably", 2, 1)))))) |>
  group_by(character, Gender) |>
  summarise(avg_rating = mean(rating, na.rm = TRUE)) |>
  ungroup() |>
  spread(Gender, avg_rating) |>
  mutate(rating_dif = Male - Female,
         character = fct_reorder(character, rating_dif)) |>
  ggplot(aes(character, rating_dif, fill = character)) + 
  geom_col(show.legend = FALSE) +
  coord_flip()+
  labs(x = "",
       y = "Difference between male & female ratings",
       title = "Do Star Wars character ratings differ by gender?",
       subtitle = "Males like bad guys more. Females like Jar Jar Binks and droids more.",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FiveThirtyEight")
```

The gender gap is revealing. Men rate villains and morally ambiguous characters (Boba Fett, Emperor Palpatine, Darth Vader) more favorably than women do. Women, on the other hand, rate the droids (R2D2, C-3PO) and — surprisingly — Jar Jar Binks higher than men. Perhaps the slapstick humor lands differently across genders.

## Saving the Final Visualization

```{r}
ggsave("star_wars.png", width = 6.25)
```
