---
title: "TidyTemplate"
date: 2025-09-05
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-09-02")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

frogs <- tt$frogID_data |>
  janitor::clean_names() |>
  # Create a Season variable based on the Australian date
  mutate(
    season = case_when(
      month(event_date) %in% c(12, 1, 2) ~ "Summer",
      month(event_date) %in% c(3, 4, 5) ~ "Autumn",
      month(event_date) %in% c(6, 7, 8) ~ "Winter",
      month(event_date) %in% c(9, 10, 11) ~ "Spring"
    )
  )

# Create a data frame of major cities in Australia with their latitudes and longitudes
major_cities <- data.frame(
  city = c("Sydney", "Melbourne", "Brisbane", "Perth", "Adelaide"),
  lat = c(-33.8688, -37.8136, -27.4698, -31.9505, -34.9285),
  lon = c(151.2093, 144.9631, 153.0251, 115.8605, 138.6007)
)

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

# create a map of the Australian frogs
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  geom_point(aes(x = decimal_longitude, y = decimal_latitude, color = season), alpha = 0.5) +
  theme_minimal() +
  labs(
    title = "Frog Observations in Australia by Season",
    x = "Longitude",
    y = "Latitude",
    color = "Season"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom"
  )

# create a density map of the Australian frogs
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  stat_density_2d(aes(x = decimal_longitude, y = decimal_latitude, fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(
    title = "Density of Frog Observations in Australia",
    x = "Longitude",
    y = "Latitude",
    fill = "Density"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom"
  )

# create a density map of the Australian frogs by season
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  stat_density_2d(aes(x = decimal_longitude, y = decimal_latitude, fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  theme_minimal() +
  labs(
    title = "Density of Frog Observations in Australia by Season",
    x = "Longitude",
    y = "Latitude",
    fill = "Density"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom"
  )

# create a hex density map of the Australian frogs by season
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  geom_hex(aes(x = decimal_longitude, y = decimal_latitude), bins = 20) +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  theme_minimal() +
  labs(
    title = "Hex Density of Frog Observations in Australia by Season",
    x = "Longitude",
    y = "Latitude",
    fill = "Count"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom"
  )

# create a hex density map of the Australian frogs by Season with major cities marked

frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  geom_hex(aes(x = decimal_longitude, y = decimal_latitude), bins = 20) +
  geom_point(data = major_cities, aes(x = lon, y = lat), color = "red", size = 3) +
  geom_text(data = major_cities, aes(x = lon, y = lat, label = city), vjust = -1, color = "red") +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  theme_minimal() +
  labs(
    title = "Hex Density of Frog Observations in Australia by Season with Major Cities",
    x = "Longitude",
    y = "Latitude",
    fill = "Count"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom"
  )

# create a hex density map of the Australian frogs by Season with major cities marked as stars; fix the map projection; remove the axis titles and text; remove grid lines
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  geom_hex(aes(x = decimal_longitude, y = decimal_latitude), bins = 20) +
  geom_point(data = major_cities, aes(x = lon, y = lat), color = "red", size = 3, shape = 8) +
  geom_text(data = major_cities, aes(x = lon, y = lat, label = city), vjust = -1, color = "red") +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  coord_fixed() +
  theme_minimal() +
  labs(
    title = "Hex Density of Frog Observations in Australia",
    subtitle = "by Season with Major Cities Marked",
    fill = "Count",
    caption = "Data: FrogID dataset 6.0 | Viz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom",
    plot.caption.position = "plot",
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank()
  )

# create a density map of the Australian frogs by season with major cities marked as stars; fix the map projection; remove the axis titles and text; remove grid lines; add a subtitle and caption
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  stat_density_2d(aes(x = decimal_longitude, y = decimal_latitude, fill = ..level..), geom = "polygon", alpha = 0.5) +
  geom_point(data = major_cities, aes(x = lon, y = lat), color = "red", size = 3, shape = 8) +
  geom_text(data = major_cities, aes(x = lon, y = lat, label = city), vjust = -1, color = "red") +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  coord_fixed() +
  theme_minimal() +
  labs(
    title = "Density of Frog Observations in Australia",
    subtitle = "by Season with Major Cities Marked",
    fill = "Density",
    caption = "Data: FrogID dataset 6.0 | Viz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom",
    plot.caption.position = "plot",
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank()
  )

# create a density map of the Australian frogs by season with major cities marked as stars; fix the map projection; remove the axis titles and text; remove grid lines; add a subtitle and caption; 
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  stat_density_2d(aes(x = decimal_longitude, y = decimal_latitude, fill = ..level..), geom = "polygon", alpha = 0.5) +
  geom_point(data = major_cities, aes(x = lon, y = lat), color = "red", size = 3, shape = 8) +
  geom_text(data = major_cities, aes(x = lon, y = lat, label = city), vjust = -1, color = "red") +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  coord_fixed() +
  theme_minimal() +
  labs(
    title = "Density of Frog Observations in Australia",
    subtitle = "by Season with Major Cities Marked",
    fill = "Density",
    caption = "Data: FrogID dataset 6.0 | Viz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom",
    plot.caption.position = "plot",
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank()
  )

# create a density map of the Australian frogs by season with major cities marked as stars; fix the map projection; remove the axis titles and text; remove grid lines; add a subtitle and caption; decrease the size of the city markers and labels
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  stat_density_2d(aes(x = decimal_longitude, y = decimal_latitude, fill = ..level..), geom = "polygon", alpha = 0.5) +
  geom_point(data = major_cities, aes(x = lon, y = lat), color = "red", size = 2, shape = 8) +
  geom_text(data = major_cities, aes(x = lon, y = lat, label = city), vjust = -1, color = "red", size = 3) +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  coord_fixed() +
  theme_minimal() +
  labs(
    title = "Density of Frog Observations in Australia",
    subtitle = "by Season with Major Cities Marked",
    fill = "Density",
    caption = "Data: FrogID dataset 6.0 | Viz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "bottom",
    plot.caption.position = "plot",
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank()
  )

# create a density map of the Australian frogs by season with major cities marked as stars; fix the map projection; remove the axis titles and text; remove grid lines; add a subtitle and caption; decrease the size of the city labels; make the city markers transparent; add lines connecting the city labels to the markers
frogs |>
  ggplot() +
  borders("world", region = "Australia", fill = "lightgrey") +
  stat_density_2d(aes(x = decimal_longitude, y = decimal_latitude, fill = ..level..), geom = "polygon", alpha = 0.5) +
  # geom_point(data = major_cities, aes(x = lon, y = lat), color = "red", size = 2, shape = 8, alpha = 0.7) +
  geom_text(data = major_cities, aes(x = lon, y = lat, label = city), vjust = -1, color = "red", size = 2.5) +
  geom_segment(data = major_cities, aes(x = lon, y = lat, xend = lon, yend = lat + 1), color = "red", linetype = "dashed", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~season) +
  coord_fixed() +
  theme_minimal() +
  labs(
    title = "Density of Frog Observations in Australia",
    subtitle = "By Season with Major Cities Marked",
    fill = "Density",
    caption = "Data: FrogID dataset 6.0 | Viz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12, face = "bold"),
    plot.caption = element_text(size = 8, hjust = 0.5),
    legend.position = "bottom",
    plot.caption.position = "plot",
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank()
  )
  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_09_02_tidy_tuesday_frogs.png",
  device = "png"
)
```
