---
title: "Women"
date: 2020-12-07
output: html_output
---

# TidyTuesday

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-12-08")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
women_raw <- tt$women

women <- women_raw
```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}
women %>%
  count(category, sort = TRUE) %>%
  mutate(category = fct_reorder(category, n)) %>%
  ggplot(aes(n, category, fill = category)) + 
  geom_col(show.legend = FALSE)

women %>%
  mutate(role = fct_lump(role,10)) %>%
  filter(role != "Other") %>%
  count(role, sort = TRUE) %>%
  mutate(role = fct_reorder(role, n)) %>%
  ggplot(aes(n, role, fill = role)) +
  geom_col(show.legend = FALSE)
```

# Create table

```{r table}
library(gt)

t <- women %>%
  select(img, name, country, category, role, description) %>%
  #group_by(category) %>%
  gt(rowname_col = "name") %>%
  text_transform(
      locations = cells_body(
        vars(img)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 50
        )
      }
    ) %>%
  cols_width(
    vars(img) ~ px(50),
    vars(country) ~ px(100),
    vars(category) ~ px(100),
    vars(role) ~ px(100),
    vars(description) ~ px(300),
    everything() ~ px(125)
  ) %>% 
  cols_label(
    img = "",
    country = "COUNTRY",
    category = "CATEGORY",
    role = "ROLE",
    description = "DESCRIPTION"
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.font.size = 12
  ) %>%
  tab_header(title = md("**WOMEN OF 2020**")) %>%
  tab_source_note(source_note = md("**Source**: BBC  |  **Designed By**: Tony Galvan (@GDataScience1)<br>**Inspiration**: @thomas_mock"))

t
```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
t %>%
  gtsave(
  "gdatascience1_women.png")

```
