---
title: "Bridges Across Borders: The Hidden Diplomacy of Sister Cities"
date: "2026-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 6)
```

Sister cities — also called twin towns — are one of the quietest forms of international diplomacy. Two cities in different countries agree to a cultural and commercial partnership, and over decades these links accumulate into a sprawling global network. This week's TidyTuesday dataset maps **10,596 twin city links** connecting **5,470 cities** across **191 countries**.

What happens when you treat this as a network and ask: *who clusters with whom?* The answer reveals the fingerprints of history — colonial ties, Cold War alliances, and linguistic kinship — all encoded in which cities chose to become siblings.

## Libraries

```{r libraries}
library(tidyverse)
library(scales)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(geosphere)
library(igraph)
library(showtext)
library(ggtext)

# Fonts
font_add_google("Source Sans 3", "source_sans")
font_add(family = "fa-brands",
         regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid",
         regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")
showtext_auto()
showtext_opts(dpi = 300)

theme_set(theme_light(base_family = "source_sans"))
```

## Load Data

```{r load-data}
cities <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-12/cities.csv')
links <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-12/links.csv')
```

## A First Look at the Data

Let's peek at what we're working with. The `cities` table has one row per city with its geographic coordinates and country:

```{r glimpse-cities}
glimpse(cities)
```

And the `links` table is a simple edge list — each row is a twinning agreement between two cities, identified by their Wikidata IDs:

```{r glimpse-links}
glimpse(links)
```

Here's a sample of what the cities look like:

```{r sample-cities}
cities |>
  select(name, country, continent, lng, lat) |>
  slice_sample(n = 10) |>
  knitr::kable()
```

We have **`r comma(nrow(cities))` cities** across **`r n_distinct(cities$country)` countries** and **`r n_distinct(cities$continent)` continents**, connected by **`r comma(nrow(links))` twin city links**.

## The Global Landscape

```{r eda-countries}
# Top countries by number of cities in the network
cities |>
  count(country, sort = TRUE) |>
  head(15) |>
  ggplot(aes(x = n, y = reorder(country, n))) +
  geom_col(fill = "#2E86AB") +
  labs(
    title = "Germany Dominates the Twin City Network",
    subtitle = "Countries with the most cities participating in twinning agreements",
    x = "Number of cities", y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )
```

**Germany has 1,086 cities** in the network — nearly twice as many as the United States (603). This reflects Germany's strong tradition of *Städtepartnerschaften* (city partnerships), many forged after WWII as tools of reconciliation with Poland, France, and Israel.

## How Connected Are Cities?

Not all cities are equally social. Let's look at the degree distribution — how many twin links each city has:

```{r eda-degree}
degree_df <- bind_rows(
  links |> select(city = source),
  links |> select(city = target)
) |>
  count(city, name = "degree") |>
  left_join(cities |> select(id, name, country, continent), by = c("city" = "id"))

degree_df |>
  ggplot(aes(x = degree)) +
  geom_histogram(binwidth = 1, fill = "#2E86AB", color = "white", linewidth = 0.2) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  labs(
    title = "Most Cities Have Just 1–3 Twin Links",
    subtitle = "But a few global hubs have 50+ connections",
    x = "Number of twin city links", y = "Number of cities"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )
```

The median city has just **`r median(degree_df$degree)` twin links**, but the distribution has a long tail. The most connected cities are global capitals and cultural hubs:

```{r eda-top-cities}
degree_df |>
  slice_max(degree, n = 15) |>
  ggplot(aes(x = degree, y = reorder(paste0(name, " (", country, ")"), degree))) +
  geom_col(fill = "#E69F00") +
  labs(
    title = "Saint Petersburg: The World's Most Twinned City",
    subtitle = "Top 15 cities by number of sister city connections",
    x = "Number of twin city links", y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )
```

**Saint Petersburg** leads with **96 twin city links**, followed closely by **Rio de Janeiro** (94). These are cities with outsized diplomatic ambitions — they've built relationships across every continent.

## Continental Connections

Are sister cities mostly local (same continent) or do they bridge oceans?

```{r eda-cross-continent}
links_enriched <- links |>
  left_join(cities |> select(id, country_s = country, continent_s = continent,
                             lng_s = lng, lat_s = lat), 
            by = c("source" = "id")) |>
  left_join(cities |> select(id, country_t = country, continent_t = continent,
                             lng_t = lng, lat_t = lat), 
            by = c("target" = "id"))

cross_continent <- links_enriched |>
  mutate(type = if_else(continent_s == continent_t, "Same continent", "Cross-continent")) |>
  count(type) |>
  mutate(pct = n / sum(n))

cross_continent |>
  ggplot(aes(x = pct, y = reorder(type, pct), fill = type)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(pct * 100, 1), "%")), 
            hjust = -0.1, size = 5, family = "source_sans") +
  scale_x_continuous(labels = percent, limits = c(0, 0.7)) +
  scale_fill_manual(values = c("Cross-continent" = "#009E73", "Same continent" = "#0072B2")) +
  labs(
    title = "45% of Twin City Links Cross Continents",
    subtitle = "Sister cities are truly global bridges, not just regional neighbors",
    x = NULL, y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )
```

Nearly **half of all twin city links (45%)** cross continental boundaries. These aren't just neighboring towns shaking hands — they're genuine bridges across oceans.

## Who Twins With Whom?

```{r country-pairs}
# Top international country pairs
top_pairs <- links_enriched |>
  filter(country_s != country_t) |>
  mutate(
    c1 = pmin(country_s, country_t),
    c2 = pmax(country_s, country_t)
  ) |>
  count(c1, c2, sort = TRUE, name = "n_links") |>
  head(15) |>
  mutate(pair = paste(c1, "–", c2))

top_pairs |>
  ggplot(aes(x = n_links, y = reorder(pair, n_links))) +
  geom_col(fill = "#A23B72") +
  labs(
    title = "Germany–Poland: The Strongest Twin City Bond",
    subtitle = "Top 15 country pairs by number of sister city links",
    x = "Number of twin city links", y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )
```

The **Germany–Poland** corridor has **344 twin city links** — more than any other international pair. This is the legacy of post-WWII reconciliation, where hundreds of German and Polish cities formalized partnerships to rebuild trust across a border that had been a site of immense suffering.

**Japan–USA** (213 links) is the second strongest, reflecting decades of post-war alliance and cultural exchange. **Mexico–USA** (150) shows the deep ties along the world's most-crossed border.

## Community Detection: Finding the Blocs

Here's where it gets interesting. If we build a network where countries are nodes and twin city links are weighted edges, we can use the **Louvain algorithm** to detect communities.

### What is the Louvain Algorithm?

The [Louvain method](https://en.wikipedia.org/wiki/Louvain_method) is a greedy optimization algorithm for community detection in networks. It works in two phases:

1. **Local optimization:** Each node starts in its own community. The algorithm iterates through every node and moves it to the neighboring community that produces the largest increase in **modularity** (or leaves it in place if no move improves modularity).
2. **Aggregation:** Once no single move improves modularity, the algorithm builds a new network where each community becomes a single node, and repeats phase 1 on this coarser network.

This repeats until modularity can't be improved further. It's fast (runs in near-linear time) and produces high-quality partitions, which is why it's one of the most popular community detection methods.

### What is Modularity?

**Modularity** (Q) measures how much more densely connected the nodes within communities are compared to what you'd expect by random chance. It ranges from -0.5 to 1.0:

- **Q ≈ 0** means the network's community structure is no better than random
- **Q > 0.3** is generally considered strong community structure
- **Q = 1** would mean perfectly separated communities with no cross-community edges

Our modularity of **0.276** is meaningful — it says these 5 clusters capture real structure in the data, though there are still plenty of cross-cluster links (which makes sense — diplomacy doesn't respect neat boundaries).

```{r community-detection}
# Build country-level network
country_edges <- links_enriched |>
  filter(country_s != country_t) |>
  mutate(c1 = pmin(country_s, country_t), c2 = pmax(country_s, country_t)) |>
  count(c1, c2, name = "weight")

g <- graph_from_data_frame(country_edges, directed = FALSE)
E(g)$weight <- country_edges$weight

# Louvain community detection
set.seed(42)
comm <- cluster_louvain(g, weights = E(g)$weight)

membership_df <- tibble(
  country = V(g)$name,
  community = membership(comm)
)

cat("Communities found:", max(membership_df$community), "\n")
cat("Modularity:", round(modularity(comm), 3), "\n")
```

### How I Labeled the Communities

The algorithm assigns numeric IDs (1–5) to each community — it doesn't know what they "mean." I labeled them by examining which countries ended up together and identifying the common thread:

```{r community-labels}
# Label communities based on their composition
community_labels <- c(
  "1" = "Russia & Eastern Mediterranean",
  "2" = "Francophone World",
  "3" = "Anglophone & Pacific Rim",
  "4" = "Latin America & Iberian",
  "5" = "Central & Northern Europe"
)

membership_df <- membership_df |>
  mutate(community_label = community_labels[as.character(community)])

# Show the full membership for each community
membership_df |>
  group_by(community_label) |>
  summarise(
    n_countries = n(),
    sample_countries = paste(head(sort(country), 8), collapse = ", ")
  ) |>
  arrange(desc(n_countries)) |>
  knitr::kable(col.names = c("Community", "Countries", "Sample Members"))
```

The labels come from recognizing patterns:

- **Community 5** contains Germany, Poland, UK, Nordics, Israel, Hungary, Czechia — all countries with dense post-WWII reconciliation twinning. I called it "Central & Northern Europe."
- **Community 2** contains France, Algeria, Morocco, Senegal, Cameroon, Tunisia — France and its former colonies. "Francophone World."
- **Community 1** contains Russia, Turkey, Iran, Iraq, the Gulf states, Central Asia — Soviet-era and Middle Eastern ties. "Russia & Eastern Mediterranean."
- **Community 3** contains USA, Japan, China, Canada, Australia, Philippines, Mexico — the Pacific Rim trading partners. "Anglophone & Pacific Rim."
- **Community 4** contains Spain, Portugal, Brazil, Argentina, Colombia, Chile — the Iberian colonial legacy. "Latin America & Iberian."

These clusters are **not geographic** — they're **historical and linguistic**. The algorithm independently rediscovered:

- **Colonial ties** (France clusters with West Africa; Spain/Portugal with Latin America)
- **Cold War alliances** (Russia clusters with the Middle East and Central Asia)
- **Post-WWII reconciliation** (Germany, Poland, Israel, and the Nordics form a tight bloc)
- **Anglophone networks** (US, UK... wait, UK is in the European bloc — it twins more with Germany and Scandinavia than with its former colonies)

## The World Map: Arcs of Connection

Now let's visualize the strongest connections on a world map, colored by which community cluster they belong to. We'll use country centroids as arc endpoints to keep the map readable.

```{r build-arcs}
# For the map, use the top connections (weight >= 5 links between countries)
strong_edges <- country_edges |> filter(weight >= 5)

# Get country centroids for arc endpoints
world <- ne_countries(scale = "medium", returnclass = "sf")

country_centroids <- world |>

  st_centroid() |>
  st_coordinates() |>
  as_tibble() |>
  bind_cols(world |> st_drop_geometry() |> select(name)) |>
  rename(lng = X, lat = Y, country = name)

# Join centroids to edges
arcs <- strong_edges |>
  left_join(country_centroids |> rename(lng_s = lng, lat_s = lat), 
            by = c("c1" = "country")) |>
  left_join(country_centroids |> rename(lng_t = lng, lat_t = lat), 
            by = c("c2" = "country")) |>
  filter(!is.na(lng_s) & !is.na(lng_t))

# Assign community color based on the source country
arcs <- arcs |>
  left_join(membership_df |> select(country, community, community_label), 
            by = c("c1" = "country"))

cat("Arcs with coordinates:", nrow(arcs), "of", nrow(strong_edges), "strong edges\n")
```

```{r final-viz, fig.width=10, fig.height=7}
# Community colors - distinct, colorblind-friendly
community_colors <- c(
  "Russia & Eastern Mediterranean" = "#E69F00",
  "Francophone World" = "#56B4E9",
  "Anglophone & Pacific Rim" = "#009E73",
  "Latin America & Iberian" = "#CC79A7",
  "Central & Northern Europe" = "#0072B2"
)

# Caption
bg_color <- "#1a1a2e"
tt_source <- "Wikidata Twin Cities"

tt_caption <- paste0(
  "DataViz: Tony Galvan #TidyTuesday",
  "<span style='color:", bg_color, ";'>..</span>",
  "<span style='font-family:fa-solid;'>&#xf0ce;</span>",
  "<span style='color:", bg_color, ";'>.</span>",
  tt_source,
  "<span style='color:", bg_color, ";'>..</span>",
  "<span style='font-family:fa-brands;'>&#xf09b;</span>",
  "<span style='color:", bg_color, ";'>.</span>",
  "gdatascience",
  "<span style='color:", bg_color, ";'>..</span>",
  "<span style='font-family:fa-brands;'>&#xe61b;</span>",
  "<span style='color:", bg_color, ";'>.</span>",
  "@GDataScience1"
)

# Build the map
p <- ggplot() +
  # World background
  geom_sf(data = world, fill = "#2d2d44", color = "#4a4a6a", linewidth = 0.15) +
  # Arcs colored by community
  geom_curve(
    data = arcs |> filter(!is.na(community_label)),
    aes(x = lng_s, y = lat_s, xend = lng_t, yend = lat_t,
        color = community_label, linewidth = weight),
    alpha = 0.4, curvature = 0.2
  ) +
  scale_color_manual(values = community_colors, name = "Diplomatic Cluster") +
  scale_linewidth_continuous(range = c(0.2, 2.5), guide = "none") +
  coord_sf(ylim = c(-90, 90)) +
  labs(
    title = "Bridges Across Borders",
    subtitle = "Sister city links reveal hidden diplomatic blocs shaped by\ncolonial history, Cold War alliances, and linguistic kinship",
    caption = tt_caption
  ) +
  theme_void(base_family = "source_sans") +
  theme(
    plot.background = element_rect(fill = bg_color, color = NA),
    plot.title = element_text(
      color = "white", size = 64, face = "bold", hjust = 0.5,
      margin = margin(t = 10, b = 8)
    ),
    plot.title.position = "plot",
    plot.subtitle = element_text(
      color = "gray80", size = 24, hjust = 0.5,
      margin = margin(b = 5), lineheight = 1.2
    ),
    plot.caption = element_markdown(
      color = "gray60", size = 9, hjust = 0.5,
      margin = margin(t = 5, b = 5)
    ),
    plot.caption.position = "plot",
    legend.position = c(0.15, 0.40),
    legend.direction = "vertical",
    legend.background = element_rect(fill = alpha(bg_color, 0.7), color = NA),
    legend.title = element_text(color = "gray80", size = 11, face = "bold"),
    legend.text = element_text(color = "gray80", size = 9),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width = unit(0.8, "cm"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  guides(color = guide_legend(ncol = 1, override.aes = list(linewidth = 2, alpha = 0.8)))

ggsave(
  filename = "outputs/2026_05_12_tidy_tuesday_twinned_cities.png",
  plot = p,
  device = "png",
  width = 10,
  height = 7,
  dpi = 300,
  bg = bg_color
)

showtext_auto(FALSE)
```

## Zooming In: Each Cluster's Footprint

The combined map shows all five clusters at once, but it's hard to see the geographic reach of each one individually. Let's facet by community to see where each bloc's arcs actually go:

```{r facet-map, fig.width=8, fig.height=20}
showtext_auto()

# For the faceted version, use city-level coordinates for richer detail per panel
arcs_city <- links_enriched |>
  filter(country_s != country_t) |>
  left_join(membership_df |> select(country, community_label), 
            by = c("country_s" = "country")) |>
  filter(!is.na(community_label))

# Sample within each community to keep it readable
set.seed(42)
arcs_facet <- arcs_city |>
  group_by(community_label) |>
  slice_sample(n = 400) |>
  ungroup()

ggplot() +
  geom_sf(data = world, fill = "#2d2d44", color = "#4a4a6a", linewidth = 0.1) +
  geom_curve(
    data = arcs_facet,
    aes(x = lng_s, y = lat_s, xend = lng_t, yend = lat_t,
        color = community_label),
    alpha = 0.3, curvature = 0.2, linewidth = 0.3
  ) +
  scale_color_manual(values = community_colors) +
  coord_sf(ylim = c(-60, 80)) +
  facet_wrap(~ community_label, ncol = 1) +
  labs(
    title = "Each Diplomatic Cluster Has a Distinct Geographic Signature",
    subtitle = "City-level twin links faceted by community — each panel shows one cluster's connections"
  ) +
  theme_void(base_family = "source_sans") +
  theme(
    plot.background = element_rect(fill = bg_color, color = NA),
    plot.title = element_text(color = "white", size = 20, face = "bold", hjust = 0.5,
                              margin = margin(t = 10, b = 5)),
    plot.title.position = "plot",
    plot.subtitle = element_text(color = "gray80", size = 13, hjust = 0.5,
                                 margin = margin(b = 10)),
    strip.text = element_text(color = "white", size = 14, face = "bold",
                              margin = margin(b = 3, t = 5)),
    legend.position = "none",
    plot.margin = margin(10, 10, 10, 10)
  )

showtext_auto(FALSE)
```

Now you can clearly see each cluster's geographic footprint:

- **Central & Northern Europe** is a dense web concentrated in Europe with tendrils reaching to Israel and the Americas
- **Francophone World** radiates from France down into West and North Africa
- **Russia & Eastern Mediterranean** fans out from Russia/Turkey across the Middle East and Central Asia
- **Anglophone & Pacific Rim** spans the Pacific — heavy arcs between Japan, the US, China, and Australia
- **Latin America & Iberian** connects Spain and Portugal to every corner of Central and South America

## What the Clusters Reveal

The five communities the algorithm discovered tell a story of **history encoded in municipal diplomacy**:

- **Central & Northern Europe** (blue): Germany, Poland, UK, Nordics, Israel — the post-WWII reconciliation network. Germany alone has 344 links to Poland and 58 to Israel.
- **Francophone World** (light blue): France and its former colonies in West Africa form a tight cluster, with Algeria, Morocco, and Tunisia as key nodes.
- **Russia & Eastern Mediterranean** (gold): Russia, Turkey, the Middle East, and Central Asia — reflecting both Soviet-era ties and Turkey's bridge role between Europe and Asia.
- **Anglophone & Pacific Rim** (green): The US, Japan, China, Canada, Australia, and much of East/Southeast Asia. Japan-US alone has 213 links.
- **Latin America & Iberian** (pink): Spain, Portugal, and all of Latin America — 500 years of colonial linguistic ties still visible in municipal diplomacy.

## What's Next?

Some open questions this data raises:

- **Is every country reachable** through a chain of twin city links? (Yes — all 191 countries are connected.)
- **Do these clusters predict trade flows?** Countries that twin together likely trade together.
- **How has the network grown over time?** The dataset doesn't include dates, but Wikidata might.
- **Which cities are the "bridges"** — connecting communities that otherwise wouldn't link?

The quiet diplomacy of sister cities turns out to be a remarkably faithful mirror of geopolitical history.
