---
title: "TidyTemplate"
date: 2023-01-25
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event! Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data. While the dataset will be "tamed", it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format. The goal of TidyTuesday is to apply your R skills, get feedback, explore other's work, and connect with the greater #RStats community! As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(showtext)

## Loading Google fonts (https://fonts.google.com/)
font_add_google("Black Ops One", "blops")

## Automatically use showtext to render text
showtext_auto()

theme_set(theme_light())

tt_caption <- "Source: {alone} package by Dan Oehm   |  DataViz: Tony Galvan  |  Twitter: @GDataScience1  |  #TidyTuesday"

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-01-24")

```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset. This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```

# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

survivalists <- tt$survivalists

prefix <- "https://cropper.watch.aetnd.com/cdn.watch.aetnd.com/sites/2/2019/06/alone-gear-"
suffix <- "-2048x1152.jpg"

loadouts <- tt$loadouts %>%
  left_join(survivalists %>% select(days_lasted, name), by = "name") %>%
  mutate(item_image = paste0(prefix, str_replace_all(str_to_lower(item), " ", "-"), suffix))

```

# Visualize

Using your processed dataset, create your unique visualization.

### GT item images

```{r GT}
library(gt)

loadouts %>%
  distinct(item, item_image) %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = c(item_image)),
    fn = function(x){
      web_image(
        url = x,
        height = 100
      )
    }
  )

```

### Bar Plot - Days Lasted by Season

```{r Visualize}

survivalists %>%
  ggplot(aes(season, days_lasted, group = season, fill = factor(season))) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none",
        plot.title = element_text(size = 32, face = "bold"),
        plot.subtitle = element_text(size = 16)) +
  labs(x = "Season",
       y = "# of days lasted",
       title = "Alone",
       subtitle = "Days Lasted by Season",
       caption = tt_caption)
  
```

### Survival Curves

```{r Visualize2}

p <- survivalists %>%
  add_count(season, name = "season_survivalists") %>%
  group_by(r=row_number()) %>%
  mutate(day_number = list(0:days_lasted)) %>%
  ungroup() %>%
  select(-r) %>%
  unnest(cols = c(day_number)) %>%
  count(season, day_number, season_survivalists) %>%
  mutate(pct_survival = n / season_survivalists) %>%
  ggplot() +
  geom_line(aes(day_number, pct_survival, color = factor(season))) +
  facet_wrap(~paste0("Season ", season))

p
  
```

Add overall survival curves

```{r Visualize2.5}

p +
  geom_line(data = survivalists %>%
              add_count(name = "total_survivalists") %>%
              group_by(r=row_number()) %>%
              mutate(day_number = list(0:days_lasted)) %>%
              ungroup() %>%
              select(-r) %>%
              unnest(cols = c(day_number)) %>%
              count(day_number, total_survivalists) %>%
              mutate(pct_survival = n / total_survivalists),
            aes(day_number, pct_survival),
            lty = 2) +
  scale_y_continuous(labels = percent_format()) +
  theme(legend.position = "none",
        plot.title = element_text(size = 32, face = "bold"),
        plot.subtitle = element_text(size = 16)) +
  labs(x = "# of day lasted",
       y = "% survived",
       title = "Alone (TV Series)",
       subtitle = "Survival Curves by Season (vs. Overall)",
       caption = tt_caption)

```

### Plot Item Average Days Lasted

```{r Visualize3}

loadouts %>%
  add_count(item, name = "total_loadouts") %>%
  count(item, total_loadouts, wt = days_lasted, name = "total_days_lasted") %>%
  mutate(days_lasted_per_loadout = total_days_lasted / total_loadouts,
         item = fct_reorder(item, days_lasted_per_loadout)) %>%
  ggplot(aes(days_lasted_per_loadout, item, color = item)) +
  geom_point(show.legend = FALSE)

```

### Boxplot Item Days Lasted

```{r Visualize4}

loadouts %>%
  mutate(item = fct_reorder(item, days_lasted)) %>%
  ggplot(aes(days_lasted, item)) +
  geom_boxplot(aes(fill = item)) +
  geom_jitter(alpha = 0.5) +
  theme(legend.position = "none",
        plot.title = element_text(size = 32, face = "bold"),
        plot.subtitle = element_text(size = 16)) +
  labs(x = "# of day lasted",
       y = "",
       title = "Alone (TV Series)",
       subtitle = "Which items resulted in more days lasted?",
       caption = tt_caption)

```

## Modeling using Tidymodels

### Data Setup

Create sparse data for each survivalists' items

```{r DataPrep}

alone <- loadouts %>%
  distinct(name, days_lasted, season, item) %>%
  mutate(n = 1) %>%
  pivot_wider(id_cols = c(name, days_lasted, season), 
              names_from = item, 
              values_from = n,
              names_prefix = "item_") %>%
  janitor::clean_names() %>%
  mutate(across(starts_with("item_"), ~replace_na(., 0))) %>%
  select(-name, -season)

glimpse(alone)

```

Split data into training/testing data sets

```{r SplitData}

library(tidymodels)

set.seed(1977)

alone_split <- initial_split(alone, strata = days_lasted)
alone_train <- training(alone_split)
alone_test <- testing(alone_split)

set.seed(2003)

alone_folds <- vfold_cv(alone_train, strata = days_lasted)
alone_folds

```

### Model Setup

Create the recipe

```{r Recipe}

alone_rec <- recipe(days_lasted ~ ., data = alone_train)

alone_rec

```

Set up the Lasso (glmnet) model

```{r LassoModel}

lasso_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lasso_wf <- workflow() %>%
  add_recipe(alone_rec) %>%
  add_model(lasso_spec)

lasso_wf

```

### Tune the model

```{r TuneModel}

lambda_grid <- grid_regular(penalty(range = c(-3, 0)), levels = 20)

doParallel::registerDoParallel()

set.seed(2010)

lasso_rs <- tune_grid(
  lasso_wf,
  resamples = alone_folds,
  grid = lambda_grid
)

lasso_rs

```

```{r EvaluateModel}

autoplot(lasso_rs)

```

```{r BestModel}

show_best(lasso_rs, "rmse")

```

```{r FinalizeModel}

best_rmse <- select_best(lasso_rs, "rmse")

final_lasso <- finalize_workflow(lasso_wf, best_rmse)
final_lasso

```

### Train & Evaluate the Model

The function last_fit() fits this finalized lasso model one last time to the training data and evaluates one last time on the testing data. The metrics are computed on the testing data.

```{r FinalFit}

alone_final <- last_fit(final_lasso, alone_split)
collect_metrics(alone_final)

```

### Explain the Results of the Model

```{r VariableImportance}

library(vip)

alone_vip <- extract_fit_parsnip(alone_final$.workflow[[1]]) %>%
  vi()


alone_vip %>%
  mutate(
    Variable = str_remove(Variable, "item_"),
    Variable = str_replace_all(Variable, "_", " "),
    Variable = str_to_title(Variable),
    Importance = if_else(Sign == "POS", abs(Importance), -Importance),
    Variable = fct_reorder(Variable, Importance),
  ) %>%
  ggplot(aes(Importance, Variable, fill = Sign)) +
  geom_col(show.legend = FALSE) +
  geom_vline(xintercept = 0, color = "white", size = 1.25, lty = 2) +
  scale_fill_manual(values = c("#FF784F", "#2364AA")) +
  theme(plot.background = element_rect(fill = "gray15"),
        panel.background = element_rect(fill = "gray15"),
        plot.title = element_text(size = 160, family = "blops", color = "white"),
        plot.subtitle = element_text(size = 68, color = "white"),
        axis.text.y = element_text(size = 64, face = "bold", color = "white"),
        axis.title.x = element_text(color = "white", size = 36),
        axis.text.x = element_text(color = "white", size = 36),
        plot.caption = element_text(color = "white", size = 32),
        # panel.grid.major.x = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = "Importance",
       y = "",
       title = "Alone (TV Series)",
       subtitle = "Which items should survivalists take with them?",
       caption = tt_caption)

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter!

```{r}

# This will save your most recent plot
ggsave(
  filename = "2023_01_24_tidy_tuesday_alone2.png",
  device = "png")

```
