---
title: "Science Foundation Ireland Grants"
date: 2026-03-02
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be "tamed", it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other's work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(lubridate)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2026-02-24")
```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```

# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt |>
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
# Get the grants data
grants <- tt$sfi_grants

# In 2013, SFI shifted from basic research to applied research focused on
# industrial applications due to economic recession (Irish Times, 2018)
# Source: https://www.irishtimes.com/news/science/funding-priorities-of-science-foundation-ireland-need-re-balancing-1.3388727

# Categorize programmes based on actual programme names in the data
# Basic Research: Individual investigator awards, fellowships, research frontiers
# Applied Research: Centres, industry partnerships, strategic programmes
viz_data <- grants |>
  mutate(
    year = year(start_date),
    research_type = case_when(
      # Basic Research - Individual investigator-led research
      programme_name %in% c(
        "Principal Investigator Programme",
        "Investigator Programme",
        "Research Frontiers Programme",
        "Frontiers for the Future",
        "Starting Investigator Research Grant",
        "Research Professorship Programme",
        "Stokes Professorship & Lectureship Programme",
        "Career Development Award",
        "President of Ireland Young Researcher Award",
        "President of Ireland Future Research Leaders Programme",
        "Royal Society University Research Fellowship",
        "ERC Support Programme",
        "ERC Development Programme",
        "Fellowship Programme"
      ) ~ "Basic Research",
      
      # Applied Industrial Research - Centres, partnerships, industry focus
      programme_name %in% c(
        "Research Centres Programme",
        "Centres for Science Engineering and Technology",
        "Strategic Research Cluster Programme",
        "Strategic Partnership Programme",
        "Spokes Programme",
        "Technology Innovation Development Award",
        "Industry Fellowship Programme",
        "Industry RD&I Fellowship Programme",
        "Industry Research Partnership",
        "Centres for Research Training (CRT) Programme",
        "Co-Centre Programme",
        "National Challenge Fund"
      ) ~ "Applied Industrial Research",
      
      # Infrastructure and support (separate category)
      programme_name %in% c(
        "Research Infrastructure Programme",
        "Infrastructural Support",
        "Equipment Call"
      ) ~ "Infrastructure & Support",
      
      # Everything else
      TRUE ~ "Other Programmes"
    )
  ) |>
  filter(!is.na(year), year >= 2000, year <= 2024) |>
  group_by(year, research_type) |>
  summarise(
    total_funding = sum(current_total_commitment, na.rm = TRUE),
    n_grants = n(),
    .groups = "drop"
  )

# Show the categorization summary
grants |>
  mutate(
    research_type = case_when(
      programme_name %in% c(
        "Principal Investigator Programme",
        "Investigator Programme",
        "Research Frontiers Programme",
        "Frontiers for the Future",
        "Starting Investigator Research Grant",
        "Research Professorship Programme",
        "Stokes Professorship & Lectureship Programme",
        "Career Development Award",
        "President of Ireland Young Researcher Award",
        "President of Ireland Future Research Leaders Programme",
        "Royal Society University Research Fellowship",
        "ERC Support Programme",
        "ERC Development Programme",
        "Fellowship Programme"
      ) ~ "Basic Research",
      programme_name %in% c(
        "Research Centres Programme",
        "Centres for Science Engineering and Technology",
        "Strategic Research Cluster Programme",
        "Strategic Partnership Programme",
        "Spokes Programme",
        "Technology Innovation Development Award",
        "Industry Fellowship Programme",
        "Industry RD&I Fellowship Programme",
        "Industry Research Partnership",
        "Centres for Research Training (CRT) Programme",
        "Co-Centre Programme",
        "National Challenge Fund"
      ) ~ "Applied Industrial Research",
      programme_name %in% c(
        "Research Infrastructure Programme",
        "Infrastructural Support",
        "Equipment Call"
      ) ~ "Infrastructure & Support",
      TRUE ~ "Other Programmes"
    )
  ) |>
  group_by(research_type) |>
  summarise(
    n_grants = n(),
    total_funding = sum(current_total_commitment, na.rm = TRUE),
    pct_of_total = total_funding / sum(grants$current_total_commitment, na.rm = TRUE) * 100
  ) |>
  arrange(desc(total_funding))
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize, fig.width=12, fig.height=8}
# Focus on Basic vs Applied research only for clearer story
viz_data_focused <- viz_data |>
  filter(research_type %in% c("Basic Research", "Applied Industrial Research"))

# Create an area chart showing the shift from basic to applied research
viz_data_focused |>
  ggplot(aes(x = year, y = total_funding / 1e6, fill = research_type)) +
  geom_area(alpha = 0.85, position = "stack") +
  geom_vline(
    xintercept = 2013,
    linetype = "dashed",
    color = "white",
    linewidth = 1.2
  ) +
  annotate(
    "text",
    x = 2013,
    y = max(aggregate(
      total_funding ~ year,
      viz_data_focused,
      sum
    )$total_funding) / 1e6 * 0.92,
    label = "2013\nPolicy Shift",
    hjust = 0.5,
    vjust = 1,
    size = 4.5,
    fontface = "bold",
    color = "white"
  ) +
  scale_y_continuous(
    labels = label_dollar(prefix = "€", suffix = "M"),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    breaks = seq(2000, 2024, 2),
    expand = c(0, 0)
  ) +
  scale_fill_manual(
    values = c(
      "Basic Research" = "#2E86AB",
      "Applied Industrial Research" = "#A23B72"
    ),
    name = "Research Type"
  ) +
  labs(
    title = "Ireland's Research Pivot: From Basic Science to Industry",
    subtitle = paste0(
      "After 2013, Science Foundation Ireland dramatically shifted ",
      "funding priorities toward applied industrial\nresearch in ",
      "response to economic recession, fundamentally changing the ",
      "research landscape"
    ),
    x = NULL,
    y = "Annual Grant Commitments",
    caption = paste0(
      "Data: Science Foundation Ireland | ",
      "Context: Irish Times (2018) | ",
      "DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
    )
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 20,
      margin = margin(b = 8)
    ),
    plot.subtitle = element_text(
      size = 11,
      color = "gray30",
      margin = margin(b = 20),
      lineheight = 1.3
    ),
    plot.caption = element_text(
      size = 9,
      color = "gray50",
      hjust = 0,
      margin = margin(t = 20)
    ),
    legend.position = c(0.15, 0.85),
    legend.title = element_text(face = "bold", size = 13),
    legend.text = element_text(size = 11),
    legend.background = element_rect(
      fill = "white",
      color = "gray80",
      linewidth = 0.5
    ),
    legend.margin = margin(8, 8, 8, 8),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "gray90"),
    plot.margin = margin(20, 20, 20, 20),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter!

```{r}
# Save the visualization
ggsave(
  filename = "2026_02_24_tidy_tuesday_sfi_grants.png",
  device = "png",
  width = 12,
  height = 8,
  dpi = 300,
  bg = "white"
)
```
