---
title: "The Atomic Age in Numbers: Mapping Every Nuclear Weapons Test from 1945 to 1998"
author: "Tony Galvan"
date: "8/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Between 1945 and 1998, the world's nuclear powers detonated over 2,000 nuclear weapons in tests — from the deserts of Nevada to remote Pacific atolls to the frozen steppes of Kazakhstan. The Stockholm International Peace Research Institute (SIPRI) cataloged every one of them. This dataset lets us map the geography of nuclear testing, trace the arms race decade by decade, and see which U.S. presidents presided over the most tests.

## Loading and Preparing the Data

We'll recode country names for readability, parse dates, assign decades, and map each U.S. test to the sitting president.

```{r}
library(tidyverse)
library(lubridate)
theme_set(theme_light())

nuclear_explosions <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-20/nuclear_explosions.csv") |>
  mutate(country = fct_recode(country, 
                              "United States" = "USA",
                              "Russia" = "USSR",
                              "France" = "FRANCE",
                              "United Kingdom" = "UK",
                              "China" = "CHINA",
                              "India" = "INDIA",
                              "Pakistan" = "PAKIST"),
         date  = ymd(date_long),
         decade = 10 * (year %/% 10),
         day_of_decade = yday(date) * (year - decade + 1),
         president = case_when(date < "1953-01-20" ~ "Truman",
                               date < "1961-01-20" ~ "Eisenhower",
                               date < "1963-11-22" ~ "Kennedy",
                               date < "1969-01-20" ~ "Johnson",
                               date < "1974-08-09" ~ "Nixon",
                               date < "1977-01-20" ~ "Ford",
                               date < "1981-01-20" ~ "Carter",
                               date < "1989-01-20" ~ "Reagan",
                               date < "1993-01-20" ~ "Bush (H. W.)",
                               TRUE ~ "Clinton")) |>
  select(-date_long)

summary(nuclear_explosions)
```

## Which Countries Conducted Nuclear Tests?

The nuclear club is small but its members tested prolifically. Let's see the total count by country.

```{r}
nuclear_explosions |>
  count(country, sort = TRUE) |>
  mutate(country = fct_reorder(country, n)) |>
  ggplot(aes(country, n)) +
  geom_col(aes(fill = country), show.legend = FALSE) + 
  coord_flip() + 
  geom_text(aes(label = n, color = country), show.legend = FALSE,
            nudge_y = 40, size = 4, fontface = "bold") + 
  labs(x = "",
       y = "# of nuclear weapons tests",
       title = "Nuclear weapons tests conducted 1945 - 1998",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: SIPRI")
```

The United States and Russia (USSR) conducted the vast majority of tests — over 1,000 each — reflecting the Cold War arms race that defined the second half of the 20th century.

## World Map of Test Locations

Where on Earth were these weapons detonated? The geographic concentration is striking.

```{r}
nuclear_explosions |>
  ggplot(aes(longitude, latitude, size = yield_upper)) +
  borders("world", colour = "gray75", fill = "gray75") +
  geom_point(aes(fill = country), pch = 21, color = "black", alpha = 0.5) +
  coord_quickmap() + 
  theme_void() + 
  labs(size = "Blast yield (in tons)",
       fill = "Country",
       title = "Nuclear weapons tests conducted 1945 - 1998",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: SIPRI")
```

The clusters are unmistakable: Nevada (USA), Semipalatinsk and Novaya Zemlya (USSR), the Pacific atolls (USA, UK, France), the Sahara (France), Lop Nur (China), and Rajasthan (India/Pakistan). These remote locations were chosen to minimize civilian exposure — though the long-term health effects on nearby populations remain controversial.

## The Arms Race by Decade

An overlayed step plot shows how testing intensity evolved decade by decade. Each line traces the cumulative number of tests within its decade.

```{r}
nuclear_explosions |>
  filter(date > "1949-12-31" & date < "1990-01-01") |>
  group_by(day_of_decade, decade) |>
  summarise(n = n()) |>
  ungroup() |>
  mutate(counter = 1) |>
  group_by(decade) |>
  mutate(counter = cumsum(n)) |>
  ungroup() |>
  ggplot(aes(day_of_decade, counter, group = decade, color = factor(decade))) + 
  geom_step() + 
  scale_color_discrete(labels = c("1950's", "1960's", "1970's", "1980's")) +
  scale_x_continuous(breaks = c(750, 3000),
                     labels = c("Beginning of the decade", "End of the decade")) +
  labs(x = "",
       y = "# of nuclear weapons tests",
       color = "Decade",
       title = "Nuclear weapons tests conducted by decade: 1950's - 1980's",
       subtitle = "The number of tests flattens out toward the end of every decade",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: SIPRI")
```

```{r}
ggsave("outputs/nuclear_explosions.png", width = 5.75)
```

A fascinating pattern emerges: every decade shows a burst of testing early on that flattens toward the end. This likely reflects test ban negotiations and political pressure that built throughout each decade.

## Nuclear Tests by U.S. President

Which American presidents oversaw the most nuclear weapons tests? The answer reflects both Cold War intensity and individual policy choices.

```{r}
nuclear_explosions |>
  filter(country == "United States" & 
           president != "Truman" & president != "Clinton") |>
  group_by(president) |>
  summarise(n = n()) |>
  mutate(president = fct_reorder(president, n)) |>
  ggplot(aes(president, n, fill = president)) + 
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "",
       y = "# of nuclear weapons tests",
       title = "Nuclear weapons tests conducted by U.S. Presidents",
       subtitle = "Between 1/20/1953 and 1/20/1993",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: SIPRI")
```

## Including the Full Presidential Timeline

Let's expand the view to include all presidents from Truman through Clinton — notably, Clinton conducted zero tests during this period.

```{r}
nuclear_explosions |>
  filter(country == "United States") |>
  group_by(president) |>
  summarise(n = n()) |>
  ungroup() |>
  bind_rows(data.frame(president = "Clinton", n = 0)) |>
  mutate(president = fct_reorder(president, n)) |>
  ggplot(aes(president, n, fill = president)) + 
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "",
       y = "# of nuclear weapons tests",
       title = "Nuclear weapons tests conducted by U.S. Presidents 1945 - 1998",
       subtitle = "Bill Clinton did not conduct any nuclear weapons tests during this time",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: SIPRI")
```

```{r}
ggsave("outputs/nuclear_presidents.png", width = 6.5)
```

Reagan conducted the most tests of any president — over 200 during his eight years — reflecting his administration's "peace through strength" philosophy. Clinton's zero tests mark the end of an era: the Comprehensive Nuclear-Test-Ban Treaty was signed in 1996, and no country has conducted a confirmed nuclear test since 1998 (with the exception of North Korea).
