---
title: "TidyTemplate"
date: 2025-08-15
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(plotly)
library(htmlwidgets)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-08-12")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

studies <- tt$attribution_studies

```


# Visualize

Using your processed dataset, create your unique visualization.

Inspired by https://dviz.manishdatt.com/posts/Weather_attribution/weather_attribution.html

```{r Visualize}

studies_agg <- studies |>
  count(study_focus, classification, event_type, name = "count")

# Define a function to prepare and plot for a single study focus
create_sunburst_data <- function(data, focus_type) {
  # Filter data for the specific study focus
  plot_data <- data |>
    filter(study_focus == focus_type)

  # Calculate total count for the title
  total_count <- sum(plot_data$count)

  # --- Prepare data for a single sunburst plot with explicit ids and parents ---
  # 1. Create data for the top level (study_focus)
  top_level_df <- data.frame(
    ids = focus_type,
    labels = paste0(focus_type, " Focus"),
    parents = "",
    values = total_count
  )

  # 2. Create data for the classification level
  class_level_df <- plot_data |>
    group_by(classification) |>
    summarise(count = sum(count), .groups = 'drop') |>
    mutate(
      ids = paste(focus_type, classification, sep = "-"),
      labels = classification,
      parents = focus_type,
      values = count
    )

  # 3. Create data for the event type level
  event_level_df <- plot_data |>
    mutate(
      ids = paste(focus_type, classification, event_type, sep = "-"),
      labels = event_type,
      parents = paste(focus_type, classification, sep = "-"),
      values = count
    )

  # Combine all levels into a single data frame for plotting
  combined_df <- bind_rows(top_level_df, class_level_df, event_level_df) |>
    # Ensure columns are of the correct type
    mutate(
      ids = as.character(ids),
      labels = as.character(labels),
      parents = as.character(parents)
    )
    
  return(combined_df)
}

# Generate data for each study focus
data_event <- create_sunburst_data(studies_agg, "Event")
data_trend <- create_sunburst_data(studies_agg, "Trend")

# Create a single plotly figure
fig <- plot_ly() |>
  add_trace(
    type = "sunburst",
    data = data_event,
    ids = ~ids,
    labels = ~labels,
    parents = ~parents,
    values = ~values,
    domain = list(column = 0), # Place this trace in the first column
    branchvalues = 'total'
  ) |>
  add_trace(
    type = "sunburst",
    data = data_trend,
    ids = ~ids,
    labels = ~labels,
    parents = ~parents,
    values = ~values,
    domain = list(column = 1), # Place this trace in the second column
    branchvalues = 'total'
  ) |>
  # Update the layout to specify subplots in a 1x2 grid
  layout(
    grid = list(rows = 1, columns = 2),
    title = list(
      text = "Climate Change and Extreme Weather Attribution",
      x = 0.5,
      xanchor = "center"
    )
  )

# Save the interactive plot as an HTML file
saveWidget(fig, file = "2025_08_12_tidy_tuesday_studies.html", selfcontained = TRUE)
```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_08_12_tidy_tuesday_studies.png",
  device = "png"
)
```
