---
title: "TidyTemplate"
date: 2025-02-11
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(tidytext)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-02-11")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

cdc <- tt$cdc_datasets

top_9_categories <- cdc |>
  filter(category != "This dataset has not been categorized") |>
  count(category, name = "total_category_datasets") |>
  # top_n(9, wt = total_category_datasets) |> # this is the same as slice_max
  slice_max(total_category_datasets, n = 9, with_ties = FALSE)

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

cdc |>
  filter(category != "This dataset has not been categorized") |>
  count(category) |>
  top_n(9, wt = n) |>
  ggplot(aes(n, fct_reorder(category, n))) +
  geom_col()

cdc |>
  filter(tags != "This dataset does not have any tags") |>
  separate_rows(tags, sep = ", ") |>
  count(tags) |>
  top_n(9, wt = n) |>
  ggplot(aes(n, fct_reorder(tags, n))) +
  geom_col()

cdc |>
  mutate(publisher = fct_lump(publisher, 4)) |>
  count(publisher) |>
  ggplot(aes(n, fct_reorder(publisher, n))) +
  geom_col()

cdc |>
  mutate(contact_name = fct_lump(contact_name, 8)) |>
  count(contact_name) |>
  ggplot(aes(n, fct_reorder(contact_name, n))) +
  geom_col()

cdc |>
  filter(category != "This dataset has not been categorized") |>
  separate_rows(tags, sep = ", ") |>
  filter(tags != "...") |>
  mutate(
    tags = case_when(
      tags == "centers for disease control and prevention" ~ "cdc",
      tags == "coronavirus" ~ "covid-19",
      tags == "covid 19" ~ "covid-19",
      tags == "covid19" ~ "covid-19",
      tags == "This dataset does not have any tags" ~ "no tags",
      TRUE ~ tags
    )
  ) |>
  unique() |>
  count(category, tags) |>
  group_by(category) |>
  slice_max(n, n = 10, with_ties = FALSE) |>
  ungroup() |>
  inner_join(
    top_9_categories,
    by = "category"
  ) |>
  mutate(
    category = fct_reorder(category, -total_category_datasets),
    tags = reorder_within(tags, n, category)
  ) |>
  ggplot(aes(n, tags, fill = category)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(~category, scales = "free") +
  labs(
    x = "# of datasets",
    y = NULL,
    title = "CDC Datasets Recently Purged",
    subtitle = "Top tags within top 9 categories",
    caption = "Data: archive.org  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )

cdc |>
  filter(category != "This dataset has not been categorized") |>
  mutate(
    # pull out the file name from the dataset_url
    file_name = str_extract(dataset_url, "([^/]+)$"),
    file_name = str_remove(file_name, ".csv")
  ) |>
  transmute(category, word = file_name) |>
  # separate_rows(file_name, sep = "_") |>
  separate_longer_delim(
    cols = word,
    delim = "_"
  ) |>
  # separate_rows(file_name, sep = "-") |>
  separate_longer_delim(
    cols = word,
    delim = "-"
  ) |>
  filter(!str_detect(word, "^[0-9]+$")) |>
  # remove all punctuation
  mutate(
    word = str_remove_all(word, "[[:punct:]]"),
    word = str_remove_all(word, "[0-9]"),
    word = str_to_lower(word)
  ) |>
  filter(
    word != "",
    !word %in% c("data", "meta", "table")
  ) |>
  anti_join(stop_words) |>
  count(category, word) |>
  group_by(category) |>
  slice_max(n, n = 10, with_ties = FALSE) |>
  ungroup() |>
  inner_join(
    top_9_categories,
    by = "category"
  ) |>
  mutate(
    category = fct_reorder(category, -total_category_datasets),
    word = reorder_within(word, n, category)
  ) |>
  ggplot(aes(n, word, fill = category)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(~category, scales = "free") +
  labs(
    x = "# of datasets",
    y = NULL,
    title = "CDC Datasets Recently Purged",
    subtitle = "Top URL words within top 9 categories",
    caption = "Data: archive.org  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )
  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_02_11_tidy_tuesday_cdc.png",
  device = "png"
)
```
