---
title: "Can We Predict If Your Broken Gadget Will Be Fixed?"
date: 2026-04-14
output: html_document
---

Every year, thousands of people carry their broken toasters, cracked phones, and torn jeans to community repair cafés — volunteer-run events where skilled fixers try to breathe new life into discarded items. But not everything can be saved. Some gadgets walk out the door humming again; others leave in the same sorry state they arrived.

This week's TidyTuesday dataset contains **178,749 repair attempts** from **448 repair cafés** across **25 countries**, spanning 2015 to 2026. The big question: can the words people use to describe a problem actually predict whether the item gets fixed? We built a machine learning model to find out.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)
library(tidytuesdayR)
library(tidytext)
library(tidymodels)
library(textrecipes)
library(scales)
library(cowplot)
library(magick)

theme_set(theme_light())
```

```{r load-data}
cache_path <- ".kiro/specs/2026_04_07_tidy_tuesday_repairs/tt_cache.rds"

if (file.exists(cache_path)) {
  tt <- readRDS(cache_path)
} else {
  tt <- tt_load("2026-04-07")
  saveRDS(tt, cache_path)
}

repairs <- tt$repairs
repairs_text <- tt$repairs_text
```

## The Repair Café Movement Is Growing Fast

Before we dive into modeling, let's look at the scale of this movement. Repair cafés have exploded in popularity — the number of logged repair attempts has grown from just 28 in 2015 to over **45,000 in 2025**.

```{r growth-over-time}
repairs_by_year <- repairs %>%
  mutate(year = year(repair_date)) %>%
  filter(year >= 2017, year <= 2025) %>%
  count(year)

ggplot(repairs_by_year, aes(x = year, y = n)) +
  geom_col(fill = "#2E86AB", width = 0.7) +
  geom_text(aes(label = comma(n)), vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_x_continuous(breaks = 2017:2025) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Repair Cafés Are Booming",
    subtitle = "Annual repair attempts logged worldwide, 2017–2025",
    x = NULL, y = "Repair Attempts",
    caption = "Data Source: Open Repair Alliance | DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  )
```

The dip in 2020–2021 is no mystery — COVID shut down in-person events globally. But the rebound has been remarkable, with volumes nearly tripling from pre-pandemic levels.

## What Gets Brought In — and What Gets Fixed?

Overall, **63% of items are fully repaired**, with another **13% partially fixed**. But the success rate varies wildly by category. Textiles and hand tools are almost always fixable, while electronics and computers are a coin flip.

```{r repair-rates-by-category}
cat_rates <- repairs %>%
  filter(!is.na(category), !is.na(repaired)) %>%
  mutate(
    outcome = case_when(
      repaired %in% c("yes", "ja") ~ "Repaired",
      repaired == "half" ~ "Partial",
      repaired == "no" ~ "Not Repaired"
    ),
    outcome = factor(outcome, levels = c("Repaired", "Partial", "Not Repaired"))
  ) %>%
  count(category, outcome) %>%
  group_by(category) %>%
  mutate(
    total = sum(n),
    pct = n / total
  ) %>%
  ungroup() %>%
  filter(total >= 100)

cat_order <- cat_rates %>%
  filter(outcome == "Repaired") %>%
  arrange(pct) %>%
  pull(category)

ggplot(
  cat_rates %>% mutate(category = factor(category, levels = cat_order)),
  aes(x = pct, y = category, fill = outcome)
) +
  geom_col(position = "stack") +
  scale_x_continuous(labels = percent, expand = expansion(mult = c(0, 0.02))) +
  scale_fill_manual(values = c("Repaired" = "#2E86AB", "Partial" = "#F6AE2D", "Not Repaired" = "#E84855")) +
  labs(
    title = "Textiles Are Easy Fixes; Electronics Are a Gamble",
    subtitle = "Repair outcomes by product category (categories with 100+ attempts)",
    x = "Share of Repair Attempts", y = NULL, fill = NULL,
    caption = "Data Source: Open Repair Alliance | DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(legend.position = "top")
```

Textiles lead at **92% repaired** — a torn seam or missing button is straightforward work. On the other end, display/sound equipment and computers hover around **46%**. The complexity of modern electronics makes them harder to crack open and diagnose.

## Can Words Predict Repair Success?

Here's where it gets interesting. When people bring items to a repair café, volunteers log a description of the problem and the defect found. We combined these text fields and fed them into a **lasso logistic regression** model to see if the language alone can predict whether an item gets repaired.

```{r prepare-ml-data}
set.seed(42)

df <- repairs %>%
  left_join(repairs_text, by = "repair_id", relationship = "many-to-many") %>%
  distinct(repair_id, .keep_all = TRUE) %>%
  mutate(
    was_repaired = case_when(
      repaired %in% c("yes", "ja") ~ "repaired",
      repaired %in% c("no", "half") ~ "not_repaired",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(was_repaired)) %>%
  mutate(
    combined_text = paste(
      coalesce(problem_description, ""),
      coalesce(defect_found, "")
    ),
    category = factor(coalesce(category, "Unknown"))
  ) %>%
  filter(nchar(trimws(combined_text)) > 5) %>%
  select(repair_id, was_repaired, combined_text, category)

df$was_repaired <- factor(df$was_repaired, levels = c("repaired", "not_repaired"))
```

We use all **`r comma(nrow(df))`** records with text descriptions, split 75/25 into training and test sets, then tokenize the text into TF-IDF features alongside the product category.

```{r train-model}
split <- initial_split(df, prop = 0.75, strata = was_repaired)
train <- training(split)
test <- testing(split)

rec <- recipe(was_repaired ~ combined_text + category, data = train) %>%
  step_tokenize(combined_text) %>%
  step_stopwords(combined_text) %>%
  step_tokenfilter(combined_text, max_tokens = 500) %>%
  step_tfidf(combined_text) %>%
  step_dummy(category) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())

lasso_spec <- logistic_reg(penalty = 0.01, mixture = 1) %>%
  set_engine("glmnet")

wf <- workflow() %>%
  add_recipe(rec) %>%
  add_model(lasso_spec)

fit <- fit(wf, data = train)
```

```{r evaluate-model}
preds <- predict(fit, test, type = "prob") %>%
  bind_cols(predict(fit, test)) %>%
  bind_cols(test %>% select(was_repaired))

acc <- accuracy(preds, truth = was_repaired, estimate = .pred_class)
auc_val <- roc_auc(preds, truth = was_repaired, .pred_repaired)
```

The model achieves **`r percent(acc$.estimate, accuracy = 0.1)` accuracy** and an **AUC of `r round(auc_val$.estimate, 3)`** — well above the 50% baseline for balanced classes. The text people use to describe problems genuinely carries signal about whether the fix will succeed.

## The Language of Fixable vs. Unfixable

The lasso model zeroes out irrelevant features, leaving us with the words that matter most. Let's look at what language predicts success versus failure.

```{r extract-coefficients}
model_engine <- extract_fit_engine(fit)
coefs <- coef(model_engine, s = 0.01)
coef_df <- data.frame(
  term = rownames(coefs),
  estimate = as.vector(coefs[, 1])
) %>%
  filter(term != "(Intercept)", estimate != 0) %>%
  mutate(
    word = str_remove(term, "tfidf_combined_text_") %>%
      str_remove("category_"),
    clean_term = str_replace_all(word, "\\.", " / "),
    direction = if_else(estimate > 0, "Predicts Repair", "Predicts Failure")
  ) %>%
  arrange(desc(abs(estimate)))
```

```{r top-words-plot, fig.width=10, fig.height=7}
top_words <- coef_df %>%
  filter(str_detect(term, "tfidf_")) %>%
  slice_head(n = 20) %>%
  mutate(clean_term = fct_reorder(clean_term, abs(estimate)))

ggplot(top_words, aes(x = abs(estimate), y = clean_term, fill = direction)) +
  geom_col() +
  scale_fill_manual(values = c("Predicts Repair" = "#2E86AB", "Predicts Failure" = "#E84855")) +
  labs(
    title = "What Words Predict Whether Your Gadget Gets Fixed?",
    subtitle = "Top 20 text features from a lasso logistic regression model",
    x = "Absolute Coefficient (Importance)", y = NULL, fill = NULL,
    caption = "Data Source: Open Repair Alliance | DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(legend.position = "top")
```

Words like **"motor"** and **"element"** predict successful repairs — these point to identifiable, replaceable components. Meanwhile, **"blunt"** (dull blades that can't be resharpened) and **"los"** (Dutch for "loose," suggesting structural issues) predict failure. The word **"snoer"** (Dutch for "cord") also signals trouble — frayed power cords on older appliances are often not worth the safety risk.

## Lost in Translation? Decoding the Multilingual Signal

Here's the thing — this dataset spans **25 countries**, and the repair descriptions are written in Dutch, French, German, Spanish, and English. Our lasso model picked up on words from all of these languages, but the top-words plot above is hard to interpret if you don't speak Dutch. Let's fix that.

We built a translation lookup for every non-English word the model selected and re-labeled the chart. Now the story becomes much clearer.

```{r translate-top-words, fig.width=10, fig.height=9}
# Translation lookup for non-English words in the model
translations <- tribble(
  ~word,                ~english,              ~lang,
  "los",               "loose",               "NL",
  "niet",              "not",                 "NL",
  "snoer",             "cord",                "NL",
  "bot",               "blunt/dull",          "NL",
  "vuil",              "dirty",               "NL",
  "printplaat",        "circuit board",       "NL",
  "zat",               "was stuck",           "NL",
  "draad",             "wire",                "NL",
  "lang",              "long/slow",           "NL",
  "stekker",           "plug",                "NL",
  "gevonden",          "found",               "NL",
  "laadt",             "charges",             "NL",
  "contacten",         "contacts",            "NL",
  "accu",              "battery",             "NL",
  "meer",              "more",                "NL",
  "doorgebrand",       "burned out",          "NL",
  "waarschijnlijk",    "probably",            "NL",
  "verwarmingselement","heating element",     "NL",
  "onbekend",          "unknown",             "NL",
  "schoongemaakt",     "cleaned",             "NL",
  "vervuild",          "contaminated",        "NL",
  "vast",              "stuck",               "NL",
  "kabel",             "cable",               "NL",
  "versleten",         "worn out",            "NL",
  "elektronica",       "electronics",         "NL",
  "vervangen",         "replaced",            "NL",
  "maar",              "but",                 "NL",
  "schakelaar",        "switch",              "NL",
  "losse",             "loose (adj.)",        "NL",
  "ketting",           "chain",               "NL",
  "draadbreuk",        "wire break",          "NL",
  "voeding",           "power supply",        "NL",
  "goed",              "good",                "NL",
  "oplader",           "charger",             "NL",
  "eraf",              "off (it)",            "NL",
  "komt",              "comes",               "NL",
  "zit",               "sits/is stuck",       "NL",
  "kabelbreuk",        "cable break",         "NL",
  "pomp",              "pump",                "NL",
  "band",              "belt/band",           "NL",
  "fil",               "wire",                "FR",
  "nettoyage",         "cleaning",            "FR",
  "carte",             "board/card",          "FR",
  "moteur",            "motor",               "FR",
  "résistance",       "resistor",            "FR",
  "mais",              "but",                 "FR",
  "pas",               "not/step",            "FR",
  "ne",                "not",                 "FR",
  "und",               "and",                 "DE",
  "trafo",             "transformer",         "DE"
)

# Annotate the full coefficient table with translations
coef_annotated <- coef_df %>%
  left_join(translations, by = "word") %>%
  mutate(
    is_foreign = !is.na(lang),
    label = case_when(
      !is.na(english) ~ paste0(english, " [", lang, ": ", word, "]"),
      TRUE ~ word
    )
  )

# Plot top 30 words with translations
top30 <- coef_annotated %>%
  slice_head(n = 30) %>%
  mutate(label = fct_reorder(label, abs(estimate)))

ggplot(top30, aes(x = abs(estimate), y = label, fill = direction)) +
  geom_col() +
  geom_text(
    data = top30 %>% filter(is_foreign),
    aes(x = abs(estimate) + 0.002, label = paste0("(", lang, ")")),
    hjust = 0, size = 3, color = "#888888"
  ) +
  scale_fill_manual(values = c("Predicts Repair" = "#2E86AB", "Predicts Failure" = "#E84855")) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "The Multilingual Vocabulary of Repair",
    subtitle = "Top 30 model features with non-English words translated to English",
    x = "Absolute Coefficient (Importance)", y = NULL, fill = NULL,
    caption = "Data Source: Open Repair Alliance | DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme(legend.position = "top")
```

Now the patterns jump out. The model learned the same concepts across languages:

- **Fixable signals**: "circuit board" (NL: *printplaat*), "motor," "heating element" (NL: *verwarmingselement*), "charges" (NL: *laadt*), "battery" (NL: *accu*) — these all point to **identifiable, replaceable components**. If a volunteer can name the broken part, they can probably swap it out.
- **Failure signals**: "dirty" (NL: *vuil*), "cord" (NL: *snoer*), "wire" (NL: *draad*, FR: *fil*), "plug" (NL: *stekker*), "loose" (NL: *los*), "contacts" (NL: *contacten*) — these describe **diffuse electrical problems** that are harder to isolate and often indicate deeper damage.
- **The sharpening dead-end**: "blunt" and "sharpening" both predict failure. Dull blades brought to repair cafés often can't be resharpened with the tools on hand.

The Dutch dominance makes sense — the Netherlands has **90,839 logged repairs**, more than any other country. The repair café movement was literally [founded in Amsterdam in 2009](https://repaircafe.org/en/about/), so Dutch volunteers have been logging data the longest.

## The Final Verdict: The Words That Fix (and Fail) Your Gadgets

Let's bring it all together with a single visualization that tells the whole story — which words in a repair description signal hope, and which signal doom?

```{r final-viz, fig.width=14, fig.height=12, dpi=300}
# Repair Café brand palette (from repaircafe.org)
rc_navy   <- "#2D2E82"
rc_orange <- "#ED6A42"
rc_bg     <- "#F2F2FF"
rc_dark   <- "#1C1C57"
rc_light  <- "#E8E8FF"

# Prepare the logo: add navy background to the transparent PNG
logo_path <- ".kiro/specs/2026_04_07_tidy_tuesday_repairs/rc_logo.png"
logo_branded_path <- ".kiro/specs/2026_04_07_tidy_tuesday_repairs/rc_logo_branded.png"

logo_raw <- image_read(logo_path)
logo_info <- image_info(logo_raw)
navy_bg <- image_blank(logo_info$width, logo_info$height, color = rc_navy)
logo_branded <- image_composite(navy_bg, logo_raw, operator = "over")
logo_branded <- image_border(logo_branded, color = rc_navy, geometry = "40x20")
image_write(logo_branded, logo_branded_path)

# Build the final dataset: top translated words, signed by direction
final_words <- coef_annotated %>%
  filter(str_detect(term, "tfidf_")) %>%
  slice_head(n = 30) %>%
  mutate(
    signed_est = if_else(direction == "Predicts Repair", abs(estimate), -abs(estimate)),
    display_label = case_when(
      !is.na(english) & word != english ~ paste0(word, " (", lang, ": ", english, ")"),
      TRUE ~ word
    ),
    display_label = fct_reorder(display_label, signed_est),
    bar_color = if_else(direction == "Predicts Repair", rc_navy, rc_orange),
    # Decide: all labels go outside the bar
    label_x = if_else(signed_est > 0, signed_est + 0.004, signed_est - 0.004),
    label_hjust = if_else(signed_est > 0, 0, 1),
    label_color = bar_color
  )

# Custom Repair Café theme
theme_repaircafe <- theme_minimal(base_size = 13) +
  theme(
    text = element_text(family = "Helvetica Neue", color = rc_dark),
    plot.background = element_rect(fill = rc_bg, color = NA),
    panel.background = element_rect(fill = rc_bg, color = NA),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = rc_light, linewidth = 0.4),
    plot.title = element_text(
      size = 22, face = "bold", color = rc_dark,
      margin = margin(b = 12), hjust = 0.5
    ),
    plot.title.position = "plot",
    plot.subtitle = element_text(
      size = 11.5, color = "#555577", lineheight = 1.3,
      margin = margin(t = 4, b = 8), hjust = 0.5
    ),
    plot.caption = element_text(
      size = 9, color = "#8888AA", hjust = 0.5,
      margin = margin(t = 12)
    ),
    plot.margin = margin(10, 20, 12, 10),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size = 10, color = "#555577"),
    axis.title.x = element_text(size = 11, color = "#555577", margin = margin(t = 8)),
    legend.position = "none"
  )

n_words <- nrow(final_words)

base_plot <- ggplot(final_words, aes(x = signed_est, y = display_label)) +
  # Bars
  geom_col(aes(fill = bar_color), width = 0.7) +
  scale_fill_identity() +
  # Word labels outside the bars
  geom_text(
    aes(x = label_x, label = display_label,
        hjust = label_hjust, color = label_color),
    size = 3.5, fontface = "bold", family = "Helvetica Neue"
  ) +
  scale_color_identity() +
  # Coefficient values inside the bars, near the end
  geom_text(
    aes(
      x = signed_est - if_else(signed_est > 0, 0.004, -0.004),
      label = sprintf("%+.3f", signed_est),
      hjust = if_else(signed_est > 0, 1, 0)
    ),
    color = "white", size = 2.7, family = "Helvetica Neue"
  ) +
  # Zero line
  geom_vline(xintercept = 0, linewidth = 0.7, color = rc_dark) +
  # Direction labels above the data
  annotate(
    "label", x = 0.08, y = n_words + 1.5,
    label = "SIGNALS FIXABLE -->",
    fill = rc_navy, color = "white",
    family = "Helvetica Neue",
    size = 3.8, fontface = "bold",
    label.padding = unit(0.5, "lines"),
    label.r = unit(0.3, "lines")
  ) +
  annotate(
    "label", x = -0.08, y = n_words + 1.5,
    label = "<-- SIGNALS FAILURE",
    fill = rc_orange, color = "white",
    family = "Helvetica Neue",
    size = 3.8, fontface = "bold",
    label.padding = unit(0.5, "lines"),
    label.r = unit(0.3, "lines")
  ) +
  scale_x_continuous(
    breaks = seq(-0.15, 0.15, 0.05),
    labels = function(x) sprintf("%+.2f", x),
    expand = expansion(mult = c(0.12, 0.12))
  ) +
  scale_y_discrete(expand = expansion(add = c(0.5, 2.5))) +
  labs(
    title = "The Words That Predict Whether Your Gadget Gets Fixed",
    subtitle = paste0(
      "Lasso logistic regression on ", comma(nrow(df)), " multilingual repair descriptions from 25 countries\n",
      "AUC = ", round(auc_val$.estimate, 2),
      " | Accuracy = ", percent(acc$.estimate, accuracy = 0.1),
      " | Non-English words shown as: word (translation)"
    ),
    x = "Model Coefficient (Predictive Strength)",
    y = NULL,
    caption = "Data Source: Open Repair Alliance via TidyTuesday | Languages: NL = Dutch, FR = French, DE = German, EN = English | DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme_repaircafe

# Stack logo above the main plot
logo_strip <- ggdraw() +
  draw_image(logo_branded_path, x = 0.5, y = 0.5, width = 0.22,
             hjust = 0.5, vjust = 0.5) +
  theme(plot.background = element_rect(fill = rc_bg, color = NA))

p <- plot_grid(
  logo_strip,
  base_plot,
  ncol = 1,
  rel_heights = c(0.06, 0.94)
)

p
```

## What's Next?

This analysis only scratches the surface. A few open threads worth pulling:

- **Beyond bag-of-words**: We translated the model's top features, but a true multilingual embedding model (e.g., multilingual BERT) could capture semantic meaning across languages simultaneously — likely pushing accuracy well past 70%.
- **Repairability scores**: The dataset includes a 0–10 repairability rating. Could we build a model that predicts this continuous score, giving manufacturers feedback on design choices?
- **The economics of repair**: With estimated production years available, we could study whether newer products are harder or easier to fix — a direct test of the "planned obsolescence" hypothesis.

The repair café movement is a quiet revolution. Every fixed toaster is a small act of defiance against throwaway culture. And as this data shows, the odds are in your favor — bring your broken stuff, and more often than not, it'll walk out the door working again.

```{r save-image, include=FALSE}
# Full desktop version saved to specs folder only (not repo root)
ggsave(
  filename = ".kiro/specs/2026_04_07_tidy_tuesday_repairs/2026_04_07_tidy_tuesday_repairs_full.png",
  plot = p,
  device = "png",
  width = 14,
  height = 12,
  dpi = 300,
  bg = "#F2F2FF"
)
```

Here's a mobile-friendly version of the same chart — fewer words, bigger text, designed to be readable on a phone screen.

```{r mobile-viz, fig.width=9, fig.height=9, dpi=300}
# Mobile version: top 10 words, larger fonts, two-line labels for translations
mobile_words <- coef_annotated %>%
  filter(str_detect(term, "tfidf_")) %>%
  slice_head(n = 10) %>%
  mutate(
    signed_est = if_else(direction == "Predicts Repair", abs(estimate), -abs(estimate)),
    display_label = case_when(
      !is.na(english) & word != english ~ paste0(word, "\n(", english, ")"),
      TRUE ~ word
    ),
    display_label = fct_reorder(display_label, signed_est),
    bar_color = if_else(direction == "Predicts Repair", rc_navy, rc_orange),
    label_x = if_else(signed_est > 0, signed_est + 0.004, signed_est - 0.004),
    label_hjust = if_else(signed_est > 0, 0, 1),
    label_color = bar_color
  )

n_mobile <- nrow(mobile_words)

theme_mobile <- theme_minimal(base_size = 16) +
  theme(
    text = element_text(family = "Helvetica Neue", color = rc_dark),
    plot.background = element_rect(fill = rc_bg, color = NA),
    panel.background = element_rect(fill = rc_bg, color = NA),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = rc_light, linewidth = 0.4),
    plot.title = element_text(
      size = 18, face = "bold", color = rc_dark,
      margin = margin(b = 10), hjust = 0.5
    ),
    plot.title.position = "plot",
    plot.subtitle = element_text(
      size = 12, color = "#555577", lineheight = 1.3,
      margin = margin(t = 2, b = 10), hjust = 0.5
    ),
    plot.caption = element_text(
      size = 9, color = "#8888AA", hjust = 0.5,
      margin = margin(t = 10)
    ),
    plot.margin = margin(10, 15, 10, 10),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "#555577"),
    axis.title.x = element_text(size = 13, color = "#555577", margin = margin(t = 8)),
    legend.position = "none"
  )

mobile_base <- ggplot(mobile_words, aes(x = signed_est, y = display_label)) +
  geom_col(aes(fill = bar_color), width = 0.85) +
  scale_fill_identity() +
  # Word labels outside — with lineheight for two-line labels
  geom_text(
    aes(x = label_x, label = display_label,
        hjust = label_hjust, color = label_color),
    size = 4.2, fontface = "bold", family = "Helvetica Neue",
    lineheight = 0.85
  ) +
  scale_color_identity() +
  # Coefficient values inside
  geom_text(
    aes(
      x = signed_est - if_else(signed_est > 0, 0.004, -0.004),
      label = sprintf("%+.3f", signed_est),
      hjust = if_else(signed_est > 0, 1, 0)
    ),
    color = "white", size = 3.5, family = "Helvetica Neue"
  ) +
  geom_vline(xintercept = 0, linewidth = 0.8, color = rc_dark) +
  annotate(
    "label", x = 0.08, y = n_mobile + 1,
    label = "FIXABLE -->",
    fill = rc_navy, color = "white",
    family = "Helvetica Neue",
    size = 4.5, fontface = "bold",
    label.padding = unit(0.5, "lines"),
    label.r = unit(0.3, "lines")
  ) +
  annotate(
    "label", x = -0.08, y = n_mobile + 1,
    label = "<-- FAILURE",
    fill = rc_orange, color = "white",
    family = "Helvetica Neue",
    size = 4.5, fontface = "bold",
    label.padding = unit(0.5, "lines"),
    label.r = unit(0.3, "lines")
  ) +
  scale_x_continuous(
    breaks = seq(-0.15, 0.15, 0.05),
    labels = function(x) sprintf("%+.2f", x),
    expand = expansion(mult = c(0.18, 0.18))
  ) +
  scale_y_discrete(expand = expansion(add = c(0.5, 2))) +
  labs(
    title = "Words That Predict If Your Gadget Gets Fixed",
    subtitle = paste0(
      "Top 10 lasso model features | AUC = ", round(auc_val$.estimate, 2),
      " | Accuracy = ", percent(acc$.estimate, accuracy = 0.1)
    ),
    x = "Coefficient",
    y = NULL,
    caption = "Data: Open Repair Alliance | Translations from Dutch | DataViz: Tony Galvan (@GDataScience1) | #TidyTuesday"
  ) +
  theme_mobile

# Stack logo above
mobile_logo <- ggdraw() +
  draw_image(logo_branded_path, x = 0.5, y = 0.5, width = 0.35,
             hjust = 0.5, vjust = 0.5) +
  theme(plot.background = element_rect(fill = rc_bg, color = NA))

p_mobile <- plot_grid(
  mobile_logo,
  mobile_base,
  ncol = 1,
  rel_heights = c(0.06, 0.94)
)

p_mobile
```

```{r save-mobile}
ggsave(
  filename = "2026_04_07_tidy_tuesday_repairs.png",
  plot = p_mobile,
  device = "png",
  width = 9,
  height = 9,
  dpi = 300,
  bg = "#F2F2FF"
)
```

```{r cleanup, include=FALSE}
if (file.exists("ml_results.rds")) file.remove("ml_results.rds")
```
