---
title: "TidyTemplate"
date: 2025-01-21
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(janitor)

theme_set(theme_light())

blue <- "#0099F8"
red <- "#F06449"
black <- "#36382E"
grey <- "#DADAD9"
purple <- "#624CAB"
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-01-21")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

exped_tidy <- tt$exped_tidy |>
  clean_names()

peaks_tidy <- tt$peaks_tidy |>
  clean_names()

exped <- exped_tidy |>
  left_join(peaks_tidy, by = "peakid") |>
  mutate(total_deaths = mdeaths + hdeaths)

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

# Reimagine Faline Rezvani's plot
exped |>
  filter(total_deaths > 0) |>
  count(pkname) |>
  mutate(pkname = fct_reorder(pkname, n)) |>
  ggplot(aes(n, pkname)) +
  geom_col(color = black, fill = blue) +
  labs(
    x = "Number of Fatal Expeditions", 
    y = "Himalayan Peak", 
    title = "36 Fatal Expeditions by Peak",
    caption = "Source: The Himalayan Database\nALT TEXT: This bar chart shows the 36 fatal expeditions distributed over peaks. The Everest peak saw 21 fatal expeditions."
  )

exped |>
  filter(total_deaths > 0) |>
  count(pkname, wt = total_deaths) |>
  mutate(pkname = fct_reorder(pkname, n)) |>
  ggplot(aes(n, pkname)) +
  geom_col(color = black, fill = red) +
  labs(
    x = "Number of Fatalities", 
    y = "Himalayan Peak", 
    title = "53 Fatalities by Peak",
    caption = "Source: The Himalayan Database\nALT TEXT: This bar chart shows the 53 fatalities distributed over peaks. The Everest peak saw 34 fatalities."
  )

exped |>
  count(year) |>
  ggplot(aes(year, n)) +
  geom_col(color = black, fill = purple) +
  labs(
    x = NULL, 
    y = "# of expeditions", 
    title = "Expeditions by Year",
    caption = "Source: The Himalayan Database\nALT TEXT: This bar chart shows the number of expeditions by year. The year 2022 saw the most expeditions."
  )

exped_bcmonth <- exped |>
  filter(!is.na(bcdate)) |>
  mutate(
    base_camp_month = month(bcdate)
  ) |>
  count(base_camp_month, year) |>
  bind_rows(
    data.frame(base_camp_month = 2, year = 2020:2024, n = 0)
  ) |>
  bind_rows(
    data.frame(base_camp_month = 7, year = 2020:2024, n = 0)
  ) |>
  left_join(
    tibble(month_abb = month.abb) |>
      mutate(month_num = 1:12),
    by = c("base_camp_month" = "month_num")
  ) |> 
  select(-base_camp_month) |>
  pivot_wider(
    names_from = month_abb,
    values_from = n,
    values_fill = 0
  ) |>
  pivot_longer(
    cols = -year,
    names_to = "month",
    values_to = "n"
  ) |>
  mutate(month = factor(month, levels = month.abb)) 

exped_bcmonth |>
  ggplot(aes(month, n, color = factor(year), group = year)) +
  geom_line() +
  labs(
    x = "Month arrived at base camp", 
    y = "# of expeditions", 
    color = NULL,
    title = "Expeditions by Month",
    subtitle = "Most expeditions arrive at base camp in April.",
    caption = "Source: The Himalayan Database\nALT TEXT: This line chart shows the number of expeditions by month. The year 2022 saw the most expeditions."
  )

exped_bcmonth |>
  ggplot(aes(factor(year), month, fill = n)) +
  geom_tile(color = grey) +
  scale_fill_viridis_c() +
  scale_y_discrete(limits = rev(month.abb)) +
  scale_x_discrete(position = "top") +
  labs(
    x = NULL, 
    y = NULL, 
    fill = "# of expeditions", 
    title = "Expeditions by Base Camp Arrival Month and Year",
    caption = "Source: The Himalayan Database\nALT TEXT: This heatmap shows the number of expeditions by month and year. The year 2022 saw the most expeditions."
  )

exped_smtmonth <- exped |>
  filter(!is.na(smtdate)) |>
  mutate(
    summit_month = month(smtdate)
  ) |>
  count(summit_month, year) |>
  bind_rows(
    data.frame(summit_month = 2, year = 2020:2024, n = 0)
  ) |>
  left_join(
    tibble(month_abb = month.abb) |>
      mutate(month_num = 1:12),
    by = c("summit_month" = "month_num")
  ) |> 
  select(-summit_month) |>
  pivot_wider(
    names_from = month_abb,
    values_from = n,
    values_fill = 0
  ) |>
  pivot_longer(
    cols = -year,
    names_to = "month",
    values_to = "n"
  ) |>
  mutate(month = factor(month, levels = month.abb)) 

exped_smtmonth |>
  ggplot(aes(month, n, color = factor(year), group = year)) +
  geom_line() +
  labs(
    x = "Month reached summit", 
    y = "# of expeditions", 
    color = NULL,
    title = "Expeditions by Month",
    subtitle = "Most expeditions reached summit in May.",
    caption = "Source: The Himalayan Database\nALT TEXT: This line chart shows the number of expeditions by month. The year 2022 saw the most expeditions."
  )

exped_smtmonth |>
  ggplot(aes(factor(year), month, fill = n)) +
  geom_tile(color = grey) +
  scale_fill_viridis_c() +
  scale_y_discrete(limits = rev(month.abb)) +
  scale_x_discrete(position = "top") +
  labs(
    x = NULL, 
    y = NULL, 
    fill = "# of expeditions", 
    title = "Expeditions by Summit Month and Year",
    caption = "Source: The Himalayan Database\nALT TEXT: This heatmap shows the number of expeditions by month and year. The year 2022 saw the most expeditions."
  )

# library(ggdist)
# 
# peaks_tidy |>
#   ggplot(aes(heightf, pkname, fill = pkname)) +
#   stat_slab(aes(thickness = after_stat(n)), scale = 0.7) +
#   stat_dotsinterval(side = "bottom", scale = 0.7, slab_linewidth = NA) #+
#   # scale_fill_brewer(palette = "Set2")
# 
# set.seed(12345) # for reproducibility
# 
# tibble(
#   abc = rep(c("a", "b", "b", "c"), 50),
#   value = rnorm(200, c(1, 8, 8, 3), c(1, 1.5, 1.5, 1))
# ) %>%
#   ggplot(aes(y = abc, x = value, fill = abc)) +
#   stat_slab(aes(thickness = after_stat(pdf*n)), scale = 0.7) +
#   stat_dotsinterval(side = "bottom", scale = 0.7, slab_linewidth = NA) +
#   scale_fill_brewer(palette = "Set2") +
#   ggtitle(
#     paste0(
#       'stat_slab(aes(thickness = after_stat(pdf*n)), scale = 0.7) +\n',
#       'stat_dotsinterval(side = "bottom", scale = 0.7, slab_linewidth = NA)'
#     ),
#     'aes(fill = abc)'
#   )

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_01_21_tidy_tuesday_exped.png",
  device = "png"
)
```
