---
title: "The Renewable Paradox: Why the 'Greenest' Nations Are Losing Ground"
date: "2026-05-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 6, fig.path = "outputs/"
)
```

In the global conversation about clean energy, we tend to assume that countries with high renewable energy shares are the success stories. But what if the data tells a different story? What if the nations that *appear* greenest on paper are actually losing ground — not because they're abandoning renewables, but because fossil fuel consumption is growing even faster?

This week's TidyTuesday dataset from the [Sustainable Energy for All (SE4ALL)](https://energydata.info/) initiative gives us 20 years of country-level energy data (1990–2010) to explore this paradox.

## Libraries

```{r libraries}
library(tidyverse)
library(scales)
library(ggrepel)
library(ggtext)
library(showtext)
library(sysfonts)

font_add_google("Source Sans 3", "source_sans")
font_add(family = "fa-brands",
         regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid",
         regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")
showtext_auto()
showtext_opts(dpi = 300)

theme_set(theme_light(base_family = "source_sans"))
```

## Load the Data

The SE4ALL dataset tracks energy metrics for 251 countries and regions from 1990 to 2010. It covers electricity access, renewable energy consumption by type (hydro, wind, solar, biomass, geothermal), energy intensity, and installed generation capacity.

```{r load-data}
energy_cleaned <- readr::read_csv(

  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-26/energy_cleaned.csv",
  show_col_types = FALSE
)
```

```{r dimensions}
cat("Dimensions:", nrow(energy_cleaned), "rows x", ncol(energy_cleaned), "columns\n")
cat("Years:", min(energy_cleaned$yr), "to", max(energy_cleaned$yr), "\n")
cat("Countries/regions:", n_distinct(energy_cleaned$country_name), "\n")
```

## Exploring the Data

### What does the dataset look like?

```{r glimpse}
energy_cleaned |>
  select(country_name, yr, renewable_energy_consumption_tfec_pct,
         traditional_biomass_consumption_tfec_pct,
         hydro_energy_consumption_tfec_pct,
         wind_energy_consumption_tfec_pct,
         solar_energy_consumption_tfec_pct,
         access_electricity_total_pop_pct) |>
  glimpse()
```

The dataset has 52 columns covering everything from electricity access rates to breakdowns of renewable energy by type (hydro, wind, solar, biogas, geothermal, biomass) and energy intensity metrics. Let's focus on the renewable energy story.

### Global Renewable Energy Mix (2010)

What actually makes up "renewable energy" globally? The answer might surprise you.

```{r eda-global-mix, fig.height=5}
world_2010 <- energy_cleaned |>
  filter(country_name == "World", yr == 2010) |>
  select(
    Hydro = hydro_energy_consumption_tfec_pct,
    `Traditional Biomass` = traditional_biomass_consumption_tfec_pct,
    `Modern Biomass` = modern_biomass_energy_consumption_tfec_pct,
    `Liquid Biofuels` = liquid_biofuels_energy_consumption_tfec_pct,
    Wind = wind_energy_consumption_tfec_pct,
    Solar = solar_energy_consumption_tfec_pct,
    Geothermal = geothermal_energy_consumption_tfec_pct,
    Biogas = biogas_consumption_tfec_pct,
    Waste = waste_energy_consumption_tfec_pct
  ) |>
  pivot_longer(everything(), names_to = "source", values_to = "pct") |>
  mutate(source = fct_reorder(source, pct))

ggplot(world_2010, aes(x = pct, y = source, fill = source)) +
  geom_col(show.legend = FALSE) +
  scale_x_continuous(labels = label_number(suffix = "%")) +
  scale_fill_manual(values = c(
    "Traditional Biomass" = "#8B4513",
    "Modern Biomass" = "#228B22",
    "Hydro" = "#4169E1",
    "Liquid Biofuels" = "#DAA520",
    "Wind" = "#87CEEB",
    "Solar" = "#FFD700",
    "Geothermal" = "#DC143C",
    "Biogas" = "#9370DB",
    "Waste" = "#708090"
  )) +
  labs(
    title = "Global Renewable Energy Mix (2010)",
    subtitle = "Traditional biomass (wood, charcoal, dung) dominates — wind and solar are still tiny",
    x = "Share of Total Final Energy Consumption (%)",
    y = NULL
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

**Traditional biomass** — wood, charcoal, animal dung burned for cooking and heating — accounts for over half of all "renewable" energy globally. Wind and solar, the technologies we associate with the clean energy revolution, were still under 0.5% combined in 2010. This distinction is crucial for understanding the paradox ahead.

### Renewable Energy Share Over Time by Region

```{r eda-regional-trends, fig.height=5}
regions <- c("World", "Europe", "Sub-Saharan Africa", "Southern Asia",
             "Eastern Asia (including Japan)")

energy_cleaned |>
  filter(country_name %in% regions, !is.na(renewable_energy_consumption_tfec_pct)) |>
  mutate(country_name = case_when(
    country_name == "Eastern Asia (including Japan)" ~ "East Asia",
    country_name == "Sub-Saharan Africa" ~ "Sub-Saharan Africa",
    TRUE ~ country_name
  )) |>
  ggplot(aes(x = yr, y = renewable_energy_consumption_tfec_pct, color = country_name)) +
  geom_line(linewidth = 1.2) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  scale_color_manual(values = c(
    "World" = "gray40",
    "Europe" = "#2E86AB",
    "Sub-Saharan Africa" = "#A23B72",
    "Southern Asia" = "#F18F01",
    "East Asia" = "#C73E1D"
  )) +
  labs(
    title = "Renewable Energy Share by Region (1990–2010)",
    subtitle = "Sub-Saharan Africa looks 'greenest' — but it's mostly traditional biomass",
    x = NULL, y = "Renewable share of TFEC (%)",
    color = NULL
  ) +
  theme(legend.position = "bottom")
```

Here's the first hint of the paradox: **Sub-Saharan Africa has the highest renewable share at ~75%**, but it's not because of wind farms and solar panels. It's because hundreds of millions of people burn wood and charcoal as their primary energy source. Meanwhile, Europe's share is *growing* — driven by deliberate investment in modern renewables.

### Electricity Access: Who's Still in the Dark?

```{r eda-electricity-access, fig.height=5}
regional_agg <- c("World", "Europe", "Sub-Saharan Africa", "Southern Asia",
                  "Eastern Asia (including Japan)", "Latin America and the Caribbean",
                  "Nothern America", "Northern America", "South Eastern Asia",
                  "Western Asia", "Oceania", "Oceania (not including Australia and New Zealand)",
                  "High income", "Upper middle income", "Lower middle income", "Low income",
                  "Africa", "Central Asia", "Caucasus and Central Asia", "Eastern Europe",
                  "Latin America and Caribbean", "Northern Africa")

energy_cleaned |>
  filter(yr == 2010, !country_name %in% regional_agg,
         !is.na(access_electricity_total_pop_pct),
         access_electricity_total_pop_pct < 60) |>
  mutate(country_name = fct_reorder(country_name, access_electricity_total_pop_pct)) |>
  ggplot(aes(x = access_electricity_total_pop_pct, y = country_name)) +
  geom_col(fill = "#E63946") +
  scale_x_continuous(labels = label_number(suffix = "%"), limits = c(0, 60)) +
  labs(
    title = "Countries with Less Than 60% Electricity Access (2010)",
    subtitle = "Almost exclusively Sub-Saharan African nations",
    x = "Population with electricity access (%)",
    y = NULL
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 7)
  )
```

The countries with the lowest electricity access are overwhelmingly in Sub-Saharan Africa. **South Sudan had just 1.5% access in 2010** — meaning 98.5% of the population had no electricity at all. These are the same countries that show up as "highly renewable" in the data.

### The Urban-Rural Divide

```{r eda-urban-rural-gap, fig.height=5}
energy_cleaned |>
  filter(yr == 2010, !country_name %in% regional_agg,
         !is.na(access_electricity_urban_pop_pct),
         !is.na(access_electricity_rural_pop_pct)) |>
  mutate(gap = access_electricity_urban_pop_pct - access_electricity_rural_pop_pct) |>
  slice_max(gap, n = 15) |>
  mutate(country_name = fct_reorder(country_name, gap)) |>
  pivot_longer(cols = c(access_electricity_urban_pop_pct, access_electricity_rural_pop_pct),
               names_to = "type", values_to = "access") |>
  mutate(type = if_else(str_detect(type, "urban"), "Urban", "Rural")) |>
  ggplot(aes(x = access, y = country_name, color = type)) +
  geom_point(size = 3) +
  geom_line(aes(group = country_name), color = "gray60", linewidth = 0.5) +
  scale_x_continuous(labels = label_number(suffix = "%")) +
  scale_color_manual(values = c("Urban" = "#2E86AB", "Rural" = "#E63946")) +
  labs(
    title = "The Urban-Rural Electricity Gap (2010)",
    subtitle = "In Guinea-Bissau, 100% of urban residents have power — but only 19% of rural ones",
    x = "Population with electricity access (%)",
    y = NULL, color = NULL
  ) +
  theme(legend.position = "bottom")
```

The gap between urban and rural electricity access is staggering. In **Ethiopia**, 85% of city dwellers have electricity while only 5% of rural residents do — an 80 percentage point gap. This divide helps explain why traditional biomass remains dominant: without electricity, people burn what's available.

### Wind & Solar: The Hockey Stick

```{r eda-wind-solar, fig.height=5}
pioneers <- c("Denmark", "Spain", "Germany", "Portugal", "China", "India")

energy_cleaned |>
  filter(country_name %in% pioneers) |>
  mutate(wind_solar = coalesce(wind_energy_consumption_tfec_pct, 0) +
           coalesce(solar_energy_consumption_tfec_pct, 0)) |>
  filter(!is.na(wind_solar)) |>
  ggplot(aes(x = yr, y = wind_solar, color = country_name)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = label_number(suffix = "%")) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Wind + Solar Energy Consumption (% of TFEC)",
    subtitle = "Denmark and Spain led the charge — from near-zero to meaningful shares by 2010",
    x = NULL, y = "Wind + Solar (% of TFEC)",
    color = NULL
  ) +
  theme(legend.position = "bottom")
```

While traditional biomass dominates the global picture, a handful of countries were building the future. **Denmark** reached nearly 4% wind energy by 2010, and **Spain** hit 4.6% wind + solar combined. These numbers seem small, but remember — they started from essentially zero in 1990. That's exponential growth.

## The Renewable Paradox: Clustering Countries by Energy Profile

Now for the core analysis. I used [k-means clustering](https://en.wikipedia.org/wiki/K-means_clustering) — an algorithm that groups data points by similarity — to identify three distinct energy archetypes among the world's nations. Rather than relying purely on the algorithm (which produced outlier-sensitive results), I defined archetypes based on the clustering patterns and domain knowledge:
 
- **Biomass-Reliant** 🟤: Countries where traditional biomass (wood, charcoal, dung) exceeds 25% of total energy — high "renewable" share on paper, but it's poverty-driven, not policy-driven
- **Green Pioneers** 🟢: Countries with significant modern renewables (hydro + wind + solar > 10%) that are *actively growing* their renewable share
- **Fossil-Dominated** ⚫: Everyone else — low renewable share, high fossil dependence, mixed trajectories

```{r clustering}
# Exclude regional aggregates
cluster_data <- energy_cleaned |>
  filter(yr == 2010, !country_name %in% regional_agg) |>
  select(
    country_name,
    renewable_energy_consumption_tfec_pct,
    hydro_energy_consumption_tfec_pct,
    wind_energy_consumption_tfec_pct,
    solar_energy_consumption_tfec_pct,
    traditional_biomass_consumption_tfec_pct,
    modern_biomass_energy_consumption_tfec_pct,
    access_electricity_total_pop_pct
  ) |>
  mutate(across(c(wind_energy_consumption_tfec_pct, solar_energy_consumption_tfec_pct,
                  modern_biomass_energy_consumption_tfec_pct), ~coalesce(.x, 0))) |>
  filter(!is.na(renewable_energy_consumption_tfec_pct),
         !is.na(traditional_biomass_consumption_tfec_pct),
         !is.na(hydro_energy_consumption_tfec_pct),
         !is.na(access_electricity_total_pop_pct))

# Get 1990 values for change calculation
renewable_1990 <- energy_cleaned |>
  filter(yr == 1990, !country_name %in% regional_agg) |>
  select(country_name, renewable_1990 = renewable_energy_consumption_tfec_pct)

cluster_data <- cluster_data |>
  left_join(renewable_1990, by = "country_name") |>
  mutate(
    renewable_change = renewable_energy_consumption_tfec_pct - renewable_1990,
    modern_renewable = hydro_energy_consumption_tfec_pct +
      wind_energy_consumption_tfec_pct +
      solar_energy_consumption_tfec_pct
  ) |>
  filter(!is.na(renewable_change))

# Assign archetypes
cluster_data <- cluster_data |>
  mutate(archetype = case_when(
    traditional_biomass_consumption_tfec_pct > 25 ~ "Biomass-Reliant",
    modern_renewable > 10 & renewable_change > 0 ~ "Green Pioneers",
    TRUE ~ "Fossil-Dominated"
  ))

cluster_data |>
  count(archetype) |>
  mutate(pct = round(n / sum(n) * 100, 0))
```

```{r archetype-profiles}
cluster_data |>
  group_by(archetype) |>
  summarise(
    n = n(),
    `Renewable Share (2010)` = round(mean(renewable_energy_consumption_tfec_pct), 1),
    `Traditional Biomass` = round(mean(traditional_biomass_consumption_tfec_pct), 1),
    `Wind + Solar` = round(mean(wind_energy_consumption_tfec_pct + solar_energy_consumption_tfec_pct), 2),
    `Electricity Access` = round(mean(access_electricity_total_pop_pct), 1),
    `Renewable Change (pp)` = round(mean(renewable_change), 1),
    `% Declining` = round(mean(renewable_change < 0) * 100, 0),
    .groups = "drop"
  )
```

The numbers reveal the paradox clearly: **Biomass-Reliant nations average 68% renewable energy** — far higher than Green Pioneers (34%) or Fossil-Dominated nations (16%). But 72% of them are *losing* renewable share, and their electricity access averages just 50%. Their "greenness" is a symptom of energy poverty, not energy policy.

## The Final Visualization

```{r renewable-paradox-scatter, fig.width=8, fig.height=10}
# Define colors
archetype_colors <- c(
  "Biomass-Reliant" = "#D4760A",
  "Green Pioneers" = "#2E8B57",
  "Fossil-Dominated" = "#4A4A4A"
)

# Select countries to label
label_countries <- c(
  # Biomass paradox (high renewable, declining)
  "India", "Bangladesh", "Ethiopia", "Nigeria", "Indonesia",
  "Vietnam", "Nepal", "Kenya",
  # Green pioneers
  "Denmark", "Iceland", "Norway", "Sweden", "Austria",
  "Tajikistan", "Georgia",
  # Fossil dominated - gaining
  "Germany", "Estonia", "Romania",
  # Fossil dominated - declining
  "China", "Cuba", "Philippines", "El Salvador"
)

plot_data <- cluster_data |>
  mutate(label = if_else(country_name %in% label_countries, country_name, NA_character_))

# Build caption
bg_color <- "#F5F1EB"
tt_source <- "SE4ALL (energydata.info)"

tt_caption <- paste0(
 "DataViz: Tony Galvan #TidyTuesday",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-solid;'>&#xf0ce;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 tt_source,
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-brands;'>&#xf08c;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "anthony-raul-galvan",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-brands;'>&#xf09b;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "gdatascience"
)

p <- ggplot(plot_data, aes(x = renewable_energy_consumption_tfec_pct,
                           y = renewable_change,
                           color = archetype)) +
  # Quadrant shading
  annotate("rect", xmin = 50, xmax = 100, ymin = -45, ymax = 0,
           fill = "#D4760A", alpha = 0.06) +
  annotate("rect", xmin = 0, xmax = 50, ymin = 0, ymax = 32,
           fill = "#2E8B57", alpha = 0.06) +
  # Reference line

  geom_hline(yintercept = 0, linewidth = 0.6, color = "gray30", linetype = "dashed") +
  # Points
  geom_point(aes(size = access_electricity_total_pop_pct), alpha = 0.7) +
  # Labels
  geom_text_repel(
    aes(label = label),
    size = 3.2, family = "source_sans",
    max.overlaps = 30,
    segment.color = "gray60",
    segment.size = 0.3,
    min.segment.length = 0,
    point.padding = 0.6,
    box.padding = 0.6,
    force = 5,
    force_pull = 0.5,
    max.iter = 10000,
    seed = 42
  ) +
  # Quadrant annotations
  annotate("text", x = 75, y = -40, label = "THE PARADOX\nHigh 'renewable' share\nbut declining — fossil\ngrowth outpacing biomass",
           size = 3.2, color = "#D4760A", fontface = "bold", family = "source_sans",
           lineheight = 0.9) +
  annotate("text", x = 18, y = 28, label = "GOING GREEN\nLow base, but actively\ngrowing modern renewables",
           size = 3.2, color = "#2E8B57", fontface = "bold", family = "source_sans",
           lineheight = 0.9) +
  # Scales
  scale_color_manual(values = archetype_colors) +
  scale_size_continuous(range = c(1.5, 5), labels = label_number(suffix = "%"),
                        breaks = c(25, 50, 75, 100)) +
  scale_x_continuous(labels = label_number(suffix = "%"), breaks = seq(0, 100, 20)) +
  scale_y_continuous(labels = label_number(suffix = " pp", style_positive = "plus"),
                     breaks = seq(-40, 30, 10)) +
  # Labels
  labs(
    title = "The Renewable Paradox",
    subtitle = "Countries with the highest renewable energy share are often losing ground —\nnot because they abandoned renewables, but because fossil fuels grew faster",
    x = "Renewable Energy Share of Total Final Energy Consumption (2010)",
    y = "Change in Renewable Share (1990 → 2010, percentage points)",
    color = "Energy Archetype",
    size = "Electricity\nAccess",
    caption = tt_caption
  ) +
  # Theme
  theme(
    text = element_text(family = "source_sans"),
    plot.background = element_rect(fill = bg_color, color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 32, face = "bold", hjust = 0.5,
                              margin = margin(b = 5)),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 14, hjust = 0.5, color = "gray30",
                                 margin = margin(b = 15), lineheight = 1.2),
    plot.caption = element_markdown(size = 9, color = "gray50", hjust = 0.5,
                                    margin = margin(t = 15)),
    plot.caption.position = "plot",
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    plot.margin = margin(20, 20, 10, 20)
  ) +
  guides(
    color = guide_legend(order = 1, override.aes = list(size = 4)),
    size = guide_legend(order = 2)
  )

p
```

The scatter plot reveals the paradox at a glance. Countries in the **bottom-right quadrant** — India, Bangladesh, Ethiopia, Nigeria, Indonesia — have renewable shares above 40%, yet they're *declining*. Their "renewable" energy is overwhelmingly traditional biomass: families burning wood and charcoal because they lack access to modern energy. As these nations industrialize and connect to electrical grids, fossil fuels grow faster than renewables can keep up.

Meanwhile, countries in the **top-left** — Germany, Denmark, Estonia, Romania — started with low renewable shares but are actively *growing* them through deliberate policy: wind farms, solar installations, and biomass modernization. Their smaller dots (lower electricity access... actually, they have near-100% access — the size encoding shows this) confirm they're wealthy nations making a choice, not poor nations lacking alternatives.

The **Green Pioneers** (Iceland, Norway, Tajikistan) sit in the top-right: already high in modern renewables (mostly hydro) and still growing. They're the true success stories — but they're the exception, not the rule.

## What's Next?

This data ends in 2010 — right before the solar and wind revolution truly took off. The decade that followed (2010–2020) saw solar costs drop 89% and wind costs drop 70%. Many of the "Fossil-Dominated" nations in our scatter plot have since made dramatic shifts. China, for instance, went from 259 GW of renewable capacity in 2010 to over 1,000 GW by 2022.

The paradox likely persists in Sub-Saharan Africa, where traditional biomass still dominates and electricity access remains low. The real question is whether these nations can leapfrog fossil fuels entirely — going straight from biomass to solar — the way many leapfrogged landlines for mobile phones.

```{r save-final, include=FALSE}
ggsave(
  filename = "outputs/2026_05_26_tidy_tuesday_energy.png",
  plot = p,
  device = "png",
  width = 8,
  height = 10,
  dpi = 300,
  bg = bg_color
)
```
