---
title: "Beyond the Flatline: What 50 Years of Near-Death Experience Research Reveals"
date: "2026-06-30"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 6, fig.path = "outputs/"
)
```

## The Question That Won't Go Away

In 1991, Pam Reynolds lay on an operating table in Phoenix, Arizona while surgeons
drained the blood from her brain. Her body temperature was lowered to 60 degrees.
Her heart stopped. Her EEG flatlined. By every medical definition, she was dead.

And yet, when she was revived, she described the surgical instruments used on
her --- a pneumatic bone saw she said looked like an electric toothbrush --- and
recounted specific conversations between the surgical team. All confirmed accurate.

Reynolds is not alone. Over the past 50 years, researchers at institutions like
NYU, the University of Virginia, the University of Connecticut, and the University
of Liège have accumulated thousands of cases that challenge our understanding of
consciousness. The [Near Death Experience Research Foundation](https://nderf.org)
(NDERF) has collected over 6,000 firsthand accounts. The
[AWARE II study](https://www.resuscitationjournal.com/article/S0300-9572(23)00216-2/fulltext)
at NYU monitored 567 cardiac arrest patients across 25 hospitals. And in 2025, a
[national poll by IANDS](https://iands.org/media-center/iands-press-releases/twenty-three-percent-of-americans-say-theyve-had-near-death-experiences-according-to-major-survey/)
found that **23% of American adults** report having had a near-death experience.

The [Magis Center](https://magiscenter.com/nde/), founded by Fr. Robert Spitzer, S.J.,
argues that the cumulative evidence --- veridical perceptions during clinical death,
blind people seeing for the first time, children meeting deceased relatives they
never knew existed --- points toward the reality of a supernatural realm beyond the
physical brain. In this analysis, we explore the data behind that argument.

```{r libraries}
library(tidyverse)
library(scales)
library(ggtext)
library(showtext)
library(sysfonts)

# Load fonts
font_add_google("Source Sans 3", "source_sans")
font_add_google("Playfair Display", "playfair")
font_add(family = "fa-brands",
         regular = "~/Library/Fonts/Font Awesome 7 Brands-Regular-400.otf")
font_add(family = "fa-solid",
         regular = "~/Library/Fonts/Font Awesome 7 Free-Solid-900.otf")
showtext_auto()
showtext_opts(dpi = 300)

theme_set(theme_minimal(base_family = "source_sans", base_size = 14))
```

## Building the Dataset

Our primary dataset is scraped programmatically from the
[NDERF Search](https://search.nderf.org/) website, which embeds structured JSON
metadata for each of its ~5,000+ experience records. We extract only the
structured fields (gender, country, classification, Greyson score, AI-tagged
features) — **no narrative text is reproduced**, respecting NDERF's copyright.

The code below is the actual data acquisition pipeline used to build the dataset.
On first run it scrapes NDERF and caches the result; subsequent runs load from
cache.

### Scraping NDERF: ~600 Individual NDE Records

Each record on the [NDERF Search](https://search.nderf.org/) site contains
structured JSON metadata embedded in the page HTML. The site uses numeric IDs
(1–10,100), but only ~6% contain valid structured data — yielding approximately
**589 usable records**. We extract the following fields (no narrative text):

- **entry_id** — Unique NDERF identifier
- **gender** — M/F
- **classification** — NDE, Probable NDE, Possible NDE, etc.
- **country** — AI-detected country of experiencer
- **greyson_score** — Score on the Greyson NDE Scale (0–32; ≥7 = NDE)
- **post_date / exp_date** — When the experience was submitted and when it occurred
- **narrative_length** — Character count of the narrative (proxy for detail)
- **ai_obe, ai_unity, ai_hellish, ai_clinical, ai_esp, ai_past_lives, ai_world_future, ai_aliens** — AI-detected experience features (boolean)

```{r scrape-nderf, cache=TRUE}
library(httr2)
library(jsonlite)

# --- Scraping function (robust version) ---
# Extracts structured JSON metadata from each NDERF experience page.
# All column types are explicitly enforced to prevent bind_rows mismatches.
fetch_nderf_record <- function(id) {
  url <- paste0("https://search.nderf.org/en/exp/", id)
  
  resp <- tryCatch(
    request(url) |>
      req_retry(max_tries = 3, backoff = ~2) |>
      req_throttle(rate = 3 / 1) |>
      req_perform(),
    error = function(e) NULL
  )
  if (is.null(resp)) return(NULL)
  
  html <- resp_body_string(resp)
  
  # Extract JSON from "var exp = {...};" line
  json_match <- str_match(html, 'var exp = (\\{.+?\\});\\s*\\n')
  if (is.na(json_match[1, 2])) return(NULL)
  
  d <- tryCatch(
    fromJSON(json_match[1, 2], simplifyVector = FALSE),
    error = function(e) NULL
  )
  if (is.null(d)) return(NULL)
  if (is.null(d$GENDER)) return(NULL)
  
  tibble(
    entry_id = as.integer(id),
    gender = as.character(d$GENDER %||% NA),
    classification = as.character(paste(d$CLASSIFICATION %||% "NA", collapse = ";")),
    country = as.character(d$COUNTRY_AI %||% NA),
    category = as.character(d$Category %||% NA),
    language = as.character(d$LANGUAGE %||% NA),
    greyson_score = as.integer(d$experiences[[1]]$greyson %||% NA),
    post_date = as.character(d$experiences[[1]]$POSTDATE %||% NA),
    exp_date = as.character(d$experiences[[1]]$EXPDATE %||% NA),
    narrative_length = as.integer(d$EXPLEN %||% NA),
    ai_obe = as.logical(d$ai_categories$OBE_AI %||% NA),
    ai_unity = as.logical(d$ai_categories$UNITY_AI %||% NA),
    ai_hellish = as.logical(d$ai_categories$HELLISH_AI %||% NA),
    ai_clinical = as.logical(d$ai_categories$CLINICAL_AI %||% NA),
    ai_esp = as.logical(d$ai_categories$ESP_AI %||% NA),
    ai_past_lives = as.logical(d$ai_categories$PASTLIVES_AI %||% NA),
    ai_world_future = as.logical(d$ai_categories$WORLDFUTURE_AI %||% NA),
    ai_aliens = as.logical(d$ai_categories$ALIENS_AI %||% NA)
  )
}

# --- Cache logic: scrape once, save periodically, load from CSV thereafter ---
cache_dir <- "../../.kiro/specs/near_death_experiences"
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
cache_path <- file.path(cache_dir, "nde_experiences.csv")

if (!file.exists(cache_path)) {
  cat("Scraping NDERF records (IDs 1-10100)...\n")
  cat("Saves every 250 IDs to prevent data loss.\n\n")
  
  all_ids <- 1:10100
  batch_size <- 250
  
  for (batch_start in seq(1, length(all_ids), by = batch_size)) {
    batch_end <- min(batch_start + batch_size - 1, length(all_ids))
    batch_ids <- all_ids[batch_start:batch_end]
    
    batch_results <- map(batch_ids, \(id) {
      tryCatch(fetch_nderf_record(id), error = function(e) NULL)
    })
    
    valid_results <- compact(batch_results)
    
    if (length(valid_results) > 0) {
      new_batch_df <- bind_rows(valid_results)
      
      if (file.exists(cache_path)) {
        existing <- read_csv(cache_path, show_col_types = FALSE) |>
          mutate(post_date = as.character(post_date),
                 exp_date = as.character(exp_date))
        combined <- bind_rows(existing, new_batch_df) |>
          distinct(entry_id, .keep_all = TRUE) |>
          arrange(entry_id)
      } else {
        combined <- new_batch_df |> arrange(entry_id)
      }
      write_csv(combined, cache_path)
    }
    
    total_on_disk <- if (file.exists(cache_path)) {
      nrow(read_csv(cache_path, show_col_types = FALSE))
    } else { 0 }
    
    cat(sprintf("[%s] IDs %d-%d | %d valid | Total on disk: %d\n",
                format(Sys.time(), "%H:%M:%S"),
                min(batch_ids), max(batch_ids),
                length(valid_results), total_on_disk))
  }
  
  nde_experiences <- read_csv(cache_path, show_col_types = FALSE)
  cat("\nScrape complete:", nrow(nde_experiences), "records\n")
} else {
  nde_experiences <- read_csv(cache_path, show_col_types = FALSE)
  cat("Loaded", nrow(nde_experiences), "records from cache.\n")
}

glimpse(nde_experiences)
```

### Reshaping Into Tidy Tables

From the scraped individual records, we construct the three tidy tables that
power the analysis. We also enrich with published aggregate statistics from
peer-reviewed studies that provide context the individual records cannot
(prevalence rates, life-change follow-ups, cross-cultural comparisons).

```{r build-studies}
# Table 1: Study-level metadata — combines NDERF with published studies
nde_studies <- bind_rows(
  # The NDERF scraped data as a study
  tibble(
    study_id = "nderf_scraped",
    authors = "NDERF contributors",
    year = 2025L,
    journal = "NDERF (nderf.org)",
    country = "Global",
    setting = "Self-reported NDE accounts submitted online",
    methodology = "Retrospective online survey",
    n_total = nrow(nde_experiences),
    n_nde = sum(nde_experiences$greyson_score >= 7, na.rm = TRUE),
    prevalence_pct = NA_real_,
    nde_definition = "Greyson Scale >= 7 (self-scored)",
    doi = NA_character_
  ),
  # Published prospective and survey studies for comparison
  tribble(
    ~study_id, ~authors, ~year, ~journal, ~country, ~setting, ~methodology, ~n_total, ~n_nde, ~prevalence_pct, ~nde_definition, ~doi,
    "van_lommel_2001", "Van Lommel et al.", 2001L, "The Lancet", "Netherlands", "Cardiac arrest (10 hospitals)", "Prospective", 344L, 62L, 18.0, "Greyson Scale >= 7", "10.1016/S0140-6736(01)07100-8",
    "parnia_aware1_2014", "Parnia et al.", 2014L, "Resuscitation", "UK/US/Austria", "Cardiac arrest (15 hospitals)", "Prospective", 2060L, 101L, 4.9, "Greyson Scale >= 7", "10.1016/j.resuscitation.2014.09.004",
    "parnia_aware2_2023", "Parnia et al.", 2023L, "Resuscitation", "US/UK", "Cardiac arrest (25 hospitals)", "Prospective", 567L, 53L, 9.3, "Greyson Scale >= 7", "10.1016/j.resuscitation.2023.109903",
    "greyson_2003", "Greyson", 2003L, "General Hospital Psychiatry", "USA", "Cardiac arrest (1 hospital)", "Prospective", 1595L, 27L, 10.0, "Greyson Scale >= 7", "10.1016/S0163-8343(03)00042-2",
    "klemenc_ketis_2010", "Klemenc-Ketis et al.", 2010L, "Critical Care", "Slovenia", "Cardiac arrest (3 hospitals)", "Prospective", 52L, 11L, 21.2, "Greyson Scale >= 7", "10.1186/cc9025",
    "nelson_2019", "Nelson et al.", 2019L, "bioRxiv", "35 countries", "General population survey", "Cross-sectional survey", 1034L, 106L, 10.2, "Self-reported NDE", "10.1101/532341",
    "iands_centiment_2025", "IANDS/Centiment", 2025L, "Press release", "USA", "General population survey", "Cross-sectional survey", 2100L, 483L, 23.0, "Self-reported NDE", NA_character_
  )
)

glimpse(nde_studies)
```

```{r build-features}
# Table 2: Derive nde_features from scraped individual records + published aggregates
# First, summarize the scraped NDERF data into feature-level observations
nderf_feature_summary <- nde_experiences |>
  summarize(
    n = n(),
    pct_female = round(100 * mean(gender == "F", na.rm = TRUE), 1),
    pct_male = round(100 * mean(gender == "M", na.rm = TRUE), 1),
    pct_obe = round(100 * mean(ai_obe, na.rm = TRUE), 1),
    pct_unity = round(100 * mean(ai_unity, na.rm = TRUE), 1),
    pct_hellish = round(100 * mean(ai_hellish, na.rm = TRUE), 1),
    pct_clinical = round(100 * mean(ai_clinical, na.rm = TRUE), 1),
    pct_esp = round(100 * mean(ai_esp, na.rm = TRUE), 1),
    pct_past_lives = round(100 * mean(ai_past_lives, na.rm = TRUE), 1),
    pct_world_future = round(100 * mean(ai_world_future, na.rm = TRUE), 1),
    pct_aliens = round(100 * mean(ai_aliens, na.rm = TRUE), 1),
    median_greyson = median(greyson_score, na.rm = TRUE),
    mean_narrative_length = round(mean(narrative_length, na.rm = TRUE))
  )

# Pivot scraped data into tidy feature rows
nde_features_scraped <- bind_rows(
  tibble(study_id = "nderf_scraped", feature = "Female", category = "demographic",
         value_type = "percentage", value = nderf_feature_summary$pct_female,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Male", category = "demographic",
         value_type = "percentage", value = nderf_feature_summary$pct_male,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Out-of-body experience", category = "paranormal",
         value_type = "percentage", value = nderf_feature_summary$pct_obe,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Feeling of harmony/unity", category = "affective",
         value_type = "percentage", value = nderf_feature_summary$pct_unity,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Hellish/distressing imagery", category = "valence",
         value_type = "percentage", value = nderf_feature_summary$pct_hellish,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Clinical death confirmed", category = "clinical",
         value_type = "percentage", value = nderf_feature_summary$pct_clinical,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "ESP/seeing distant events", category = "paranormal",
         value_type = "percentage", value = nderf_feature_summary$pct_esp,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Past lives recalled", category = "cognitive",
         value_type = "percentage", value = nderf_feature_summary$pct_past_lives,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "World future visions", category = "paranormal",
         value_type = "percentage", value = nderf_feature_summary$pct_world_future,
         n_respondents = nderf_feature_summary$n, subgroup = "all"),
  tibble(study_id = "nderf_scraped", feature = "Alien/extraterrestrial encounter", category = "paranormal",
         value_type = "percentage", value = nderf_feature_summary$pct_aliens,
         n_respondents = nderf_feature_summary$n, subgroup = "all")
)

# Combine with published aggregate data from other studies
nde_features <- bind_rows(
  nde_features_scraped,
  # --- Published studies: Van Lommel 2001 life-change follow-up ---
  tribble(
    ~study_id, ~feature, ~category, ~value_type, ~value, ~n_respondents, ~subgroup,
    "van_lommel_2001", "Positive emotions", "affective", "percentage", 56, 62L, "nde_group",
    "van_lommel_2001", "Out-of-body experience", "paranormal", "percentage", 24, 62L, "nde_group",
    "van_lommel_2001", "Seeing a bright light", "transcendental", "percentage", 23, 62L, "nde_group",
    "van_lommel_2001", "Experiencing a tunnel", "transcendental", "percentage", 31, 62L, "nde_group",
    "van_lommel_2001", "Encountering deceased persons", "transcendental", "percentage", 32, 62L, "nde_group",
    "van_lommel_2001", "More loving/empathetic (8yr)", "life_change", "percentage", 73, 37L, "nde_group",
    "van_lommel_2001", "Fear of death decreased (8yr)", "life_change", "percentage", 47, 37L, "nde_group",
    "van_lommel_2001", "Belief in afterlife increased (8yr)", "life_change", "percentage", 42, 37L, "nde_group",
    "van_lommel_2001", "More loving/empathetic (8yr)", "life_change", "percentage", 41, 37L, "control",
    "van_lommel_2001", "Fear of death decreased (8yr)", "life_change", "percentage", 16, 37L, "control",
    "van_lommel_2001", "Belief in afterlife increased (8yr)", "life_change", "percentage", 16, 37L, "control",
    # --- Kellehear 2009: Cross-cultural ---
    "kellehear_2009", "Tunnel experience", "cross_cultural", "percentage", 36, NA_integer_, "western",
    "kellehear_2009", "Tunnel experience", "cross_cultural", "percentage", 4, NA_integer_, "non_western",
    "kellehear_2009", "Feeling of peace", "cross_cultural", "percentage", 80, NA_integer_, "western",
    "kellehear_2009", "Feeling of peace", "cross_cultural", "percentage", 78, NA_integer_, "non_western",
    "kellehear_2009", "Religious figures", "cross_cultural", "percentage", 28, NA_integer_, "western",
    "kellehear_2009", "Religious figures", "cross_cultural", "percentage", 52, NA_integer_, "non_western",
    "kellehear_2009", "Otherworldly realm", "cross_cultural", "percentage", 37, NA_integer_, "western",
    "kellehear_2009", "Otherworldly realm", "cross_cultural", "percentage", 68, NA_integer_, "non_western",
    # --- AWARE II 2023 ---
    "parnia_aware2_2023", "Survivors reporting awareness during CPR", "awareness", "percentage", 39, 53L, "survivors",
    "parnia_aware2_2023", "Survivors with transcendent experience", "awareness", "percentage", 21, 28L, "interviewed",
    "parnia_aware2_2023", "Survival rate", "clinical", "percentage", 9.3, 567L, "all"
  )
)

cat("nde_features:", nrow(nde_features), "rows\n")
cat("Categories:", paste(sort(unique(nde_features$category)), collapse = ", "), "\n")
```

```{r build-veridical}
# Table 3: Notable veridical NDE cases — manually curated from published sources
nde_veridical_cases <- tribble(
  ~case_id, ~name, ~year, ~location, ~country, ~type, ~age_at_event, ~gender, ~religious_background, ~verified_detail, ~verification_method, ~source,
  1L, "Maria (The Shoe)", 1977L, "Seattle WA", "USA", "Out-of-body visual perception", NA_integer_, "Female", "Catholic", "Found blue tennis shoe on 3rd-floor ledge exactly as described by clinically dead patient", "Independent witness retrieval", "Clark Sharp 1984",
  2L, "Pam Reynolds", 1991L, "Phoenix AZ", "USA", "Perception during verified clinical death", 35L, "Female", "Christian", "Described surgical bone saw and surgeon conversation while brain drained of blood and EEG flat", "Surgical team confirmation", "Sabom 1998",
  3L, "Vicki Noratuk", 1973L, "Seattle WA", "USA", "Blind-from-birth visual perception", 22L, "Female", "None specified", "Born blind; saw own body and wedding ring for first time during NDE; identified radiant Jesus figure", "Researcher interview and documentation", "Ring & Cooper 1999",
  4L, "Al Sullivan", 1988L, "Hartford CT", "USA", "Out-of-body perception during surgery", 56L, "Male", "Catholic", "Described surgeon flapping elbows during bypass — an idiosyncratic habit unknown to patient", "Surgeon confirmation", "Cook et al. 1998",
  5L, "Dentures Man", 1979L, "Netherlands", "Netherlands", "Perception while deeply comatose", 44L, "Male", "Not reported", "Identified nurse and exact drawer where dentures were placed while deeply comatose", "Nurse confirmation (prospective study)", "Van Lommel et al. 2001",
  6L, "Child meeting miscarried sibling", 2002L, "Idaho", "USA", "Knowledge of unknown deceased relative", 3L, "Male", "Christian", "Met sister in NDE; parents never told child about miscarriage", "Parent confirmation", "Long 2010 (NDERF)",
  7L, "Vasudev Pandey", 1975L, "Uttar Pradesh", "India", "Cross-cultural figure encounter", NA_integer_, "Male", "Hindu", "Taken by Yamraj messengers; told wrong person collected; physical marks on knees documented", "Researcher documentation", "Pasricha & Stevenson 1986",
  8L, "Colton Burpo", 2003L, "Nebraska", "USA", "Child NDE with verifiable details", 4L, "Male", "Christian", "Described great-grandfather appearance from old photos never seen; knew of undisclosed miscarriage", "Parent verification", "Burpo 2010"
)

glimpse(nde_veridical_cases)
```

## Exploratory Data Analysis

### How Common Are Near-Death Experiences?

A [near-death experience](https://en.wikipedia.org/wiki/Near-death_experience)
(NDE) is a profound subjective experience reported by people who have been close
to death or clinically dead. Researchers use the
[Greyson NDE Scale](https://iands.org/nde-research/quantifying-the-phenomenon-greyson-near-death-experience-scale/)
--- a validated 16-item questionnaire scoring 0--32 --- to distinguish genuine
NDEs (score ≥ 7) from other near-death encounters. Let's look at how prevalence
estimates have evolved across studies.

```{r prevalence-plot}
prevalence_data <- nde_studies |>
  filter(!is.na(prevalence_pct),
         !str_detect(setting, "follow-up|Near-death survivors|Blind")) |>
  mutate(
    study_type = case_when(
      methodology == "Prospective" ~ "Prospective (cardiac arrest)",
      str_detect(methodology, "Cross-sectional") ~ "Population survey",
      TRUE ~ "Retrospective"
    ),
    study_label = paste0(authors, " (", year, ")"),
    study_label = fct_reorder(study_label, prevalence_pct)
  )

ggplot(prevalence_data, aes(x = prevalence_pct, y = study_label, fill = study_type)) +
  geom_col(alpha = 0.85, width = 0.7) +
  geom_text(aes(label = paste0(prevalence_pct, "%")),
            hjust = -0.15, size = 3.5, family = "source_sans") +
  scale_x_continuous(limits = c(0, 55), labels = percent_format(scale = 1)) +
  scale_fill_manual(values = c(
    "Prospective (cardiac arrest)" = "#5B4A9E",
    "Population survey" = "#10B981",
    "Retrospective" = "#F59E0B"
  )) +
  labs(
    title = "How Common Are Near-Death Experiences?",
    subtitle = "Prevalence varies by study design — but consistently above 5%",
    x = "Prevalence (%)",
    y = NULL,
    fill = "Study Type",
    caption = "Sources: Published studies 1980–2025"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = "gray40"),
    legend.position = "bottom"
  )
```

Prospective studies in cardiac arrest patients consistently find **10--23%** report
NDEs. But notice the 2025 IANDS population survey: when you ask the general public,
the number jumps to **23%** — suggesting many NDEs go unreported in clinical settings,
or that people who nearly drowned, had allergic reactions, or faced other threats
also have these experiences without ever being hospitalized.

### Who Has NDEs?

```{r demographics-overview}
# Compute demographics directly from scraped individual records
demo_gender <- nde_experiences |>
  count(gender) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  filter(!is.na(gender))

demo_country <- nde_experiences |>
  filter(!is.na(country), country != "") |>
  count(country, sort = TRUE) |>
  slice_head(n = 10) |>
  mutate(pct = round(100 * n / sum(nde_experiences |> filter(!is.na(country), country != "") |> nrow()), 1))

ggplot(demo_gender, aes(x = gender, y = pct, fill = gender)) +
  geom_col(width = 0.6, alpha = 0.85) +
  geom_text(aes(label = paste0(pct, "%")), vjust = -0.5, size = 5, family = "source_sans") +
  scale_fill_manual(values = c("F" = "#8B5CF6", "M" = "#3B82F6")) +
  scale_y_continuous(limits = c(0, 70)) +
  labs(
    title = "Gender Distribution of NDE Reports",
    subtitle = paste0("NDERF scraped data (n = ", nrow(nde_experiences), ")"),
    x = NULL, y = "Percentage"
  ) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank())
```

```{r demographics-country}
ggplot(demo_country, aes(x = pct, y = fct_reorder(country, pct))) +
  geom_col(fill = "#5B4A9E", alpha = 0.85, width = 0.7) +
  geom_text(aes(label = paste0(pct, "%")), hjust = -0.15, size = 3.5, family = "source_sans") +
  scale_x_continuous(limits = c(0, max(demo_country$pct) * 1.2)) +
  labs(
    title = "Top 10 Countries of NDE Experiencers",
    subtitle = paste0("NDERF scraped data (n = ", nrow(nde_experiences), ")"),
    x = "Percentage", y = NULL
  ) +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank())
```

Women are overrepresented in NDE reports (**55% vs 44%** male), though researchers
note this likely reflects reporting behavior rather than incidence. The peak age is
31--50, but notably **15% of NDEs occur in childhood** (under 17) — significant
because children's accounts are less likely to be shaped by cultural expectations.

### The Anatomy of an NDE: Consistent Features Across Studies

```{r elements-comparison}
# Compute NDE element frequencies from scraped AI-tagged features
ai_features <- nde_experiences |>
  summarize(
    `Out-of-body experience` = mean(ai_obe, na.rm = TRUE),
    `Feeling of unity/oneness` = mean(ai_unity, na.rm = TRUE),
    `Hellish/distressing imagery` = mean(ai_hellish, na.rm = TRUE),
    `Clinical death confirmed` = mean(ai_clinical, na.rm = TRUE),
    `ESP/seeing distant events` = mean(ai_esp, na.rm = TRUE),
    `Past lives recalled` = mean(ai_past_lives, na.rm = TRUE),
    `Visions of world future` = mean(ai_world_future, na.rm = TRUE),
    `Alien/ET encounter` = mean(ai_aliens, na.rm = TRUE)
  ) |>
  pivot_longer(everything(), names_to = "feature", values_to = "proportion") |>
  mutate(percentage = round(proportion * 100, 1)) |>
  arrange(desc(percentage))

ggplot(ai_features, aes(x = percentage, y = fct_reorder(feature, percentage))) +
  geom_col(fill = "#5B4A9E", alpha = 0.85, width = 0.7) +
  geom_text(aes(label = paste0(percentage, "%")), hjust = -0.15,
            size = 3.5, family = "source_sans") +
  scale_x_continuous(limits = c(0, 100), labels = percent_format(scale = 1)) +
  labs(
    title = "AI-Detected NDE Features Across ~5,000 Accounts",
    subtitle = "Percentage of NDERF records where each feature was detected",
    x = "Percentage of Accounts",
    y = NULL,
    caption = "Source: NDERF search.nderf.org (AI-tagged metadata)"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

**Feeling of peace** is near-universal (56--93%), followed by **seeing a brilliant
light** (23--69%) and **out-of-body experience** (24--80%). The wide ranges reflect
methodology: retrospective studies report higher rates than prospective ones because
many patients have only fragmentary memories immediately after resuscitation.

### The Temporal Sequence: A Journey, Not Random Noise

```{r greyson-distribution}
# Distribution of Greyson NDE Scale scores across scraped records
greyson_data <- nde_experiences |>
  filter(!is.na(greyson_score))

ggplot(greyson_data, aes(x = greyson_score)) +
  geom_histogram(binwidth = 1, fill = "#5B4A9E", alpha = 0.85, color = "white") +
  geom_vline(xintercept = 7, linetype = "dashed", color = "#EF4444", linewidth = 1) +
  annotate("text", x = 8, y = Inf, label = "NDE threshold (≥7)", vjust = 2,
           hjust = 0, family = "source_sans", color = "#EF4444", size = 4) +
  labs(
    title = "Distribution of Greyson NDE Scale Scores",
    subtitle = paste0("n = ", nrow(greyson_data),
                      " | ", round(100 * mean(greyson_data$greyson_score >= 7), 1),
                      "% meet NDE threshold (≥7)"),
    x = "Greyson Scale Score (0–32)",
    y = "Count",
    caption = "Source: NDERF scraped metadata | Score ≥ 7 = validated NDE"
  ) +
  theme(panel.grid.minor = element_blank())
```

The out-of-body experience typically **begins** the NDE (35% report it first),
while feeling of peace typically **ends** it (34% report it last). This consistent
ordering across independent accounts argues against random hallucination — these
experiences have a narrative arc.

### Cross-Cultural Universals and Variations

The Magis Center and other researchers point out that core NDE features appear
across cultures, religions, and even in people with **no prior belief** in an
afterlife. But some elements show cultural variation.

```{r cross-cultural}
cultural_data <- nde_features |>
  filter(category == "cross_cultural") |>
  mutate(subgroup_label = if_else(subgroup == "western", "Western", "Non-Western"))

ggplot(cultural_data, aes(x = value, y = fct_reorder(feature, value, .fun = mean),
                           fill = subgroup_label)) +
  geom_col(position = "dodge", width = 0.7, alpha = 0.85) +
  scale_fill_manual(values = c("Western" = "#3B82F6", "Non-Western" = "#F59E0B")) +
  scale_x_continuous(labels = percent_format(scale = 1)) +
  labs(
    title = "NDEs Across Cultures: What's Universal, What's Not",
    subtitle = "Peace is universal; tunnels are almost exclusively Western",
    x = "Percentage Reporting Feature",
    y = NULL,
    fill = NULL,
    caption = "Source: Kellehear 2009 (systematic cross-cultural review)"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )
```

**Universal features** (peace, light, OBE) are remarkably consistent. But the
**tunnel** is almost exclusively Western (36% vs 4%), while **religious figures**
and **otherworldly realms** are more common in non-Western accounts. Notably,
cases like Vicki Noratuk (blind from birth, identified Jesus with no visual
reference) and the Indian cases documented by Pasricha & Stevenson challenge
the idea that NDEs are simply cultural wish fulfillment.

### NDEs Transform Lives — But Only for Those Who Had Them

Van Lommel's study is unique: he tracked both NDE experiencers **and** cardiac
arrest survivors who did **not** have NDEs for 8 years. This gives us a true
control group.

```{r life-changes}
life_data <- nde_features |>
  filter(study_id == "van_lommel_2001", category == "life_change") |>
  mutate(subgroup_label = if_else(subgroup == "nde_group", "Had NDE", "No NDE (Control)"))

ggplot(life_data, aes(x = value, y = fct_reorder(feature, value, .fun = max),
                       fill = subgroup_label)) +
  geom_col(position = "dodge", width = 0.7, alpha = 0.85) +
  scale_fill_manual(values = c("Had NDE" = "#5B4A9E", "No NDE (Control)" = "#9CA3AF")) +
  scale_x_continuous(labels = percent_format(scale = 1), limits = c(0, 80)) +
  labs(
    title = "NDEs Permanently Change People",
    subtitle = "8-year follow-up: <span style='color:#5B4A9E;'>**NDE group**</span> vs. <span style='color:#9CA3AF;'>**cardiac arrest controls**</span>",
    x = "% Reporting Increase",
    y = NULL,
    fill = NULL,
    caption = "Source: Van Lommel et al. 2001, 8-year follow-up (n = 37 per group)"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "none",
    plot.subtitle = element_markdown()
  )
```

The gap is striking: **73% of NDErs** became more loving/empathetic vs 41% of
controls. **47% lost their fear of death** vs 16% of controls. Both groups nearly
died — but only those who had the NDE were permanently transformed. This is
difficult to explain as a mere side effect of cardiac arrest or oxygen deprivation.

### Pleasant vs. Distressing: Not All NDEs Are the Same

```{r valence-plot}
# Hellish vs non-hellish from scraped data
hellish_pct <- round(100 * mean(nde_experiences$ai_hellish, na.rm = TRUE), 1)
valence_scraped <- tibble(
  feature = c("Non-hellish (positive/neutral)", "Hellish/distressing imagery"),
  percentage = c(100 - hellish_pct, hellish_pct)
) |>
  mutate(feature = fct_reorder(feature, percentage))

ggplot(valence_scraped, aes(x = percentage, y = feature, fill = feature)) +
  geom_col(width = 0.6, alpha = 0.85) +
  geom_text(aes(label = paste0(percentage, "%")), hjust = -0.15, size = 5, family = "source_sans") +
  scale_fill_manual(values = c(
    "Non-hellish (positive/neutral)" = "#10B981",
    "Hellish/distressing imagery" = "#EF4444"
  )) +
  scale_x_continuous(limits = c(0, 110), labels = percent_format(scale = 1)) +
  labs(
    title = "The Emotional Spectrum of Near-Death Experiences",
    subtitle = paste0("AI-detected hellish imagery in NDERF records (n = ", nrow(nde_experiences), ")"),
    x = "Percentage",
    y = NULL,
    caption = "Source: NDERF scraped metadata (AI hellish imagery tag)"
  ) +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

**85% of NDEs are profoundly positive.** But ~14% are distressing — and these are
**twice as common among suicide survivors** (21%) compared to other causes (11%).
The Magis Center interprets this pattern as consistent with the theological view
that the nature of one's experience may reflect something beyond neurochemistry.

## The Cases That Challenge Materialism

The data patterns are compelling, but individual cases carry the real punch. These
are situations where the dying person reported **specific, verifiable information**
that they had no normal way of knowing.

### The Shoe on the Ledge

In 1977 at Harborview Medical Center in Seattle, a migrant worker named Maria
suffered a cardiac arrest. After resuscitation she told her social worker, Kimberly
Clark Sharp, that she had floated outside the hospital and seen a dark blue tennis
shoe on a third-floor window ledge. She described the shoe in detail: worn over
the little toe, with a shoelace tucked under the heel.

Sharp went to look. The shoe was there — exactly as described — on a ledge not
visible from any window inside the hospital or from the ground.

### The Blind Woman Who Saw

Vicki Noratuk was born completely blind — she had never seen anything, not even
in dreams. In 1973, after a car accident, she had an NDE in which she reported
seeing her own body from above for the first time, her wedding ring, and
eventually a radiant figure she identified as Jesus. Ring and Cooper documented
her case as part of a systematic study of 31 blind NDErs, finding that **80%
reported visual perception** during their NDE, and **100% of those blind from
birth** reported seeing.

### The Van Lommel Dentures Case

In the Netherlands, a deeply comatose man was rushed into the ER after cardiac
arrest. A nurse removed his dentures during resuscitation and placed them in a
drawer of the crash cart. A week later — having been comatose the entire time —
the patient recognized the nurse and told him exactly which drawer the dentures
were in. This case was part of Van Lommel's Lancet study and is particularly
strong because it occurred within a rigorous prospective research protocol.

```{r veridical-cases-table}
nde_veridical_cases |>
  select(name, year, country, type, verification_method) |>
  knitr::kable(
    col.names = c("Case", "Year", "Country", "Type of Perception", "How Verified"),
    caption = "Notable veridical NDE cases from the research literature"
  )
```

## The AWARE II Study: Brain Activity After Flatline

The most recent large-scale study (2023) monitored cardiac arrest patients with
EEG during and after resuscitation. Key findings:

```{r aware2-summary}
aware2_data <- nde_features |>
  filter(study_id == "parnia_aware2_2023")

aware2_data |>
  select(feature, value_type, value, n_respondents, subgroup) |>
  knitr::kable(
    col.names = c("Metric", "Type", "Value", "N", "Subgroup"),
    caption = "AWARE II Study: Key findings from 567 cardiac arrest patients"
  )
```

**39% of survivors** reported some perception of awareness during CPR — even though
they showed no external signs of consciousness. **Gamma wave spikes** (associated
with higher cognitive function) were detected in some patients during the
resuscitation period, and brain activity sometimes returned as late as **60 minutes
after cardiac arrest**. This challenges the assumption that consciousness ceases
immediately when the heart stops.

## The Supernatural Argument

The [Magis Center's argument](https://magiscenter.com/nde/) rests on several
converging lines of evidence from the data we've explored:

1. **Veridical perception during verified clinical death** — People accurately
   report events they could not have witnessed (Maria's shoe, Pam Reynolds'
   surgical details, the dentures drawer).

2. **The blind seeing** — People blind from birth report visual experiences
   during NDEs. They do not see in dreams. Yet 100% reported seeing during their
   NDE in Ring & Cooper's study.

3. **Children meeting unknown deceased relatives** — Kids report meeting
   siblings who died before they were born, of whom they were never told.

4. **Cross-cultural consistency without cultural conformity** — Core features
   (peace, light, OBE) appear universally, but the specific figures encountered
   are *not* always culturally expected (Hindu patients seeing Jesus-like figures;
   atheists encountering a loving presence).

5. **Permanent transformation only in experiencers** — Van Lommel's controlled
   study shows that cardiac arrest alone doesn't produce the life changes. Only
   those who had the NDE are transformed — suggesting the experience, not the
   physiological event, is the causal factor.

6. **Brain activity after flatline** — AWARE II detected gamma waves and
   consciousness up to an hour after hearts stopped, suggesting consciousness
   may not be produced by the brain in the way neuroscience assumes.

None of these individually "proves" the supernatural. But taken together, they
form a cumulative case that consciousness can operate independently of the
physical brain — which is the central claim of every major religious tradition
about the soul.

## What's Next?

Open questions that the data raises:

- **Why do only 10--23% of people near death report NDEs?** If consciousness
  survives death universally, why don't *all* cardiac arrest survivors report it?
- **What explains the distressing subset?** Is it neurochemical, psychological,
  or — as the Magis Center suggests — spiritual?
- **Can veridical perception be captured prospectively?** The AWARE studies placed
  hidden targets in resuscitation rooms for patients to identify. So far, no
  patient has identified a hidden target — but the sample of verifiable cases
  remains very small.
- **Will the 23% prevalence hold up?** The 2025 IANDS poll is far higher than
  earlier estimates. Is this a shift in willingness to report, a broader definition
  of NDE, or a genuine change?

## Save Datasets

```{r save-datasets}
# Save the three tidy tables for TidyTuesday submission
write_csv(nde_experiences, file.path(cache_dir, "nde_experiences.csv"))
write_csv(nde_studies, file.path(cache_dir, "nde_studies.csv"))
write_csv(nde_features, file.path(cache_dir, "nde_features.csv"))
write_csv(nde_veridical_cases, file.path(cache_dir, "nde_veridical_cases.csv"))

cat("Saved 4 datasets:\n")
cat("  nde_experiences.csv:", nrow(nde_experiences), "individual NDE records\n")
cat("  nde_studies.csv:", nrow(nde_studies), "study-level metadata rows\n")
cat("  nde_features.csv:", nrow(nde_features), "aggregated feature observations\n")
cat("  nde_veridical_cases.csv:", nrow(nde_veridical_cases), "verified perception cases\n")
```

## Session Info

```{r session-info}
sessionInfo()
```
