---
title: "TidyTemplate"
date: 2025-04-02
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-04-01")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

# Create a function that scales a numeric variable between 0 and 100
scale_0_100 <- function(x) {
  (x - min(x)) / (max(x) - min(x)) * 100
}

# Apply the scale_0_100 function to all columns between hp and speed
pokemon <- tt$pokemon_df |>
  mutate(
    across(hp:speed, ~ scale_0_100(.x), .names = "{.col}_scaled"),
    score_aggressive = (
      (0.35 * pmax(attack_scaled, special_attack_scaled)) +
        (0.30 * speed_scaled) +
        (0.15 * pmax(defense_scaled, special_defense_scaled)) +
        (0.20 * hp_scaled)
    ),
    score_defensive = (
      (0.20 * pmax(attack_scaled, special_attack_scaled)) +
        (0.10 * speed_scaled) +
        (0.30 * pmax(defense_scaled, special_defense_scaled)) +
        (0.40 * hp_scaled)
    ),
    score_balanced = (
      (0.25 * pmax(attack_scaled, special_attack_scaled)) +
        (0.25 * speed_scaled) +
        (0.25 * pmax(defense_scaled, special_defense_scaled)) +
        (0.25 * hp_scaled)
    )
  ) |>
  select(-ends_with("_scaled"))
  
```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

pokemon |>
  ggplot(aes(weight, height)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    x = "Weight",
    y = "Height",
    title = "Pokemon Weight vs Height",
    caption = "Source: {pokemon}  |  Analysis: Tony Galvan (@GDataScience1)  |  #TidyTuesday"
  )

pokemon |>
  select(pokemon, hp, attack, defense, special_attack, special_defense, speed) |>
  pivot_longer(
    cols = c(hp, attack, defense, special_attack, special_defense, speed),
    names_to = "stat",
    values_to = "value"
  ) |>
  ggplot(aes(value, fill = stat)) +
  geom_histogram(show.legend = FALSE) +
  facet_wrap(~ stat, scales = "free")

pokemon |>
  count(type_1) |>
  mutate(type_1 = fct_reorder(type_1, n)) |>
  ggplot(aes(n, type_1, fill = type_1)) +
  geom_col(show.legend = FALSE)

pokemon |>
  count(type_2) |>
  mutate(type_2 = fct_reorder(type_2, n)) |>
  ggplot(aes(n, type_2, fill = type_2)) +
  geom_col(show.legend = FALSE)

pokemon |>
  count(egg_group_1) |>
  mutate(egg_group_1 = fct_reorder(egg_group_1, n)) |>
  ggplot(aes(n, egg_group_1, fill = egg_group_1)) +
  geom_col(show.legend = FALSE)

pokemon |>
  count(egg_group_2) |>
  mutate(egg_group_2 = fct_reorder(egg_group_2, n)) |>
  ggplot(aes(n, egg_group_2, fill = egg_group_2)) +
  geom_col(show.legend = FALSE)

pokemon |>
  ggplot(aes(hp, base_experience, color = factor(generation_id))) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "loess", se = FALSE)
  
```

# Modeling

Create a model to predict the base experience of a Pokemon based on the other variables in the dataset. This is based on Julia Silge's blog post "Predict #TidyTuesday giant pumpkin weights with workflowsets" - https://juliasilge.com/blog/giant-pumpkins/

```{r}

library(tidymodels)

# Split the data into training and test sets
set.seed(123)
pokemon_split <- initial_split(
  pokemon |>
    select(
      pokemon, base_experience, height, weight, type_1, egg_group_1, hp, 
      attack, defense, special_attack, special_defense, speed, generation_id
    ) |>
    mutate(generation_id = as.factor(generation_id)) |>
    na.omit(), 
  prop = 0.8, 
  strata = base_experience
)
pokemon_train <- training(pokemon_split)
pokemon_test <- testing(pokemon_split)

# Create cross validation folds
set.seed(456)
pokemon_folds <- vfold_cv(pokemon_train, strata = base_experience)
pokemon_folds

# Create 3 recipes for the model
base_rec <- recipe(base_experience ~ ., data = pokemon_train) |>
  update_role(pokemon, new_role = "ID") |>
  step_other(type_1, egg_group_1, threshold = 0.02)
  
ind_rec <- base_rec |>
  step_dummy(all_nominal_predictors())

spline_rec <- ind_rec |>
  step_bs(hp)

scale_rec <- spline_rec |>
  step_center(all_numeric_predictors()) |>
  step_scale(all_numeric_predictors()) |>
  step_zv(all_predictors())
  
# Create 3 model specs
lm_spec <- linear_reg() |>
  set_engine("lm")

rf_spec <- rand_forest() |>
  set_engine("ranger") |>
  set_mode("regression")

mars_spec <- mars() |>
  set_engine("earth") |>
  set_mode("regression")

lasso_spec <- linear_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet") |>
  set_mode("regression")

# Create a workflow set
pokemon_set <- workflow_set(
  preproc = list(
    base = base_rec,
    ind = ind_rec,
    spline = spline_rec
  ),
  models = list(
    rf = rf_spec,
    mars = mars_spec,
    lm = lm_spec
  ),
  cross = FALSE
)
pokemon_set

# Fit the models
doParallel::registerDoParallel()
set.seed(789)

pokemon_rs <- workflow_map(
  pokemon_set,
  "fit_resamples",
  resamples = pokemon_folds,
  control = control_resamples(save_pred = TRUE)
)
pokemon_rs

# Evaluate the workflow set
autoplot(pokemon_rs)

# Collect the results
collect_metrics(pokemon_rs)

# Extract the workflow and fit to the training data
final_fit <- extract_workflow(pokemon_rs, "spline_lm") |>
  fit(data = pokemon_train)

# Examine the model parameters
final_fit |>
  tidy() |>
  mutate(
    term = fct_reorder(term, estimate),
    significant = if_else(p.value < 0.05, "Significant", "Not Significant")
  ) |>
  ggplot(aes(x = estimate, y = term, color = significant)) +
  geom_point() +
  geom_errorbar(
    aes(xmin = estimate - std.error, xmax = estimate + std.error), 
    width = 0.2
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  scale_color_manual(values = c("Significant" = "blue", "Not Significant" = "gray")) +
  labs(title = "Linear Regression Model Coefficients",
       x = "Estimated Coefficient",
       y = "Term",
       color = NULL)

# Plot the predictions vs actuals
final_fit |>
  predict(pokemon_test) |>
  bind_cols(pokemon_test) |>
  ggplot(aes(x = .pred, y = base_experience)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, lty = 2, linewidth = 1, color = "red") +
  labs(
    title = "Predicted vs Actual Base Experience",
    subtitle = "Linear Regression Model",
    x = "Predicted Base Experience",
    y = "Actual Base Experience"
  )

# -----------------------------------------------------------------------------
# Use the Lasso Model
# -----------------------------------------------------------------------------
# Create a workdlow for the lasso model
lassso_workflow <- workflow() |>
  add_model(lasso_spec) |>
  add_recipe(scale_rec)

# Set up a grid of penalty values to tune (SIMPLIFIED)
penalty_grid <- grid_regular(penalty(range = c(-2, 2)),
                             levels = 5) # Try 5 different penalty values

# Tune the hyperparameters (penalty) using cross-validation
tune_results <- tune_grid(
  lassso_workflow,
  resamples = pokemon_folds,
  grid = penalty_grid,
  metrics = metric_set(rmse, rsq, mae)
)

# Select the best model based on the chosen metric (e.g., RMSE)
best_penalty <- select_best(tune_results, metric = "rmse")

# Finalize the workflow with the best penalty value
final_workflow <- finalize_workflow(lassso_workflow, best_penalty)

# Train the final Lasso model on the entire training data
final_fit <- fit(final_workflow, data = pokemon_train)

# Make predictions on the test data using the final model
lasso_predictions <- predict(final_fit, new_data = pokemon_test)

# Evaluate the final Lasso model performance
lasso_results <- pokemon_test |>
  select(base_experience) |>
  bind_cols(lasso_predictions)

metrics(lasso_results, truth = base_experience, estimate = .pred)

# Examine the coefficients of the final Lasso model
final_fit |>
  extract_fit_parsnip() |>
  tidy() |>
  filter(estimate != 0) |> # Show only non-zero coefficients (features selected by Lasso)
  mutate(term = fct_reorder(term, abs(estimate))) |>
  ggplot(aes(estimate, term, fill = estimate > 0)) +
  geom_col() +
  scale_fill_manual(values = c("FALSE" = "firebrick2", "TRUE" = "forestgreen"),
                    labels = c("Negative", "Positive")) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Lasso Coefficients at Optimal Penalty",
       subtitle = paste("Lambda = 0.1"),
       x = "Coefficient Estimate",
       y = "Term",
       fill = "Direction")


# Plot the predictions vs actuals
lasso_results |>
  ggplot(aes(x = .pred, y = base_experience)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, lty = 2, linewidth = 1, color = "red") +
  labs(
    title = "Predicted vs Actual Base Experience",
    subtitle = "Lasso Model",
    x = "Predicted Base Experience",
    y = "Actual Base Experience"
  )

# Plot the predictions vs actuals and show the images for the outliers
library(ggimage)
p <- lasso_results |>
  bind_cols(pokemon_test |> select(pokemon)) |>
  inner_join(pokemon |> select(-base_experience), by = "pokemon") |>
  ggplot(aes(x = .pred, y = base_experience)) +
  # geom_point() +
  geom_image(aes(image = url_image), size = 0.05) +
  geom_abline(slope = 1, intercept = 0, lty = 2, linewidth = 1, color = "red") +
  labs(
    title = "Predicted vs Actual Base Experience",
    subtitle = "Lasso Model with Images",
    x = "Predicted Base Experience",
    y = "Actual Base Experience"
  )

library(magick)
pokemon_logo_url <- "https://1000logos.net/wp-content/uploads/2017/05/Pokemon-logo.jpg"

pokemon_logo <- image_read(pokemon_logo_url) |> 
  image_transparent(color = "white", fuzz = 10)

g1 <- ggplot() +
  annotation_custom(
    grid::rasterGrob(pokemon_logo),
    xmin = -Inf, xmax = Inf, 
    ymin = -Inf, ymax = Inf
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = NA,
      colour = NA
    ),
    panel.background = element_rect(
      fill = NA,
      colour = NA
    )
  )

g1

library(patchwork)
g <- p +
  inset_element(
    p = g1,
    left = 0.001, right = 0.55,
    top = 1, bottom = 0.6
  )
g

```

# Create a GT table

```{r}
library(gt)

pokemon |>
  slice_max(score_aggressive, n = 10) |>
  transmute(
    agg_rank = row_number(),
    agg_pokemon = pokemon, 
    agg_url_image = url_image
  ) |> 
  bind_cols(
    pokemon |>
      slice_max(score_balanced, n = 10) |>
      transmute(
        bal_rank = row_number(),
        bal_pokemon = pokemon, 
        bal_url_image = url_image
      )
  ) |>
  bind_cols(
    pokemon |>
      slice_max(score_defensive, n = 10) |>
      transmute(
        def_rank = row_number(),
        def_pokemon = pokemon, 
        def_url_image = url_image
      )
  ) |>
  gt() |>
  tab_header(
    title = "Top 10 Pokemon by Score",
    subtitle = "Aggressive, Balanced, and Defensive"
  ) |>
  cols_label(
    agg_rank = "Aggressive Rank",
    agg_pokemon = "Aggressive Pokemon",
    agg_url_image = "Aggressive Image",
    bal_rank = "Balanced Rank",
    bal_pokemon = "Balanced Pokemon",
    bal_url_image = "Balanced Image",
    def_rank = "Defensive Rank",
    def_pokemon = "Defensive Pokemon",
    def_url_image = "Defensive Image"
  ) |>
  cols_align(
    align = "center",
    columns = everything()
  ) |>
  cols_width(
    agg_url_image ~ px(100),
    bal_url_image ~ px(100),
    def_url_image ~ px(100)
  ) |>
  text_transform(
    locations = cells_body(columns = agg_url_image),
    fn = function(x) {
      web_image(url = x, height = 50)
    }
  ) |>
  text_transform(
    locations = cells_body(columns = bal_url_image),
    fn = function(x) {
      web_image(url = x, height = 50)
    }
  ) |>
  text_transform(
    locations = cells_body(columns = def_url_image),
    fn = function(x) {
      web_image(url = x, height = 50)
    }
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_title(groups = "title")
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightgreen"),
      cell_text(weight = "bold")
    ),
    locations = cells_title(groups = "subtitle")
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightyellow"),
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightgray"),
      cell_text(weight = "bold")
    ),
    locations = cells_row_groups()
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightpink"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = everything())
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = agg_rank)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightgreen"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = bal_rank)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightyellow"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = def_rank)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = agg_pokemon)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightgreen"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = bal_pokemon)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightyellow"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = def_pokemon)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = agg_url_image)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightgreen"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = bal_url_image)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightyellow"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = def_url_image)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = agg_rank)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightgreen"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = bal_rank)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "lightyellow"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = def_rank)
  )
```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_04_01_tidy_tuesday_pokemon.png",
  device = "png", width = 6, height = 4
)
```
