---
title: "Watch Out on Approach: When and Where Birds Strike Aircraft"
author: "Tony Galvan"
date: "7/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Every year, thousands of birds collide with aircraft in the United States — a hazard that costs the aviation industry hundreds of millions of dollars and occasionally threatens passenger safety. The FAA's wildlife strike database records decades of these incidents, letting us identify when strikes are most likely, which aircraft are most vulnerable, and which airports are hotspots. If you're a nervous flyer, the approach phase should worry you most.

## Loading and Preparing the Data

We'll clean up the flight phase variable and get a summary of the dataset.

```{r}
library(tidyverse)
library(lubridate)
theme_set(theme_light())

wildlife_impacts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv") |>
  mutate(flight_phase = na_if(str_to_title(phase_of_flt), "Unknown"))

summary(wildlife_impacts)
```

## Strikes Over Time

Are wildlife strikes increasing? Let's plot the annual count to see the trend.

```{r}
wildlife_impacts |>
  count(incident_year) |>
  ggplot(aes(incident_year, n)) +
  geom_line() + 
  geom_smooth(method = "lm")
```

The upward trend is dramatic — but it likely reflects improved reporting requirements and awareness rather than an actual increase in bird populations near airports.

## Which States Have the Most Strikes?

```{r}
wildlife_impacts |>
  filter(state != "N/A") |>
  count(state, sort = TRUE)
```

## Strikes by Flight Phase

This is the key safety question: during which phase of flight are strikes most dangerous and most common?

```{r}
wildlife_impacts |>
  filter(!is.na(flight_phase)) |>
  group_by(flight_phase) |>
  summarise(n = n()) |>
  mutate(flight_phase = fct_reorder(flight_phase, n)) |>
  ggplot(aes(flight_phase, n, fill = flight_phase)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip() +
  labs(x = "Phase of flight",
       y = "# of strikes",
       title = "Flying? Watch out for birds on approach!!",
       subtitle = "There are also many strikes during take-off, climb, and landing",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FAA")
```

Approach dominates — aircraft are at low altitude, moving through bird habitat, and decelerating (giving birds less time to react). Take-off and climb are also high-risk phases.

```{r}
ggsave("outputs/wildlife_strikes.png", width = 5.5)
```

## Strikes by Aircraft Type

Do certain aircraft attract more strikes? Let's look at the top 9 most-struck aircraft models and see if the flight phase pattern differs.

```{r}
top_struck_aircraft <- wildlife_impacts |>
  filter(!is.na(flight_phase)) |>
  group_by(atype) |>
  summarise(n = n()) |>
  top_n(9, n) |>
  ungroup()

wildlife_impacts |>
  inner_join(top_struck_aircraft) |>
  filter(!is.na(flight_phase)) |>
  group_by(atype, flight_phase) |>
  summarise(n = n()) |>
  ungroup() |>
  mutate(flight_phase = fct_reorder(flight_phase, n, sum),
         atype = fct_reorder(atype, n, sum)) |>
  ggplot(aes(flight_phase, n, fill = flight_phase)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~atype, scales = "free_x") +
  coord_flip() +
  labs(x = "Phase of flight",
       y = "# of strikes",
       title = "When do strikes occur for the top struck aircraft?",
       subtitle = "Why does the MD-88 get struck so much during landing roll?",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FAA")
```

Most aircraft follow the same pattern (approach dominates), but the MD-88 stands out with an unusual number of landing roll strikes — possibly related to its engine placement or the airports where it operates.

## Strikes by Airport

Which airports are the most dangerous for bird strikes? Geography, habitat, and flight volume all play a role.

```{r}
top_struck_airports <- wildlife_impacts |>
  filter(!is.na(flight_phase)) |>
  group_by(airport_id) |>
  summarise(n = n()) |>
  top_n(9, n) |>
  ungroup()

wildlife_impacts |>
  inner_join(top_struck_airports) |>
  filter(!is.na(flight_phase)) |>
  group_by(airport_id, airport, flight_phase) |>
  summarise(n = n()) |>
  ungroup() |>
  mutate(airport_name = paste0(airport, " (", airport_id, ")"),
         flight_phase = fct_reorder(flight_phase, n, sum),
         airport_name = fct_reorder(airport_name, n, sum)) |>
  ggplot(aes(flight_phase, n, fill = flight_phase)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~airport_name, scales = "free_x") +
  coord_flip() +
  labs(x = "Phase of flight",
       y = "# of strikes",
       title = "In which phase of flight do strikes occur at airports with most strikes?",
       subtitle = "Why do so many strikes occur during landing roll and take-off run in Denver?",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: FAA")
```

Denver International Airport (DEN) shows an unusual pattern with high numbers of strikes during landing roll and take-off run — likely related to its prairie location and the wildlife habitat surrounding its runways. Dallas/Fort Worth and Chicago O'Hare also rank high, reflecting both high traffic volume and proximity to bird habitats.
