---
title: "A Decade of Doctorates: Which Fields Are Growing the Fastest?"
author: "Tony Galvan"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Every year, thousands of scholars complete the grueling journey to a PhD — but the landscape of doctoral education is shifting. Some fields are booming while others plateau. Using data on PhDs awarded by field from 2008 to 2017, we can trace which disciplines are attracting more researchers and which are holding steady. The patterns reveal something about where society is directing its intellectual energy.

## Loading the Data

```{r}
library(tidyverse)

theme_set(theme_light())

phd_field <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-19/phd_by_field.csv")
```

## First Look at the Structure

```{r}
glimpse(phd_field)
```

There are 3370 observations with 5 variables. Let's look at the first few records to understand the hierarchy.

```{r}
head(phd_field)
```

This data lists the number of PhDs awarded by Broad Field, Major Field, Field, and Year.

```{r}
summary(phd_field)
```

The observations come from the years 2008 to 2017. Interestingly, there are 278 missing values for the number of PhDs. Let's see how many observations there are per year.

## Observations Per Year

```{r}
phd_field |> 
    count(year)
```

There are 337 observations (Fields) per year. Let's look at the number of missing values by year.

## Tracking Missing Values Over Time

```{r}
phd_field |>
    filter(is.na(n_phds)) |>
    group_by(year) |>
    summarise(n = n())
```

The more missing values, the fewer different fields of PhDs were awarded that year. 2008 and 2009 had the fewest different fields of PhDs awarded. 2016 and 2017 had the most different fields of PhDs awarded.

Let's remove the observations where there are missing values for the number of PhDs awarded.

```{r}
phd_processed <- phd_field |>
    filter(!is.na(n_phds))
```

## Total PhDs Awarded by Year

Which year produced the most doctorates? Let's create a bar plot to see the trend.

```{r}
phd_processed |>
    group_by(year) |>
    summarise(total_n_phds = sum(n_phds)) |>
    mutate(year = forcats::fct_reorder(as.factor(year), total_n_phds)) |>
    ggplot(aes(year, total_n_phds)) + 
    geom_col() +
    scale_y_continuous(labels = scales::comma_format()) +
    coord_flip() + 
    labs(x = "Year",
        y = "# of PhDs awarded",
        title = "2016 had the highest number of PhDs awarded.",
        subtitle = "2010 had the lowest number of PhDs awarded.")
```

The overall trend is upward — more PhDs are being awarded each year, with 2016 marking the peak in this dataset.

## Broad Field Trends Over Time

Let's see how the major categories of doctoral study have evolved. Are all fields growing equally, or are some pulling ahead?

```{r}
phd_processed |>
    group_by(year, broad_field) |>
    summarise(total_n_phds = sum(n_phds)) |>
    ggplot(aes(year, total_n_phds, color = broad_field)) + 
    geom_line() + 
    scale_x_continuous(breaks = c(2008, 2011, 2014, 2017)) +
    scale_color_discrete(name = "Broad field") +
    labs(x = "Year", 
        y = "Total # of PhDs awarded", 
        title = "Toal number of PhDs awarded for each broad field by year",
        subtitle = "From 2008 - 2017")
```

It is a little surprising that the Engineering broad field had the lowest number of PhDs awarded.

## Zooming In: Top Major Fields

Let's look at the major fields that produce the most PhDs each year to see which specific disciplines are driving the growth.

```{r}
phd_processed |>
    group_by(year, major_field) |>
    summarise(total_n_phds = sum(n_phds)) |>
    top_n(3) |>
    ggplot(aes(year, total_n_phds, color = major_field)) + 
    geom_line() + 
    scale_x_continuous(breaks = c(2008, 2011, 2014, 2017)) +
    labs(x = "Year", 
        y = "Total # of PhDs awarded", 
        title = "Toal number of PhDs awarded by major field and year",
        subtitle = "Filtered to just look at the top 3 major fields.") 
```

The top three major fields show distinct trajectories — some growing steadily while others have plateaued, reflecting shifting demand in academia and industry for different types of expertise.
