---
title: "TidyTemplate"
date: 2024-09-26
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

library(rvest)
library(janitor)
library(httr2)

timeline_df <- read_html("https://www.imo-official.org/organizers.aspx") |>
  html_table() |>
  .[[1]] |>
  clean_names() |>
  rename(
    "all_contestant" = contestants,
    "male_contestant" = contestants_2,
    "female_contestant" = contestants_3,
    "edition" = number
  ) |>
  filter(edition != "#") |>
  mutate(
    start_date = paste0(gsub("(.*)(-)(.*)", "\\1", date),year),
    end_date = paste0(gsub("(.*)(-)(.*)", "\\3", date),year),
    across(
      c(start_date, end_date),
      ~as.Date(.x, format = "%d.%m.%Y")
    ),
    across(
      c(edition, year, countries, all_contestant, male_contestant, female_contestant),
      as.integer
    )
  ) |>
  select(-date) |>
  # only keeping records till current year
  filter(year < 2025)

# circulate through country results link and rbind tables
scrape_country <- function(year) {
  paste0("https://www.imo-official.org/year_country_r.aspx?year=", year) |>
    read_html() |>
    html_table() |>
    .[[1]] |>
    clean_names() |>
    filter(country != "Country") |>
    mutate(year = year, .before = "country") 
}

country_results_df <- map_df(timeline_df$year, scrape_country) |>
  select(
    year,
    country,
    team_size_all = team_size,
    team_size_male = team_size_2,
    team_size_female = team_size_3,
    starts_with("p"),
    awards_gold = awards,
    awards_silver = awards_2,
    awards_bronze = awards_3,
    awards_honorable_mentions = awards_4,
    leader,
    deputy_leader
  ) |> 
  mutate(
    across(
      c(team_size_all:awards_honorable_mentions),
      as.integer
    )
  )


# circulate through individual results link and rbind tables
scrape_individual <- function(year) {
  # These can time out, so we'll use httr2 to retry.
  paste0("https://www.imo-official.org/year_individual_r.aspx?year=", year) |>
    httr2::request() |>
    httr2::req_retry(max_tries = 3) |>
    httr2::req_perform() |>
    httr2::resp_body_html() |>
    html_table() |>
    .[[1]] |>
    clean_names() |>
    mutate(year = year, .before = "contestant") 
}

individual_results_df <- map_df(timeline_df$year, scrape_individual) |>
  select(
    year:p6, p7, total,
    individual_rank = number_rank,
    award
  ) |>
  mutate(
    across(
      c(year, p1:individual_rank),
      as.integer
    )
  )

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

timeline_df |>
  ggplot(aes(x = start_date, y = edition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Timeline of International Mathematical Olympiad",
    x = "Year",
    y = "Edition"
  )

country_results_df |>
  ggplot(aes(x = country, y = team_size_all)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Team Size of International Mathematical Olympiad",
    x = "Country",
    y = "Team Size"
  )

individual_results_df |>
  ggplot(aes(x = year, y = individual_rank)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Individual Rank of International Mathematical Olympiad",
    x = "Year",
    y = "Rank"
  )

library(gt)
library(gtExtras)

# Create a table of the timeline dataset and use a theme from gtExtras to make it look nice
timeline_df |>
  gt() |>
  gt_theme_espn()

# Create a plot of the country results data that highlights the ratio of females to males
country_results_df |>
  group_by(year) |>
  summarize(
    total_females = sum(team_size_female, na.rm = TRUE),
    ratio = sum(team_size_female, na.rm = TRUE) / sum(team_size_all)
  ) |>
  ggplot(aes(x = year, y = ratio)) +
  geom_point() +
  geom_smooth(
    method = "loess",
    se = FALSE
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "International Mathematical Olympiad",
    subtitle = "Ratio of Females",
    x = "",
    y = "Ratio of Females"
  )

# Create a boxplot of the ratio of female team members to total team members
country_results_df |>
  mutate(ratio = team_size_female / team_size_all) |>
  ggplot(aes(x = year, y = ratio)) +
  geom_boxplot()

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2024_09_24_tidy_tuesday_imo.png",
  device = "png"
)
```
