---
title: "Money and Mentorship: Does National Wealth Buy Smaller Class Sizes?"
author: "Tony Galvan"
date: "5/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Intuitively, wealthier countries should be able to afford more teachers per student — but does the data actually support this? Using global student-teacher ratio data from UNESCO (2012–2018) combined with GDP per capita from the World Bank, we can test whether national wealth translates into smaller class sizes. The relationship turns out to be strong, but with some fascinating outliers.

## Loading the Data

```{r}
library(tidyverse)

theme_set(theme_minimal())

student_ratio <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-07/student_teacher_ratio.csv")
```

## What Years Are Covered?

```{r}
summary(student_ratio)
```

## Enriching with GDP Data

We'll pull GDP per capita from the World Bank's World Development Indicators (WDI) API and join it to our student-teacher ratio data. This gives us the economic context needed to test our hypothesis.

```{r}
library(WDI)

gdp_per_capita <- WDI(indicator = "NY.GDP.PCAP.CD", start = 2012, end = 2018, 
                      extra = TRUE) |>
  tbl_df() |>
  transmute(year, country_code = as.character(iso3c), 
           gdp_per_capita = NY.GDP.PCAP.CD)

student_ratio_GDP <- student_ratio |>
  inner_join(gdp_per_capita, by = c("country_code", "year")) |>
  filter(!is.na(student_ratio))
```

## Distribution of Student-Teacher Ratios

Before looking at relationships, let's understand the shape of the student-teacher ratio distribution across all countries and years.

```{r}
student_ratio_GDP |>
  ggplot(aes(student_ratio)) + 
  geom_histogram() + 
  scale_x_log10()
```

The distribution is right-skewed on a log scale, with most countries clustering between 10 and 30 students per teacher, but some developing nations reaching ratios above 100.

## The Key Question: GDP vs. Student-Teacher Ratio

Here's the central visualization — a log-log scatter plot showing the relationship between national wealth and class sizes, with country labels for context.

```{r}
student_ratio_GDP |>
  ggplot(aes(gdp_per_capita, student_ratio, color = indicator)) +
  geom_point() + 
  scale_x_log10() + 
  scale_y_log10() + 
  geom_text(aes(label = paste(year, country, sep = ":")), vjust = 1, hjust = 1, check_overlap = TRUE) +
  labs(x = "GDP per capita",
       y = "Student/teacher ratio",
       title = "GDP per capita and student/teacher ratio are negatively correlated",
       subtitle = "2012 to 2018")
```

The negative correlation is unmistakable: wealthier countries consistently have lower student-teacher ratios. On a log-log scale, the relationship is roughly linear, suggesting a power-law relationship between national income and educational investment. However, the scatter is wide — some middle-income countries achieve ratios comparable to wealthy nations, while others with similar GDP lag behind, pointing to differences in policy priorities.
