---
title: "Mapping America's Wind Power: A Visual Tour of 60,000+ Turbines"
date: 2021-11-08
output: html_document
---

The United States has quietly built one of the largest wind energy infrastructures on the planet. With over 60,000 turbines scattered across the country — from the windswept plains of Texas to the ridgelines of Appalachia — wind power has become a defining feature of America's energy landscape. This week's data from the USGS gives us coordinates, capacity, and installation year for every single one. Let's map the wind revolution.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")

library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())

tt_caption <- "Source: usgs.gov  |  Analysis: Tony Galvan (@GDataScience1)"
```

## Loading the Turbine Data

We'll pull the US Wind Turbine dataset from the TidyTuesday archive. This dataset contains detailed records for every utility-scale wind turbine in the country, including geographic coordinates, turbine capacity in kilowatts, and the year each project came online.

```{r Load}

tt <- tt_load("2018-11-06")

```

## A First Look at the Data

Let's explore the structure of the dataset to understand what we're working with.

```{r Glimpse}

tt |> 
  map(glimpse)

```

## Preparing the Data for Mapping

We'll filter out any records with invalid coordinates (positive longitudes would place turbines outside the continental US) and remove entries with zero or negative capacity or missing project years.

```{r Wrangle}

us_wind <- tt$us_wind |>
  filter(xlong <= 0,
         t_cap >= 0,
         p_year > 0)

```

## How Powerful Are These Turbines?

Before mapping, let's look at the distribution of turbine capacity. Modern turbines range from small community installations to massive multi-megawatt machines. Where does the typical American turbine fall?

```{r}

# Turbine capacity - Histogram
us_wind |>
  ggplot(aes(t_cap)) +
  geom_histogram()

```

The distribution is heavily right-skewed, with most turbines clustered in the 1,000-2,000 kW range (1-2 megawatts). There's a long tail of older, smaller turbines and a growing number of newer, larger ones pushing past 3,000 kW.

## Painting the Map: Where America Generates Wind Power

Now for the main event — plotting every turbine on a map of the continental US, colored by capacity. This gives us both the geographic spread and a sense of which regions have invested in the largest, most modern turbines.

```{r Visualize}

us_wind |>
  mutate(decade = p_year %/% 10 * 10) |>
  ggplot(aes(xlong, ylat, color = t_cap)) +
  geom_point(size = 0.25, alpha = 0.25) +
  #facet_wrap(~decade) +
  coord_map() +
  theme_void() +
  scale_color_distiller(palette = "Reds") +
  labs(title = "US Wind Turbine Data",
       subtitle = "By Decade",
       color = "Turbine capacity (kW)",
       caption = tt_caption)
  
```

The map tells a striking story. Wind farms cluster heavily in the Great Plains corridor — Texas, Oklahoma, Kansas, Iowa — where flat terrain and consistent winds make turbine placement ideal. California's passes and the upper Midwest also show dense concentrations. The East Coast is notably sparse, with only scattered installations along ridgelines. The color gradient reveals that the newest, highest-capacity turbines tend to be in the central states, while older, smaller installations dot the coasts.

## Saving the Final Visualization

```{r}

# This will save your most recent plot
ggsave(
  filename = "2018_11_06_tidy_tuesday.png",
  device = "png")

```
