---
title: "The Truth Is Out There: Analyzing 80,000 UFO Sighting Reports"
author: "Tony Galvan"
date: "6/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

The National UFO Reporting Center (NUFORC) has collected over 80,000 sighting reports from around the world. Whether you're a believer or a skeptic, the data itself tells a fascinating story about human perception, reporting patterns, and the language people use to describe the unexplained. Let's explore what shapes people see, how they describe their experiences, and whether sentiment varies by location.

## Loading and Preparing the Data

We'll parse dates and create an ID column for joining with sentiment analysis results later.

```{r}
library(tidyverse)
library(lubridate)
library(tidytext)
library(sentimentr)
theme_set(theme_light())

ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv") |>
  rownames_to_column(var = "id") |>
  mutate(id = as.numeric(id),
         date = mdy(date_documented),
         year = year(date))
```

## Date Range

```{r}
summary(ufo_sightings$date)
```

## What Shapes Do People Report?

The shape of a UFO is one of the most distinctive features witnesses describe. Let's see which shapes dominate the reports.

```{r}
ufo_sightings |>
  mutate(ufo_shape = if_else(is.na(ufo_shape), "unknown", ufo_shape),
         ufo_shape = fct_lump(ufo_shape, 10, other_level = "other")) |>
  group_by(ufo_shape) |>
  summarise(n = n()) |>
  mutate(ufo_shape = fct_reorder(ufo_shape, n)) |>
  ggplot(aes(ufo_shape, n, fill = ufo_shape)) +
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  labs(x = "UFO shape",
       y = "# of sightings",
       title = "Top 10 UFO Shapes",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: NUFORC")
```

"Light" is the most commonly reported shape — which makes sense, as most sightings occur at night when witnesses can see illumination but not detailed structure. The classic "disk" shape ranks surprisingly low compared to more ambiguous descriptions.

## Word Cloud from Sighting Descriptions

What words do witnesses use most frequently when describing their encounters? Let's mine the description text for common terms.

```{r eval=FALSE}
ufo_sightings |>
  unnest_tokens(tbl = ., output = word, input = description) |>
  count(word, sort = TRUE) |>
  filter(is.na(as.numeric(word))) |>
  anti_join(get_stopwords()) |>
  filter(n > 2000) |>
  na.omit() |>
  wordcloud2::wordcloud2(shape = "cardiod")
```

## Sentiment Analysis: How Do People Feel About Their Sightings?

Are UFO reports written with fear, excitement, or neutral observation? Let's apply sentiment analysis to the descriptions and see if the emotional tone varies by country.

```{r}
ufo_sentences <- ufo_sightings |>
  pull(description) |>
  get_sentences()
```

```{r}
ufo_sentiment <- sentiment_by(ufo_sentences)
```

## Sentiment by Country

```{r}
ufo_sightings |>
  inner_join(ufo_sentiment, by = c("id" = "element_id")) |>
  group_by(country) |>
  summarise(avg_sentiment = mean(ave_sentiment)) |>
  mutate(country = fct_reorder(country, avg_sentiment)) |>
  ggplot(aes(country, avg_sentiment, fill = country)) +
  geom_col(show.legend = FALSE) +
  coord_flip()
```

All countries show slightly positive average sentiment — people tend to describe their sightings with more wonder than fear.

## Sentiment by US State

Within the United States, does sentiment vary by state? Perhaps cultural differences or local attitudes toward the unexplained influence how people write their reports.

```{r}
ufo_sightings |>
  inner_join(ufo_sentiment, by = c("id" = "element_id")) |>
  filter(country == "us") |>
  group_by(state) |>
  summarise(avg_sentiment = mean(ave_sentiment)) |>
  mutate(state = fct_reorder(state, avg_sentiment)) |>
  ggplot(aes(state, avg_sentiment, fill = state)) +
  geom_col(show.legend = FALSE) +
  coord_flip()
```

## Animated Map: US Sightings Over Time

Finally, let's visualize how UFO sightings have spread across the United States over time. The animation reveals the dramatic increase in reporting that coincides with the internet age.

```{r}
ufo_sightings |>
  filter(country == "us" & state != "hi" & state != "ak" & state != "pr") |>
  ggplot() + 
  geom_map(data = map_data("state"), 
           map = map_data("state"), 
           aes(long, lat, map_id = region, group = group),
           fill = "white", color = "black", size = 0.1) + 
  geom_point(aes(longitude, latitude), 
             size = 0.75, alpha = 0.25, color = "blue") +
  theme_void() + 
  coord_map() +
  gganimate::transition_states(year) +
  #facet_wrap(~year) + 
  labs(title = "US UFO Sightings in {closest_state}",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: NUFORC")
```

```{r, eval=FALSE, echo=FALSE}
gganimate::anim_save(filename = "outputs/2019_06_25_tidy_tuesday_ufo_sightings.gif", animation = gganimate::last_animation())
```

The explosion of sightings in the 1990s and 2000s likely reflects the rise of online reporting platforms rather than an actual increase in unexplained aerial phenomena. The internet made it easy for anyone to file a report, dramatically lowering the barrier to participation.
