---
title: "TidyTemplate"
date: 2021-03-08
output: html_output
---

# Shout out to Julia Silge ()!!
This code is inspired (pretty much copy/paste/change_input) by Julia Silge's blog post:

https://juliasilge.com/blog/animal-crossing/

https://youtu.be/whE85O1XCkg

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)

theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-03-09")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

movies <- tt$movies %>%
  filter(!is.na(plot))

movies

```


# Visualize

Using your processed dataset, create your unique visualization.

First, let's compare the variables `binary` and `clean_test` to see which one we should focus on going forward.

```{r Visualize}

movies %>%
  count(clean_test, binary) %>%
  mutate(clean_test = fct_reorder(clean_test, n)) %>%
  ggplot(aes(n, clean_test, fill = binary)) +
  geom_col() +
  labs(x = "# of movies",
       title = "The binary = PASS only when clean_test = ok")
  
```

The `binary` variable is what I will focus on.

Next, let's explore the `plot` variable.

```{r}
movies %>%
  mutate(plot_length = str_length(plot)) %>%
  ggplot(aes(plot_length, fill = binary)) +
  geom_histogram(bins = 35) +
  labs(title = "Binary vs. the length of the plot description")
```

There are some interesing spikes around 180 characters, but the distributions look similar for both values of `binary`.

Instead, let's do some text mining on the plot descriptions.

```{r}
library(tidytext)

words_per_plot <- movies %>%
  unnest_tokens(word, plot) %>%
  count(imdb_id, name = "total_words")

words_per_plot %>%
  ggplot(aes(total_words)) + 
  geom_histogram() +
  labs(title = "Distribution of total words per plot description")
```

I wonder if I could create a model to predict if a movie will pass or fail the Bechdel Test.  This is a good chance to use the `tidymodels` package!

First, I will split the data into traning and testing sets.

```{r}
library(tidymodels)

set.seed(123)

plot_split <- initial_split(movies, strata = binary)
plot_train <- training(plot_split)
plot_test <- testing(plot_split)
```

Next, I will create a recipe and use the `textrecipes` package to do some preprossing on the plot descriptions (prep the recipe).

```{r}
library(textrecipes)

plot_rec <- recipe(binary ~ plot, data = plot_train) %>% # model formula
  step_tokenize(plot) %>% # tokenize the text (split into words)
  step_stopwords(plot) %>% # remove stop words ("a," "and," "the", etc.)
  step_tokenfilter(plot, max_tokens = 100) %>% # keep top 100 most-used tokens
  step_tfidf(plot) %>% # calculate term frequency inverse document frequency
  step_normalize(all_predictors()) #center and scale

plot_prep <- prep(plot_rec) # prep/evaluate everything in the recipe

plot_prep
```

#### Declare model specification

Next, I will specify the model as a lasso regression (with GLM Net computational engine).  We will use a mixture of 1, but we will tune the penalty hyperparameter.  I will utilize a workflow to hold my preprocessing recipe together with my model specification.

```{r}
lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lasso_wf <- workflow() %>%
  add_recipe(plot_rec) %>%
  add_model(lasso_spec)

lasso_wf
```

### Hyperparameter tuning
After creating a lambda grid of values to try for the regularization penalty hyperparameter, I create a set of resampled data to fit and evaluate the different models (1 model per penalty).

```{r}
lambda_grid <- grid_regular(penalty(), levels = 40)

set.seed(234)
plot_folds <- bootstraps(plot_train, strata = binary)
plot_folds
```

Tune the model hyperparameters

```{r}
# I had errors running parallel on Win10 machine
# Un-comment the following line if you want to try
#doParallel::registerDoParallel()

set.seed(345)
lasso_grid <- tune_grid(
  lasso_wf,
  resamples = plot_folds,
  grid = lambda_grid,
  metrics = metric_set(roc_auc, ppv, npv)
)
```

```{r}
lasso_grid %>%
  collect_metrics()
```

```{r}
lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_line(size = 1.5, show.legend = FALSE) +
  facet_wrap(~.metric) +
  scale_x_log10()
```

```{r}
best_auc <- lasso_grid %>%
  select_best("roc_auc")

best_auc
```

```{r}
final_lasso <- finalize_workflow(lasso_wf, best_auc)

final_lasso
```

```{r}
library(vip)

final_lasso %>%
  fit(plot_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = best_auc$penalty) %>%
  group_by(Sign) %>%
  top_n(20, wt = abs(Importance)) %>%
  ungroup() %>%
  mutate(
    Importance = abs(Importance),
    Variable = str_remove(Variable, "tfidf_plot_"),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(Importance, Variable, fill = Sign)) + 
  geom_col(show.legend = FALSE) + 
  scale_fill_manual(values = c("#27476E","#AA7BC3")) +
  facet_wrap(~ifelse(Sign == "NEG", "Fail", "Pass"), 
             scales = "free_y") +
  labs(y = "",
       title = "Bechdel Test - Movies with Quality Female Characters",
       subtitle = "Top 20 most important words from the plot description",
       caption = "Source: FiveThirtyEight  |  Analysis: Tony Galvan (@GDataScience1")
```

```{r}
plot_final <- last_fit(final_lasso, plot_split)

plot_final %>%
  collect_metrics()
```


```{r}
plot_final %>%
  collect_predictions() %>%
  conf_mat(binary, .pred_class)
```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2021_03_09_tidy_tuesday.png",
  device = "png")

```
