---
title: "The Golden Age of TV: What IMDb Ratings Reveal About Genre and Audience"
author: "Tony Galvan"
date: "8/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

We're living in what many call the "Golden Age of Television" — but does the data back that up? With over 2,200 season-level ratings from IMDb spanning nearly three decades (1990–2018), we can explore which genres consistently earn critical praise and which ones capture the largest audiences. Let's dig into the numbers and see what patterns emerge.

## Loading the Data

We'll pull the IMDb/Economist TV ratings dataset directly from the TidyTuesday repository and take a first look at the summary statistics.

```{r}
library(tidyverse)
theme_set(theme_light())

tv_ratings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv")

summary(tv_ratings)
```

This data includes 2266 television show ratings from Jan 1990 - Oct 2018.

## How Many Unique Shows Are There?

With over 2,000 ratings in the dataset, let's figure out how many distinct shows we're actually looking at.

```{r}
# unique titleId
summary(unique(tv_ratings$titleId))

# unique title
summary(unique(tv_ratings$title))
```

There are 876 unique shows, but only 868 unique titles. Some shows must share the same title.

## Shows That Share a Title

It's surprisingly common for different TV shows to reuse the same name — reboots, international versions, or just coincidence. Let's find them.

```{r}
same_title <- tv_ratings |>
  group_by(title, titleId) |>
  summarise(avg_share = mean(share)) |>
  group_by(title) |>
  summarise(n = n()) |>
  filter(n > 1) |>
  pull(title)

tv_ratings |>
  filter(title %in% same_title) |>
  group_by(titleId, title) |>
  summarise(start_date = min(date)) |>
  arrange(title, start_date) |>
  ungroup() |>
  select(-titleId)
```

## Which Genres Get the Best Ratings?

Since many shows span multiple genres (e.g., "Crime,Drama,Mystery"), we'll split them out so each genre gets its own row. Then we can compare average ratings across all genres.

```{r}
tv_ratings_tidy <- tv_ratings |>
  mutate(genre = strsplit(genres, ",")) |>
  unnest(genre)

tv_ratings_tidy |>
  group_by(genre) |>
  summarise(avg_rating = mean(av_rating)) |>
  ungroup() |>
  mutate(genre = fct_reorder(genre, avg_rating)) |>
  ggplot(aes(genre, avg_rating, fill = genre)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "",
       y = "Average rating",
       title = "What genres have the best ratings?",
       subtitle = "Reality-TV gets the worst ratings by far",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: IMDb")
```

Reality-TV sits firmly at the bottom — perhaps unsurprising given the genre's reputation for quantity over quality. Documentary and War genres lead the pack, suggesting that niche, high-effort productions tend to earn stronger critical reception.

## Which Genres Capture the Biggest Audience Share?

Ratings measure quality, but audience share measures reach. A show can be critically acclaimed but watched by few, or widely watched but poorly rated. Let's see which genres dominate in terms of viewership share.

```{r}
tv_ratings_tidy |>
  group_by(genre) |>
  summarise(avg_share = mean(share)) |>
  ungroup() |>
  mutate(genre = fct_reorder(genre, avg_share)) |>
  ggplot(aes(genre, avg_share, fill = genre)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "",
       y = "Average share",
       title = "What genres have the biggest share",
       subtitle = "Adventure and Sci-Fi shows get a 3 share on average",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: IMDb")
```

Adventure and Sci-Fi shows command the largest average audience share — these are the big-budget spectacles that draw viewers in. The gap between "best rated" and "most watched" genres tells an interesting story about the difference between critical quality and mass appeal.

## Building a Show-Level Dataset

Let's aggregate the season-level data into a show-level summary, capturing each show's total seasons, rating range, and audience share statistics.

```{r}
tv_shows <- tv_ratings |>
  group_by(titleId, title) |>
  summarise(tot_seasons = n(),
            max_season = max(seasonNumber),
            min_date = min(date),
            max_date = max(date),
            avg_rating = mean(av_rating),
            max_rating = max(av_rating),
            min_rating = min(av_rating),
            tot_share = sum(share),
            avg_share = mean(share),
            max_share = max(share),
            min_share = min(share))
```

## The Highest-Rated Seasons of All Time

Which individual seasons earned the highest IMDb ratings? These represent the absolute peaks of television quality according to viewer scores.

```{r}
tv_ratings |>
  arrange(desc(av_rating)) |>
  head(5)
```

These top-rated seasons represent some of the most celebrated moments in television history — the kind of seasons that spark water-cooler conversations and define an era of storytelling.
