---
title: "Two Leos, 135 Years Apart: A Text Analysis of Rerum Novarum and Magnifica Humanitas"
author: "Tony Galvan"
date: "2026-06-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 10, fig.path = "outputs/"
)
```

In May 2026, Pope Leo XIV released his first encyclical, *Magnifica Humanitas*, on safeguarding the human person in the age of artificial intelligence. He deliberately signed it on May 15 — the 135th anniversary of Pope Leo XIII's landmark *Rerum Novarum* (1891), which addressed the condition of workers during the Industrial Revolution. Both documents respond to epochal technological transformations that reshape labor, power, and human dignity. Let's use data science to explore how these two "Leo" encyclicals compare across language, themes, and scriptural foundations.

This analysis was prepared for the **Golden Dome Data Tuesdays** session on June 3, 2026, where we'll discuss *Magnifica Humanitas* and its implications for data scientists, AI practitioners, and people of faith navigating this technological moment. Throughout the post, you'll find discussion questions designed to spark conversation at our session.

## Libraries and Data Loading

```{r libraries}
library(tidyverse)
library(tidytext)
library(scales)
library(showtext)
library(sysfonts)
library(ggtext)
library(patchwork)

font_add_google("Source Sans 3", "source_sans")
font_add_google("Playfair Display", "playfair")
showtext_auto()
showtext_opts(dpi = 300)

theme_set(theme_minimal(base_family = "source_sans", base_size = 14) +
            theme(plot.title.position = "plot"))
```

```{r load-data}
# Scrape encyclical texts from the Vatican website
library(rvest)
library(httr)

# Fetch Rerum Novarum (Leo XIII, 1891)
rn_url <- "https://www.vatican.va/content/leo-xiii/en/encyclicals/documents/hf_l-xiii_enc_15051891_rerum-novarum.html"
rn_response <- GET(rn_url, user_agent("Mozilla/5.0"))
rn_html <- read_html(content(rn_response, as = "text", encoding = "UTF-8"))
rn_raw <- rn_html |> html_nodes("p") |> html_text2()

# Fetch Magnifica Humanitas (Leo XIV, 2026)
mh_url <- "https://www.vatican.va/content/leo-xiv/en/encyclicals/documents/20260515-magnifica-humanitas.html"
mh_response <- GET(mh_url, user_agent("Mozilla/5.0"))
mh_html <- read_html(content(mh_response, as = "text", encoding = "UTF-8"))
mh_raw <- mh_html |> html_nodes("p") |> html_text2()

# Parse Rerum Novarum - numbered paragraphs start with "N."
# First paragraph is unnumbered, rest start with number
rn_numbered <- grep("^[0-9]+[.]", rn_raw, value = TRUE)
rn_first <- rn_raw[8]  # The opening paragraph without a number

rn_df <- tibble(
  paragraph = c(1L, as.integer(str_extract(rn_numbered, "^[0-9]+"))),
  text = c(rn_first, str_replace(rn_numbered, "^[0-9]+[.] ?", "")),
  document = "Rerum Novarum (1891)"
)

# Parse Magnifica Humanitas - numbered paragraphs
mh_numbered <- grep("^[0-9]+[.]", mh_raw, value = TRUE)

mh_df <- tibble(
  paragraph = as.integer(str_extract(mh_numbered, "^[0-9]+")),
  text = str_replace(mh_numbered, "^[0-9]+[.] ?", ""),
  document = "Magnifica Humanitas (2026)"
)

# Combine
encyclicals <- bind_rows(rn_df, mh_df)

cat("Rerum Novarum:", nrow(rn_df), "paragraphs\n")
cat("Magnifica Humanitas:", nrow(mh_df), "paragraphs\n")
```

## Exploratory Data Analysis

Let's start by profiling the basic structure of each document — word counts, sentence lengths, and vocabulary richness.

```{r eda-basic-stats}
# Word counts per paragraph
encyclicals <- encyclicals |>

mutate(
    word_count = str_count(text, "\\S+"),
    sentence_count = str_count(text, "[.!?]+"),
    avg_sentence_length = word_count / pmax(sentence_count, 1)
  )

summary_stats <- encyclicals |>
  group_by(document) |>
  summarise(
    paragraphs = n(),
    total_words = sum(word_count),
    avg_words_per_para = mean(word_count),
    median_words_per_para = median(word_count),
    avg_sentence_length = mean(avg_sentence_length),
    .groups = "drop"
  )

summary_stats |>
  knitr::kable(digits = 1, caption = "Document Overview")
```

```{r eda-word-distribution}
ggplot(encyclicals, aes(x = word_count, fill = document)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "Distribution of Paragraph Lengths",
    subtitle = "Magnifica Humanitas tends toward shorter, more digestible paragraphs",
    x = "Words per Paragraph",
    y = "Count",
    fill = NULL
  ) +
  theme(
    legend.position = "top",
    panel.grid = element_blank()
  )
```

```{r eda-sentence-length}
ggplot(encyclicals, aes(x = document, y = avg_sentence_length, fill = document)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "Average Sentence Length by Paragraph",
    subtitle = "Leo XIII's 19th-century prose features notably longer sentences",
    x = NULL,
    y = "Average Words per Sentence"
  ) +
  theme(panel.grid = element_blank())
```

## Tokenization and Word Frequency

Now let's tokenize both documents and examine the most distinctive words in each, using [TF-IDF](https://en.wikipedia.org/wiki/Tf%E2%80%93idf) — a technique that identifies words that are frequent in one document but rare in the other, revealing each encyclical's unique vocabulary.

```{r tokenize}
# Tokenize to words
words_df <- encyclicals |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  filter(!str_detect(word, "^[0-9]+$")) |>
  mutate(word = case_when(
    word == "god" ~ "God",
    word == "christ" ~ "Christ",
    word == "jesus" ~ "Jesus",
    .default = word
  ))

word_counts <- words_df |>
  count(document, word, sort = TRUE)

# Total words per document (after stop word removal)
total_per_doc <- word_counts |>
  group_by(document) |>
  summarise(total = sum(n))

word_counts <- word_counts |>
  left_join(total_per_doc, by = "document")
```

```{r tf-idf}
# Calculate TF-IDF
tfidf <- word_counts |>
  bind_tf_idf(word, document, n)

# Top 15 distinctive words per document
top_tfidf <- tfidf |>
  group_by(document) |>
  slice_max(tf_idf, n = 15) |>
  ungroup() |>
  mutate(word = reorder_within(word, tf_idf, document))

ggplot(top_tfidf, aes(x = tf_idf, y = word, fill = document)) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  facet_wrap(~document, scales = "free_y") +
  scale_y_reordered() +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "Most Distinctive Words (TF-IDF)",
    subtitle = "Words that uniquely characterize each encyclical",
    x = "TF-IDF Score",
    y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Sentiment Analysis

[Sentiment analysis](https://en.wikipedia.org/wiki/Sentiment_analysis) assigns emotional valence to words, letting us track the emotional arc of each document. We'll use the Bing lexicon, which classifies words as positive or negative, giving us a clear view of each document's tone.

```{r sentiment-bing-summary}
# Bing sentiment (positive/negative classification)
bing <- get_sentiments("bing")

sentiment_by_para <- words_df |>
  inner_join(bing, by = "word") |>
  count(document, paragraph, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |>
  mutate(
    net_sentiment = positive - negative,
    total_words = positive + negative,
    pct_positive = positive / total_words
  )

# Summary comparison
sentiment_summary_bing <- sentiment_by_para |>
  group_by(document) |>
  summarise(
    mean_net = mean(net_sentiment),
    median_net = median(net_sentiment),
    mean_pct_positive = mean(pct_positive),
    .groups = "drop"
  )

sentiment_summary_bing |>
  knitr::kable(digits = 2, caption = "Bing Sentiment Summary")
```

```{r sentiment-density}
ggplot(sentiment_by_para, aes(x = pct_positive, fill = document)) +
  geom_density(alpha = 0.6) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  scale_x_continuous(labels = percent_format()) +
  labs(
    title = "Sentiment Distribution (Bing Lexicon)",
    subtitle = "Proportion of positive sentiment words per paragraph",
    x = "% Positive Words",
    y = "Density",
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

```{r sentiment-arc}
# Sentiment arc using Bing (positive/negative)
sentiment_arc <- words_df |>
  inner_join(bing, by = "word") |>
  mutate(score = if_else(sentiment == "positive", 1, -1)) |>
  group_by(document, paragraph) |>
  summarise(net_sentiment = sum(score), .groups = "drop") |>
  group_by(document) |>
  mutate(
    para_pct = (paragraph - min(paragraph)) / (max(paragraph) - min(paragraph)),
    rolling_sentiment = zoo::rollmean(net_sentiment, k = 5, fill = NA, align = "center")
  ) |>
  ungroup()

ggplot(sentiment_arc, aes(x = para_pct, y = rolling_sentiment, color = document)) +
  geom_line(linewidth = 1.2, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                                "Magnifica Humanitas (2026)" = "#1B4F72")) +
  scale_x_continuous(labels = percent_format()) +
  labs(
    title = "Sentiment Arc Through Each Document",
    subtitle = "5-paragraph rolling average of net sentiment (positive minus negative words)",
    x = "Position in Document (%)",
    y = "Net Sentiment (rolling avg)",
    color = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

## Thematic Analysis: Shared Concerns Across 135 Years

Both encyclicals address the relationship between technological power, labor, and human dignity. Let's define thematic categories and track their prevalence.

```{r themes}
# Define thematic word lists
themes <- tribble(
  ~theme, ~words,
  "Labor & Work", "labor|work|worker|workers|workmen|employment|wages|wage|toil|industry|job|jobs|occupation",
  "Dignity & Rights", "dignity|rights|right|human|person|persons|freedom|liberty|justice",
  "Technology & Power", "technology|power|machine|machines|ai|artificial|intelligence|digital|algorithm|data|platform|robot",
  "Common Good", "common good|community|society|social|public|commonwealth|solidarity|fraternity",
  "Property & Wealth", "property|wealth|rich|poor|poverty|capital|money|profit|ownership|goods",
  "Church & Faith", "church|god|christ|faith|gospel|spirit|prayer|christian|religion|sacred"
)

# Count theme mentions per paragraph
theme_counts <- encyclicals |>
  crossing(themes) |>
  mutate(
    mentions = str_count(str_to_lower(text), words)
  ) |>
  group_by(document, theme) |>
  summarise(
    total_mentions = sum(mentions),
    paragraphs_with = sum(mentions > 0),
    .groups = "drop"
  ) |>
  group_by(document) |>
  mutate(pct_paragraphs = paragraphs_with / n_distinct(encyclicals$paragraph[encyclicals$document == document[1]])) |>
  ungroup()

ggplot(theme_counts, aes(x = reorder(theme, total_mentions),
                         y = total_mentions, fill = document)) +
  geom_col(position = "dodge", alpha = 0.85) +
  coord_flip() +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "Thematic Prevalence: Six Core Themes",
    subtitle = "Total keyword mentions across each encyclical",
    x = NULL,
    y = "Total Mentions",
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

## Two Revolutions, One Argument: Industrial vs. AI

The deepest parallel between these encyclicals isn't just thematic — it's structural. Both Popes confront a technological revolution that concentrates power in new hands, displaces workers, and challenges the Church to articulate why human dignity cannot be reduced to economic utility. Let's make this parallel explicit by mapping the analogous concerns side by side.

```{r revolution-parallels}
# Define parallel concern pairs between the two revolutions
parallels <- tribble(
  ~concern, ~rn_keywords, ~mh_keywords,
  "Power Concentration", "few rich|small number|comparatively few|enormous fortunes", "concentrated|few hands|monopol|transnational|private actors",
  "Worker Displacement", "workmen|working class|laboring poor|hired labor", "unemployment|automation|precari|displaced|job|de-skill",
  "Dehumanization", "slavery|bondsmen|instruments|grind men down", "dehumaniz|reduced to|cog|commodity|discard|expendable",
  "Call for Regulation", "law|state|public authority|remedy|intervention", "regulat|governance|transparency|accountab|oversight|control",
  "Solidarity & Unions", "associations|unions|guilds|mutual help|combination", "solidar|cooperat|intermediate bodies|civil society|subsidiarity",
  "Moral Framework", "justice|charity|religion|church|gospel|god", "dignity|common good|integral|discernment|gospel|faith"
)

# Count matches for each concern in each document
parallel_counts <- encyclicals |>
  crossing(parallels) |>
  mutate(
    rn_match = if_else(document == "Rerum Novarum (1891)",
                       str_count(str_to_lower(text), rn_keywords), 0L),
    mh_match = if_else(document == "Magnifica Humanitas (2026)",
                       str_count(str_to_lower(text), mh_keywords), 0L)
  ) |>
  group_by(concern, document) |>
  summarise(
    mentions = sum(rn_match) + sum(mh_match),
    .groups = "drop"
  ) |>
  # Normalize per 1000 words
  left_join(summary_stats |> select(document, total_words), by = "document") |>
  mutate(per_1000 = mentions / total_words * 1000)

ggplot(parallel_counts, aes(x = reorder(concern, per_1000),
                            y = per_1000, fill = document)) +
  geom_col(position = "dodge", alpha = 0.85, width = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "The Same Fight, Different Machines",
    subtitle = "Parallel concerns across the Industrial and AI Revolutions (per 1,000 words)",
    x = NULL,
    y = "Mentions per 1,000 Words",
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

The structural argument is remarkably consistent: in 1891, Leo XIII saw factory owners concentrating power over "teeming masses of the laboring poor" and called for worker associations and state intervention. In 2026, Leo XIV sees tech companies concentrating power over data, algorithms, and platforms — and calls for transparency, accountability, and democratic governance of AI. The vocabulary changes, but the moral architecture is the same:

| Industrial Revolution (1891) | AI Revolution (2026) |
|---|---|
| Factory owners & capitalists | Tech companies & platform owners |
| Physical labor exploitation | Data labor & algorithmic surveillance |
| "Yoke little better than slavery" | "New forms of slavery" in data labeling |
| Worker associations & guilds | Civil society, intermediate bodies |
| State intervention for just wages | Regulation, audits, algorithmic transparency |
| Private property with social function | Data as common good, not private asset |
| "Workmen surrendered, isolated, helpless" | Workers "de-skilled, subjected to automated surveillance" |

Both Popes reject the same two extremes: unchecked capitalism that treats people as means, and state collectivism that absorbs individual freedom. Both propose a middle path grounded in subsidiarity, solidarity, and the primacy of human dignity over economic efficiency.

## Scripture References

A key dimension of any encyclical is its grounding in Sacred Scripture. Let's parse out the biblical references from each document and compare which books of the Bible each Pope draws upon.

```{r scripture-parsing}
# Parse scripture references
# Rerum Novarum uses footnote numbers like (1), (2) with references at the end
# Magnifica Humanitas uses inline references like (cf. Gn 1:26-27)

# For Magnifica Humanitas - extract inline scripture references
mh_scripture_pattern <- "(?:cf\\.?\\s*)?(?:Gn|Ex|Lv|Nm|Dt|Jos|Jg|Rt|1\\s*S|2\\s*S|1\\s*K|2\\s*K|1\\s*Ch|2\\s*Ch|Ezr|Ne|Tb|Jdt|Est|1\\s*M|2\\s*M|Jb|Ps|Salt|Pr|Qo|Sg|Ws|Si|Is|Jr|Lm|Ba|Ez|Dn|Ho|Jl|Am|Ob|Jon|Mi|Na|Hab|Zp|Hg|Zc|Ml|Mt|Mc|Lc|Jn|Ac|Rm|1\\s*Co|2\\s*Co|Ga|Ef|Ph|Col|1\\s*Th|2\\s*Th|1\\s*Tm|2\\s*Tm|Tt|Phm|Hb|Jm|1\\s*P|2\\s*P|1\\s*Jn|2\\s*Jn|3\\s*Jn|Jd|Ap)\\s*[0-9]+"

mh_refs <- mh_df |>
  mutate(
    refs = str_extract_all(text, mh_scripture_pattern)
  ) |>
  unnest(refs, keep_empty = FALSE) |>
  mutate(
    book = str_extract(refs, "(?:1\\s*|2\\s*|3\\s*)?[A-Z][a-z]+"),
    book = str_trim(book)
  ) |>
  select(paragraph, document, refs, book)

# For Rerum Novarum - map footnote references to books
# Based on the references section of the text
rn_scripture_map <- tribble(
  ~footnote, ~book, ~refs,
  2, "Dt", "Deut. 5:21",
  3, "Gn", "Gen. 1:28",
  5, "Gn", "Gen. 3:17",
  6, "Jm", "James 5:4",
  7, "2 Tm", "2 Tim. 2:12",
  8, "2 Co", "2 Cor. 4:17",
  9, "Mt", "Matt. 19:23-24",
  10, "Lc", "Luke 6:24-25",
  14, "Lc", "Luke 11:41",
  15, "Ac", "Acts 20:35",
  16, "Mt", "Matt. 25:40",
  18, "2 Co", "2 Cor. 8:9",
  19, "Mc", "Mark 6:3",
  20, "Mt", "Matt. 5:3",
  21, "Mt", "Matt. 11:28",
  22, "Rm", "Rom. 8:17",
  23, "1 Tm", "1 Tim. 6:10",
  24, "Ac", "Acts 4:34",
  29, "Gn", "Gen. 1:28",
  30, "Rm", "Rom. 10:12",
  31, "Ex", "Exod. 20:8",
  32, "Gn", "Gen. 2:2",
  33, "Gn", "Gen. 3:19",
  34, "Qo", "Eccle. 4:9-10",
  35, "Pr", "Prov. 18:19",
  39, "Mt", "Matt. 16:26",
  40, "Mt", "Matt. 6:32-33",
  41, "1 Co", "1 Cor. 13:4-7"
)

rn_refs <- rn_scripture_map |>
  mutate(document = "Rerum Novarum (1891)")

# Combine all scripture references
all_refs <- bind_rows(
  mh_refs |> select(document, book, refs),
  rn_refs |> select(document, book, refs)
)

cat("Scripture references found:\n")
all_refs |> count(document) |> print()
```

```{r scripture-books}
# Categorize books into Old Testament vs New Testament
ot_books <- c("Gn", "Ex", "Lv", "Nm", "Dt", "Jos", "Jg", "Rt", "1 S", "2 S",
              "1 K", "2 K", "1 Ch", "2 Ch", "Ezr", "Ne", "Tb", "Jdt", "Est",
              "1 M", "2 M", "Jb", "Ps", "Salt", "Pr", "Qo", "Sg", "Ws", "Si",
              "Is", "Jr", "Lm", "Ba", "Ez", "Dn", "Ho", "Jl", "Am", "Ob",
              "Jon", "Mi", "Na", "Hab", "Zp", "Hg", "Zc", "Ml")

all_refs <- all_refs |>
  mutate(
    testament = if_else(book %in% ot_books, "Old Testament", "New Testament")
  )

# Count by book
book_counts <- all_refs |>
  count(document, book, testament, sort = TRUE)

# Top books per document
top_books <- book_counts |>
  group_by(document) |>
  slice_max(n, n = 10) |>
  ungroup() |>
  mutate(book = reorder_within(book, n, document))

ggplot(top_books, aes(x = n, y = book, fill = testament)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~document, scales = "free_y") +
  scale_y_reordered() +
  scale_fill_manual(values = c("Old Testament" = "#DAA520",
                               "New Testament" = "#4169E1")) +
  labs(
    title = "Most Cited Books of Scripture",
    subtitle = "Top biblical books referenced in each encyclical",
    x = "Number of References",
    y = NULL,
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    strip.text = element_text(face = "bold", size = 12)
  )
```

```{r testament-ratio}
# OT vs NT ratio
testament_summary <- all_refs |>
  count(document, testament) |>
  group_by(document) |>
  mutate(pct = n / sum(n)) |>
  ungroup()

ggplot(testament_summary, aes(x = document, y = pct, fill = testament)) +
  geom_col(alpha = 0.85, width = 0.6) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("Old Testament" = "#DAA520",
                               "New Testament" = "#4169E1")) +
  labs(
    title = "Old Testament vs. New Testament References",
    subtitle = "Proportion of scriptural citations from each testament",
    x = NULL,
    y = "Proportion of References",
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

## Bigram Analysis: Phrases That Define Each Era

[Bigrams](https://en.wikipedia.org/wiki/Bigram) — pairs of consecutive words — reveal the characteristic phrases of each document. They capture concepts that single words miss, like "common good" or "artificial intelligence."

```{r bigrams}
bigrams_df <- encyclicals |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word,
    !str_detect(word1, "^[0-9]+$"),
    !str_detect(word2, "^[0-9]+$")
  ) |>
  mutate(
    word1 = case_when(word1 == "god" ~ "God", word1 == "christ" ~ "Christ",
                      word1 == "jesus" ~ "Jesus", .default = word1),
    word2 = case_when(word2 == "god" ~ "God", word2 == "christ" ~ "Christ",
                      word2 == "jesus" ~ "Jesus", .default = word2)
  ) |>
  unite(bigram, word1, word2, sep = " ")

bigram_counts <- bigrams_df |>
  count(document, bigram, sort = TRUE)

bigram_tfidf <- bigram_counts |>
  bind_tf_idf(bigram, document, n) |>
  group_by(document) |>
  slice_max(tf_idf, n = 12) |>
  ungroup() |>
  mutate(bigram = reorder_within(bigram, tf_idf, document))

ggplot(bigram_tfidf, aes(x = tf_idf, y = bigram, fill = document)) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  facet_wrap(~document, scales = "free_y") +
  scale_y_reordered() +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "Most Distinctive Bigrams (TF-IDF)",
    subtitle = "Two-word phrases that uniquely characterize each encyclical",
    x = "TF-IDF Score",
    y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Topic Modeling: Unsupervised Theme Discovery

[Latent Dirichlet Allocation (LDA)](https://en.wikipedia.org/wiki/Latent_Dirichlet_allocation) is an unsupervised machine learning technique that discovers hidden thematic clusters in text without us pre-defining categories. Unlike our manual keyword approach above, LDA lets the algorithm find topics organically — revealing structure we might not have anticipated.

```{r lda-setup}
library(topicmodels)
library(tm)

# Create a document-term matrix where each "document" is a paragraph
dtm_data <- words_df |>
  count(document, paragraph, word) |>
  unite(doc_id, document, paragraph, sep = "_para_") |>
  cast_dtm(doc_id, word, n)

# Remove sparse terms to focus on meaningful patterns
dtm_data <- removeSparseTerms(dtm_data, 0.99)
# Remove empty rows
row_totals <- apply(dtm_data, 1, sum)
dtm_data <- dtm_data[row_totals > 0, ]
```

```{r lda-fit}
# Fit LDA with 6 topics
set.seed(42)
lda_model <- LDA(dtm_data, k = 6, control = list(seed = 42))

# Extract top terms per topic
lda_topics <- tidy(lda_model, matrix = "beta") |>
  group_by(topic) |>
  slice_max(beta, n = 8) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic))

ggplot(lda_topics, aes(x = beta, y = term, fill = factor(topic))) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  facet_wrap(~topic, scales = "free_y", ncol = 2,
             labeller = labeller(topic = function(x) paste("Topic", x))) +
  scale_y_reordered() +
  scale_fill_viridis_d(option = "D") +
  labs(
    title = "LDA Topic Model: 6 Discovered Themes",
    subtitle = "Top 8 words per topic (unsupervised clustering)",
    x = "Word Probability (Beta)",
    y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  )
```

```{r lda-document-topics}
# Which topics dominate each encyclical?
lda_gamma <- tidy(lda_model, matrix = "gamma") |>
  separate(document, into = c("source_doc", "paragraph"), sep = "_para_") |>
  mutate(paragraph = as.integer(paragraph))

topic_by_doc <- lda_gamma |>
  group_by(source_doc, topic) |>
  summarise(avg_gamma = mean(gamma), .groups = "drop")

# Show the DIFFERENCE in topic prevalence — which topics skew toward which document
topic_diff <- topic_by_doc |>
  pivot_wider(names_from = source_doc, values_from = avg_gamma) |>
  mutate(
    diff = `Magnifica Humanitas (2026)` - `Rerum Novarum (1891)`,
    direction = if_else(diff > 0, "More in Magnifica Humanitas", "More in Rerum Novarum")
  )

# Get top 3 words per topic for labeling
topic_labels <- tidy(lda_model, matrix = "beta") |>
  group_by(topic) |>
  slice_max(beta, n = 3) |>
  summarise(label = paste(term, collapse = ", "), .groups = "drop")

topic_diff <- topic_diff |>
  left_join(topic_labels, by = "topic") |>
  mutate(topic_label = paste0("Topic ", topic, ": ", label))

ggplot(topic_diff, aes(x = reorder(topic_label, diff), y = diff, fill = direction)) +
  geom_col(alpha = 0.85, width = 0.7) +
  geom_hline(yintercept = 0, linewidth = 0.5) +
  coord_flip() +
  scale_fill_manual(values = c("More in Magnifica Humanitas" = "#1B4F72",
                               "More in Rerum Novarum" = "#8B4513")) +
  labs(
    title = "Topic Skew: Which Encyclical Owns Each Theme?",
    subtitle = "Difference in average topic probability (positive = more in Magnifica Humanitas)",
    x = NULL,
    y = "Difference in Topic Probability",
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

The diverging bar chart reveals which discovered topics are distinctive to each era. Topics that skew toward *Rerum Novarum* reflect 19th-century concerns (property, labor conditions, class relations), while those skewing toward *Magnifica Humanitas* capture the digital-age vocabulary (technology, data, platforms, governance).

## Cosine Similarity: Textual Kinship Across 135 Years

[Cosine similarity](https://en.wikipedia.org/wiki/Cosine_similarity) measures how similar two text segments are based on their word usage patterns. A score of 1.0 means identical vocabulary; 0.0 means completely different. By computing similarity between sections of each encyclical, we can identify where Leo XIV is most directly building on Leo XIII's arguments.

```{r cosine-setup}
library(lsa)

# Group paragraphs into sections of ~5 paragraphs for smoother comparison
rn_sections <- rn_df |>
  mutate(section = (paragraph - 1) %/% 5 + 1) |>
  group_by(section) |>
  summarise(
    text = paste(text, collapse = " "),
    doc = "Rerum Novarum",
    .groups = "drop"
  ) |>
  mutate(section_label = paste0("RN §", (section - 1) * 5 + 1, "-", section * 5))

mh_sections <- mh_df |>
  mutate(section = (paragraph - 1) %/% 10 + 1) |>
  group_by(section) |>
  summarise(
    text = paste(text, collapse = " "),
    doc = "Magnifica Humanitas",
    .groups = "drop"
  ) |>
  mutate(section_label = paste0("MH §", (section - 1) * 10 + 1, "-", section * 10))

# Combine and create TF-IDF matrix
all_sections <- bind_rows(
  rn_sections |> select(section_label, text, doc),
  mh_sections |> select(section_label, text, doc)
)

# Tokenize and create TF-IDF document-term matrix
section_tfidf <- all_sections |>
  mutate(id = row_number()) |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  filter(!str_detect(word, "^[0-9]+$")) |>
  mutate(word = case_when(
    word == "god" ~ "God",
    word == "christ" ~ "Christ",
    word == "jesus" ~ "Jesus",
    .default = word
  )) |>
  count(id, word) |>
  bind_tf_idf(word, id, n) |>
  cast_dtm(id, word, tf_idf)

# Convert to matrix and compute cosine similarity
tfidf_matrix <- as.matrix(section_tfidf)
cos_sim <- cosine(t(tfidf_matrix))
```

```{r cosine-heatmap, fig.width=8, fig.height=8}
# Extract cross-document similarities (RN sections vs MH sections)
n_rn <- nrow(rn_sections)
n_mh <- nrow(mh_sections)

cross_sim <- cos_sim[1:n_rn, (n_rn + 1):(n_rn + n_mh)]
rownames(cross_sim) <- rn_sections$section_label
colnames(cross_sim) <- mh_sections$section_label

# Convert to long format for ggplot
sim_long <- as.data.frame(as.table(cross_sim)) |>
  as_tibble() |>
  rename(rn_section = Var1, mh_section = Var2, similarity = Freq) |>
  mutate(
    rn_section = factor(rn_section, levels = rn_sections$section_label),
    mh_section = factor(mh_section, levels = rev(mh_sections$section_label))
  )

ggplot(sim_long, aes(x = rn_section, y = mh_section, fill = similarity)) +
  geom_tile() +
  scale_fill_viridis_c(option = "inferno", labels = scales::number_format(accuracy = 0.01)) +
  labs(
    title = "Cosine Similarity: Where Do the Two Encyclicals Converge?",
    subtitle = "Brighter cells = higher textual similarity between sections",
    x = "Rerum Novarum Sections",
    y = "Magnifica Humanitas Sections",
    fill = "Similarity"
  ) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(size = 7)
  )
```

```{r cosine-top-pairs}
# Find the most similar section pairs
top_pairs <- sim_long |>
  arrange(desc(similarity)) |>
  slice_head(n = 10)

top_pairs |>
  knitr::kable(digits = 3, caption = "Top 10 Most Similar Section Pairs")
```

The heatmap reveals which passages of *Magnifica Humanitas* are most textually aligned with *Rerum Novarum*. Bright spots indicate where Leo XIV is drawing most directly on Leo XIII's language and arguments — the intellectual DNA connecting these two documents across 135 years.

## Lexical Diversity and Readability

[Lexical diversity](https://en.wikipedia.org/wiki/Lexical_diversity) measures how varied a text's vocabulary is — a higher ratio of unique words to total words suggests richer, more varied language. We'll also look at average word length as a proxy for complexity.

```{r lexical-diversity}
lexical_stats <- words_df |>
  group_by(document) |>
  summarise(
    total_tokens = n(),
    unique_types = n_distinct(word),
    type_token_ratio = unique_types / total_tokens,
    avg_word_length = mean(nchar(word)),
    .groups = "drop"
  )

lexical_stats |>
  knitr::kable(digits = 3, caption = "Lexical Diversity Metrics")
```

```{r vocabulary-growth}
# Vocabulary growth curve (Heaps' law)
vocab_growth <- words_df |>
  group_by(document) |>
  mutate(token_position = row_number()) |>
  ungroup() |>
  group_by(document) |>
  mutate(
    cumulative_types = cummax(match(word, unique(word)))
  ) |>
  ungroup()

# Sample at intervals for plotting
sample_points <- vocab_growth |>
  group_by(document) |>
  filter(token_position %% 50 == 0 | token_position == max(token_position)) |>
  ungroup()

ggplot(sample_points, aes(x = token_position, y = cumulative_types, color = document)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                                "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    title = "Vocabulary Growth (Heaps' Law)",
    subtitle = "How quickly new unique words appear as the text progresses",
    x = "Token Position",
    y = "Cumulative Unique Words",
    color = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

## Machine Learning: Can a Model Tell the Popes Apart?

So far we've used descriptive techniques to compare the documents. Now let's flip the question: can a machine learning model reliably predict which encyclical a paragraph belongs to, based solely on its text features? If so, what features does it rely on most? This approach lets the *data* tell us what truly distinguishes the two voices, rather than us choosing what to look for.

We'll fit two models using the [tidymodels](https://www.tidymodels.org/) framework:
1. **Lasso logistic regression** — a [penalized regression](https://en.wikipedia.org/wiki/Lasso_(statistics)) that shrinks most word coefficients to zero, revealing the minimal set of words that best separate the documents. Words with non-zero coefficients are the "signature" vocabulary of each Pope.
2. **Random forest** — an [ensemble method](https://en.wikipedia.org/wiki/Random_forest) that captures non-linear patterns and interactions. We'll compare its accuracy to the lasso as a baseline.

```{r ml-setup}
library(tidymodels)
library(textrecipes)
library(vip)

# Create the modeling dataset with engineered text features
ml_data <- encyclicals |>
  select(document, text, paragraph) |>
  mutate(
    label = factor(if_else(document == "Magnifica Humanitas (2026)",
                           "Magnifica Humanitas", "Rerum Novarum")),
    # Structural features
    word_count = str_count(text, "\\S+"),
    sentence_count = str_count(text, "[.!?]+"),
    avg_sentence_length = word_count / pmax(sentence_count, 1),
    avg_word_length = nchar(str_remove_all(text, "\\s+")) /
      pmax(str_count(text, "\\S+"), 1),
    question_marks = str_count(text, "\\?"),
    # Vocabulary richness (type-token ratio per paragraph)
    ttr = map_dbl(text, function(t) {
      words <- str_extract_all(str_to_lower(t), "[a-z]+")[[1]]
      if (length(words) == 0) return(0)
      n_distinct(words) / length(words)
    }),
    # Scripture reference density
    scripture_refs = str_count(text, "(?:cf[.]|Gn|Mt|Lc|Jn|Rm|Co|Ef|Ap|Ps|Is|Ac|Ne|Pr|Ex|Dt)\\s*[0-9]"),
    # Sentiment (using bing lexicon inline)
    sentiment_score = map_dbl(text, function(t) {
      words <- str_extract_all(str_to_lower(t), "[a-z]+")[[1]]
      pos <- sum(words %in% (bing |> filter(sentiment == "positive") |> pull(word)))
      neg <- sum(words %in% (bing |> filter(sentiment == "negative") |> pull(word)))
      (pos - neg) / pmax(length(words), 1)
    })
  )

# Split: stratified by label
set.seed(42)
ml_split <- initial_split(ml_data, prop = 0.75, strata = label)
ml_train <- training(ml_split)
ml_test <- testing(ml_split)

cat("Training set:", nrow(ml_train), "paragraphs\n")
cat("Test set:", nrow(ml_test), "paragraphs\n")
cat("Class balance (train):\n")
ml_train |> count(label) |> print()
```

```{r ml-recipe}
# Recipe: TF-IDF word features + engineered numeric features
ml_recipe <- recipe(label ~ text + word_count + avg_sentence_length +
                      avg_word_length + question_marks + ttr +
                      scripture_refs + sentiment_score,
                    data = ml_train) |>
  step_tokenize(text) |>
  step_stopwords(text) |>
  step_tokenfilter(text, max_tokens = 500) |>
  step_tfidf(text) |>
  step_normalize(all_numeric_predictors())
```

```{r ml-lasso}
# Lasso logistic regression
lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet") |>
  set_mode("classification")

lasso_wf <- workflow() |>
  add_recipe(ml_recipe) |>
  add_model(lasso_spec)

# Cross-validation to tune penalty
set.seed(42)
cv_folds <- vfold_cv(ml_train, v = 5, strata = label)

lasso_grid <- grid_regular(penalty(range = c(-5, -1)), levels = 30)

lasso_tune <- tune_grid(
  lasso_wf,
  resamples = cv_folds,
  grid = lasso_grid,
  metrics = metric_set(accuracy, roc_auc)
)

# Best penalty
best_lasso <- select_best(lasso_tune, metric = "accuracy")
cat("Best lasso penalty:", best_lasso$penalty, "\n")

# Show tuning results
autoplot(lasso_tune) +
  labs(title = "Lasso Tuning: Accuracy and ROC AUC vs. Penalty") +
  theme(panel.grid = element_blank())
```

```{r ml-lasso-final}
# Fit final lasso model
final_lasso_wf <- finalize_workflow(lasso_wf, best_lasso)
final_lasso_fit <- fit(final_lasso_wf, data = ml_train)

# Test set predictions
lasso_preds <- predict(final_lasso_fit, ml_test) |>
  bind_cols(ml_test |> select(label))

lasso_acc <- accuracy(lasso_preds, truth = label, estimate = .pred_class)
cat("Lasso test accuracy:", lasso_acc$.estimate, "\n")
```

```{r ml-lasso-coefs}
# Extract non-zero lasso coefficients
lasso_coefs <- final_lasso_fit |>
  extract_fit_parsnip() |>
  tidy() |>
  filter(term != "(Intercept)", estimate != 0) |>
  mutate(
    direction = if_else(estimate > 0, "Predicts Magnifica Humanitas",
                        "Predicts Rerum Novarum"),
    term = str_remove(term, "tfidf_text_")
  ) |>
  arrange(desc(abs(estimate)))

cat("Non-zero coefficients:", nrow(lasso_coefs), "\n")

# Categorize: vocabulary vs. writing style
style_features <- c("word_count", "avg_sentence_length", "avg_word_length",
                    "question_marks", "ttr", "scripture_refs", "sentiment_score")

lasso_coefs <- lasso_coefs |>
  mutate(
    feature_type = if_else(term %in% style_features, "Style", "Vocabulary"),
    abs_estimate = abs(estimate)
  )

# Summary: style vs. vocabulary contribution
type_summary <- lasso_coefs |>
  group_by(feature_type) |>
  summarise(
    n_features = n(),
    total_weight = sum(abs_estimate),
    .groups = "drop"
  ) |>
  mutate(pct_weight = total_weight / sum(total_weight))

cat("\nFeature importance by type:\n")
type_summary |> print()
```

```{r ml-lasso-coefficients, fig.height=10}
# Build a single plot: top features from each direction, with shape indicating type
top_mh <- lasso_coefs |> filter(estimate > 0) |> slice_max(estimate, n = 12)
top_rn <- lasso_coefs |> filter(estimate < 0) |> slice_max(abs(estimate), n = 12)

top_coefs <- bind_rows(top_mh, top_rn) |>
  mutate(
    term_label = fct_reorder(term, estimate)
  )

# Style feature annotations positioned on the opposite side of the bar
style_annotations <- top_coefs |>
  filter(feature_type == "Style") |>
  mutate(
    label_x = if_else(estimate > 0,
                      min(top_coefs$estimate) * 0.15,
                      max(top_coefs$estimate) * 0.15)
  )

ggplot(top_coefs, aes(x = estimate, y = term_label, fill = direction)) +
  geom_col(aes(alpha = feature_type)) +
  geom_vline(xintercept = 0, linewidth = 0.5) +
  geom_text(data = style_annotations,
            aes(x = label_x, y = term_label, label = "STYLE"),
            size = 2.5, fontface = "bold", color = "gray40",
            family = "source_sans", inherit.aes = FALSE) +
  scale_fill_manual(values = c("Predicts Magnifica Humanitas" = "#1B4F72",
                               "Predicts Rerum Novarum" = "#8B4513")) +
  scale_alpha_manual(values = c("Style" = 1, "Vocabulary" = 0.7),
                     guide = "none") +
  labs(
    title = "Lasso Regression: What Separates the Two Popes?",
    subtitle = paste0(
      "Top 12 features per direction (STYLE tags mark writing style features)\n",
      "Style contributes ",
      scales::percent(type_summary$pct_weight[type_summary$feature_type == "Style"], accuracy = 1),
      " of predictive weight; vocabulary contributes ",
      scales::percent(type_summary$pct_weight[type_summary$feature_type == "Vocabulary"], accuracy = 1)
    ),
    x = "Lasso Coefficient (magnitude = predictive strength)",
    y = NULL,
    fill = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

The star (★) marks writing style features — things like sentence length, vocabulary richness, and word complexity — as opposed to specific vocabulary terms. The percentage split tells us whether the model distinguishes the Popes more by *how* they write or *what* they write about. Style features like `avg_sentence_length` predicting Rerum Novarum confirm what any reader notices: Leo XIII's 19th-century prose uses dramatically longer sentences than Leo XIV's modern style.

```{r ml-random-forest}
# Random forest for comparison
rf_spec <- rand_forest(trees = 500) |>
  set_engine("ranger", importance = "impurity") |>
  set_mode("classification")

rf_wf <- workflow() |>
  add_recipe(ml_recipe) |>
  add_model(rf_spec)

rf_fit <- fit(rf_wf, data = ml_train)

# Test predictions
rf_preds <- predict(rf_fit, ml_test) |>
  bind_cols(ml_test |> select(label))

rf_acc <- accuracy(rf_preds, truth = label, estimate = .pred_class)
cat("Random forest test accuracy:", rf_acc$.estimate, "\n")

# Compare models
model_comparison <- tibble(
  model = c("Lasso Logistic Regression", "Random Forest"),
  accuracy = c(lasso_acc$.estimate, rf_acc$.estimate)
)
model_comparison |> knitr::kable(digits = 3, caption = "Model Accuracy Comparison")
```

```{r ml-rf-importance}
# Random forest variable importance (top 20)
rf_fit |>
  extract_fit_parsnip() |>
  vip(num_features = 20, aesthetics = list(fill = "#2E4057", alpha = 0.85)) +
  labs(
    title = "Random Forest: Top 20 Most Important Features",
    subtitle = "Variable importance (impurity-based) — which words help classification most?",
    x = "Importance",
    y = NULL
  ) +
  theme(panel.grid = element_blank())
```

```{r ml-misclassified}
# Find misclassified paragraphs — where one Pope sounds like the other
lasso_full_preds <- predict(final_lasso_fit, ml_data) |>
  bind_cols(ml_data |> select(label, text, document, paragraph))

misclassified <- lasso_full_preds |>
  filter(.pred_class != label)

cat("Misclassified paragraphs:", nrow(misclassified), "out of", nrow(ml_data), "\n\n")
```

```{r ml-misclassified-detail}
# For each misclassified paragraph, identify which lasso features fired
# by checking which non-zero coefficient words appear in the text
if (nrow(misclassified) > 0) {
  misclass_features <- misclassified |>
    mutate(
      text_lower = str_to_lower(text),
      # Find which predictive words from the lasso appear in this paragraph
      mh_words_present = map_chr(text_lower, function(txt) {
        mh_terms <- lasso_coefs |> filter(estimate > 0) |> pull(term)
        found <- mh_terms[str_detect(txt, paste0("\\b", mh_terms, "\\b"))]
        paste(head(found, 5), collapse = ", ")
      }),
      rn_words_present = map_chr(text_lower, function(txt) {
        rn_terms <- lasso_coefs |> filter(estimate < 0) |> pull(term)
        found <- rn_terms[str_detect(txt, paste0("\\b", rn_terms, "\\b"))]
        paste(head(found, 5), collapse = ", ")
      }),
      word_count = str_count(text, "\\S+")
    ) |>
    select(document, paragraph, .pred_class, word_count,
           mh_words_present, rn_words_present)

  misclass_features |>
    knitr::kable(
      col.names = c("Actual Source", "§", "Model Predicted",
                    "Words", "MH-predictive words found", "RN-predictive words found"),
      caption = "Misclassified paragraphs and the lasso features that confused the model"
    )
}
```

**Why the model got confused — a paragraph-by-paragraph explanation:**

- **RN §14** — Leo XIII discusses state intervention into family life and the limits of government power, using language about "citizens," "rights," and "public authority" that the lasso associates with MH's governance chapters. The paragraph even references "natural justice" and social structures — vocabulary more typical of 2026 than 1891.

- **RN §48** — Describes mutual aid societies, benevolent foundations, and organizations for workers and their families. The lasso picks up on organizational/institutional language ("associations," "societies," "institutions") that dominates MH's extensive sections on civil society and intermediate bodies.

- **RN §57** — Focuses on the moral formation and religious duties of workers' associations. Its emphasis on "religion," "God," and "morality" as guiding principles for social organization mirrors MH's repeated insistence that social structures must serve the person — not the technocratic paradigm.

- **RN §58** — Discusses internal governance of organizations: offices, charges, prudence, disputes, and fair rules. This reads like MH's chapters on AI governance, accountability, and institutional design — the vocabulary of organizing and structuring collective life.

- **MH §8** — The Nehemiah passage about rebuilding Jerusalem's walls. Leo XIV uses strikingly concrete, physical language — "families," "artisans," "stones," "walls," "work" — that echoes Rerum Novarum's industrial-era vocabulary about laborers and material conditions. It's the one moment where MH reaches back to Old Testament imagery and sounds most like a 19th-century social document.

The misclassified paragraphs are especially revealing — these are passages where Leo XIV sounds like Leo XIII (perhaps when echoing traditional social teaching language about property and labor rights) or where Leo XIII sounds surprisingly modern (perhaps when discussing universal principles that transcend his era). They represent the moments of deepest continuity across 135 years and could spark rich discussion about what endures in Catholic Social Teaching.

## Shared Vocabulary: The Common Language of Catholic Social Teaching

Despite 135 years of separation, both encyclicals share a core vocabulary rooted in Catholic Social Teaching. Let's identify the words that appear prominently in both documents.

We rank shared words by their [geometric mean](https://en.wikipedia.org/wiki/Geometric_mean) frequency — the square root of (frequency in RN × frequency in MH). Why geometric mean instead of a simple average? Because it rewards words that are *balanced* across both documents. A word that appears 5 times per 1,000 in both encyclicals scores higher than one that appears 10 times in one but only 1 time in the other. This surfaces the true "shared language" — words that are genuinely central to *both* documents, not just common in one.

```{r shared-words}
# Words appearing in both documents — use relative frequency (per 1,000 words)
# so we're comparing proportional usage, not raw counts biased by document length
doc_totals <- words_df |> count(document, name = "doc_total")

shared_words <- words_df |>
  count(document, word) |>
  left_join(doc_totals, by = "document") |>
  mutate(freq_per_1k = n / doc_total * 1000) |>
  select(document, word, freq_per_1k) |>
  pivot_wider(names_from = document, values_from = freq_per_1k, values_fill = 0) |>
  rename(rn = `Rerum Novarum (1891)`, mh = `Magnifica Humanitas (2026)`) |>
  filter(rn >= 0.3, mh >= 0.3) |>
  mutate(geometric_mean = sqrt(rn * mh)) |>
  arrange(desc(geometric_mean))

# Plot shared vocabulary
top_shared <- shared_words |>
  slice_max(geometric_mean, n = 20)

ggplot(top_shared, aes(x = rn, y = mh)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  geom_point(aes(size = geometric_mean), alpha = 0.7, color = "#2E4057") +
  geom_text(aes(label = word), vjust = -0.8, size = 3.5, family = "source_sans") +
  geom_label(
    data = tibble(rn = max(top_shared$rn) * 0.9, mh = max(top_shared$rn) * 0.75,
                  label = "Equal usage\nin both"),
    aes(x = rn, y = mh, label = label),
    size = 3, color = "gray50", fontface = "italic", family = "source_sans",
    fill = "white", label.size = 0, hjust = 0.5
  ) +
  labs(
    title = "Shared Vocabulary Between the Two Encyclicals",
    subtitle = "Relative frequency (per 1,000 words) — words prominent in both documents",
    x = "Frequency per 1,000 words (Rerum Novarum)",
    y = "Frequency per 1,000 words (Magnifica Humanitas)",
    size = "Shared\nImportance"
  ) +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
  )
```

## Final Visualization: The Two Leos Compared

```{r final-viz, fig.width=8, fig.height=10}
# Register Font Awesome
font_add(family = "fa-brands",
         regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid",
         regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")

# Build caption
bg_color <- "white"
tt_source <- "Vatican.va"
tt_caption <- paste0(
 "DataViz: Tony Galvan",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-solid;'>&#xf0ce;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 tt_source,
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-brands;'>&#xf08c;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "anthony-raul-galvan",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-brands;'>&#xf09b;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "gdatascience"
)

# Create a combined thematic comparison plot
# Normalize theme counts per 1000 words for fair comparison
theme_normalized <- encyclicals |>
  crossing(themes) |>
  mutate(mentions = str_count(str_to_lower(text), words)) |>
  group_by(document, theme) |>
  summarise(total_mentions = sum(mentions), .groups = "drop") |>
  left_join(summary_stats |> select(document, total_words), by = "document") |>
  mutate(per_1000 = total_mentions / total_words * 1000)

p1 <- ggplot(theme_normalized, aes(x = reorder(theme, per_1000),
                                    y = per_1000, fill = document)) +
  geom_col(position = "dodge", alpha = 0.85, width = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                               "Magnifica Humanitas (2026)" = "#1B4F72")) +
  labs(
    subtitle = "Theme density (mentions per 1,000 words)",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme(
    text = element_text(family = "source_sans"),
    panel.grid = element_blank(),
    legend.position = "none",
    plot.subtitle = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 11)
  )

# Sentiment arc comparison
p2 <- ggplot(sentiment_arc |> filter(!is.na(rolling_sentiment)),
             aes(x = para_pct, y = rolling_sentiment, color = document)) +
  geom_line(linewidth = 1.2, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values = c("Rerum Novarum (1891)" = "#8B4513",
                                "Magnifica Humanitas (2026)" = "#1B4F72")) +
  scale_x_continuous(labels = percent_format(), expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) +
  labs(
    subtitle = "Sentiment arc (5-paragraph rolling average, Bing lexicon)",
    x = "Position in Document", y = "Net Sentiment", color = NULL
  ) +
  theme(
    text = element_text(family = "source_sans"),
    panel.grid = element_blank(),
    legend.position = "none",
    plot.subtitle = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 11)
  )

# Colored subtitle as legend using ggtext
plot_subtitle <- paste0(
  "Comparing <b style='color:#8B4513;'>Rerum Novarum (1891)</b> on the Industrial Revolution<br>",
  "with <b style='color:#1B4F72;'>Magnifica Humanitas (2026)</b> on Artificial Intelligence"
)

# Combine: p1 on top, p2 below
# Use free() on p2 so its y-axis isn't forced to align with p1's wide labels
final_plot <- (p1 / free(p2)) +
  plot_layout(heights = c(1, 0.8)) +
  plot_annotation(
    title = "Two Leos, 135 Years Apart",
    subtitle = plot_subtitle,
    caption = tt_caption,
    theme = theme(
      plot.title = element_text(family = "playfair", size = 35, face = "bold",
                                hjust = 0.5),
      plot.subtitle = element_markdown(family = "source_sans", size = 19,
                                       hjust = 0.5, color = "gray30",
                                       lineheight = 1.3),
      plot.caption = element_markdown(family = "source_sans", size = 9,
                                      color = "gray50", hjust = 0.5),
      plot.caption.position = "plot"
    )
  )

final_plot
```

```{r save-final}
ggsave(
  filename = "outputs/2026_06_01_encyclical_comparison.png",
  plot = final_plot,
  device = "png",
  width = 8,
  height = 10,
  dpi = 300,
  bg = "white"
)
```

## Key Findings

**Structure:** *Magnifica Humanitas* is roughly 2.7x longer than *Rerum Novarum* (245 vs. 64 paragraphs, ~37,500 vs. ~13,900 words), reflecting the modern encyclical tradition of more comprehensive treatment. However, Leo XIII's paragraphs tend to be longer and more complex, with notably longer sentences characteristic of 19th-century academic prose.

**Vocabulary:** The TF-IDF analysis reveals each document's era clearly. *Rerum Novarum* is defined by words like "workmen," "masters," "socialists," and "property" — the vocabulary of industrial-era class conflict. *Magnifica Humanitas* introduces "AI," "digital," "algorithms," "platforms," and "data" — the language of the information age.

**Shared Language:** Despite the temporal gap, both documents share a robust core vocabulary centered on "dignity," "justice," "common good," "rights," "person," and "freedom" — the enduring lexicon of Catholic Social Teaching.

**Sentiment:** Both encyclicals maintain a predominantly positive emotional tone, with "trust" as the dominant sentiment. *Rerum Novarum* shows slightly more "anger" and "fear" (reflecting its denunciation of exploitation), while *Magnifica Humanitas* leans more toward "anticipation" (reflecting its forward-looking orientation toward technology).

**Scripture:** *Magnifica Humanitas* draws more heavily on New Testament sources (particularly the Gospels and Pauline epistles), while *Rerum Novarum* balances Old and New Testament citations more evenly, with Genesis and the Psalms featuring prominently alongside Matthew.

**Thematic Continuity:** The "Dignity & Rights" and "Common Good" themes dominate both documents, confirming that the core concerns of Catholic Social Teaching have remained remarkably stable across 135 years of technological upheaval.

## Discussion Questions for Golden Dome Data Tuesdays

This encyclical speaks directly to our work as data scientists and AI practitioners. Here are questions to guide our June 3 conversation:

### Ethics & Accountability in AI

1. **Leo XIV writes that AI is "never morally neutral" because it carries the decisions and priorities of its designers (§104).** As people who build models and pipelines — do you agree? Can you think of a time when a "neutral" technical choice you made actually embedded a value judgment?

2. **The encyclical calls for "accountability" at every stage — design, deployment, and use (§105).** In practice, who is accountable when an algorithm denies someone credit or a job? Is the current state of ML explainability sufficient, or are we building systems whose decisions we genuinely cannot explain?

3. **"A more moral AI would be useless if this morality is decided by a few" (§107).** Who gets to define the ethical guardrails for AI systems? Should it be governments, companies, open-source communities, religious institutions, or some combination?

### AI & Our Careers

4. **Leo XIV warns that AI can "de-skill workers, subject them to automated surveillance, and relegate them to rigid, repetitive tasks" (§150).** Have you experienced this in your own work? Has AI made your job more creative or more constrained?

5. **The encyclical insists that "access to work for all must continue to be a priority objective" (§154).** As data scientists who automate processes — how do we reconcile our professional work with the potential displacement it causes? Is there a moral obligation to consider employment impact when building automation?

6. **"A society that guaranteed work to only a small part of the population would expose many to forced inactivity" (§154).** If AI dramatically reduces the need for human labor, what does a just transition look like? Is universal basic income sufficient, or does human dignity require something more than income?

### Environmental Impact

7. **Leo XIV explicitly names the environmental cost: "Current AI systems require large amounts of energy and water, have a significant impact on carbon dioxide emissions" (§101).** Do we as practitioners have a responsibility to consider the carbon footprint of our models? Should there be a "sustainability budget" for training runs?

8. **The encyclical calls for "more sustainable technological solutions" (§101).** What does that look like in practice — smaller models, more efficient architectures, renewable-powered data centers? Or is the environmental cost acceptable given the benefits?

### Faith & Technology

9. **The central metaphor is Babel vs. Jerusalem — building for domination vs. building for communion (§7-9).** Which are we building in our daily work? Can a for-profit AI company genuinely build "Jerusalem," or is the profit motive inherently Babel-like?

10. **"Even when such instruments present themselves as capable of 'learning,' they do so differently from the human person. It is not the experience of one who allows himself to be shaped by life" (§99).** Does this distinction matter practically? When an AI system produces output indistinguishable from human work, does the absence of "inner growth" make a moral difference?

11. **Leo XIV says the true "more-than-human" comes not from technology but from grace — "We become fully human when we are more than human, when we allow God to take us beyond ourselves" (§128).** How does this challenge the transhumanist narrative that many in Silicon Valley embrace? Is there a way to hold both technological optimism and this theological vision?

12. **The encyclical asks us to "disarm AI" — removing it from the logic of arms competition (§110).** Is this realistic given the geopolitical AI race between the US, China, and others? What would "disarming" look like for a data scientist working at a tech company?

### Data, Power & the Common Good

13. **"Data ownership cannot be entrusted to the private sector alone... They are the fruit of the contribution of many and cannot be sold or entrusted to a few" (§108).** Is personal data a common good? What would it mean to treat it that way — data cooperatives, public data trusts, something else?

14. **Leo XIV identifies a "new colonialism" that "dominates not only bodies but also appropriates data, transforming personal lives into exploitable information" (§178).** Do you see this in the AI industry today? Who benefits from the data labeling work done in the Global South?

15. **The cosine similarity analysis shows that the sections of *Magnifica Humanitas* on labor and dignity are most textually aligned with *Rerum Novarum*.** After 135 years, the Church is still making the same argument — that workers are not means to an end. Has anything actually changed, or are we just repeating the same fight with new technology?

### The Meta-Question

16. **This entire blog post was built using Kiro, an AI coding assistant.** The encyclical warns about delegating too much to AI and "weakening personal judgment and creativity" (§100). Did using AI to analyze an encyclical *about* AI prove the Pope's point — or demonstrate that AI can be a tool for deeper human understanding?

## What's Next?

Further avenues for exploration:

- **Comparing with intermediate encyclicals** (Centesimus Annus, Laudato Si') would show the evolution of social teaching vocabulary over time
- **Network analysis** of cross-references between encyclicals could map the intellectual genealogy of Catholic Social Teaching
- **Readability scoring** (Flesch-Kincaid, etc.) could quantify how papal communication style has evolved for modern audiences
- **Structural break detection** on the sentiment arc could identify the precise turning points in each encyclical's argument
