---
title: "Press Start to Profit: Can Machine Learning Predict Video Game Movie Hits?"
date: 2026-06-09
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE,
  fig.width = 8, 
  fig.height = 6, 
  fig.path = "outputs/"
)
```

Hollywood spent decades proving that turning a beloved video game into a movie was essentially a cheat code for **losing money**. From the 1993 *Super Mario Bros.* disaster to the Uwe Boll cinematic universe nobody asked for, video game movies were the industry's most reliable punchline. But something changed. In the 2020s, Sonic started sprinting to the bank, Mario collected a billion-dollar coin, and even *Five Nights at Freddy's* — a game about security cameras — outearned most Marvel movies per dollar spent.

So what separates a game movie that prints money from one that belongs in the bargain bin? Let's find out — with **machine learning**, some **exploratory data wizardry**, and a healthy appreciation for the fact that *Alone in the Dark* got a 1% on Rotten Tomatoes.

## Loading the Quarters (Libraries & Data)

```{r libraries}
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(showtext)
library(ggtext)
library(tidymodels)
library(vip)
library(patchwork)

# Custom fonts
font_add_google("Press Start 2P", "press_start")
font_add_google("Source Sans 3", "source_sans")
showtext_auto()
showtext_opts(dpi = 300)

# Font Awesome for caption
font_add(family = "fa-brands",
         regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid",
         regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")

theme_set(theme_minimal(base_family = "source_sans", base_size = 14))
```

```{r load-data}
game_films <- readr::read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-06-09/game_films.csv'
)
```

## Insert Coin: What Are We Working With?

The dataset contains **`r nrow(game_films)` films** based on video games, scraped from Wikipedia. It covers everything from big-budget theatrical releases to obscure direct-to-video anime and documentaries about speedrunning.

```{r data-overview}
glimpse(game_films)
```

```{r category-breakdown}
game_films |>
  count(category, sort = TRUE) |>
  mutate(category = fct_reorder(category, n)) |>
  ggplot(aes(n, category, fill = category)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), hjust = -0.2, size = 5, family = "source_sans") +
  scale_fill_viridis_d(option = "plasma", end = 0.85) +
  labs(
    title = "Film Categories in the Dataset",
    subtitle = "Theatrical releases dominate, but there's a surprising number of documentaries",
    x = "Number of Films", y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  ) +
  xlim(0, 280)
```

## Narrowing the Focus: Theatrical Releases with USD Financials

For our machine learning model, we need films with comparable financial data. Let's filter to theatrical releases with USD box office figures and known budgets — the films where we can actually calculate whether they made or lost money.

```{r prepare-ml-data}
theatrical <- game_films |>
  filter(
    category == "Theatrical releases",
    worldwide_box_office_currency == "$",
    !is.na(worldwide_box_office),
    !is.na(budget_low)
  ) |>
  mutate(
    budget_mid = (budget_low + budget_high) / 2,
    roi = (worldwide_box_office - budget_mid) / budget_mid * 100,
    year = year(release_date),
    decade = paste0(floor(year / 10) * 10, "s"),
    hit = factor(ifelse(roi > 100, "Hit", "Flop"), levels = c("Hit", "Flop")),
    has_sequel = str_detect(title, "2|3|II|III|Annihilation|Apocalypse|Extinction|Afterlife|Retribution|Final Chapter|Welcome|Revelation|Return|Galaxy"),
    budget_millions = budget_mid / 1e6,
    # Map iconic game characters
    game_character = case_when(
      str_detect(title, "Mario") ~ "Mario",
      str_detect(title, "Sonic") ~ "Sonic",
      str_detect(title, "Pok.mon|Pikachu") ~ "Pikachu",
      str_detect(title, "Resident Evil") ~ "Jill Valentine",
      str_detect(title, "Tomb Raider|Lara Croft") ~ "Lara Croft",
      str_detect(title, "Mortal Kombat") ~ "Scorpion",
      str_detect(title, "Street Fighter") ~ "Ryu",
      str_detect(title, "Angry Birds") ~ "Red (Angry Birds)",
      str_detect(title, "Minecraft") ~ "Steve",
      str_detect(title, "Silent Hill") ~ "Pyramid Head",
      str_detect(title, "Hitman") ~ "Agent 47",
      str_detect(title, "Uncharted") ~ "Nathan Drake",
      str_detect(title, "Assassin") ~ "Ezio",
      str_detect(title, "Prince of Persia") ~ "The Prince",
      str_detect(title, "Warcraft") ~ "Thrall",
      str_detect(title, "Five Nights") ~ "Freddy Fazbear",
      str_detect(title, "Doom") ~ "Doomguy",
      str_detect(title, "Rampage") ~ "George (Rampage)",
      str_detect(title, "Need for Speed") ~ "Racer",
      str_detect(title, "Max Payne") ~ "Max Payne",
      str_detect(title, "Ratchet") ~ "Ratchet",
      str_detect(title, "Until Dawn") ~ "Sam",
      str_detect(title, "Borderlands") ~ "Lilith",
      str_detect(title, "Iron Lung") ~ "Submarine Pilot",
      TRUE ~ "Other"
    )
  )

cat("Films for analysis:", nrow(theatrical), "\n")
cat("Hits (ROI > 100%):", sum(theatrical$hit == "Hit"), "\n")
cat("Flops (ROI ≤ 100%):", sum(theatrical$hit == "Flop"), "\n")
```

We're defining a "**Hit**" as a film that more than doubled its production budget at the worldwide box office (ROI > 100%). This is a generous threshold — studios typically need 2-3x their budget to truly break even after marketing costs — but it gives us a clean binary classification.

## EDA: The Evolution of Game Movies

### Budget vs. Box Office: The Money Trail

```{r eda-budget-vs-box-office, fig.height=8}
theatrical |>
  ggplot(aes(budget_millions, worldwide_box_office / 1e6, color = hit)) +
  geom_point(aes(size = abs(roi)), alpha = 0.7) +
  geom_abline(slope = 2, intercept = 0, linetype = "dashed", color = "gray40") +
  geom_text(
    data = theatrical |> filter(roi > 1000 | roi < -80 | worldwide_box_office > 800e6),
    aes(label = str_wrap(title, 15)),
    size = 3, vjust = -1, show.legend = FALSE, family = "source_sans"
  ) +
  scale_x_continuous(labels = dollar_format(suffix = "M")) +
  scale_y_continuous(labels = dollar_format(suffix = "M")) +
  scale_color_manual(values = c("Hit" = "#2ECC71", "Flop" = "#E74C3C")) +
  scale_size_continuous(range = c(2, 12), guide = "none") +
  annotate("text", x = 100, y = 150, label = "Break-even line (2x budget)", 
           angle = 28, color = "gray40", size = 4, family = "source_sans") +
  labs(
    title = "Budget vs. Worldwide Box Office",
    subtitle = "Dashed line = 2x budget (our hit threshold). Bigger dots = bigger ROI (absolute value)",
    x = "Production Budget", y = "Worldwide Box Office",
    color = "Outcome"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )
```

The plot reveals a fascinating pattern: **the biggest hits aren't necessarily the biggest budgets**. *Five Nights at Freddy's* and *Iron Lung* achieved astronomical ROI on modest budgets, while *Borderlands* cratered despite a $100M+ budget. The dashed line represents our break-even threshold — everything above it made at least twice its money back.

### The Decade of Redemption

```{r eda-decade-improvement, fig.height=7}
decade_stats <- theatrical |>
  group_by(decade) |>
  summarise(
    n = n(),
    mean_rt = mean(rotten_tomatoes, na.rm = TRUE),
    mean_roi = mean(roi, na.rm = TRUE),
    hit_rate = mean(hit == "Hit") * 100,
    .groups = "drop"
  )

p1 <- decade_stats |>
  ggplot(aes(decade, mean_rt, fill = decade)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(mean_rt), "%")), vjust = -0.5, size = 5, 
            family = "source_sans") +
  scale_fill_viridis_d(option = "turbo", end = 0.85) +
  labs(title = "Average Rotten Tomatoes Score", y = "Score (%)", x = NULL) +
  ylim(0, 60) +
  theme(panel.grid = element_blank(), axis.ticks = element_blank())

p2 <- decade_stats |>
  ggplot(aes(decade, hit_rate, fill = decade)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(hit_rate), "%")), vjust = -0.5, size = 5,
            family = "source_sans") +
  scale_fill_viridis_d(option = "turbo", end = 0.85) +
  labs(title = "Hit Rate (ROI > 100%)", y = "% of Films", x = NULL) +
  ylim(0, 85) +
  theme(panel.grid = element_blank(), axis.ticks = element_blank())

p1 + p2 +
  plot_annotation(
    title = "Video Game Movies: From Joke to Juggernaut",
    subtitle = "Critical reception and financial success have both improved dramatically by decade",
    theme = theme(
      plot.title = element_text(size = 18, face = "bold"),
      plot.subtitle = element_text(size = 13)
    )
  )
```

The numbers tell a clear story: game movies went from **19% average RT score** in the 1990s to **48%** in the 2020s, while the hit rate climbed from **50%** to a stunning **76%**. Hollywood finally learned to respect the source material.

### Publisher Performance: Who Wins the Adaptation Game?

```{r eda-publisher-performance, fig.height=7}
publisher_stats <- theatrical |>
  group_by(original_game_publisher) |>
  summarise(
    n = n(),
    mean_roi = mean(roi),
    total_box_office = sum(worldwide_box_office),
    hit_rate = mean(hit == "Hit") * 100,
    .groups = "drop"
  ) |>
  filter(n >= 2) |>
  arrange(desc(mean_roi))

publisher_stats |>
  mutate(original_game_publisher = fct_reorder(original_game_publisher, mean_roi)) |>
  ggplot(aes(mean_roi, original_game_publisher, fill = hit_rate)) +
  geom_col() +
  geom_vline(xintercept = 100, linetype = "dashed", color = "white", linewidth = 0.8) +
  geom_text(aes(label = paste0(n, " films")), hjust = -0.1, size = 4, 
            family = "source_sans") +
  scale_fill_gradient2(low = "#E74C3C", mid = "#F39C12", high = "#2ECC71", 
                       midpoint = 50, limits = c(0, 100)) +
  labs(
    title = "Average ROI by Game Publisher",
    subtitle = "Color = hit rate (% of films with ROI > 100%). Dashed line = our hit threshold.",
    x = "Average ROI (%)", y = NULL,
    fill = "Hit Rate (%)"
  ) +
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  )
```

**Nintendo** and **ScottGames** (Five Nights at Freddy's) lead the pack with extraordinary average returns. Meanwhile, **Capcom** is the volume king — 10 Resident Evil films! — with consistently strong performance. The dashed line marks our "hit" threshold of 100% ROI.

### Rotten Tomatoes vs. ROI: Do Critics Even Matter?

```{r eda-rt-vs-roi, fig.height=7}
theatrical |>
  filter(!is.na(rotten_tomatoes)) |>
  ggplot(aes(rotten_tomatoes, roi, color = hit)) +
  geom_point(size = 4, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "gray30", linetype = "dashed") +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray50") +
  geom_text(
    data = theatrical |> filter(!is.na(rotten_tomatoes), roi > 1000 | (rotten_tomatoes > 80 & roi < 0)),
    aes(label = str_wrap(title, 15)),
    size = 3, vjust = -1, show.legend = FALSE, family = "source_sans"
  ) +
  scale_color_manual(values = c("Hit" = "#2ECC71", "Flop" = "#E74C3C")) +
  labs(
    title = "Do Critics Predict Box Office Success?",
    subtitle = "Weak positive correlation — but some beloved films flop and some hated films print money",
    x = "Rotten Tomatoes Score (%)", y = "ROI (%)",
    color = "Outcome"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )
```

Here's the kicker: **critics barely matter** for video game movies. The correlation between RT score and ROI is weak. *Werewolves Within* scored 86% on RT but lost 86% of its budget. Meanwhile, *Silent Hill: Revelation* scored 8% and still more than doubled its money. Fans show up for their characters regardless of what critics say.

### Timeline: The Renaissance in Real Time

```{r eda-timeline, fig.height=7}
theatrical |>
  ggplot(aes(release_date, roi, color = hit)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray50") +
  geom_point(aes(size = budget_millions), alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, color = "gold", linewidth = 1.2) +
  scale_color_manual(values = c("Hit" = "#2ECC71", "Flop" = "#E74C3C")) +
  scale_size_continuous(range = c(2, 10), labels = dollar_format(suffix = "M")) +
  labs(
    title = "The Video Game Movie Renaissance",
    subtitle = "ROI over time — dot size = budget. The gold trend line tells the story.",
    x = NULL, y = "ROI (%)",
    color = "Outcome", size = "Budget"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )
```

The gold [LOESS trend line](https://en.wikipedia.org/wiki/Local_regression) — a smoothing technique that captures the local trend without assuming a linear relationship — shows the unmistakable upward trajectory. After 2018, flops became the exception rather than the rule.

## Boss Fight: Training a Random Forest Classifier

Time for the main event. We'll train a [random forest](https://en.wikipedia.org/wiki/Random_forest) — an ensemble method that builds hundreds of decision trees and lets them vote on the outcome — to predict whether a video game movie will be a **Hit** or a **Flop**.

Our features:
- **Budget** (in millions): How much was invested?
- **Year**: When was it released? (capturing the "renaissance" effect)
- **Has sequel**: Is it a sequel or part of a franchise?
- **Rotten Tomatoes score**: Critics' verdict
- **Publisher track record**: How many films has this publisher made?

```{r random-forest-model}
# Add publisher track record
publisher_counts <- theatrical |>
  count(original_game_publisher, name = "publisher_film_count")

ml_data <- theatrical |>
  left_join(publisher_counts, by = "original_game_publisher") |>
  filter(!is.na(rotten_tomatoes)) |>
  select(hit, budget_millions, year, has_sequel, rotten_tomatoes, publisher_film_count) |>
  mutate(has_sequel = as.numeric(has_sequel))

cat("Training data size:", nrow(ml_data), "films\n")
cat("Hit/Flop split:\n")
print(table(ml_data$hit))

# Tidymodels workflow
set.seed(42)  # The answer to everything

# Define the recipe
rf_recipe <- recipe(hit ~ ., data = ml_data)

# Define the model specification
rf_spec <- rand_forest(trees = 500, mtry = 2, min_n = 3) |>
  set_engine("ranger", importance = "impurity") |>
  set_mode("classification")

# Create and fit the workflow
rf_workflow <- workflow() |>
  add_recipe(rf_recipe) |>
  add_model(rf_spec)

rf_fit <- rf_workflow |>
  fit(data = ml_data)

# Model performance on training data
rf_preds <- augment(rf_fit, ml_data)
cat("\nModel Accuracy:\n")
rf_preds |>
  metrics(truth = hit, estimate = .pred_class) |>
  print()

cat("\nConfusion Matrix:\n")
rf_preds |>
  conf_mat(truth = hit, estimate = .pred_class) |>
  print()
```

### Feature Importance: What Predicts a Hit?

```{r feature-importance, fig.height=6}
# Extract the fitted model and use vip for importance
importance_df <- rf_fit |>
  extract_fit_parsnip() |>
  vip::vi() |>
  mutate(
    feature_label = case_when(
      Variable == "budget_millions" ~ "Budget ($M)",
      Variable == "year" ~ "Release Year",
      Variable == "has_sequel" ~ "Is a Sequel?",
      Variable == "rotten_tomatoes" ~ "RT Score",
      Variable == "publisher_film_count" ~ "Publisher Track Record"
    ),
    feature_label = fct_reorder(feature_label, Importance)
  )

importance_df |>
  ggplot(aes(Importance, feature_label, fill = Importance)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.2, size = 5,
            family = "source_sans") +
  scale_fill_gradient(low = "#3498DB", high = "#8E44AD") +
  labs(
    title = "What Makes a Video Game Movie a Hit?",
    subtitle = "Random Forest feature importance (Mean Decrease in Gini impurity)",
    x = "Importance Score", y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  ) +
  xlim(0, max(importance_df$Importance) * 1.3)
```

The random forest reveals that **release year** and **budget** are the most predictive features. This makes intuitive sense — newer films benefit from the industry's improved approach to adaptations, and budget signals studio confidence. Interestingly, whether a film is a sequel matters less than you'd think.

[Mean Decrease in Gini impurity](https://en.wikipedia.org/wiki/Decision_tree_learning#Gini_impurity) measures how much each feature contributes to correctly splitting hits from flops across all 500 decision trees. Higher values = more predictive power.

## The Hero Visualization: Character Power Rankings

Now for the fun part. Let's rank the most iconic video game characters by how their movies performed — complete with **movie poster art** pulled directly from [The Movie Database (TMDB)](https://www.themoviedb.org/) image CDN, exactly like we did in our [horror movies analysis](https://github.com/gdatascience/tidytuesday/tree/master/2022/2022_11_01).

The trick: TMDB hosts poster images on a public CDN at `https://image.tmdb.org/t/p/{size}/{poster_path}`. Once you know the `poster_path` for a movie (findable by browsing [themoviedb.org](https://www.themoviedb.org/)), the images load with no API key, no authentication, and no expiration. We use `gt::web_image()` to render them directly in the table.

```{r final-table}
library(gt)
library(gtExtras)

# TMDB image CDN prefix (no API key required to load images)
tmdb_prefix <- "https://image.tmdb.org/t/p/w154"

# Poster paths for each character's best film (found via themoviedb.org)
poster_lookup <- c(
 "Pikachu" = "/6YPzBcMH0aPNTvdXNCDLY0zdE1g.jpg",
 "Freddy Fazbear" = "/7BpNtNfxuocYEVREzVMO75hso1l.jpg",
 "Mario" = "/qNBAXBIQlnOThrVvA6mA2B5ggV6.jpg",
 "Submarine Pilot" = "/rkGgWXKPXR2q2TJpiMnzbKj8q4I.jpg",
 "Sonic" = "/d8Ryb8AunYAuycVKDp5HpdWPKgC.jpg",
 "Jill Valentine" = "/7glPlA0xPpxPxBu0TnY4ulQVCV1.jpg",
 "Scorpion" = "/9YhpvhhvLrtL1wlwEfpVL16GI9d.jpg",
 "Steve" = "/yFHHfHcUgGAxziP1C3lLt0q2T4s.jpg",
 "Red (Angry Birds)" = "/iOH0fEFtV9z9rZp9zmBFGGeWicv.jpg",
 "Agent 47" = "/skHqceAFYee0JZuYd9MVk2IQggi.jpg",
 "Nathan Drake" = "/rJHC1RUORuUhtfNb4Npclx0xnOf.jpg",
 "Sam" = "/bLY5yN4MKVynZ2HMZWElTOGBgBe.jpg",
 "George (Rampage)" = "/MGADip4thVSErP34FAAfzFBTZ5.jpg",
 "Lara Croft" = "/s4Qn5LF6OwK4rIifmthIDtbqDSs.jpg",
 "Thrall" = "/eGi5aoxaZveqNLtE7BZJCuWwR3G.jpg",
 "Racer" = "/2CgzPAdh7JFkjPoQlU6fP1gCEyw.jpg",
 "Max Payne" = "/5LrI4GiCSrChgkdskVZiwv643Kg.jpg",
 "Ryu" = "/tMrfMq01XcSthIqwqDGSaMw8iPI.jpg"
)

# Build character stats
character_stats <- theatrical |>
  filter(game_character != "Other") |>
  group_by(game_character) |>
  summarise(
    n_films = n(),
    mean_roi = mean(roi),
    total_box_office = sum(worldwide_box_office),
    best_film = title[which.max(roi)],
    mean_rt = mean(rotten_tomatoes, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(mean_roi)) |>
  slice_head(n = 16)

# Build the table data — drop tier, add % to ROI
table_data <- character_stats |>
  mutate(
    rank = row_number(),
    box_office_label = case_when(
      total_box_office >= 1e9 ~ paste0("$", round(total_box_office / 1e9, 1), "B"),
      total_box_office >= 1e6 ~ paste0("$", round(total_box_office / 1e6, 0), "M"),
      TRUE ~ paste0("$", round(total_box_office / 1e3, 0), "K")
    ),
    poster_url = paste0(tmdb_prefix, poster_lookup[game_character]),
    roi_pct = round(mean_roi)
  ) |>
  select(rank, poster_url, game_character, n_films, roi_pct, mean_roi,
         box_office_label, best_film)

# Get the global max ROI so bars are scaled consistently across both halves
global_max_roi <- max(table_data$mean_roi)

# Helper function to build a styled gt table from a subset
build_half_table <- function(data, global_max) {
  data |>
    gt() |>
    # Render poster images from TMDB CDN URLs
    text_transform(
      locations = cells_body(columns = poster_url),
      fn = function(x) {
        web_image(url = x, height = 55)
      }
    ) |>
    # Custom inline SVG bars scaled to global max
    text_transform(
      locations = cells_body(columns = mean_roi),
      fn = function(x) {
        values <- as.numeric(x)
        purrr::map(values, function(val) {
          pct <- val / global_max * 100
          gt::html(glue::glue(
            '<div style="width:100px;height:12px;background:#21262D;border-radius:2px;">
              <div style="width:{pct}%;height:100%;background:gold;border-radius:2px;"></div>
            </div>'
          ))
        })
      }
    ) |>
    # Column labels
    cols_label(
      rank = "#",
      poster_url = "",
      game_character = "CHARACTER",
      n_films = "FILMS",
      roi_pct = "ROI",
      mean_roi = "",
      box_office_label = "GROSS",
      best_film = "BEST FILM"
    ) |>
    # Format numbers with %
    fmt_number(columns = roi_pct, decimals = 0, pattern = "{x}%") |>
    # Dark theme
    tab_options(
      table.background.color = "#0D1117",
      heading.background.color = "#0D1117",
      column_labels.background.color = "#161B22",
      column_labels.font.weight = "bold",
      table.font.color = "white",
      table.font.size = px(8),
      table.border.top.style = "hidden",
      table.border.bottom.style = "hidden",
      column_labels.border.bottom.color = "#30363D",
      table_body.border.bottom.style = "hidden",
      table_body.hlines.color = "#21262D",
      data_row.padding = px(5)
    ) |>
    tab_style(
      style = cell_text(weight = "bold", size = px(9)),
      locations = cells_body(columns = game_character)
    ) |>
    tab_style(
      style = cell_text(color = "#58A6FF", size = px(8)),
      locations = cells_column_labels()
    ) |>
    tab_style(
      style = cell_text(weight = "bold", size = px(9), color = "#FFD700"),
      locations = cells_body(columns = roi_pct)
    ) |>
    # 8-bit font throughout
    opt_table_font(
      font = list(
        google_font(name = "Press Start 2P"),
        default_fonts()
      )
    ) |>
    cols_width(
      rank ~ px(25),
      poster_url ~ px(45),
      game_character ~ px(125),
      n_films ~ px(45),
      roi_pct ~ px(65),
      mean_roi ~ px(100),
      box_office_label ~ px(75),
      best_film ~ px(145)
    )
}

# Split into two halves and build each table
left_data <- table_data |> filter(rank <= 8)
right_data <- table_data |> filter(rank > 8)

left_table <- build_half_table(left_data, global_max_roi)
right_table <- build_half_table(right_data, global_max_roi)

# Save each half as a temporary PNG, then combine side-by-side with magick
library(magick)

specs_dir <- "../../.kiro/specs/2026_06_09_tidy_tuesday_game_films"
gtsave(left_table, file.path(specs_dir, "_left.png"), vwidth = 700, vheight = 750)
gtsave(right_table, file.path(specs_dir, "_right.png"), vwidth = 700, vheight = 750)

# Compose the final image: title bar + two tables side by side
left_img <- image_read(file.path(specs_dir, "_left.png"))
right_img <- image_read(file.path(specs_dir, "_right.png"))

# Trim white borders and replace with dark background
left_img <- image_trim(left_img)
right_img <- image_trim(right_img)
left_img <- image_border(left_img, color = "#0D1117", geometry = "10x10")
right_img <- image_border(right_img, color = "#0D1117", geometry = "10x10")

# Make both the same height
max_h <- max(image_info(left_img)$height, image_info(right_img)$height)
left_img <- image_extent(left_img, 
  paste0(image_info(left_img)$width, "x", max_h), 
  gravity = "north", color = "#0D1117")
right_img <- image_extent(right_img, 
  paste0(image_info(right_img)$width, "x", max_h), 
  gravity = "north", color = "#0D1117")

# Join side by side
combined <- image_append(c(left_img, right_img), stack = FALSE)

# Add title bar on top
title_bar <- image_blank(
  width = image_info(combined)$width, 
  height = 280, 
  color = "#0D1117"
) |>
  image_annotate(
    "CHARACTER SELECT: VIDEO GAME MOVIE POWER RANKINGS",
    font = "Courier", size = 84, weight = 700,
    color = "#00FF41", gravity = "center", location = "+0-30"
  ) |>
  image_annotate(
    "Avg ROI by game character | Posters via TMDB | #TidyTuesday",
    font = "Courier", size = 40, color = "#8B949E", 
    gravity = "center", location = "+0+50"
  )

# Add caption bar on bottom
caption_bar <- image_blank(
  width = image_info(combined)$width,
  height = 80,
  color = "#0D1117"
) |>
  image_annotate(
    "DataViz: Tony Galvan | Data: Wikipedia | Posters: TMDB | github.com/gdatascience",
    font = "Courier", size = 28, color = "#666666",
    gravity = "center"
  )

# Stack: title + tables + caption
final_img <- image_append(c(title_bar, combined, caption_bar), stack = TRUE)

# Add green border on top and bottom
border_line <- image_blank(
  width = image_info(final_img)$width, height = 4, color = "#00FF41"
)
final_img <- image_append(c(border_line, final_img, border_line), stack = TRUE)

image_write(final_img, "outputs/2026_06_09_tidy_tuesday_game_films.png")
cat("Final image:", image_info(final_img)$width, "x", image_info(final_img)$height, "\n")

# Clean up temp files
file.remove(file.path(specs_dir, "_left.png"))
file.remove(file.path(specs_dir, "_right.png"))
```

## Saving the Final Boss

```{r save-final}
# Final table already saved by gtsave above
showtext_auto(FALSE)
```

## Game Over: Key Takeaways

**The machine learning model confirms what the data shows visually:**

1. **Timing is everything** — Release year is the strongest predictor. Post-2018 game movies are dramatically more likely to succeed.
2. **Budget signals confidence** — Higher budgets correlate with hits, but low-budget horror games (*Five Nights at Freddy's*, *Iron Lung*) can achieve insane ROI.
3. **Critics don't matter much** — The correlation between RT scores and financial success is weak. Fans show up for their characters.
4. **Some characters are just built different** — Pikachu, Mario, Sonic, and Freddy Fazbear are money-printing machines.

The video game movie went from Hollywood's biggest joke to its most reliable investment. The cheat code? Stop treating source material with contempt and actually hire people who played the games.

## What's Next?

- Could we predict the box office of upcoming adaptations (Zelda, Metroid, Mass Effect) based on their franchise characteristics?
- How do Japanese anime game films (which dominate the Pokémon franchise) compare to Western live-action adaptations?
- Is there a "Uwe Boll effect" — did his string of disasters in the 2000s permanently damage the genre's reputation?

The data says the curse is broken. Player 2 has entered the chat, and this time, they're winning.
