---
title: "Goals and Glory: Visualizing the Women's World Cup Through Eight Tournaments"
author: "Tony Galvan"
date: "7/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

The FIFA Women's World Cup has grown from a 12-team experiment in 1991 to a global sporting spectacle. With data from all eight tournaments through 2019, we can trace the expansion of the competition, identify the dominant nations, and visualize the scoring relationships between teams. The United States' dominance is clear in the numbers — but the competition is getting fiercer.

## Loading the Data

We'll work with three datasets: country codes, match outcomes, and squad rosters.

```{r}
library(tidyverse)
theme_set(theme_light())

codes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/codes.csv")
wwc_outcomes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/wwc_outcomes.csv") |>
  left_join(codes, by = "team")
squads <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/squads.csv")
```

```{r}
glimpse(wwc_outcomes)
```

## Tournament Growth: Games Per Round by Year

The Women's World Cup has expanded significantly — from 26 matches in 1991 to 52 in 2019. Let's visualize how the tournament structure has grown.

```{r}
wwc_outcomes |>
  group_by(year, round) |>
  summarise(games = n() / 2) |>
  ungroup() |>
  mutate(round = fct_relevel(round, "Final", "Third Place Playoff", 
                             "Semi Final", "Quarter Final", "Round of 16")) |>
  ggplot(aes(year, games, fill = round)) +
  geom_col()
```

The addition of the Round of 16 in 1999 (when the tournament expanded to 16 teams) and further growth to 24 teams in 2015 are clearly visible in the stacked bars.

## The Winningest Nations

Which countries have accumulated the most victories across all Women's World Cup tournaments?

```{r}
wwc_outcomes |>
  filter(win_status == "Won") |>
  mutate(country = fct_lump(country, 9)) |>
  group_by(country) |>
  summarise(n = n()) |>
  mutate(country = fct_reorder(country, n, .desc = TRUE)) |>
  ggplot(aes(country, n, fill = country)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  labs(x = "",
       y = "# of wins",
       title = "Countries with the most Women's World Cup wins",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: Wikipedia")
```

The United States leads comfortably — a testament to decades of investment in women's soccer through Title IX and professional leagues.

## The Most-Defeated Nations

Losses tell a different story — they often indicate teams that consistently qualify but face tough competition in later rounds.

```{r}
wwc_outcomes |>
  filter(win_status == "Lost") |>
  mutate(country = fct_lump(country, 9)) |>
  group_by(country) |>
  summarise(n = n()) |>
  mutate(country = fct_reorder(country, n, .desc = TRUE)) |>
  ggplot(aes(country, n, fill = country)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  labs(x = "",
       y = "# of losses",
       title = "Countries with the most Women's World Cup losses",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: Wikipedia")
```

## Scoring Trends Over Time

Has the Women's World Cup become higher-scoring as the tournament has grown? More games means more goals, but what about the rate?

```{r}
wwc_outcomes |>
group_by(year) |>
summarise(goals = sum(score)) |>
ggplot(aes(year, goals)) + 
geom_col() +
geom_smooth(method = "loess", se = FALSE)
```

Total goals have increased with tournament expansion, and the smoothed trend suggests the rate of scoring has remained relatively stable.

## Does Squad Age Predict Success?

Is there a relationship between a team's average age and their win rate? Older squads might have more experience, but younger ones might have more energy.

```{r}
avg_age <- squads |>
  group_by(country) |>
  summarise(avg_age = mean(age)) |>
  mutate(country = fct_recode(country, "United States" = "US"))

wwc_outcomes |>
  left_join(avg_age, by = "country") |>
  group_by(win_status) |>
  summarise(avg_age = mean(avg_age, na.rm = TRUE)) |>
  ggplot(aes(win_status, avg_age, fill = win_status)) + 
  geom_col(show.legend = FALSE) + 
  coord_flip()
```

## Score Distribution

Let's look at how goals are distributed across all matches — most games are low-scoring affairs, but blowouts do happen.

```{r}
wwc_outcomes |>
  ggplot(aes(score)) +
  geom_histogram()
```

## Preparing Data for the Chord Diagram

To visualize head-to-head scoring relationships, we need to pair winners and losers from each match.

```{r}
wwc_winners <- wwc_outcomes |>
  filter(win_status == "Won") |>
  select(year, yearly_game_id, round, w_country = country, w_score = score) 

wwc_losers <- wwc_outcomes |>
  filter(win_status == "Lost") |>
  select(year, yearly_game_id, l_country = country, l_score = score) 
```

## Scoring Differentials: A Chord Diagram

This chord diagram shows the average scoring differential between pairs of countries that have played each other more than twice. Thicker connections indicate larger average margins of victory.

```{r eval=FALSE}
wwc_winners |>
  left_join(wwc_losers, by = c("year" = "year", 
                               "yearly_game_id" = "yearly_game_id")) |>
  group_by(w_country, l_country) |>
  summarise(games = n(),
            avg_score_diff = mean(w_score - l_score)) |>
  filter(games > 2) |>
  select(w_country, l_country, avg_score_diff) |>
  circlize::chordDiagram(order = c("United States", "China PR", "Brazil", "South Korea", "Norway", "France", "Japan", "Germany", "Nigeria", "Canada", "Sweden", "North Korea"))

title("Women's World Cup: scoring differentials chord diagram 
      Designer: Tony Galvan @gdatascience1  |  Source: Wikipedia")
```

The chord diagram reveals the web of competitive relationships in women's international soccer. The United States' connections are consistently thick, reflecting their dominance across multiple opponents over decades of competition.
