---
title: "Lights Out for Birds: How Building Illumination Drives Collision Deaths in Chicago"
author: "Tony Galvan"
date: "April 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Every year, hundreds of millions of birds die from collisions with buildings in the United States. In Chicago, researchers have been tracking these collisions at McCormick Place — one of the largest convention centers in North America — since 1978. What makes this dataset special is that it pairs collision counts with detailed lighting data, letting us directly test whether turning off lights saves bird lives. The answer is striking.

## Loading the Data

We have two datasets: bird collision records spanning nearly four decades, and nightly light scores at McCormick Place measuring what fraction of window bays were illuminated.

```{r}
library(tidyverse)

theme_set(theme_minimal())

bird_collisions <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-30/bird_collisions.csv")
mp_light <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-30/mp_light.csv")
```

## Exploring the Collision Data

```{r}
glimpse(bird_collisions)
```

```{r}
summary(bird_collisions$date)
```

The "bird_collisions" data contains observations of birds colliding with various lighted structures in Chicago from 9/15/1978 to 11/30/2016.

## Understanding the Light Score Data

```{r}
mp_light |>
  summary()
```

The "mp_light" data contains light scores (proportion of the 17 window bays that were illuminated) at McCormick Place in Chicago between 3/6/2000 and 5/26/2018.

Are there dates with multiple light_scores?

```{r}
mp_light_processed <- mp_light |>
  group_by(date) |>
  summarise(n = n()) |>
  filter(n == 1) |>
  select(date) |>
  left_join(mp_light)
```

## Illumination by Day of the Week

Let's see if there's a pattern in how brightly the building is lit based on the day of the week — convention schedules likely drive this.

```{r}
mp_light_processed |>
  mutate(weekday = lubridate::wday(date, label = TRUE)) |>
  group_by(weekday) |>
  summarize(pct_light_score = mean(light_score)/100) |>
  ggplot(aes(weekday, pct_light_score, fill = weekday)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Weekday", 
       y = "Average % of window bays illuminated",
       title = "Thursday was the most illuminated day of the week",
       subtitle = "McCormick Place Between 3/6/2000 and 5/26/2018")
```

Thursday stands out as the brightest day — likely because conventions often run Thursday through the weekend, with setup lighting on Wednesday nights.

## Illumination by Month

Seasonal patterns in building use should show up in the monthly averages.

```{r}
mp_light_processed |>
  mutate(month = lubridate::month(date, label = TRUE)) |>
  group_by(month) |>
  summarize(pct_light_score = mean(light_score)/100) |>
  ggplot(aes(month, pct_light_score, fill = month)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Month", 
       y = "Average % of window bays illuminated",
       title = "September was the most illuminated month",
       subtitle = "McCormick Place Between 3/6/2000 and 5/26/2018")
```

September — peak convention season — is also peak illumination. Unfortunately, this coincides with fall bird migration, creating a deadly combination.

## Joining Collisions with Light Data

Now let's connect the dots: on nights when more windows were lit, did more birds die?

```{r}
mp_bird_collisions <- bird_collisions |>
  filter(locality == "MP") |>
  left_join(mp_light_processed) |>
  filter(date > "2000-03-05" & !is.na(light_score)) |>
  mutate(light_score = light_score/100) |>
  group_by(date, light_score) |>
  summarise(n = n())

summary(mp_bird_collisions$date)
```

## Collisions Over Time, Colored by Light Score

```{r}
mp_bird_collisions |>
  ggplot(aes(date, light_score, color = n)) +
  geom_point(aes(size = n), alpha = 0.25) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Date", 
       y = "Average % of window bays illuminated",
       title = "More bird collisions occur when more windows are illuminated",
       subtitle = "McCormick Place Between 3/6/2000 and 5/26/2018")
```

## The Direct Relationship: Light Score vs. Collision Count

This is the key plot — a direct scatter of illumination level against bird deaths, with a smoothed trend line.

```{r}
mp_bird_collisions |>
  ggplot(aes(light_score, n)) +
  geom_point() +
  scale_x_continuous(labels = scales::percent_format()) +
  geom_smooth(method = "loess") +
  labs(x = "Average % of window bays illuminated", 
       y = "# of bird collisions",
       title = "More bird collisions occur when more windows are illuminated",
       subtitle = "McCormick Place Between 3/6/2000 and 5/26/2018")
```

The relationship is clear and compelling: as illumination increases, so do bird collisions. The loess curve shows the effect accelerating at higher light levels.

## Statistical Modeling: Poisson Regression

Since we're modeling count data (number of collisions), a Poisson regression is a natural starting point.

```{r eval=FALSE}
mod_pois <- glm(n ~ light_score, data = mp_bird_collisions, family = poisson)
summary(mod_pois)
pcount_pois <- colSums(pscl::predprob(mod_pois))[0:10]
```

## Negative Binomial Regression

Count data often exhibits overdispersion (variance exceeding the mean), which violates the Poisson assumption. A negative binomial model relaxes this constraint.

```{r}
mod_negb <- MASS::glm.nb(n ~ light_score, data = mp_bird_collisions)
summary(mod_negb)
```

## Comparing Model Fit

Let's compare the observed counts with the predicted probabilities from both models to see which fits better.

```{r eval=FALSE}
ocount <- table(mp_bird_collisions$n)[0:10]
pcount_negb <- colSums(pscl::predprob(mod_negb))[0:10]
data.frame(ocount, pcount_pois, pcount_negb)
```

## Conclusion

The Negative Binomial Regression model fits better than the Poisson model, confirming that the collision data is overdispersed. More importantly, both models confirm the same story: building illumination is a statistically significant predictor of bird collisions. The "Lights Out" programs that many cities have adopted are backed by solid evidence — turning off unnecessary lighting during migration seasons genuinely saves bird lives.
