---
title: "Repeat Champions: Which Countries Produce Multiple Nobel Prize Winners?"
author: "Tony Galvan"
date: "5/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

Winning a Nobel Prize is one of the highest honors in science, literature, and peace — but some individuals have won more than once. Where do these exceptional repeat laureates come from? Using data on all Nobel Prize winners and their publications, we can map the geography of extraordinary achievement and see which nations have produced the most multi-time winners.

## Loading the Data

```{r}
library(tidyverse)

theme_set(theme_minimal())

nobel_winners <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-14/nobel_winners.csv")
nobel_winner_all_pubs <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-14/nobel_winner_all_pubs.csv")
```

## Multiple Nobel Prize Winners by Birth Country

Let's identify laureates who have won more than once and see which countries they were born in. We'll lump smaller countries into an "Other" category to focus on the top producers.

```{r}
nobel_winners |>
  group_by(laureate_id, birth_country) |>
  summarise(n = n()) |>
  filter(n > 1) |>
  ungroup() |>
  mutate(birth_country = fct_lump(birth_country, n = 7)) |>
  group_by(birth_country) |>
  summarise(n2 = n()) |>
  filter(!is.na(birth_country)) |>
  mutate(birth_country = fct_reorder(birth_country, n2)) |>
  ggplot(aes(birth_country, n2, fill = birth_country)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "",
       y = "# of multiple Nobel Prize winners", 
       title = "The USA has the most multiple Nobel Prize winners")
```

The United States leads decisively in producing repeat Nobel laureates — a reflection of its massive research infrastructure, university system, and ability to attract top talent from around the world. The presence of several European nations (UK, France, Germany) reflects the historical concentration of scientific institutions in the West, though the landscape of Nobel-caliber research is gradually becoming more global.
