---
title: "TidyTemplate"
date: 2023-07-19
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(camcorder)
library(ggtext)
library(showtext)

# start recording
gg_record(
  dir = "~/Downloads/camcorder/2023_07_18_tidy_tuesday_detectors", # where to save the recording
  device = "png", # device to use to save images
  width = 6, # width of saved image
  height = 6, # height of saved image
  units = "in", # units for width and height
  dpi = 300 # dpi to use when saving image
)

# Font selected from http://fonts.google.com
tt_family <- "prompt"
font_add_google("Prompt", tt_family)

theme_set(theme_light(base_size = 32))

font_add(family = "fa-brands", regular = "~/Library/Fonts/Font Awesome 6 Brands-Regular-400.otf")
font_add(family = "fa-solid", regular = "~/Library/Fonts/Font Awesome 6 Free-Solid-900.otf")
showtext_auto()

# Colors selected using https://coolors.co/
tt_source <- "{detectors} Package"
bg_color <- "#605B56"
txt_color <- "#F2FBE0"
ltgrn <- "#DAFEB7"  
dkgrn <- "#ACC18A"
gray <- "#837A75"

tt_caption <- paste0("DataViz: Tony Galvan #TidyTuesday<span style='color:", bg_color, ";'>..</span><span style='font-family:fa-solid;color:", txt_color, ";'>&#xf0ce;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>", tt_source, "</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf099;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>@GDataScience1</span><span style='color:", bg_color, ";'>..</span><span style='font-family:fa-brands;color:", txt_color, ";'>&#xf09b;</span><span style='color:", bg_color, ";'>.</span><span style='color:", txt_color, ";'>GDataScience</span><span style='color:", bg_color, ";'>..</span>")

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-07-18")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Initial Data Wrangling

Read the data into memory and clean up the column names

```{r Wrangle}

detectors_raw <- tt$detectors |> 
  janitor::clean_names()

```


# Initial Exploration

Determine if there are any duplicate observations

```{r Visualize}

detectors_raw |> 
  count(document_id, name = "document_count") |> 
  count(document_count) |> 
  mutate(document_count = fct_reorder(factor(document_count), n)) |> 
  ggplot(aes(n, document_count, fill = document_count)) +
  geom_col(show.legend = FALSE) +
  scale_x_log10() +
  labs(title = "Why do some documents have so many observations?",
       subtitle = "Are there duplicates?")

detectors_raw |> 
  add_count(document_id) |> 
  filter(n > 7) |> 
  arrange(document_id, pred_ai) |> 
  View()
  
```

# Wrangle

Explore the data and process it into a nice format for plotting!

```{r Wrangle}

detectors <- detectors_raw |> 
  distinct()

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

detectors |> 
  count(document_id, name = "document_count") |> 
  count(document_count) |> 
  mutate(document_count = fct_reorder(factor(document_count), n)) |> 
  ggplot(aes(n, document_count, fill = document_count)) +
  geom_col(show.legend = FALSE) +
  scale_x_log10() +
  labs(title = "Did we fix the duplicate observation issues?")
  
```


```{r Visualize}

detectors |> 
  filter(!is.na(native)) |> 
  ggplot() +
  aes(x = detector, y = pred_ai, fill = native) +
  geom_violin(bw = .05) +
  labs(
    x = "GPT Detector Tool",
    y = "Predicted Probability That Sample Was Written by AI",
    fill = "Native English Writer"
  ) +
  scale_fill_brewer(type = "qual") +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "top"
  )
  
```


```{r Visualize}

detectors |> 
  filter(!is.na(native)) |> 
  ggplot(aes(pred_ai, fill = native)) +
  geom_histogram(show.legend = FALSE) +
  facet_wrap(~native, ncol = 1, scales = "free_y")

```




```{r Visualize}

gg_resize_film(
  width = 10, # width of saved image
  height = 6, # height of saved image
  units = "in", # units for width and height
  dpi = 300 # dpi to use when saving image
)

detectors |> 
  filter(!is.na(native)) |> 
  mutate(native = if_else(
    native == "Yes", 
    "Native English",
    "Non-Native English"
    ),
    native = fct_rev(native)
  ) |> 
  ggplot(aes(
    pred_ai, 
    fill = native
  )) +
  geom_histogram(
    show.legend = FALSE, 
    bins = 10
  ) +
  scale_x_continuous(
    labels = percent_format(),
    breaks = c(0, 0.5, 1)
  ) +
  scale_fill_manual(values = c(ltgrn, dkgrn)) +
  facet_grid(
    cols = vars(detector), 
    rows = vars(native), 
    scales = "free_y"
  ) +
  theme(
    panel.background = element_rect(fill = bg_color),
    plot.background = element_rect(fill = bg_color), 
    strip.background = element_rect(fill = bg_color), 
    plot.title = element_textbox_simple(
      family = tt_family,
      color = txt_color, 
      face = "bold",
      size = 230
    ),
    plot.subtitle = element_textbox_simple(
      family = tt_family,
      color = txt_color, 
      size = 56
    ),
    plot.caption = element_textbox_simple(
      family = tt_family,
      color = txt_color, 
      size = 32,
      margin = margin(t = 10, r = 0, b = 0, l = 0)
    ),
    axis.title = element_text(
      family = tt_family,
      color = txt_color,
      size = 26
    ),
    axis.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 26
    ),
    strip.text = element_text(
      family = tt_family,
      color = txt_color,
      size = 32
    ),
    plot.caption.position = "plot",
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    x = "Predicted probability of being classified as AI",
    y = "# of observations",
    title = "GPT DETECTORS",
    subtitle = "Which models are most biased against non-native English writers?",
    caption = tt_caption
  )

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2023_07_18_tidy_tuesday_detectors.png",
  device = "png")

gg_stop_recording()

gg_playback(
  name = "2023_07_18_tidy_tuesday_detectors.gif",
  first_image_duration = 4,
  last_image_duration = 20,
  frame_duration = .25
)

```
