---
title: "TidyTemplate"
date: 2021-06-29
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(lubridate)
library(prophet) # forecasting

theme_set(theme_light())

tt_caption <- "Source: London.gov  |  Analysis: Tony Galvan (@GDataScience1)"

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-06-29")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

animal_rescues <- tt$animal_rescues %>%
  mutate(date_time_of_call = dmy_hm(date_time_of_call),
         date_of_call = as.Date(date_time_of_call),
         month_of_call = floor_date(date_of_call, unit = "months"),
         month_name_of_call = month(month_of_call, label = TRUE),
         pandemic = ifelse(month_of_call >= ymd(20200301), "Pandemic", "Other"),
         animal = fct_lump(animal_group_parent, 3))

```

```{r}
summary(animal_rescues$date_of_call)
```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

animal_rescues %>%
  count(month_of_call, pandemic) %>%
  ggplot(aes(month_of_call, n)) +
  #geom_point(aes(color = pandemic), alpha = 0.25) +
  geom_col(aes(fill = pandemic), alpha = 0.25) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_fill_manual(values = c("gray50", "darkblue")) +
  theme(legend.position = "top") +
  labs(x = "",
       y = "# of animal rescues",
       color = "",
       fill = "",
       title = "The London Fire Brigade: Animal Rescues 2009 - 2021",
       subtitle = "Are animal rescues on the rise during the pandemic?",
       caption = tt_caption)
  
```

```{r}
animal_rescues %>%
  mutate(cal_year = as.factor(cal_year)) %>%
  count(cal_year, month_name_of_call, pandemic) %>%
  ggplot(aes(month_name_of_call, n, color = pandemic, group = cal_year)) +
  geom_line() +
  scale_color_manual(values = c("gray85", "darkblue")) +
  theme(legend.position = "top") +
  labs(x = "",
       y = "# of animal rescues",
       color = "",
       fill = "",
       title = "The London Fire Brigade: Animal Rescues 2009 - 2021",
       subtitle = "Are animal rescues on the rise during the pandemic?",
       caption = tt_caption)
```

```{r}
animal_rescues %>%
  count(month_of_call, animal, pandemic) %>%
  ggplot(aes(month_of_call, n, color = animal)) +
  geom_point(alpha = 0.15) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_fill_manual(values = c("gray50", "darkblue")) +
  facet_wrap(~animal, scale = "free_y") +
  theme(legend.position = "top") +
  labs(x = "",
       y = "# of animal rescues",
       color = "",
       fill = "",
       title = "The London Fire Brigade: Animal Rescues 2009 - 2021",
       subtitle = "Are animal rescues on the rise during the pandemic?",
       caption = tt_caption)
```


```{r}
animal_rescues %>%
  mutate(cal_year = as.factor(cal_year),
         line_color = if_else(pandemic == "Pandemic", paste0(animal, pandemic), pandemic)) %>%
  count(cal_year, month_name_of_call, animal, pandemic, line_color) %>%
  ggplot(aes(month_name_of_call, n, color = line_color, group = cal_year)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(~animal, scale = "free_y") +
  scale_color_manual(values = c("cyan", "green", "red", "gray85", "violet")) +
  theme(legend.position = "top") +
  labs(x = "",
       y = "# of animal rescues",
       color = "",
       fill = "",
       title = "The London Fire Brigade: Animal Rescues 2009 - 2021",
       subtitle = "This pandemic is for the dogs! All other animals need more rescues since March 2020.\n*Highlighted lines are during the pandemic",
       caption = tt_caption)
```

# Create a daily forecast

Prophet

```{r}
daily_animal_rescues <- animal_rescues %>%
  count(ds = date_of_call, name = "y")

fit_prophet <- prophet(daily_animal_rescues)

future <- make_future_dataframe(fit_prophet, periods = 12, freq = 'month')
fcast_prophet <- predict(fit_prophet, future)

fcast_prophet %>%
  ggplot(aes(ds, yhat,
             color = ifelse(
               ds <= ymd("2021-05-31"), "actual", "forecast"
             ))) +
  geom_ribbon(
    data = fcast_prophet %>%
      filter(ds > today()),
    aes(ymin = yhat_lower, ymax = yhat_upper),
    fill = "gray75",
    color = NA,
    size = 0,
    alpha = 0.8
  ) +
  geom_line() +
  scale_color_manual(values = c("gray50", "red")) +
  theme(legend.position = "bottom") +
  labs(
    x = "",
    y = "# of animal rescues",
    color = "",
    title = "The London Fire Brigade: Animal Rescues 2009 - 2021",
    subtitle = "Forecast created using a Prophet model",
    caption = tt_caption
  )
```

# Create a monthly forecast

```{r}
monthly_animal_rescues <- animal_rescues %>%
  count(ds = month_of_call, name = "y")

fit_prophet <- prophet(monthly_animal_rescues)

future <- make_future_dataframe(fit_prophet, periods = 12, freq = 'month')
fcast_prophet <- predict(fit_prophet, future)

fcast_prophet %>%
  ggplot(aes(ds, yhat,
             color = ifelse(
               ds <= ymd("2021-05-31"), "actual", "forecast"
             ))) +
  geom_ribbon(
    data = fcast_prophet %>%
      filter(ds > today()),
    aes(ymin = yhat_lower, ymax = yhat_upper),
    fill = "gray75",
    color = NA,
    size = 0,
    alpha = 0.8
  ) +
  geom_line() +
  scale_color_manual(values = c("gray50", "red")) +
  theme(legend.position = "bottom") +
  labs(
    x = "",
    y = "# of animal rescues",
    color = "",
    title = "The London Fire Brigade: Animal Rescues 2009 - 2021",
    subtitle = "Forecast created using a Prophet model",
    caption = tt_caption
  )
```

# Conclusion 

Was it the pandemic or just a trend that started in 2018? Or is there a 4 year cycle?

# Other ideas

### Do forecast by animals

### Add forecast to yearly lines plot

```{r}

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2021_06_29_tidy_tuesday.png",
  device = "png")

```
