---
title: "TidyTemplate"
date: 2024-06-21
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(shiny)
library(shinydashboard)
library(lubridate)

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2024-06-18")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

holidays <- tt$federal_holidays

```


# Shiny App

Using your processed dataset, create your unique visualization.

```{r ShinyAppUI}

# Define the list of holidays with their names and dates
holidays <- list(
  "New Year's Day" = "2025-01-01 00:00:00",
  "Independence Day" = "2024-07-04 00:00:00",
  "Labor Day" = "2024-09-02 00:00:00",
  "Thanksgiving" = "2024-11-28 00:00:00",
  "Christmas Day" = "2024-12-25 00:00:00"
)

# Find the closest future holiday
closest_holiday <- function(holidays) {
  now <- Sys.time()
  future_holidays <- sapply(holidays, function(date) as.POSIXct(date, format="%Y-%m-%d %H:%M:%S", tz="America/Los_Angeles"))
  future_holidays <- future_holidays[future_holidays > now]
  closest <- names(future_holidays)[which.min(as.numeric(future_holidays) - as.numeric(now))]
  return(closest)
}

default_holiday <- closest_holiday(holidays)

# Define the UI
ui <- dashboardPage(
  dashboardHeader(title = "Holiday Countdown Timer"),
  dashboardSidebar(
    sidebarMenu(
      selectInput(
        "selected_date",
        "Select a Holiday:",
        choices = holidays,
        selected = holidays[[default_holiday]]
      )
    )
  ),
  dashboardBody(
    fluidRow(
      box(
        title = textOutput("countdown_title"), 
        status = "primary", 
        solidHeader = TRUE, 
        width = 12,
        textOutput("countdown_timer"),
        tags$head(tags$script(HTML("
          setInterval(function() {
            var now = new Date();
            Shiny.onInputChange('current_time', now.toISOString());
          }, 1000);
        ")))
      )
    )
  )
)

```

```{r ShinyAppServer}

# Define the server logic
server <- function(input, output, session) {
  
  # Reactive expression to get the current time every second
  current_time <- reactive({
    invalidateLater(1000, session)
    if (is.null(input$current_time)) return(NULL)
    as.POSIXct(input$current_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="America/Los_Angeles")
  })
  
  # Get the name of the selected holiday
  selected_holiday <- reactive({
    names(holidays)[sapply(holidays, function(date) date == input$selected_date)]
  })
  
  # Update the title of the box dynamically
  output$countdown_title <- renderText({
    paste("Countdown to", selected_holiday())
  })
  
  output$countdown_timer <- renderText({
    if (is.null(current_time()) || is.null(input$selected_date)) {
      return("Calculating...")
    }
    
    current <- current_time()
    target_date <- as.POSIXct(input$selected_date, format="%Y-%m-%d %H:%M:%S", tz="America/Los_Angeles")
    diff <- as.numeric(difftime(target_date, current, units = "secs"))
    
    if (diff <= 0) {
      return("The countdown is over!")
    }
    
    days <- as.integer(diff / (60 * 60 * 24))
    hours <- as.integer((diff %% (60 * 60 * 24)) / (60 * 60))
    minutes <- as.integer((diff %% (60 * 60)) / 60)
    seconds <- as.integer(diff %% 60)
    
    paste0(
      "Time remaining: ", 
      sprintf("%02d", days), " days, ",
      sprintf("%02d", hours), " hours, ",
      sprintf("%02d", minutes), " minutes, ",
      sprintf("%02d", seconds), " seconds"
    )
  })
}

# Run the application 
shinyApp(ui = ui, server = server)

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "2024_06_18_tidy_tuesday_holidays.png",
  device = "png")

```
