---
title: "The Steam Economy: How PC Game Prices and Revenue Have Shifted Over Time"
author: "Tony Galvan"
date: "8/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "outputs/")
```

The PC gaming market has transformed dramatically over the past decade. Free-to-play games, indie titles, and massive sales have reshaped how games are priced and how revenue flows. Using data from Steam Spy — which estimates ownership and pricing for games on Valve's Steam platform — we can trace these shifts and see which years and months generated the most revenue. The story of PC gaming economics is one of falling prices but rising volume.

## Loading and Preparing the Data

We'll do some preprocessing: standardize game titles, parse release dates, extract the lower bound of the ownership range, and calculate estimated revenue.

```{r}
library(tidyverse)
library(lubridate)
theme_set(theme_light())

video_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv") |>
  mutate(game = str_to_title(game),
         release_date = str_replace(release_date, "8 Oct, 2014", "Oct 8, 2014"),
         release_date = mdy(release_date),
         release_year = year(release_date),
         release_month = month(release_date, label = TRUE),
         owners = as.numeric(gsub(",","",str_extract(owners,"[0-9]+(,[0-9]+)*"))),
         owners = if_else(owners == 0, 1, owners),
         revenue = owners * price) |>
  select(-number)
```

## The Most-Owned Games

Which games have the largest player bases on Steam?

```{r}
video_games |>
  arrange(desc(owners)) |>
  select(game, release_date, price, owners, revenue) |>
  top_n(7, owners)
```

Why do so many of the top games have a missing price? It turns out that these may be "free to play" games.

## The Missing Price Problem

How many games lack price data? This is important because it affects our revenue calculations.

```{r}
paste0(sum(is.na(video_games$price)), " out of ", nrow(video_games), " = ", round(100*sum(is.na(video_games$price))/nrow(video_games), 2), "%")
```

Are the missing prices "free to play" games or truly missing data?

## Missing Prices Over Time

If free-to-play games are becoming more common, we'd expect the number of missing prices to increase over time.

```{r}
video_games |>
  filter(is.na(price)) |>
  group_by(release_year) |>
  summarise(n = n()) |>
  ggplot(aes(release_year, n)) + 
  geom_col()
```

The sharp increase in missing prices after 2015 strongly suggests these are free-to-play titles — a business model that has exploded in popularity.

## Price Trends Over Time

For games that do have a price, how have prices changed?

```{r}
video_games |>
  ggplot(aes(release_date, price)) + 
  geom_point(alpha = 0.25) + 
  geom_smooth(method = "lm")
```

Prices are decreasing over time. Could it be because of the increase in indie games and mobile-style pricing?

## Average Price by Year

Let's look at the trend more clearly with yearly averages.

```{r}
video_games |>
  group_by(release_year) |>
  summarise(avg_price = mean(price, na.rm = TRUE)) |>
  ggplot(aes(release_year, avg_price)) + 
  geom_col() +
  geom_smooth(method = "lm")
```

The downward trend is unmistakable — the average PC game price has fallen steadily as the market has shifted toward lower price points and free-to-play models.

## Revenue Patterns by Month and Year

When do the highest-revenue games get released? Let's look at average revenue by release month, faceted by year.

```{r}
avg_video_game_revenue <- mean(video_games$revenue, na.rm = TRUE)

video_games |>
  group_by(release_year, release_month) |>
  summarise(avg_revenue = mean(revenue, na.rm = TRUE)) |>
  ggplot(aes(release_month, avg_revenue, fill = as.factor(release_month))) + 
  geom_col() + 
  facet_wrap(~release_year, scales = "free_y") +
  geom_hline(yintercept = avg_video_game_revenue, linetype = 2)
```

## Top Revenue Games

Which individual games generated the most estimated revenue?

```{r}
video_games |>
  arrange(desc(revenue)) |>
  select(game, release_date, price, owners, revenue)
```

## Revenue Over Time

Is the overall revenue per game increasing or decreasing?

```{r}
video_games |>
  ggplot(aes(release_date, revenue)) + 
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm")
```

Revenue is decreasing over time on a per-game basis — but that doesn't mean the market is shrinking. Let's look at total revenue by year.

## Total Revenue by Year

```{r}
video_games |>
  group_by(release_year) |>
  summarise(total_revenue = sum(revenue, na.rm = TRUE)) |>
  ggplot(aes(release_year, total_revenue, fill = as.factor(release_year))) + 
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::dollar_format(scale = 0.000000001, suffix = "B")) + 
  labs(x = "Release year",
       y = "Total revenue (in billions)",
       title = "2017 PC video games generated over $4 billion",
       subtitle = "Based on price multiplied by the number of owners",
       caption = "Designer: Tony Galvan @gdatascience1  |  Source: Steam Spy")
```

Despite falling per-game prices and revenue, the total market has grown — 2017 generated over $4 billion in estimated revenue. The PC gaming economy is healthy, just distributed across many more titles at lower individual price points.

## Word Cloud from Game Names

What words appear most frequently in PC game titles? This gives us a sense of the themes and genres that dominate the platform.

```{r eval=FALSE}
video_games |>
  tidytext::unnest_tokens(tbl = ., output = word, input = game) |>
  count(word, sort = TRUE) |>
  filter(is.na(as.numeric(word))) |>
  anti_join(get_stopwords()) |>
  filter(n > 100) |>
  na.omit() |>
  wordcloud2::wordcloud2(shape = "cardiod")
```

Words like "dark," "war," "world," and "dead" dominate — reflecting the gaming industry's well-known preference for action, conflict, and fantasy themes in their titles.
