---
title: "The Changing Face of the American Priesthood"
author: "Tony Galvan"
date: "2026-06-12"
output:
  html_document:
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 6, fig.path = "outputs/"
)
```

## A Church in Transition

The American Catholic priesthood is undergoing one of the most dramatic
transformations in its history. A [viral data visualization](https://x.com/ryanburge/status/2046569709009555937) 
by religion researcher Ryan Burge — using data from the 
[National Study of Catholic Priests (NSCP)](https://catholicproject.catholic.edu/national-study-of-catholic-priests/) — 
revealed a striking theological shift: priests ordained today are overwhelmingly 
conservative, a complete inversion from 1970 when progressives dominated.

A [reply to that post](https://x.com/SeamusNua/status/2046581326610895348) by 
@SeamusNua took things further by combining Burge's theological orientation 
percentages with ordination count data from [CARA](https://cara.georgetown.edu/frequently-requested-church-statistics/) 
(the Center for Applied Research in the Apostolate at Georgetown University).
The result was even more dramatic: in raw numbers, progressive ordinations 
collapsed from ~451 per year in 1970 to just ~9 in 2025.

This analysis brings together both data sources — CARA's longitudinal Church 
statistics and the NSCP's theological orientation data — to tell the full story 
of a priesthood in transition. We'll explore the theological inversion, the 
workforce crisis, the parish consolidation wave, and what machine learning can 
tell us about the future.

```{r libraries}
library(tidyverse)
library(readxl)
library(scales)
library(showtext)
library(ggtext)
library(tidymodels)
library(patchwork)

# Fonts
font_add_google("Source Sans 3", "source_sans")
font_add(family = "fa-brands",
         regular = "~/Library/Fonts/Font Awesome 7 Brands-Regular-400.otf")
font_add(family = "fa-solid",
         regular = "~/Library/Fonts/Font Awesome 7 Free-Solid-900.otf")
showtext_auto()
showtext_opts(dpi = 300)

theme_set(theme_light(base_family = "source_sans", base_size = 14))
```

```{r load-cara-data}
# Download CARA data directly from Georgetown
url <- "https://cara.georgetown.edu/s/CARAFrequentlyRequested-y3mw.xlsx"
tmp <- tempfile(fileext = ".xlsx")
download.file(url, tmp, mode = "wb", quiet = TRUE)

cara_raw <- read_excel(tmp, sheet = "US")

# Extract years from the header row
years_raw <- as.character(cara_raw[1, 2:ncol(cara_raw)])
years_clean <- parse_number(years_raw)

# Helper function to extract a row by label (uses regex, not fixed)
extract_row <- function(data, pattern, years) {
  row_idx <- which(str_detect(data[[1]], pattern))
  if (length(row_idx) == 0) return(NULL)
  values <- as.character(data[row_idx[1], 2:ncol(data)])
  parsed <- case_when(
    str_detect(values, "m$") ~ parse_number(values) * 1e6,
    values %in% c("--", "na", "NA", "TBA") ~ NA_real_,
    TRUE ~ parse_number(values)
  )
  tibble(year = years, value = parsed)
}

# Extract key series using regex patterns
total_priests <- extract_row(cara_raw, "^Total priests$", years_clean) |>
  rename(total_priests = value)

diocesan_priests <- extract_row(cara_raw, "^Diocesan priests$", years_clean) |>
  rename(diocesan_priests = value)

religious_priests <- extract_row(cara_raw, "^Religious priests$", years_clean) |>
  rename(religious_priests = value)

ordinations <- extract_row(cara_raw, "^Priestly ordinations$", years_clean) |>
  rename(ordinations = value)

parishes <- extract_row(cara_raw, "^Parishes$", years_clean) |>
  rename(parishes = value)

no_priest_parishes <- extract_row(cara_raw, "^Parishes without a resident priest", years_clean) |>
  rename(no_priest_parishes = value)

catholic_pop <- extract_row(cara_raw, "^Catholic population \\(The Official", years_clean) |>
  rename(catholic_pop = value)

mass_attendance <- extract_row(cara_raw, "^Catholics who attend Mass every week", years_clean) |>
  rename(mass_pct = value)

marriages <- extract_row(cara_raw, "^Marriages in previous year", years_clean) |>
  rename(marriages = value)

baptisms <- extract_row(cara_raw, "^Baptisms of infants", years_clean) |>
  rename(baptisms = value)

active_pct <- extract_row(cara_raw, "^Percentage of diocesan priests active", years_clean) |>
  rename(active_pct = value)

# Combine into a single dataset
cara_data <- total_priests |>
  left_join(diocesan_priests, by = "year") |>
  left_join(religious_priests, by = "year") |>
  left_join(ordinations, by = "year") |>
  left_join(parishes, by = "year") |>
  left_join(no_priest_parishes, by = "year") |>
  left_join(catholic_pop, by = "year") |>
  left_join(mass_attendance, by = "year") |>
  left_join(marriages, by = "year") |>
  left_join(baptisms, by = "year") |>
  left_join(active_pct, by = "year") |>
  mutate(
    priests_per_parish = total_priests / parishes,
    catholics_per_priest = catholic_pop / total_priests,
    pct_parishes_no_priest = no_priest_parishes / parishes
  )

cara_data
```

```{r theological-data}
# Theological orientation data from the NSCP (via Ryan Burge's visualization)
# These represent the % of priests in each ordination cohort who self-identify
# as progressive, moderate, or conservative
theological <- tibble(
  year = c(1970, 1975, 1980, 1985, 1990, 1995, 2000, 2005, 2010, 2015, 2020, 2021, 2025),
  ordinations = c(805, 771, 593, 533, 595, 511, 442, 454, 459, 515, 495, 441, 426),
  pct_progressive = c(0.56, 0.49, 0.44, 0.29, 0.28, 0.18, 0.17, 0.13, 0.09, 0.08, 0.02, 0.02, 0.02),
  pct_moderate = c(0.27, 0.33, 0.34, 0.39, 0.37, 0.35, 0.35, 0.31, 0.23, 0.24, 0.14, 0.14, 0.14),
  pct_conservative = c(0.17, 0.18, 0.22, 0.32, 0.35, 0.47, 0.48, 0.56, 0.68, 0.68, 0.84, 0.84, 0.84)
) |>
  mutate(
    total_progressive = round(ordinations * pct_progressive),
    total_moderate = round(ordinations * pct_moderate),
    total_conservative = round(ordinations * pct_conservative)
  )
```

## Exploring the Data

Let's start by understanding what we're working with. The CARA dataset covers 
the U.S. Catholic Church from 1965 to 2025 — six decades of demographic change 
captured in a single spreadsheet.

```{r eda-cara-overview}
cara_data |>
  select(year, total_priests, ordinations, parishes, catholic_pop, mass_pct) |>
  mutate(catholic_pop = catholic_pop / 1e6) |>
  rename(
    Year = year,
    `Total Priests` = total_priests,
    Ordinations = ordinations,
    Parishes = parishes,
    `Catholics (millions)` = catholic_pop,
    `Weekly Mass %` = mass_pct
  ) |>
  knitr::kable(digits = 1)
```

A few things jump out immediately:

- **Total priests dropped 44%** — from 59,192 in 1970 to 33,462 in 2025
- **The Catholic population grew 38%** — from 47.9 million to 66.3 million (OCD-reported)
- **Weekly Mass attendance halved** — from 55% in 1970 to under 18% today
- **Ordinations nearly halved** — from 805 in 1970 to 426 in 2025

```{r eda-priest-decline}
cara_data |>
  filter(!is.na(total_priests)) |>
  ggplot(aes(x = year, y = total_priests)) +
  geom_line(linewidth = 1.2, color = "#2C3E50") +
  geom_point(size = 2.5, color = "#2C3E50") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1965, 2025, 5)) +
  labs(
    title = "U.S. Catholic Priests: A Steady 60-Year Decline",
    subtitle = "Total priests (diocesan + religious), 1965–2025",
    x = NULL, y = "Total Priests",
    caption = "Source: CARA, Georgetown University"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

```{r eda-supply-demand}
cara_data |>
  filter(!is.na(catholics_per_priest)) |>
  ggplot(aes(x = year, y = catholics_per_priest)) +
  geom_line(linewidth = 1.2, color = "#E74C3C") +
  geom_point(size = 2.5, color = "#E74C3C") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1965, 2025, 5)) +
  labs(
    title = "The Growing Gap: Catholics Per Priest Has Tripled",
    subtitle = "From ~800 Catholics per priest in 1965 to nearly 2,000 in 2025",
    x = NULL, y = "Catholics Per Priest",
    caption = "Source: CARA, Georgetown University"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

```{r eda-mass-attendance}
cara_data |>
  filter(!is.na(mass_pct)) |>
  ggplot(aes(x = year, y = mass_pct)) +
  geom_line(linewidth = 1.2, color = "#8E44AD") +
  geom_point(size = 2.5, color = "#8E44AD") +
  scale_y_continuous(labels = percent, limits = c(0, 0.6)) +
  scale_x_continuous(breaks = seq(1965, 2025, 5)) +
  labs(
    title = "Weekly Mass Attendance: From Majority to Minority",
    subtitle = "Percentage of Catholics attending Mass every week, 1970–2025",
    x = NULL, y = "Weekly Mass Attendance",
    caption = "Source: CARA, Georgetown University"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

```{r eda-parishes-no-priest}
cara_data |>
  filter(!is.na(pct_parishes_no_priest)) |>
  ggplot(aes(x = year, y = pct_parishes_no_priest)) +
  geom_line(linewidth = 1.2, color = "#D35400") +
  geom_point(size = 2.5, color = "#D35400") +
  scale_y_continuous(labels = percent, limits = c(0, 0.25)) +
  scale_x_continuous(breaks = seq(1965, 2025, 5)) +
  labs(
    title = "Parishes Without a Resident Priest: 1 in 4",
    subtitle = "Percentage of U.S. Catholic parishes without a resident priest pastor",
    x = NULL, y = NULL,
    caption = "Source: CARA, Georgetown University"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

```{r eda-theological-pct}
theological |>
  select(year, pct_progressive, pct_moderate, pct_conservative) |>
  pivot_longer(-year, names_to = "orientation", values_to = "pct") |>
  mutate(orientation = str_remove(orientation, "pct_") |> str_to_title()) |>
  mutate(orientation = factor(orientation, levels = c("Conservative", "Moderate", "Progressive"))) |>
  ggplot(aes(x = year, y = pct, color = orientation)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1970, 2025, 5)) +
  scale_color_manual(values = c(
    "Conservative" = "#1B4F72",
    "Moderate" = "#27AE60",
    "Progressive" = "#3498DB"
  )) +
  labs(
    title = "The Theological Inversion: 55 Years of Shift",
    subtitle = "Self-identified theological orientation of priests by ordination cohort",
    x = NULL, y = NULL, color = "Orientation",
    caption = "Source: National Study of Catholic Priests (2022) via Ryan Burge"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot",
    legend.position = "top"
  )
```

## The Theological Inversion

Ryan Burge's original chart showed percentages. But as @SeamusNua demonstrated,
the raw numbers tell an even more dramatic story. When you multiply theological 
orientation percentages by actual ordination counts, you see a priesthood being 
completely remade:

```{r theological-inversion}
theological_long <- theological |>
  select(year, total_progressive, total_moderate, total_conservative) |>
  pivot_longer(-year, names_to = "orientation", values_to = "count") |>
  mutate(
    orientation = str_remove(orientation, "total_") |> str_to_title(),
    orientation = factor(orientation, levels = c("Conservative", "Moderate", "Progressive"))
  )

ggplot(theological_long, aes(x = year, y = count, color = orientation)) +
  geom_line(linewidth = 1.3) +
  geom_point(size = 3) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1970, 2025, 5)) +
  scale_color_manual(values = c(
    "Conservative" = "#1B4F72",
    "Moderate" = "#27AE60",
    "Progressive" = "#3498DB"
  )) +
  annotate("text", x = 2020, y = 380, label = "Conservative", 
           color = "#1B4F72", fontface = "bold", size = 5, family = "source_sans") +
  annotate("text", x = 2020, y = 80, label = "Moderate", 
           color = "#27AE60", fontface = "bold", size = 5, family = "source_sans") +
  annotate("text", x = 2020, y = 20, label = "Progressive", 
           color = "#3498DB", fontface = "bold", size = 5, family = "source_sans") +
  labs(
    title = "The Great Inversion: Ordinations by Theological Orientation",
    subtitle = "Estimated number of priests ordained per year by self-identified theology",
    x = NULL, y = "Priests Ordained",
    caption = "Sources: CARA (ordination counts) + NSCP via Ryan Burge (orientation %)"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot",
    legend.position = "none"
  )
```

**The crossover happened around 1990–1995.** Before that, progressive ordinations 
outnumbered conservative ones. After that, conservatives rapidly dominated — and 
by 2020, the ratio was approximately **40:1** (conservative to progressive).

This isn't just a percentage shift. In absolute terms:
- **Progressive ordinations collapsed** from 451 (1970) to 9 (2025) — a 98% decline
- **Conservative ordinations grew** from 137 (1970) to 358 (2025) — a 161% increase
- **Moderate ordinations fell** from 217 (1970) to 60 (2025) — a 72% decline

## The Workforce Crisis

While the theological composition shifted, the overall workforce contracted 
dramatically. The Church has fewer priests serving more Catholics across fewer 
but larger parishes.

```{r workforce-crisis}
workforce <- cara_data |>
  filter(!is.na(total_priests), year >= 1970) |>
  select(year, total_priests, parishes, catholic_pop) |>
  mutate(
    priests_idx = total_priests / total_priests[1] * 100,
    parishes_idx = parishes / parishes[1] * 100,
    pop_idx = catholic_pop / catholic_pop[1] * 100
  ) |>
  select(year, priests_idx, parishes_idx, pop_idx) |>
  pivot_longer(-year, names_to = "metric", values_to = "index") |>
  mutate(metric = case_when(
    metric == "priests_idx" ~ "Total Priests",
    metric == "parishes_idx" ~ "Parishes",
    metric == "pop_idx" ~ "Catholic Population"
  ))

ggplot(workforce, aes(x = year, y = index, color = metric)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray60") +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = seq(1970, 2025, 5)) +
  scale_color_manual(values = c(
    "Total Priests" = "#C0392B",
    "Parishes" = "#F39C12",
    "Catholic Population" = "#2980B9"
  )) +
  labs(
    title = "Supply vs. Demand: A Growing Gap",
    subtitle = "Indexed to 1970 = 100. Population grew while priests and parishes declined.",
    x = NULL, y = "Index (1970 = 100)", color = NULL,
    caption = "Source: CARA, Georgetown University"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot",
    legend.position = "top"
  )
```

The scissors chart tells the story clearly: Catholic population grew ~38% since 
1970 while the number of priests dropped ~44%. That means the workload per 
priest roughly **doubled** — and that's before accounting for the decline in 
active ministry rates (from 90% to 65%).

## Parish Consolidation

With fewer priests, parishes have been closing or merging at an accelerating rate.
The number of parishes without a resident priest pastor grew from 571 in 1970 to 
over 3,600 in 2025 — more than a six-fold increase.

```{r parish-consolidation}
parish_data <- cara_data |>
  filter(!is.na(parishes), !is.na(no_priest_parishes), year >= 1970)

ggplot(parish_data) +
  geom_col(aes(x = year, y = parishes), fill = "#BDC3C7", width = 4) +
  geom_col(aes(x = year, y = no_priest_parishes), fill = "#E74C3C", width = 4) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1970, 2025, 5)) +
  labs(
    title = "Parishes Without a Priest (Red) vs. Total Parishes (Gray)",
    subtitle = "The red segment has grown from 3% to 23% of all parishes",
    x = NULL, y = "Number of Parishes",
    caption = "Source: CARA, Georgetown University"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

## Machine Learning: Forecasting the Future

Can we use the historical trends to forecast where the priesthood is headed? 
The CARA data gives us a rich multivariate time series. Let's use the 
[tidymodels](https://www.tidymodels.org/) framework to build predictive models.

### Predicting Total Priest Count

We'll model total priest counts as a function of year and related Church metrics,
then use the models to understand which factors best explain the decline and 
what the trajectory looks like.

```{r ml-data-prep}
# Create a modeling dataset with year as numeric predictor
# and derived features
ml_data <- cara_data |>
  filter(!is.na(total_priests), !is.na(ordinations), year >= 1970) |>
  mutate(
    years_since_1970 = year - 1970,
    decade = floor(year / 10) * 10,
    ordination_rate = ordinations / total_priests,
    parish_coverage = total_priests / parishes,
    pop_millions = catholic_pop / 1e6
  ) |>
  filter(!is.na(pop_millions))

ml_data |>
  select(year, total_priests, ordinations, ordination_rate, parish_coverage, pop_millions) |>
  knitr::kable(digits = 3)
```

### Model 1: Linear Regression — The Baseline

A simple linear model tells us the average rate of decline:

```{r ml-linear}
# Linear model using tidymodels
lm_spec <- linear_reg() |>
  set_engine("lm") |>
  set_mode("regression")

lm_recipe <- recipe(total_priests ~ years_since_1970, data = ml_data)

lm_workflow <- workflow() |>
  add_recipe(lm_recipe) |>
  add_model(lm_spec)

lm_fit <- lm_workflow |> fit(data = ml_data)

# Extract coefficients
lm_fit |> tidy()
```

```{r ml-linear-predict}
# Predict and visualize
future_years <- tibble(
  years_since_1970 = seq(0, 65, 5),
  year = 1970 + seq(0, 65, 5)
)

lm_preds <- predict(lm_fit, future_years) |>
  bind_cols(future_years)

ggplot() +
  geom_point(data = ml_data, aes(x = year, y = total_priests), 
             size = 3, color = "#2C3E50") +
  geom_line(data = lm_preds, aes(x = year, y = .pred), 
            linewidth = 1, color = "#E74C3C", linetype = "dashed") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1970, 2035, 5)) +
  labs(
    title = "Linear Forecast: Total U.S. Priests",
    subtitle = "At the current rate, the priesthood will be under 30,000 by 2030",
    x = NULL, y = "Total Priests",
    caption = "Linear model: tidymodels + lm engine"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

### Model 2: Polynomial Regression — Capturing the Curve

The decline isn't perfectly linear — it accelerated from 1970–2000, then 
decelerated slightly as the priest population aged out. A polynomial model 
captures this curvature:

```{r ml-polynomial}
poly_recipe <- recipe(total_priests ~ years_since_1970, data = ml_data) |>
  step_poly(years_since_1970, degree = 2)

poly_workflow <- workflow() |>
  add_recipe(poly_recipe) |>
  add_model(lm_spec)

poly_fit <- poly_workflow |> fit(data = ml_data)

poly_preds <- predict(poly_fit, future_years) |>
  bind_cols(future_years)

ggplot() +
  geom_point(data = ml_data, aes(x = year, y = total_priests), 
             size = 3, color = "#2C3E50") +
  geom_line(data = poly_preds, aes(x = year, y = .pred), 
            linewidth = 1, color = "#8E44AD", linetype = "dashed") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1970, 2035, 5)) +
  labs(
    title = "Polynomial Forecast: Capturing the Deceleration",
    subtitle = "A quadratic model suggests the decline rate is slowing — but still declining",
    x = NULL, y = "Total Priests",
    caption = "Polynomial (degree 2) model: tidymodels + lm engine"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

### Model 3: Random Forest — What Drives the Decline?

With multiple predictors, a [random forest](https://en.wikipedia.org/wiki/Random_forest) 
— an ensemble of decision trees that makes predictions by averaging many 
individual trees — can identify which factors are most important. We'll use 
ordination rates, parish coverage, Catholic population, and year:

```{r ml-rf}
rf_spec <- rand_forest(trees = 500) |>
  set_engine("ranger", importance = "impurity") |>
  set_mode("regression")

rf_recipe <- recipe(
  total_priests ~ years_since_1970 + ordinations + pop_millions + parishes,
  data = ml_data
)

rf_workflow <- workflow() |>
  add_recipe(rf_recipe) |>
  add_model(rf_spec)

rf_fit <- rf_workflow |> fit(data = ml_data)

# Variable importance
rf_fit |>
  extract_fit_parsnip() |>
  vip::vip(num_features = 4) +
  labs(
    title = "What Drives the Priest Decline?",
    subtitle = "Variable importance from a random forest model"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

### Model 4: Modeling Theological Orientation as a Continuous Trend

Rather than a binary classification (which has near-perfect separation), let's
model the *percentage* of conservative priests as a continuous outcome using 
[logistic-shaped regression](https://en.wikipedia.org/wiki/Logistic_function).
This is a nonlinear least-squares fit of a sigmoid curve — capturing the S-shaped
transition from minority to supermajority:

```{r ml-theological-trend}
# Model pct_conservative as a logistic curve (nls)
theo_model_data <- theological |>
  mutate(years_since_1970 = year - 1970)

# Fit a logistic growth curve: y = L / (1 + exp(-k*(x - x0)))
nls_fit <- nls(
  pct_conservative ~ 0.85 / (1 + exp(-k * (years_since_1970 - x0))),
  data = theo_model_data,
  start = list(k = 0.1, x0 = 20)
)

cat("Logistic curve parameters:\n")
summary(nls_fit)

# Predict across time range
future_theo <- tibble(
  years_since_1970 = 0:65,
  year = 1970 + 0:65
)

future_theo$pred_conservative <- predict(nls_fit, newdata = future_theo)

ggplot() +
  geom_point(data = theo_model_data, 
             aes(x = year, y = pct_conservative),
             size = 3, color = "#E74C3C") +
  geom_line(data = future_theo, aes(x = year, y = pred_conservative),
            linewidth = 1.2, color = "#1B4F72") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1970, 2035, 5)) +
  annotate("text", x = 1978, y = 0.55, label = "50% threshold", 
           color = "gray50", size = 4, family = "source_sans") +
  labs(
    title = "The S-Curve of Theological Change",
    subtitle = "Logistic growth model fit to % conservative by ordination cohort",
    x = NULL, y = "% Conservative",
    caption = "Nonlinear least squares: logistic growth model"
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.title.position = "plot"
  )
```

The logistic curve tells us the inflection point — where the rate of change was
fastest — occurred around the mid-1990s. The transition is now in its saturation 
phase, approaching an asymptote near 84–85% conservative.

### Model Performance

Let's assess our models with leave-one-out cross-validation (appropriate for 
small time series datasets):

```{r ml-cv}
# LOO-CV for the linear and polynomial models
loo_results <- ml_data |>
  mutate(row_id = row_number()) |>
  rowwise() |>
  mutate(
    lm_pred = {
      train <- ml_data |> slice(-row_id)
      fit <- lm_workflow |> fit(data = train)
      predict(fit, ml_data |> slice(row_id))$.pred
    },
    poly_pred = {
      train <- ml_data |> slice(-row_id)
      fit <- poly_workflow |> fit(data = train)
      predict(fit, ml_data |> slice(row_id))$.pred
    }
  ) |>
  ungroup()

# Calculate RMSE for each model
lm_rmse <- sqrt(mean((loo_results$total_priests - loo_results$lm_pred)^2))
poly_rmse <- sqrt(mean((loo_results$total_priests - loo_results$poly_pred)^2))

tibble(
  Model = c("Linear", "Polynomial (degree 2)"),
  RMSE = c(lm_rmse, poly_rmse),
  `RMSE as % of mean` = c(lm_rmse, poly_rmse) / mean(ml_data$total_priests) * 100
) |>
  knitr::kable(digits = 1)
```

## The Combined Picture

```{r final-viz, fig.width = 8, fig.height = 10}
# Build the final shareable visualization
bg_color <- "white"

tt_caption <- paste0(
 "DataViz: Tony Galvan",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-solid;'>&#xf0ce;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "CARA + NSCP",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-brands;'>&#xf08c;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "anthony-raul-galvan",
 "<span style='color:", bg_color, ";'>..</span>",
 "<span style='font-family:fa-brands;'>&#xf09b;</span>",
 "<span style='color:", bg_color, ";'>.</span>",
 "gdatascience"
)

# Panel 1: Theological inversion (raw counts)
p1 <- ggplot(theological_long, aes(x = year, y = count, color = orientation)) +
  geom_line(linewidth = 1.3) +
  geom_point(size = 2) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1970, 2025, 10)) +
  scale_color_manual(values = c(
    "Conservative" = "#1B4F72",
    "Moderate" = "#27AE60",
    "Progressive" = "#3498DB"
  )) +
  labs(
    subtitle = "Ordinations by Theological Orientation",
    x = NULL, y = "Priests Ordained", color = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.subtitle = element_text(size = 14, face = "bold"),
    legend.position = "top"
  )

# Panel 2: Supply vs demand index
p2 <- ggplot(workforce, aes(x = year, y = index, color = metric)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray60", linewidth = 0.5) +
  geom_line(linewidth = 1.3) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1970, 2025, 10)) +
  scale_color_manual(values = c(
    "Total Priests" = "#C0392B",
    "Parishes" = "#F39C12",
    "Catholic Population" = "#2980B9"
  )) +
  labs(
    subtitle = "Supply vs. Demand (Indexed: 1970 = 100)",
    x = NULL, y = "Index", color = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.subtitle = element_text(size = 14, face = "bold"),
    legend.position = "top"
  )

# Panel 3: Parish consolidation
p3 <- cara_data |>
  filter(!is.na(pct_parishes_no_priest), year >= 1970) |>
  ggplot(aes(x = year, y = pct_parishes_no_priest)) +
  geom_area(fill = "#E74C3C", alpha = 0.3) +
  geom_line(linewidth = 1.3, color = "#E74C3C") +
  geom_point(size = 2, color = "#E74C3C") +
  scale_y_continuous(labels = percent, limits = c(0, 0.25)) +
  scale_x_continuous(breaks = seq(1970, 2025, 10)) +
  labs(
    subtitle = "Parishes Without a Resident Priest",
    x = NULL, y = NULL,
    caption = tt_caption
  ) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.subtitle = element_text(size = 14, face = "bold"),
    plot.caption = element_markdown(size = 9, color = "gray50", hjust = 0.5),
    plot.caption.position = "plot"
  )

# Combine with patchwork
final_plot <- p1 / p2 / p3 +
  plot_annotation(
    title = "The Changing Face of the American Priesthood",
    subtitle = "55 years of theological, demographic, and structural transformation",
    theme = theme(
      plot.title = element_text(size = 24, face = "bold", family = "source_sans", hjust = 0.5),
      plot.subtitle = element_text(size = 16, color = "gray40", family = "source_sans", hjust = 0.5),
      plot.margin = margin(20, 20, 10, 20)
    )
  )

final_plot
```

```{r save-final, include=FALSE}
ggsave(
  filename = "outputs/catholic_priests.png",
  plot = final_plot,
  device = "png",
  width = 8,
  height = 10,
  dpi = 300,
  bg = "white"
)
```

## Key Takeaways

1. **The theological inversion is nearly complete.** In 1970, 56% of new priests 
   were progressive. In 2025, 84% are conservative. In raw numbers, conservative 
   ordinations grew from 137 to 358, while progressive ordinations collapsed from 
   451 to just 9.

2. **The workforce crisis is structural.** With 44% fewer priests serving 38% 
   more Catholics, the workload per priest has roughly doubled. Nearly 1 in 4 
   parishes now lacks a resident priest.

3. **The decline is decelerating but not reversing.** Machine learning models 
   confirm the decline follows a curve, not a straight line — suggesting the 
   steepest losses are behind us, but the priesthood will continue shrinking.

4. **Year is the strongest predictor.** Random forest variable importance shows 
   that time itself (capturing the broad cultural and institutional forces) 
   explains more of the variance than any single Church metric.

## What's Next?

Open questions for further analysis:

- How does the theological shift interact with parish-level outcomes? Are 
  conservative-led parishes growing while progressive-led ones shrink?
- What's the age distribution of the current priesthood, and when does the 
  "retirement cliff" hit hardest?
- How do international priest transfers (from Africa, Asia) change the 
  workforce projections?
- Can the NSCP Wave 2 burnout data predict which priests are likely to leave?

## Data Sources

- **CARA Frequently Requested Church Statistics:** [cara.georgetown.edu](https://cara.georgetown.edu/frequently-requested-church-statistics/) — 
  U.S. Catholic Church data from 1965–2025
- **National Study of Catholic Priests (NSCP):** [The Catholic Project](https://catholicproject.catholic.edu/national-study-of-catholic-priests/) 
  at The Catholic University of America — theological orientation data via 
  [Ryan Burge](https://x.com/ryanburge/status/2046569709009555937)
- **Combined analysis inspired by:** [@SeamusNua](https://x.com/SeamusNua/status/2046581326610895348)
