---
title: "TidyTemplate"
date: 2025-07-15
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-07-08")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

answers <- tt$answers

# Add RGB columns
answers_rgb <- answers |>
  rowwise() |>
  mutate(rgb_values = list(col2rgb(hex))) |>
  ungroup() |>
  mutate(
    red = map_dbl(rgb_values, ~ .["red", ]),
    green = map_dbl(rgb_values, ~ .["green", ]),
    blue = map_dbl(rgb_values, ~ .["blue", ])
  ) |>
  select(-rgb_values)

users_rgb <- tt$users |>
  left_join(answers_rgb, by = "user_id")

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

answers_rgb |>
  sample_n(1000) |>
  ggplot(aes(x = red, y = green, color = hex)) +
  geom_point(aes(size = rank, alpha = 1 / rank), show.legend = TRUE) + # Size by rank, alpha inversely by rank
  scale_color_identity() + # Use actual hex colors
  labs(
    title = "RGB Color Distribution with Rank",
    x = "Red Component (0-255)",
    y = "Green Component (0-255)",
    caption = "Size and opacity scaled by rank (lower rank = larger/more opaque)"
  ) +
  theme_minimal() +
  # Add a Z-axis for Blue by adjusting point shape or a third dimension if using 3D plot (beyond basic ggplot)
  # For 2D, we can indicate blue with another aesthetic or facet
  # geom_text(aes(label = name, y = green + 10), size = 3, color = "black", vjust = 0, check_overlap = TRUE) +
  # You can also use facets for blue component, or map it to shape
  # facet_wrap(~ cut(blue, breaks = 3, labels = c("Low Blue", "Medium Blue", "High Blue")))
  # geom_point(aes(shape = factor(cut(blue, breaks = 3))))
  scale_size_continuous(range = c(0.1, 0.5)) # Adjust size range for better visualization

# create a 3d plot so that we can see the blue component
library(plotly)
answers_rgb |>
  mutate(alpha = 1/rank) |>
  sample_n(1000) |>
  plot_ly(
    x = ~red,
    y = ~green,
    z = ~blue,
    color = ~hex,
    colors = ~hex,
    size = ~rank,
    # alpha = ~alpha,
    type = "scatter3d",
    mode = "markers"
  ) |>
  layout(
    title = "3D RGB Color Distribution with Rank",
    scene = list(
      xaxis = list(title = "Red Component (0-255)"),
      yaxis = list(title = "Green Component (0-255)"),
      zaxis = list(title = "Blue Component (0-255)")
    )
  ) |>
  # hide the legend
  hide_legend()

users_rgb |>
  sample_n(1000) |>
  plot_ly(
    x = ~red,
    y = ~green,
    z = ~blue,
    color = ~colorblind,
    # colors = ~monitor,
    size = ~rank,
    # alpha = 1 / answers_rgb$rank,
    type = "scatter3d",
    mode = "markers"
  ) |>
  layout(
    title = "3D RGB Color Distribution with Rank",
    scene = list(
      xaxis = list(title = "Red Component (0-255)"),
      yaxis = list(title = "Green Component (0-255)"),
      zaxis = list(title = "Blue Component (0-255)")
    )
  )
  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "2025_07_08_tidy_tuesday_colors.png",
  device = "png"
)
```
